---
title: "regen_and_plot_input_2023_cleaning.Rmd"
author: "Peter Smalldige"
date: "`r Sys.Date()`"
output: html_document
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


<!-- README

File Description: The purpose of this file is to:
- input data from the 2023 growing season regeneration plots (code 01, code 02)
    note, the plot dimensions have changed since the 2022 data were collected and processed

- input the plot site characteristics (code xx)
- Distinct from previous growing seasons, there are no seedling height data. 


- input the overstory measurements of dbh and perhaps silvah variables (code xx)


- standardize the names of the harvests (code xx)

- count the number of sample plots and compare to number of veg plots (code xx)

- insert "empty" veg plots where no veg occurred, using spp code "001" (code xx)

- convert stem counts to stems per acre (code 05)
- convert harvest names to consistent usage across years (code 05)
- convert species numeric to alphabetic abbrev (code 05)
- convert untidy to tidy data, alphabetic species (code 06)


exported data files include:
plot_data_2023_untidy.csv (line 237)
regen_per_acre_2023_numeric_point_untidy.csv (line 353)
regen_point_per_acre_2023_alpha_tidy.csv (line 530)

 -->

### Analysis of slash wall regeneation plot data - 2023

###Initialize Libraries

```{r}

#library(ggplot2)
#library(ggpubr)
library(expss)
library(readr)
library(readxl)
library(skimr)
library(tidyverse)  # includes: tidyr, dplyr, haven, readr, readxl, forcats, stringer, ggplot2. See all with tidyverse_packages()
library(knitr)
library(janitor)
#install.packages("openxlsx")
library(openxlsx)

```

#
# ALT plus hypen = <- (within chunk)
# ctrl plus shift plus m = %>% 
# ctrl + ALT + I = insert chunk
# ALT + - = insert assignment operator
# filter rows, select columns
# RMD comment = <!-- comment -->
# 
# 




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### read data from csv file, check variables and data structure

<!-- note to self: in the "read_csv" code, after the file location the code
for 'col..." etc ensures that the specified column is adjusted to correct data type
the file "cheat sheet data-import" has details -->

<!--     -->

### code 01
###Input Regen Data for north, south, rectangle, gas_line (code 01) 

```{r, input regen data from from multiple files and combine rows}

#new raw data at "regen_2023_feb_1_2024(north, south, rectangle, gasline)_final.xlsx"
# the file "feb_1_2024" is the final version
# includes Recknagel North, Recknagel South, Rectangle, and Gas_line

#future input files need to interchange the first two rows, and delete the comments column.
#
# seed01 @ 3.7 ft radius
# seed02, seed03 @ 6 ft radius
# sapl @ 16 ft radius


input_veg01 <-  read_excel("slash.wall.data.2023_1_feb_2024_regen(north, south, rectangle, gasline)_final.xlsx", 
    sheet = "regen", col_types = c("text", 
        "text", "text", "text", "text", "text", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric"), 
    skip = 1) %>% 
            # replace missing values before changing character variables to factor variables
#    filter(!harvest =="recknagel_north") %>% 
    replace_na(list(date = "other", count = 1,
                  sup.seed01 = 0, sup.seed02 = 0, sup.seed03 = 0, sup.sapl = 0,
                  exp.seed01 = 0, exp.seed02 = 0, exp.seed03 = 0, exp.sapl = 0, wall = "99"
                  )) %>%   #need to change missing values to real values while variable is a character type
    mutate(
     location = point,
      origin = "clump", 
      stand = as.factor(stand), 
      season = as.numeric(season),
      status = "vacant"
      ) %>% 
   mutate(stand = recode(stand,
      "recknagel" = "recknagel_south"
      )) %>% 
    select(stand, wall, season, date, point, location, spp, origin, count, everything()) %>% 
   arrange(stand, point, location, spp)


input_veg01$origin[input_veg01$count < 1.1] <- "single" #in "replace_na" above, missing values of "count" changed to 1. Here
#    if the value of count is less than 1.1, then the variable origin is assigned the value of "single"
input_veg01$location[input_veg01$stand == "recknagel_north"] <- "p"
input_veg01$status[input_veg01$stand == "recknagel_north"] <- "protected"

# the above code changes the value of the variable "origin" based on the value of the variable "count"
# if "count" is less than 1, then "origin" is changed to "single"
```


```{r assign treatment class to each point}
location_by_point <- read_excel("location_by_point.xlsx") |> 
    mutate(
        stand = harvest,
        status = wall,
        wall = "99",
        point = as.character(point)
    ) |> 
    select(stand, status, point, location) |> 
     mutate(stand = fct_recode(stand,
          "gas_line" = "gasline",
          "recknagel" = "rectangle1",
          "recknagel" = "rectangle2",
          "recknagel" = "rectangle3",
          "recknagel" = "rectangle4"
   )) |> 
   filter(stand %in% c("gas_line", "campridge", "recknagel")) |> 
   mutate(
   stand = fct_drop(stand)  
   )


```



Merge the veg01 data with the locationxpoint codes

```{r}
walls01 = left_join(input_veg01, location_by_point, by = c("stand", "point")) |> 
    select(!c(location.x, status.x )) |> 
    rename(
        status = status.y,
        location = location.y
    )

```



code 02 - input regen data for campridge and cafri

```{r}


input_veg02 <- read_excel("slash.wall.data.2023.23july2024(maybe FINAL).xlsx", 
    sheet = "regen", col_types = c("text", 
        "text", "numeric", "text", "text", 
        "text", "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "skip"), skip = 2) |> 
    replace_na(list(date = "other", count = 1,
                  sup.seed01 = 0, sup.seed02 = 0, sup.seed03 = 0, sup.sapl = 0,
                  exp.seed01 = 0, exp.seed02 = 0, exp.seed03 = 0, exp.sapl = 0, wall = "99"
                  )) |>   #need to change missing values to real values while variable is a character type 
   mutate(
        origin = "clump",
        location = point, 
        stand = as.factor(stand),
        status = "vacant"
        ) |> 
   select(stand, wall, season, date, point, location, spp, origin, count, everything())
   
   input_veg02$origin[input_veg01$count < 1.1] <- "single" #in "replace_na" above, missing values of "count" changed to 1. Here

```

#read in the cafri categories and treatments for joining to the cafri veg data
```{r}

cafri_label <- read_excel("CAFRI Treatment & Plot Location 8-2024.xlsx", 
    sheet = "merged", skip = 4) |> 
    mutate(
    wall = as.character(wall),
    height = as.character(height),
    overstory = as.character(overstory),
    understory = as.character(understory),
    center_tree = as.character(center_tree),
    point = as.character(point),
    location = "p",
    status = "protected"
    )
    
cafri_label$status[cafri_label$height == "0"] <- "control"
    
```

```{r}
cafri01 = left_join(input_veg02, cafri_label, by = c("stand", "wall", "point")) |> 
    drop_na(height) |> 
    select(!c("status.x", "location.y")) |> 
    rename(
    status = status.y,
    location = location.x) |> 
    mutate(
    location = "p")
  


cafri01$location[cafri01$status == "control"] <- "c"

write.xlsx(cafri01, "raw_data_cafri2023_point_location_regen_trmt_assignment.xlsx")  
    
```

```{r}
campridge = left_join(input_veg02, location_by_point, by = c("stand", "point")) |> 
    filter(wall == "99") |> 
    select(!c("location.x", "status.x")) |> 
    rename(
        status = status.y,
        location = location.y
    )


```

```{r}
## merge the datafiles and retain the regen data

data01 = bind_rows(walls01, cafri01, campridge) |> 
    select(stand, wall, season, date, point, spp, origin, count,
        sup.seed01, sup.seed02, sup.seed03, sup.sapl,
        exp.seed01, exp.seed02, exp.seed03, exp.sapl,
        status, location) |> 
mutate(spp = fct_recode(spp, 
                          "rp" = "125",
                          "wp" = "129",
                          "scp" = "130",
                          "he" = "261",
                          "svb" = "356",
                          "sm" = "310",
                          "stm" = "315",
                          "rm" = "316",
                          "sm" = "318",
                          "ah" = "341",
                          "ab" = "351", # 6-9, wall 3, point 11
                          "svb" = "355",
                          "svb" = "356",
                          "yb" = "371",
                          "sb" = "372",
                          "pb" = "375",
                          #"gb" = "379",
                          "ah" = "391",
                          "bih" = "402",
                          "sh" = "407",
                          "pc" = "461", # campridge, point 483, typo for 761
                          "unk" = "492",
                          "haw" = "500",
                          "ab" = "521", # 6-9, wall 5, point 26, typo as spp 521
                          "ab" = "53",  # 6-9, wall 10, point 80, typo 53 as 531
                          "ab" = "531",
                          "wa" = "541",
                          "wh" = "585",
                          "yp" = "621",
                          "ab" = "631", # 6-9, wall 9, point 478, typo 631 for 531
                          #"cu" = "651",
                          "app" = "660",
                          "la" = "70",
                          "eh" = "701",
                          "unk" = "706",
                          #"asp" = "740",
                          "bta" = "743",
                          "qa" = "746",
                          "pc" = "761",
                          "bc" = "762",
                          "cc" = "763",
                          #"xx6" = "793",
                          "wo" = "802",
                          "co" = "832",
                          "ro" = "833",
                          "smc" = "866",
                          "ns" = "91",
                          "bl" = "901",
                          "wi" = "920",
                          "sa" = "931",
                          "ba" = "951",
                          "ToH" = "998",
                          "unk" = "999",
                          "brs" = "Devil's Walking Stick"
                          #"sp" = "90", 
                          #"au" = "535",
                          #"ec" = "742",
                          ))

write.xlsx(data01, "regen_data_dummy_print_check_species.xlsx")  
```





join dataframes of the early slash walls with the cafri slash walls
seed01 radius = 3.7 ft
seed02 radius = 6.0 ft
seed03 radius = 6.0 ft
sapl radius = 16.0 ft
pi = 3.141593

```{r}

veg2023raw = data01 %>% 
group_by(origin) |> 
summarize(
test_clump = mean(count)
)

mutate(
  sup.seed01ac = sup.seed01 * pi *(3.7^2) * count,
  sup.seed02ac = sup.seed02 * pi *(6^2) * count,
  sup.seed03ac = sup.seed03 * pi *(6^2) * count,
  sup.saplac = sup.sapl * pi *(16^2) * count,
  exp.seed01ac = exp.seed01 * pi *(3.7^2) * count,
  exp.seed02ac = exp.seed02 * pi *(6^2) * count,
  exp.seed03ac = exp.seed03 * pi *(6^2) * count,
  exp.saplac = exp.sapl * pi *(16^2) * count
) %>% 
select(
  !c(sup.seed01, sup.seed02, sup.seed03, sup.sapl,
  exp.seed01, exp.seed02, exp.seed03, exp.sapl)
) %>% 
mutate(clumped = origin) %>% 
mutate(clumped = fct_recode(clumped,
  "1" = "clump",
  "0" = "single"
)) %>% 
mutate(clumped = as.character(clumped)) %>%  #going straight from factor to as.numeric resulted in singles = 2 and clumps = 1
mutate(clumped = as.numeric(clumped)) %>% 
mutate(harvest = fct_recode(harvest,
                              "gasline" = "gas_line"
             )) 



veg2023summ = veg2023raw %>% 
arrange(harvest, wall, point, spp) %>% 
group_by(harvest, wall, point, spp) %>% 
summarize(
  sup.seed01ac = sum(sup.seed01ac),
  sup.seed02ac = sum(sup.seed02ac),
  sup.seed03ac = sum(sup.seed03ac),
  sup.saplac = sum(sup.saplac),
  exp.seed01ac = sum(exp.seed01ac),
  exp.seed02ac = sum(exp.seed02ac),
  exp.seed03ac = sum(exp.seed03ac),
  exp.saplac = sum(exp.saplac),
  clumps = sum(clumped),
  avg_ramet_size = mean(count)  #count is the numnber of stems per rament
) %>% 
mutate(
type = "empty"
)


veg2023summ$type[veg2023summ$spp %in% c("ba", "bc", "bih", "bl", "co", "he", "rm", "ro", "sa", 
"sb", "sh", "sm", "wa", "wo", "wp", "yb", "yp")] <- "commercial"   #n=17

veg2023summ$type[veg2023summ$spp %in% c("ab", "eh", "pc", "stm", "toh" )] <- "interfere"   #n=5

veg2023summ$type[veg2023summ$spp %in% c("app", "bta", "haw", "dws", "la", "ns", "pb", "qa", "rp", "scp", "smc", "svb", "wh", "wi" )] <- "diversity" #n=14

veg2023summ$type[veg2023summ$spp %in% c("ah", "cc", "unk" )] <- "other"  #n=3

#commercial
# "ba", "bc", "bih", "bl", "co", "he", "rm", "ro", "rp", "sa",
# "sb", "sh", "sm", "wa", "wo", "wp", "yb", "yp"

#interfere
#"ab", "eh", "pc", "stm", "toh"

#diversity
#"app", "bta", "haw", "dws", "la", "ns", "pb", "qa", "rp", "scp", "smc", "wh", "wi"

#other
#"ah", "cc", "unk"

```

```{r}

#code to export the point labels for the 2023 growing season veg data
#an update to "location_by_point.xlsx" file
# use the location_by_point file to update the "location" for the early slash walls


old_wall_points <- read_excel("points2023excel.xlsx", 
    col_types = c("text", "text", "text", 
        "text", "text")) %>% 
        mutate(harvest = fct_recode(harvest,
                              "recknagel_south" = "recknagel"
             )) 

        
cafri_trmt_points_2023 <- read_excel("cafri_trmt_points_2023.xlsx", 
    col_types = c("text", "text", "text", 
        "text", "text"))

points2023 <- bind_rows(old_wall_points, cafri_trmt_points_2023)



```


```{r}


```

<!-- comment 


-->

    # Full join created duplicate variables of "location.x" and "location.y"


```{r}
#need to double check this code (15 july 2024)
veg_data1a <- full_join(veg2023summ, points2023, by = (c("harvest", "wall", "point"))) %>% 
    select(!"location.x") %>% 
    drop_na(season) %>%    # this removes harvests without data collected in 2023
    rename("location" = "location.y") %>% 
    mutate(
    cut = 0, 
    year = as.numeric(season)
          ) %>% 
    arrange(harvest, point, spp) %>% 
    select(harvest, season, cut, year, date, point, location, wall, everything())
   
veg_data1a$cut[veg_data1a$harvest %in% c("gasline")] <- 2017
veg_data1a$cut[veg_data1a$harvest %in% c("campridge", "recknagel_south", "recknagel_north")] <- 2019
veg_data1a$cut[veg_data1a$harvest %in% c("6_6")] <- 2021
veg_data1a$cut[veg_data1a$harvest %in% c("6_9")] <- 2021
# veg_data1a$wall[veg_data1a$harvest %in% c("recknagel_north")] = "protected"
# veg_data1a$location[veg_data1a$harvest %in% c("recknagel_north")] = "perimeter"


veg_data1 <- veg_data1a %>% 
      mutate(
          numb_yrs_grow = year - cut
          ) %>% 
    mutate(location = recode(location,
    "c" = "control", "i" = "interior", "p" = "perimeter"
    )) %>% 
    select(harvest, season, cut, year, date, numb_yrs_grow, point, location, wall, everything())

  
  
```

<!-- 
The next chunk reads the site level data. Currently this is set to 
eval = false.  Change at some future point in time
-->

```{r, , include=T, message = F, eval = FALSE}

plot_data01  <- read_excel("~/R/slash-wall-vegetation/slash.wall.data.2022.16may2023_final.xlsx", 
    sheet = "site-plot", col_types = c("text", 
        "text", "text", "text", "numeric", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "numeric", "skip"), skip = 6) %>% 
        select(harvest, season, date, location, point, comm_sdl) %>% 
  mutate(
    location = as.character(location),
    point = as.character(point)
) %>% 
  mutate(location = point) %>% 
  mutate(harvest = fct_recode(harvest,
                              "gas_line" = "gasline",
                              "recknagel" = "rectangle1",
                              "recknagel" = "rectangle2",
                              "recknagel" = "rectangle3",
                              "recknagel" = "rectangle4",
                              "circle" = "circle1",
                              "circle" = "circle2",
                              "circle" = "circle3",
                              "sta_rd" = "boot",
                              "sta_rd" = "wedge"
  )) %>% 
  arrange(harvest, point)
  
  
  plot_data <- full_join(plot_data01, point_location, by = (c("harvest", "point"))) %>% 
      select(!"location.x") %>% 
      rename("location" = "location.y") %>% 
      arrange(harvest, point, spp) %>% 
      mutate(
      season = as.numeric(season)
)
      # select(harvest, date, season, point, location, spp, origin, comm_sdl)

# the missing values in df=plot_data are created because some plot data haven't been entered ("plot_data")
# but there are data for the location assignment of the point ("point_location"), In full_join
# all entries are included.

#write_csv(plot_data, "plot_data_2022_untidy.csv")

```



<!--


-->




#####
#####
begin code review, below. Above code created 30 october 2023
#####
#####


### code 03
###Count Number of Sample Points (code 03)

```{r, count number of sample points}

reck_north = input_veg %>% 
  select(harvest, point) %>% 
  filter(harvest == "recknagel_north") %>% 
  group_by(harvest, point)%>% 
  slice_head(n=1)%>% 
  mutate(
  location = "perimeter",
  wall = "protected"
  )


# create variable for total plots for denominator of average density
point_count_01 <-  bind_rows(point_location, reck_north) %>% 
  mutate(
  n = 1
    ) %>% 
  mutate(location = recode(location,
    "c" = "control", "i" = "interior", "p" = "perimeter"
    )) %>% 
  arrange(point) %>% 
  select(harvest, location, point, wall, n)


point_count_01$wall[point_count_01$location %in% c("interior", "perimeter")] = "protected"


# counts the number of points by each location: interior, perimeter, control
point_count_location <- point_count_01 %>%    #denominator for harvest x location, number of plots
  group_by(harvest, location) %>% 
  summarize(
    location_denom = sum(n))

# counts the number of points by inside the wall ("protected") vs. outside the wall ("control")
point_count_wall = point_count_01 %>%
    group_by(harvest, wall) %>% 
    summarize(
    wall_denom = sum(n)
    )


write_csv(point_count_location, "count_of_points_harvest_x_location.csv")
write_csv(point_count_wall, "count_of_points_harvest_x_wall.csv")


# for each inventory point, the value of "n" = 1
```


### code 04
###Create plots for regen with no observations (code 04)

```{r, filling out empty sample points, eval=FALSE}

#There are 486 plots locations in 2020. 

#The missing points would cause an error for calculation
#of the number of stems per acre as an average of points.
#
#  IGNORE IGNORE
#  code 04 not needed. Expansion of veg_data occurred at prior step. 486 unique plots
#  there are now no missing point
#  assess using console "unique(veg_data$point)
#  IGNORE dataframe = site_data, rather use dataframe = "veg_data"

site_data01 <- full_join(point_location, veg_data, by = "point") %>% 
  select(!c(harvest.y, location.y)) %>% 
  rename(harvest = harvest.x, location = location.x)   

site_data <- site_data01 %>% 
  # select(!c(season.y, date.y, location.y, spp.y, origin.y, comm_sdl.y)) %>% 
  # rename( location = location.x, date = date.x, season = season.x, spp = spp.x, comm_sdl = comm_sdl.x, origin = origin.x) %>% 
  replace_na(list(date = "other",  origin = "u", spp = "001",
                  sup.seed01 = 0, sup.seed02 = 0, sup.seed03 = 0, sup.sapl = 0,
                  exp.seed01 = 0, exp.seed02 = 0, exp.seed03 = 0, exp.sapl = 0,
                  top4 = 0, comm_sdl = 0
                  )) %>%    #need to change missing values to real values with variable as character
    mutate(location = recode(location,
    "c" = "control", "i" = "interior", "p" = "perimeter"
    ))  #"old value" = "new value"


#renames a variable label from new = old

# 2022 data
# confirmed that all plot numbers are unique
#date = character
#spp, origin, location and point are all character
#all other = numeric


```

```{r, eval=FALSE}

### this code is not needed, do not run

# inside <- veg_data %>% 
#     filter(location %in% c("interior", "perimeter")) %>% 
#     droplevels %>% 
#     mutate(
#       wall = "protected"
#     )
# 
# outside <- veg_data %>% 
#     filter(location %in% c("control")) %>% 
#     droplevels %>% 
#     mutate(
#       wall = "control"
#     )

```


### code 05
###data manipulations - per acre calculations & spp names
### output csv files of untidy data with spp numeric and spp alphabetic
```{r, regen per acre at points - calculations}
#
#height classes: 
#   01 = 4" to 12", within 3.7 ft radius
#   02 = 12" to 5 ft, within 6 ft radius
#   03 = 5 ft to 1", within 6 ft radius
#   sapl = 1" - 5" dbh, within 16 ft radius
  #following code calculates the number of seedlings per acre
  #by species for each height class. Not yet tidy.


regen_per_acre <- veg_data1 %>%
  mutate(  #convert plot count of stems to stems/acre
    sup_short = sup.seed01 * count * (43560 /((pi) * (3.7 * 3.7))),
    sup_med = sup.seed02 * count * (43560 /((pi) * (6 * 6))),
    sup_tall = sup.seed03 * count * (43560 /((pi) * (6 * 6))),
    sup_sapl = sup.sapl * count * (43560 /((pi) * (16 * 16))),
    exp_short = exp.seed01 * count * (43560 /((pi) * (3.7 * 3.7))),
    exp_med = exp.seed02 * count * (43560 /((pi) * (6 * 6))),
    exp_tall = exp.seed03 * count * (43560 /((pi) * (6 * 6))),
    exp_sapl = exp.sapl * count * (43560 /((pi) * (16 * 16))),
    ) %>% 
  mutate(
    suppressed = sup_short + sup_med + sup_tall + sup_sapl,
    exposed = exp_short + exp_med + exp_tall + exp_sapl,
    total = suppressed + exposed
    ) %>% 
    select(!c(sup.seed01, sup.seed02, sup.seed03, sup.sapl,
            exp.seed01, exp.seed02, exp.seed03, exp.sapl,
            date))

write_csv(regen_per_acre, "regen_per_acre_2023_numeric_point_untidy.csv")

  
  regen_alpha <- regen_per_acre %>% 
   mutate(spp = fct_recode(spp,   #comment out species not present in data
                         # "empty" = "1",
                          # "empty" = "001",
                          # "ohw" = "004",
                          "ab" = "531",
                          "stm" = "315",
                          "sm" = "318",
                          "rm" = "316",
                          "rp" = "125",
                          # "wp" = "129",
                          # "he" = "261",
                          "svb" = "356",
                          "svb" = "355",
                          # "yb" = "371",
                          "sb" = "372",
                          # "pb" = "375",
                          #"gb" = "379",
                          "ah" = "391",
                          # "haw" = "500",
                          "wa" = "541",
                          # "yp" = "621",
                          #"cu" = "651",
                          # "app" = "660",
                          # "la" = "70",
                          "eh" = "701",
                          #"asp" = "740",
                          "bta" = "743",
                          "qa" = "746",
                          "pc" = "761",
                          "bc" = "762",
                          # "cc" = "763",
                          # "wo" = "802",
                          # "co" = "832",
                          "ro" = "833",
                          "smc" = "866",
                          #"bl" = "901",
                          "wi" = "920",
                          "ba" = "951",
                          #"sp" = "90", 
                          # "sa" = "931",
                          # "wh" = "585",
                          # "sh" = "407",
                          #"ns" = "91",
                           "bh" = "402",
                          #"au" = "535",
                          #"ec" = "742",
                          #"unk" = "999",
                          
#recode data entry errors (avoids repeated fixes with data file updates)
# These numeric codes were entered, and because the raw data file is repeatedly updated, the
# errors would have to be correct each time. This lacks finesse but is more effecient.
                          # "sb" = "273",
                          # "ab" = "331",
                          # "ab" = "351"
                          ))

table(regen_alpha$spp)
table(regen_alpha$harvest)
#view(regen.per.acre)

write_csv(regen_alpha, "regen_per_acre_2023_alpha_point_untidy.csv")





```


### code 06



### making data tidy - comm/regen
```{r, making data tidy commercial species}
# commercial regeneration species
# 316 318 372 541 621 762  802 832 833 951 
# 129, 261, 371,407, 402
# maple, yeb, swb, wa, yp, oaks, ba , bh, sh
# no populus 740, 743, 746, no pab 375


comm <- regen_alpha %>% 
  
  filter(spp %in% c("wp", "he", "rm", "sm","yb", "sb", "wa", "yp",
                     "bc", "wo", "co", "ro", "ba", "bl", "rp", "cu", "sh", "bh", "ohw"
                    )) %>% 
  mutate(
    type = as.factor("commercial")
  ) %>% 
  pivot_longer(c( sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))


```

###making data tidy - diverse/regen
```{r, making data tidy diverse}

# desirable diversity species (no interfering, no commercial, but includes Populus, paper birch, not pin cherry)
# 129 261 316 318 356 372 375 500 541 621 743 746 762 802 832 833  90 920 951, 407, 931

diverse <- regen_alpha %>% 
  
  filter(spp %in% c("svb", "pb", "gb", "haw", "asp", "bta", "qa",
                    "wi", "sh", "wh","sa","ec", "smc"
                    )) %>% 
  #

  mutate(
    type = as.factor("diversity")
  ) %>% 
  pivot_longer(c(sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))


```

###making data tidy - interfering species/regen
```{r, tidy for interfering species}

# interfering species

interfere <- regen_alpha %>% 
  
  filter(spp %in% c("ab", "stm", "eh", "pc", "au")) %>% 
  mutate(
    type = as.factor("interfering")
  ) %>% 
  pivot_longer(c(sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))


```

###making data tidy - other species
```{r}
# other species

other <- regen_alpha %>% 
  
  filter(spp %in% c("empty", "ah", "la", "sp", "unk", "cc", "ns"
                    )) %>% 
  mutate(
    type = as.factor("other")
  ) %>% 
  pivot_longer(c(sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))


```

 

###Write tidy data for each point
```{r, tidy regen and plot data, assign stocking label to each point}

tidy_regen_point <- 
  bind_rows(comm, diverse, interfere, other)
 
 
write_csv(tidy_regen_point, "regen_point_per_acre_2023_alpha_tidy.csv")

```




###Make Plot Data Tidy



```{r}

# not ready to run this code
# not ready to run this code
# not ready to run this code
# not ready to run this code

# 
plot_data_tidy <- #read the csv file having basal area calculations %>% 

# (need to input the calculated values of plot variables 
#  such as basal area per acre.  
  
  select(!c(baf,
            low_spp, inv_spp, mid_spp, upper_spp, vine_spp)) %>% 
  
  pivot_longer(c(disturb, slash, fern, herb, 
                 low_cov, inv_cov, mid_cov, upper_cov, vine_cov,
                  comm_stock, pre_comm_stock, diver_stock,
                  pre_diver_stock,  interf_woody_stock, fern_stock, all_intf,
                 winner_stock
                 ),
               names_to = "plot_attrib",
               values_to = "attrib_val") %>% 
  
  replace_na(list("attrib_val" = "no", comm_per_acre = 0, pre_comm_per_acre = 0,
                  diver_per_acre = 0, pre_diver_per_acre = 0, interfere_per_acre = 0,
                  pre_interfere_per_acre = 0))

write_csv(plot.data_tidy, "tidy_plot_2019.csv")




```



<!-- 
 includes factor and number variables...can't blend via "pivot_longer"
 pivot_longer(c(disturb, slash, fern, herb, 
                 low_cov, inv_cov, mid_cov, upper_cov, vine_cov,
                 ba, seedling_numb, comm_per_acre, comm_stock,
                 pre_comm_per_acre, pre_comm_stock,diver_per_acre, diver_stock,
                 pre_diver_per_acre, pre_diver_stock, interfere_per_acre, interfere_stock,
                 pre_interfere_per_acre, pre_interfere_stock
                 ),
               names_to = "plot_attrib",
               values_to = "attrib_val")
-->
