---
title: "Slash Wall Plot Analysis"
author: "Peter Smallidge"
date: "10/14/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# code to make variables into factors
  mutate_at(
    vars(harvest, season, location, point, spp, origin),
    list(factor)) %>%
    
# code to re-label the levels of a factor, newname=oldname
   mutate(harvest = fct_recode(harvest,
                             "gas_line" = "gl",
                              "red_pine" = "rp",
                              "sta_rd" = "wb

```{r, include=FALSE, message=FALSE}
# Library: dplyr, expss, ggplot2, ggpubr, labeling, plyr,
#library(ggplot2)
#library(ggpubr)
library(expss)
library(readr)
library(readxl)
library(writexl)
library(skimr)
library(tidyverse)  # includes: tidyr, dplyr, haven, readr, readxl, forcats, stringer, ggplot2. See all with tidyverse_packages()
library(knitr)
library(janitor)
#install.packages("openxlsx")
library(openxlsx)
# install.packages("shinyaddins")
# library(shinyaddins)

```


#
# ALT plus hypen = <- (within chunk)
# ctrl plus shift plus m = %>% 
# ctrl + ALT + I = insert chunk
# filter rows, select columns
# 


```

<!-- Comments 




-->



#file paths, input and output
```{r}
#Input data files
file_path_site20192023 <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\site_plot_data_2019_2023(19feb2025)(readme).xlsx"
file_path_BA2020 <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\site_summary_2020_basal_area.xlsx"
file_path_cafriBA2023 <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\overstory dbh data for 6-9 and 6-6 CAFRI_2023_oct_30.xlsx"

#Output data files
output_folder <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\"


# sample input code
#   raw2019 <- read_excel(file_path_2019, sheet = "seedling", skip = 0 ) %>% 
# 
#sample output code
#   output_file <- file.path(output_folder, paste0("site_plot_data_2019_2023(19feb2025)", ".xlsx"))
#   write_xlsx(site_data, output_file)




```


```{r}
#site_basal <- read_excel(file_path_site20192023, sheet = "19feb2025", skip = 0)
```

 










##
##
## old code below
```{r}
overstory_raw <- read_excel("site_2023_march_19_2024 (Recknagel and Gas Line).xlsx", 
    sheet = "baf10", skip = 1) %>% 
    slice(-2) %>% 
  rename (
    spp = species,
    harvest = site
          ) %>% 
  mutate(
    TRMT = "missing",
    sprout = 0,
    spp = as.character(spp),
    ba = 10
  ) %>% 
     mutate(spp = fct_recode(spp,   #comment out species not present in data
                          #"ab" = "531",
                          "sm" = "318",
                          "rm" = "316",
                          # "wp" = "129",
                          # "he" = "261",
                          # "yb" = "371",
                          #"sb" = "372",
                          #"wa" = "541",
                          # "yp" = "621",
                          #"cu" = "651",
                          #"bta" = "743",
                          #"qa" = "746",
                          "bc" = "762",
                          # "wo" = "802",
                          # "co" = "832",
                          "ro" = "833",
                          #"bl" = "901",
                          "ba" = "951",
                          # "sa" = "931",
                          # "sh" = "407",
                          # "bh" = "402",
                          )) %>% 
  mutate(
    tpa = 43560 / (((2.75 * dbh2023) * (2.75 * dbh2023)) * pi),
    vol_int = ((1.52968 * (numb_logs * numb_logs)) + (9.58615 * numb_logs) - 13.35212) +
          ((1.79620 - (0.27465 * (numb_logs * numb_logs))  - 2.59995* (numb_logs)) * dbh) +
          ((0.04482 - (0.00961 * (numb_logs * numb_logs)) + (0.44957 * numb_logs) ) * (dbh * dbh))
    # formula INT 1/4" from Purdue fact sheet, confirmed 11/28/2023, slightly lower than the table 6 values
  )

overstory_raw$TRMT[overstory_raw$harvest %in% c("recknagel_north")] = "protected"
overstory_raw$sprout[overstory_raw$wi_16ft %in% c("y")] = 1

```




```{r}
overstory_points <- overstory_raw %>% 
  group_by(harvest, TRMT) %>% 
    summarize(
      point_count = n_unique(point)
    )

overstory_summ <- left_join(overstory_raw, overstory_points, by = c("harvest", "TRMT")) %>% 
  group_by(harvest, TRMT, spp) %>% 
  summarise(
    point_count = mean(point_count),
    sprouts_ac = sum(sprout) / point_count,
    ba_a = sum(ba)/ point_count,
    tpa = sum(tpa)/ point_count,
    INT_vol_ac = sum(vol_int) / point_count
  ) %>% 
  mutate(
    dbh_q = sqrt((ba_a/tpa)/0.005454),
    spp = as.character(spp)
  ) %>% 
   mutate_if(is.numeric, round, 1) %>% 
  select(harvest, TRMT, spp, point_count, dbh_q, ba_a, tpa, INT_vol_ac, sprouts_ac) %>% 
  arrange(harvest, TRMT, spp) # spp wouldn't sort as a factor, so changed to character


```



```{r}


plot_data_2020_untidy <- read_csv("plot_data_2020_untidy.csv", 
    col_types = cols(invasive_spp = col_character(), 
        mid_spp = col_character(), upper_spp = col_character()))
  
  
  #calculate basal area per acre
  plot_2020 <- plot_data_2020_untidy %>% 
      replace_na(list(oakhick = 0, beech = 0, maple = 0, birch = 0, hdwd = 0, conifer = 0
                  )) %>%   #need to change missing values to real values 
      mutate(
        ba_acre = baf * (oakhick + beech + maple + birch + hdwd + conifer)
        )
  
  cut2017 <- plot_2020 %>% 
    filter(harvest %in% c("boot", "gas_line", "red_pine", "wedge", "sta_rd")) %>% 
    mutate(
      cut = 2017,
      year = season,
      numb_yrs_grow = year - cut +1
    )
  
  cut2019 <- plot_2020 %>% 
    filter(harvest %in% c("campridge", "circle", 
                          "decker","recknagel")) %>% 
      mutate(
      cut = 2019,
      year = season,
      numb_yrs_grow = year - cut +1
      )

  
plot_summary_2020 <- 
  bind_rows(cut2017, cut2019) %>% 

  
  arrange(harvest, location) %>% 
  group_by(harvest,location) 

```


<br>

### Subset data, group and summarize for harvest x location

```{r}
# Sort and group data by site (harvest name + interior vs. control)
# https://cmdlinetips.com/2020/08/dplyr-groupby-and-summarize-group-by-one-or-more-variables/ 
# Pat Sullivan suggested adding "as.data.frame()" which worked for him but not for me.
# Pat may have been using a different version of dyplr or maybe base R.
# Nina Thirkildsen suggested adding "dplyr::" before each verb. Including the package name
# is a way to ensure that it is the package's version of the verb that is used, 
# not a different package's verb.
# 
# UPDATE
# *** the problem was loading too many libraries with conflicts. 
# Currently just load tidyverse and skimr




site <- plot_summary_2020 %>% 
 #filter(location =="interior" | location=="control") %>% 
  select(harvest, location, ba_acre, cut, numb_yrs_grow) %>% 
  mutate(harvest = recode(
    harvest, "boot" = "sta_rd", "wedge" = "sta_rd", "wedge_boot" = "sta_rd",
    "rect1" = "campridge", "rect2" = "campridge", "rect3" = "campridge", "rect4" = "campridge",
    "circle1" = "circle", "circle2" = "circle", "circle3" = "circle")) %>% 
  group_by(harvest, location) %>% 
  summarize(ba_sqft = mean(ba_acre),
            harvested = mean(cut),
            n = n(),
            Number_Growing_Seasons = mean(numb_yrs_grow)
            ) %>% 
  mutate(
    ba_sqm = ba_sqft * 0.22956) %>% 
  arrange(harvest, location)
  
#site
#write.table(site, "site_BA_means.txt", sep="\t")
write_csv(plot_summary_2020, "plot_calc_2020_untidy.csv")
write_csv(site, "site_summary_2020_basal_area.csv")
# (df not found)   write_csv(tidy_regen_point, "regen_point_per_acre_2020_alpha_tidy.csv")

```
mutate(mid_spp = recode(mid_spp, "swm" = "rhus")) %>%  #old = new


```{r}

#write.table(site, "site_BA_means.txt", sep="\t")

```



```{r}
# site %>% 
# pivot_wider(names_from = location,
#             values_from = basal_area)
  
#view(ba_summ)

```

