str---
title: "plot summary for all years"
author: "Peter Smallidge"
date: "3/9/2023"
output: github_document
```

<!-- README

File Description: The purpose of this file is to:
- input data from the 2019-2022 growing season site-level plot data (code 01)

plot-data-2019.csv
plot_data_2020_untidy.csv
plot_data_2021_untidy(revised).csv
plot_data_2022_untidy.csv



exported data files include:


 -->




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### input of slash wall site plot data 

###Initialize Libraries
```{r, cars, include=FALSE}


#library(ggplot2)
#library(ggpubr)
library(readr)
library(dplyr)
library(readxl)
#library(skimr)
library(tidyr)
library(tidyverse)

#library(readr)
#library(skimr)
#library(tidytable)  ## CAUTION this may conflict with dplyr functions
library(ggplot2)
library(ggpubr)


#
# ALT plus hypen = <- (within chunk)
# ctrl plus shift plus m = %>% 
# ctrl + ALT + I = insert chunk
# ALT + - = insert assignment operator
# filter rows, select columns
# RMD comment = <!-- comment -->
# 
# 

```


### read data from csv file, check variables and data structure

<!-- note to self: in the "read_csv" code, after the file location the code
for 'col..." etc ensures that the specified column is adjusted to correct data type
the file "cheat sheet data-import" has details -->

### code 01
###Input Plot Data (code 01)


```{r}

plot2019 <- read_excel("~/R/slash-wall-vegetation/old files but do not discard/plot-data-2019.xlsx") %>% 
    mutate(
    mid_spp = as.character(mid_spp),
    upper_spp = as.character(upper_spp),
    inv_spp = as.character(inv_spp)
    ) %>% 
    rename(
    invasive_spp = inv_spp
    ) %>% 
    mutate(
    spp = "004",
    origin = "s",
#    season = 2019
    )
    
plot2020 <- read_csv("~/R/slash-wall-vegetation/old files but do not discard/plot_calc_2020_untidy.csv") %>% 
    mutate(
    mid_spp = as.character(mid_spp),
    upper_spp = as.character(upper_spp),
    invasive_spp = as.character(invasive_spp)
    ) %>% 
    mutate(
    spp = "004",
    origin = "s"
#    season = 2020
    )
    
    
plot2021 <- read_csv("~/R/slash-wall-vegetation/old files but do not discard/plot_data_2021_untidy(revised).csv") %>% 
    mutate(
    mid_spp = as.character(mid_spp),
    upper_spp = as.character(upper_spp),
    invasive_spp = as.character(invasive_spp)
    ) %>% 
    mutate(
    spp = "004",
    origin = "s",
#    season = 2021
    )
    
    
plot2022 <- read_csv("~/R/slash-wall-vegetation/old files but do not discard/plot_data_2022_untidy.csv") 

plotdata_all <- bind_rows(plot2019, plot2020, plot2021, plot2022) %>% 
           mutate(harvest = fct_recode(harvest,
                             "gas_line" = "gl",
                              "red_pine" = "rp",
                              #"sta_rd" = "stationrd",
                              #"recknagel" = "rectangle1",
                              #"recknagel" = "rectangle2",
                              #"recknagel" = "rectangle3",
                              #"recknagel" = "rectangle4",
                              #"circle" = "circle1",
                              #"circle" = "circle2",
                              #"circle" = "circle3",
                              "sta_rd" = "boot",
                              "sta_rd" = "wedge",
                              "red_pine" = "crp",
                              "gas_line" = "cgl",
                              "sta_rd" = "cwb"
       ),
       point = as.character(point),
       season = as.character(season)
       ) %>% 
       filter(!(harvest == "gas_line" & point %in% c(866, 864)))  %>%  #corrected code selects for either point at the intersection of harvest and point
        mutate(
            location = case_when(
            location == "perimeter" ~ "p",
            location == "interior" ~ "i",
            location == "control" ~ "c",
      TRUE ~ location  # Keeps other values unchanged
        )
        ) %>% 
        mutate(
        filename = "plot_summary_all_years.Rmd",
        df = "plotdata_final"
)
  

location_by_point <- read_excel("~/R/slash-wall-vegetation/summary output/point_labels_all(readme).xlsx")
          


plotdata_final <- left_join(plotdata_all, location_by_point, by = (c("harvest", "point", "location"))) %>% 
    mutate(
        year = as.numeric(season),
        firstgrow = 2035, # default value for the first growing season
        firstgrow = case_when(
          harvest %in% c("gas_line", "red_pine") ~ 2017,
          harvest %in% c("sta_rd") ~ 2018,
          harvest %in% c("campridge", "recknagel", "recknagel_north", "circle", "decker") ~ 2021,
          harvest %in% c("6_6", "6_9") ~ 2022,
          TRUE ~ firstgrow  # Keeps the existing value of "firstgrow" for all other cases
    ),
        numb_yrs_grow = (year - firstgrow) + 1
     ) %>% 
     mutate(
        cut = 2035, # default value for the first growing season
        cut = case_when(
          harvest %in% c("gas_line", "red_pine", "sta_rd") ~ 2017,
          harvest %in% c("campridge","circle") ~ 2019,
          harvest %in% c("recknagel", "recknagel_north","decker") ~ 2020,
          harvest %in% c("6_6", "6_9") ~ 2022,
          TRUE ~ cut  # Keeps the existing value of "firstgrow" for all other cases
    ))


    

write_csv(plotdata_final, "plotdata_all_years(revised).csv")


```


