---
title: "Regen_Tables_2025"
author: "Peter Smalldige"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# activate library functions
```{r message=FALSE}

library(ggplot2)
library(ggpattern)
#library(ggpubr)
library(expss)
library(readr)
library(readxl)
library(writexl)
library(skimr)
library(tidyverse)  # includes: tidyr, dplyr, haven, readr, readxl, forcats, stringer, ggplot2. See all with tidyverse_packages()
library(knitr)
library(janitor)
#install.packages("openxlsx")
library(openxlsx)
# install.packages("shinyaddins")
# library(shinyaddins)

```


### Create file path input and the output folder (code 01) 

file paths, input and output
```{r}
#complete data
file_path_veg_complete <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\veg_data_2019to2023_complete(19feb2025)(read_me).xlsx"

file_path_spp_harvest <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\spp_x_harvest_x_trmt_x_season(5march2025).xlsx"

file_path_denom <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\sample_size_table(16feb2025).xlsx"


output_folder <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\"
```


# read denominators, but they mostly aren't used
```{r}
denom <- read_excel(file_path_denom)

harvest_summ <- read_excel(file_path_spp_harvest)

```



```{r}
species <- read_excel(file_path_spp_harvest) %>% 
    select(harvest, season, trmt, growingseasons, spp, type, pct_points_any, pct_points_GTR1ft, pct_points_GTR5ft,
           all_sizes, total, GTR5ft, firstgrow, year, wall_count, stems.ramet, denom, height)
    
## Step 1: Select top 10 species within each group
## there are no calculations here, just sorting the top five spp for each group_by
top_spp <- species %>%  #stem count variables
    select(harvest, season, trmt, growingseasons, spp, type, all_sizes, total, GTR5ft,pct_points_any, pct_points_GTR1ft, pct_points_GTR5ft, 
           firstgrow, year, wall_count, denom, height, stems.ramet) %>% 
    group_by(season, harvest, trmt) %>%
    arrange(desc(total), .by_group = TRUE) %>%
    slice_head(n = 10) %>%
    ungroup()

# Step 2: Calculate remainder for all other species
remainder <- species_by_harvest %>%
    select(harvest, season, trmt, growingseasons, spp, type, all_sizes, total, GTR5ft,pct_points_any, pct_points_GTR1ft, pct_points_GTR5ft, 
           firstgrow, year, wall_count, denom, height, stems.ramet) %>% 
  group_by(season, harvest, trmt) %>%
  filter(!spp %in% top_spp$spp) %>%  # Exclude top 5 spp
  summarize(
    wall_count = first(wall_count),
    spp = "Remainder",  # Label for remaining species
    across(where(is.numeric), sum, na.rm = TRUE),  # Sum numeric variables for the total remainder
    #(average?)  stems.ramet = mean(stems.ramet)
    .groups = "drop"
  )

# Step 3: Calculate total (sum of top 5 + remainder)
total_row <- bind_rows(top_spp, remainder) %>%
    select(harvest, season, trmt, growingseasons, spp, type, all_sizes, total, GTR5ft,pct_points_any, pct_points_GTR1ft, pct_points_GTR5ft, 
           firstgrow, year, wall_count, denom, height, stems.ramet) %>% 
  group_by(season, harvest, trmt) %>%
  summarize(
    wall_count = first(wall_count),
    spp = "TOTAL",  # Label for total
    across(where(is.numeric), sum, na.rm = TRUE),
    .groups = "drop"
  )

# Step 4: Combine top 5 species, remainder,  total and add a blank row.
# use a dummy variable to force the sort given there is no existing variable that
# would sort ascending or descending.

# Combine and sort Remainder and TOTAL first
remain_total <- bind_rows(remainder, total_row) %>%
  arrange(season, harvest, trmt, spp) %>%   # Sort by spp to ensure Remainder comes before TOTAL
  # mutate(spp = case_when(
  #       spp %in% c("TOTAL") ~ "TOTAL", # see step 3 above where spp defined as "Total" but could change to "TOTAL"
  #       TRUE ~ spp,  # Keeps the existing value of "spp" for all other cases
  #   )) %>% 
  mutate(dummy = case_when(
        spp == "Remainder" ~ 11,
        spp == "TOTAL" ~ 12,
        TRUE ~ NA_real_  # Assign NA for other cases
  ))

# Assign dummy values to top_spp
top_spp <- top_spp %>%
  group_by(season, harvest, trmt) %>%
  mutate(dummy = row_number()) %>%  # Assign 1 to 10
  ungroup()

blank_rows <- top_spp %>%
  group_by(season, harvest, trmt) %>%
  summarize(spp = "", total = NA, dummy = 13, .groups = "drop")

# Now combine with top_spp
final_result <- bind_rows(top_spp, remain_total, blank_rows) %>%
  arrange(harvest, season, trmt, dummy) %>%  # Sort by dummy for correct order
  mutate(
    date = "5 March 2025",
      filename = "regen_tables_2025.Rmd",
      df = "species",
    across(
      c(wall_count, growingseasons, stems.ramet, firstgrow, year, denom, height, 
        pct_points_any, pct_points_GTR1ft, pct_points_GTR5ft),
      ~ if_else(dummy %in% c(11, 12), NA_real_, .x)  # Correct syntax
    )
  ) %>%  
 rename(
     GTR1ft = total
 ) %>% 
      select(dummy, filename,	df	,date ,	harvest,	trmt,	season,growingseasons,	spp,	type,	all_sizes,	GTR1ft,	GTR5ft,	pct_points_any	,
             pct_points_GTR1ft	,pct_points_GTR5ft,	firstgrow,	year,	denom,	height,	stems.ramet,    wall_count
) %>% 
    arrange(harvest, desc(season), trmt, dummy)
```


```{r}
output_file <- file.path(output_folder, paste0("top_spp_by_harvest_x_trmt_x_season(5march2025)", ".xlsx"))
write_xlsx(final_result, output_file)

```




#####
#####  Assumes data are tidy  
#####  Code taken from "summary_regen_veg_data.Rmd" line 371
regen_exp_GT_4.5ft_spp_initial <- regen_all %>% 
  filter(tpa>0) %>% 
  filter(size == "exp_med" | size == "exp_tall" | size == "exp_sapl") %>% 
  mutate(
    describ = "exp_GT_4.5ft"  #labeling data to be clear about the variables selected
  ) %>% 
  group_by(describ, harvest, season, TRMT, spp) %>% 
    summarize(
      tpa = sum(tpa_x_wall),  # the average tpa for each spp is aggregated across points
      tpha = sum(tpha_x_wall),
      wall_denom = mean(wall_denom),
      points_occupied = n_distinct(point),
      age = mean(numb_yrs_grow)
    ) %>% 
  mutate(
    PCT_pts_present = (points_occupied / wall_denom )*100
     ) %>% 
      arrange(harvest, season, TRMT, (desc(tpa)))

  regen_exp_GT_4.5ft_head <- regen_exp_GT_4.5ft_spp_initial %>% 
    slice_head(n=5) %>% 
    mutate(spp_count = 1)
  
  regen_exp_GT_4.5ft_tail <- regen_exp_GT_4.5ft_spp_initial %>% 
    slice_tail(n=-5) %>% 
    filter(tpa >0) %>% 
    mutate(
      spp = "remainder",
      numb = 1
      ) %>% 
    group_by(describ, harvest, season, TRMT, spp) %>% 
    summarize(
      tpa = sum(tpa),  # the average tpa for each spp is aggregated across points
      tpha = sum(tpha),
      wall_denom = mean(wall_denom),
      age = mean(age),
      spp_count = sum(numb)
    )
 
  #this code provides the sum of stem density by harvest, giving totals plus spp summ in next dataframe 
  regen_summ_exp_GT_4.5ft <- bind_rows(regen_exp_GT_4.5ft_head, regen_exp_GT_4.5ft_tail) %>% 
    select(describ, harvest, season, TRMT, spp, tpa, tpha, PCT_pts_present, wall_denom, age, spp_count) %>% 
    group_by(describ, harvest, season, TRMT) %>% 
    summarize(
      tpa = sum(tpa),
      tpha=sum(tpha),
      spp_count = sum(spp_count),
      age = mean(age)
    ) %>% 
    mutate(spp = "TOTAL")

    
regen_exp_GT_4.5ft_spp <- bind_rows(regen_exp_GT_4.5ft_head, regen_exp_GT_4.5ft_tail, regen_summ_exp_GT_4.5ft) %>% 
    select(describ, harvest, season, TRMT, spp, tpa, tpha, PCT_pts_present, wall_denom, age, spp_count) %>% 
    rename(
      TOTAL_tpa = tpa,
      TOTAL_tpha = tpha
    ) %>% 
    #ungroup %>% 
    arrange(describ, harvest, desc(season), TRMT) %>% 
    dplyr::mutate_if(is.numeric, round, 0)  #changes all numeric variable to round to 0 significant digits
