---
title: "regen and plot analysis"
author: "Peter Smallidge"
date: "5/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Initialize Libraries
```{r, include=FALSE}

library(tidyverse)
library(skimr)        ## install.packages("skimr")
library(readxl)
library(readr)
#
# ALT plus hypen = <- (within chunk)
# ctrl plus shift plus m = %>% 
# ctrl + ALT + I = insert chunk
# ALT + - = insert assignment operator
# filter rows, select columns
# RMD comment = <!-- comment -->
# 
# 

```


### read data from csv file, check variables and data structure

```{r, data input}
# "tidy_plot_2019.csv" created in file "silvah_stocking_levels_assigned.Rmd"

plot_tidy <- read_csv("tidy_plot_2019.csv") %>% 
        # replace missing values before changing variables to factors.
   mutate(
    attrib_val=replace_na(attrib_val, "not")
    ) %>% 

  mutate_at(vars(harvest, year, point, location, plot_attrib, attrib_val),
            list(factor))


plot_untidy <- read_csv("untidy_plot_2019_stocking.csv") %>% 
       # replace missing values before changing variables character to factors.

   mutate(
    silvah25=replace_na(silvah25, "not"),
    silvah5 = replace_na(silvah5, "not"),
    silvah2 = replace_na(silvah2, "not"),
    silvah25a=replace_na(silvah25a, "not"),
    silvah5a = replace_na(silvah5a, "not"),
    silvah2a = replace_na(silvah2a, "not"),

    div_stock = replace_na(div_stock, "not"),
    woody_stock = replace_na(woody_stock, "not"),
    fern_stock = replace_na(fern_stock, "not"),
    all_intf = replace_na(all_intf, "not")
    ) %>% 
  
  mutate(
    silvah01_per_acre = replace_na(silvah01_per_acre, 0),
    silvah02_per_acre = replace_na(silvah02_per_acre, 0),
    silvah03_per_acre = replace_na(silvah03_per_acre, 0),
    silvah01a_per_acre = replace_na(silvah01a_per_acre, 0),
    silvah02a_per_acre = replace_na(silvah02a_per_acre, 0),
    silvah03a_per_acre = replace_na(silvah03a_per_acre, 0),

    silvah01div_per_acre = replace_na(silvah01div_per_acre, 0),
    interfere_per_acre = replace_na(interfere_per_acre, 0),
 ) %>% 

  mutate_at(vars(harvest, year, point, location, 
                 disturb, slash, fern, herb, 
                 low_cov, inv_cov, mid_cov, upper_cov, vine_cov,
                 silvah25, silvah5, silvah2, 
                 silvah25a, silvah5a, silvah2a, 
                 div_stock, woody_stock,
                 fern_stock, all_intf),
            list(factor)) 
  


regen_tidy <- read_csv("tidy_regen_point_2019.csv")   %>% 
  mutate_at(vars(harvest, location, point, spp, origin, year, type, size),
             list(factor))
  
  


```

### base R table functions, see dyplr below
```{r base R table functions}
# the table function (base R) provides the count of occurrences of two 
# categorical (factor) variables. In the first example the number of 
# plots that had commercial stocking (no vs. stocked) among each level
# of ratings of low % cover (absent, present, low, medium, high)

table(plot_untidy$comm_stock, plot_untidy$low_cov,useNA = "ifany")
table(plot_untidy$comm_stock, plot_untidy$mid_cov,useNA = "ifany")
table(plot_untidy$comm_stock, plot_untidy$upper_cov,useNA = "ifany")
table(plot_untidy$comm_stock, plot_untidy$disturb,useNA = "ifany")
table(plot_untidy$comm_stock, plot_untidy$slash,useNA = "ifany")
table(plot_untidy$comm_stock, plot_untidy$fern,useNA = "ifany")
table(plot_untidy$comm_stock, plot_untidy$herb,useNA = "ifany")
table(plot_untidy$comm_stock, plot_untidy$inv_cov,useNA = "ifany")




```

### base R proportion table functions, see dyplr below
```{r base R proportion table functions}

prop.table(table(plot_untidy$comm_stock, plot_untidy$mid_cov,useNA = "ifany"),margin=2)

prop.table(table(plot_untidy$comm_stock, plot_untidy$harvest, plot_untidy$location
                 ,useNA = "ifany"),margin=c(2,3))

#margin 2 is sum of second variable, and "3" is sum of third variable

```



###Frequency Table Plot Attributs_5  (gas line, red pine, boot, wedge, controls)
```{r}


freq_plot_attrib_5 <- plot_tidy %>% 
  filter(plot_attrib %in% c(
    "winner_stock", "comm_stock", "pre_comm_stock", "diver_stock",
    "pre_diver_stock", "interf_woody_stock","fern_stock", 
    "all_intr")
  ) %>% 
  group_by(harvest, location, plot_attrib, attrib_val) %>% 
  summarise(n = n()) %>% 
  mutate(freq = n / sum(n)) %>% 
  arrange(harvest, plot_attrib, location,  attrib_val)

write_csv(freq_plot_attrib_5, "freq_plot_attrib_5.csv")




```

###Frequency Stocking Attributs_4 sites (gas line, red pine, station road, controls)
```{r}

# frequency table of stocking indicies
# 
freq_stocking_attrib_4 <- plot_tidy %>% 
  mutate(harvest = fct_recode(harvest,
                              "sta_rd" = "boot",
                              "sta_rd" = "wedge")) %>% 
  filter(plot_attrib %in% c(
   "silvah25", "silvah5", "silvah2", 
   "silvah25a", "silvah5a", "silvah2a", 
   "div_stock", "woody_stock"
    )
  ) %>% 
  group_by(harvest, plot_attrib, location, attrib_val) %>% 
  summarise(n = n()) %>% 
  mutate(freq = (n / sum(n)) * 100) %>% 
  arrange(harvest , plot_attrib, location,  attrib_val)

write_csv(freq_stocking_attrib_4, "freq_stocking_attrib_4_all.csv")


```

###Frequency % Cover Attributs_4 sites (gas line, red pine, station road, controls)
```{r}

# frequency table of stocking indicies
# 
freq_cover <- plot_tidy %>% 
  mutate(harvest = fct_recode(harvest,
                              "sta_rd" = "boot",
                              "sta_rd" = "wedge")) %>% 
  filter(plot_attrib %in% c(
   "disturb", "slash", "fern", "herb", "low_cov", "inv_cov", "mid_cov", "upper_cov", "vine_cov"
    )
  ) %>% 
  group_by(harvest, location, plot_attrib, attrib_val) %>% 
  summarise(n = n()) %>% 
  mutate(freq = (n / sum(n)) * 100) %>% 
  arrange(harvest , plot_attrib, location,  attrib_val)

write_csv(freq_cover, "freq_cover_attrib_4.csv")


```


