str---
title: "Regen and Plot Input of 2020 Data and Cleaning/Tidy"
author: "Peter Smallidge"
date: "10/17/2020"
output: github_document
```

<!-- README

File Description: The purpose of this file is to:
- input data from the 2020 growing season regeneration plots (code 01)
- input the plot site characteristics (code 02)
- A separate file will read/input the seedling height data. 
- standardize the names of the harvests (code 02)
- count the number of sample plots and compare to number of veg plots (code 03)
- insert "empty" veg plots where no veg occurred, using spp code "001" (code 04)
- convert stem counts to stems per acre (code 05)
- convert harvest names to consistent usage across years (code 05)
- convert species numeric to alphabetic abbrev (code 05)
- convert untidy to tidy data, alphabetic species (code 06)


exported data files include:
plot_data_2020_untidy.csv
regen_per_acre_2020_numeric_point_untidy.csv
regen_per_acre_2020_alpha_untidy_point.csv
regen_point_per_acre_2020_alpha_tidy.csv

 -->


```{r}
# code from one of Maria's "notes" pages
# TOC may only work with html output

# title: "Lesson 13: Factors"
# output: 
#  html_document:
#    keep_md: yes 
#    toc: true
#    toc_depth: 2 
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Analysis of slash wall regeneation plot data - 2020

###Initialize Libraries
```{r, include=FALSE}

#library(ggplot2)
#library(ggpubr)
#library(readr)
library(readxl)
library(skimr)
library(tidyr)
library(tidyverse)

#
# ALT plus hypen = <- (within chunk)
# ctrl plus shift plus m = %>% 
# ctrl + ALT + I = insert chunk
# ALT + - = insert assignment operator
# filter rows, select columns
# RMD comment = <!-- comment -->
# 
# 

```


### read data from csv file, check variables and data structure

<!-- note to self: in the "read_csv" code, after the file location the code
for 'col..." etc ensures that the specified column is adjusted to correct data type
the file "cheat sheet data-import" has details -->

### code 01
###Input Regen Data (code 01)

```{r, input regen data}
# Seedling data, located in the same folder as this 
# the initial read of the data revealed that some columns were incorrectly
# recognized when inputted. the code for "col_types=" was added in a second pass
# of "read_csv" to clarify the desired data type.

veg_data1_2020 <- regen_data_2020_cleaned <- read_excel("regen.data.2020_cleaned.xlsx", 
    sheet = "regen", col_types = c("text", 
        "numeric", "text", "text", "text", 
        "text", "text", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "skip"))  %>% 
  
  # replace missing values before changing character variables to factor variables
  
    replace_na(list(date = "other",  origin = "u", 
                  sup.seed01 = 0, sup.seed02 = 0, sup.seed03 = 0, sup.sapl = 0,
                  exp.seed01 = 0, exp.seed02 = 0, exp.seed03 = 0, exp.sapl = 0,
                  top4 = 0
                  )) %>%   #need to change missing values to real values with variable as character
  mutate_at(
    vars(harvest, location, point),
    list(factor)) %>% 
  
   mutate(harvest = fct_recode(harvest,
                             "gas_line" = "gl",
                              "red_pine" = "rp",
                              "sta_rd" = "wb",
                              "red_pine" = "rp",
                              "gas_line" = "gl",
                              "boot" = "b",
                              "wedge" = "w",
                              "sta_rd" = "wb",
                              "recknagel" = "rect1",
                              "recknagel" = "rect2",
                              "recknagel" = "rect3",
                              "recknagel" = "rect4",
                              "recknagel" = "recksouth",
                              "circle" = "circle1",
                              "circle" = "circle2",
                              "circle" = "circle3"
   ))
   
   

   
   harvested2017 <- veg_data1_2020 %>% 
    filter(harvest %in% c("boot", "gas_line", "red_pine", "wedge", "sta_rd")) %>% 
    mutate(
      cut = 2017,
      year = season,
      numb_yrs_grow = year - cut
    )


  harvested2019 <- veg_data1_2020 %>% 
    filter(harvest %in% c("campridge", "circle",
                          "decker","recknagel")) %>% 
      mutate(
      cut = 2019,
      year = season,
      numb_yrs_grow = year - cut
      )


veg_data <- 
  bind_rows(harvested2017, harvested2019) %>%
    
    mutate(location = recode(location,
    "c" = "control", "i" = "interior", "p" = "perimeter"
))



```


### code 02
###Input Plot Data (code 02)



```{r, input plot data}

plot_data <- read_excel("regen.data.2020_cleaned.xlsx", 
    sheet = "site-plot", skip = 6) %>% 

    replace_na(list(date = "other",  origin = "u", 
                  oakhick = 0, beech = 0, maple = 0, birch = 0, hdwd = 0, conifer = 0
                  )) %>%    #need to change missing values to real values with variable as character
 
   mutate_at(
    vars(harvest, location, point,disturb, slash, fern, herb, native_cov,
          invasive_cov, mid_cov, upper_cov, vine_cov),
    list(factor)) %>% 

 
   mutate(harvest = fct_recode(harvest,
                             "gas_line" = "gl",
                              "red_pine" = "rp",
                              "sta_rd" = "wb",
                              "red_pine" = "rp",
                              "gas_line" = "gl",
                              "boot" = "b",
                              "wedge" = "w",
                              "sta_rd" = "wb",
                              "recknagel" = "rect1",
                              "recknagel" = "rect2",
                              "recknagel" = "rect3",
                              "recknagel" = "rect4",
                              "recknagel" = "recksouth",
                              "circle" = "circle1",
                              "circle" = "circle2",
                              "circle" = "circle3"
   ))
  


write_csv(plot_data, "plot_data_2020_untidy.csv")

```

#  mutate(harvest = recode(
#   harvest, "boot" = "boot", "wedge" = "wedge", "wedge_boot" = "sta_rd",
#    "rect1" = "campridge", "rect2" = "campridge", "rect3" = "campridge", "rect4" = "campridge",
#    "circle1" = "circle", "circle2" = "circle", "circle3" = "circle"))


### code 03
###Count Number of Sample Points (code 03)
```{r, count number of sample points}

point_count_01 <-  plot_data %>% # create variable for total plots for denominator of average density
  group_by(harvest, location, point) %>% 
  summarize(
    n =n()
    ) %>% 
  mutate(count = n/n) %>% 
  arrange(point) %>% 
  select(harvest, location, point, count) 
  
point_count_sum <- point_count_01 %>% 
  group_by(harvest, location) %>% 
  summarize(
    numplot = sum(count)
    )

write_csv(point_count_sum, "count_of_points_harvest_x_location.csv")


# for each point, the value of "count" = 1
```


### code 04
###Create plots for regen with no observations (code 04)

```{r, filling out empty sample points}

#There are 486 plots locations in 2020. 

#The missing points would cause an error for calculation
#of the number of stems per acre as an average of points.
#
#

site_data <- full_join(point_count_01, veg_data, by = "point") %>% 
  select(!c(harvest.y, location.y)) %>% 
  rename(harvest = harvest.x, location = location.x) %>%     
  replace_na(list(date = "other",  origin = "u", spp = "001",
                  sup.seed01 = 0, sup.seed02 = 0, sup.seed03 = 0, sup.sapl = 0,
                  exp.seed01 = 0, exp.seed02 = 0, exp.seed03 = 0, exp.sapl = 0,
                  top4 = 0
                  )) %>%    #need to change missing values to real values with variable as character
    mutate(location = recode(location,
    "c" = "control", "i" = "interior", "p" = "perimeter"
    ))


#renames a variable label from new = old

# confirmed that all plot numbers are unique
#date = character
#spp, origin = factor
#all other = numeric


```

```{r}
inside <- site_data %>% 
    filter(location %in% c("interior", "perimeter")) %>% 
    droplevels %>% 
    mutate(
      wall = "protected"
    )

outside <- site_data %>% 
    filter(location %in% c("control")) %>% 
    droplevels %>% 
    mutate(
      wall = "control"
    )

```



### code 05
###data manipulations - per acre calculations & spp names
### output csv files of untidy data with spp numeric and spp alphabetic
```{r, regen per acre at points - calculations}
#
#height classes: 
#   01 = 12" to 4.5ft, within 3.7 ft radius
#   02 = 4.5 ft to 9ft, within 6 ft radius
#   03 = 9 ft. to 1" dbh, withing 6 ft radius
#   winner = top4, identification of the 4 stems deemed winners, within 6 ft radius

regen_per_acre <- bind_rows(inside, outside) %>% 

  #following code calculates the number of seedlings per acre
  #by species for each height class. Not yet tidy.

    mutate(  #convert plot count of stems to stems/acre
    sup_short = sup.seed01 * (43560 /((3.14156) * (3.7 * 3.7))),
    sup_med = sup.seed02 * (43560 /((3.14156) * (6 * 6))),
    sup_tall = sup.seed03 * (43560 /((3.14156) * (6 * 6))),
    sup_sapl = sup.sapl * (43560 /((3.14156) * (6 * 6))),
    exp_short = exp.seed01 * (43560 /((3.14156) * (3.7 * 3.7))),
    exp_med = exp.seed02 * (43560 /((3.14156) * (6 * 6))),
    exp_tall = exp.seed03 * (43560 /((3.14156) * (6 * 6))),
    exp_sapl = exp.sapl * (43560 /((3.14156) * (6 * 6))),
  ) %>% 

  mutate(
    suppressed = sup_short + sup_med + sup_tall + sup_sapl,
    exposed = exp_short + exp_med + exp_tall + exp_sapl,
    total = suppressed + exposed, 
        #exclude winners, which would double count
    winners = top4 * (43560 /((3.14156) * (6 * 6))) 
  ) %>% 
  
  select(!c(sup.seed01, sup.seed02, sup.seed03, sup.sapl,
            exp.seed01, exp.seed02, exp.seed03, exp.sapl,
            date, top4)) %>% 
  
  #unify/aggregate values of a factor, "new" = "old"
  mutate(origin = fct_recode(origin,
                             "veg" = "r",
                             "veg" = "t",
                             "singl/unk" = "s",
                             "singl/unk" = "u"
    )) 
    
        write_csv(regen_per_acre, "regen_per_acre_2020_numeric_point_untidy.csv")

  
  regen_alpha <- regen_per_acre %>% 
   mutate(spp = fct_recode(spp,   #comment out species not present in data
                          "empty" = "1",
                          "ab" = "531",
                          "stm" = "315",
                          "sm" = "318",
                          "rm" = "316",
                          "rp" = "125",
                          "wp" = "129",
                          "he" = "261",
                          #"svb" = "356",
                          "svb" = "355",
                          "yb" = "371",
                          "sb" = "372",
                          "pb" = "375",
                         # "gb" = "379",
                          "ah" = "391",
                          "haw" = "500",
                          "wa" = "541",
                          "yp" = "621",
                          #"cu" = "651",
                          #"la" = "70",
                          "eh" = "701",
                          #"asp" = "740",
                          "bta" = "743",
                          "qa" = "746",
                          "pc" = "761",
                          "bc" = "762",
                          "cc" = "763",
                          "wo" = "802",
                          "co" = "832",
                          "ro" = "833",
                          "bl" = "901",
                          "wi" = "920",
                          "ba" = "951",
                          "sp" = "90", 
                          "sa" = "931",
                          "wh" = "585",
                          "sh" = "407",
                          "ns" = "91",
                          "bh" = "402",
                          "au" = "535",
                          "ec" = "742"
                          #"unk" = "999"
                          ))

#view(regen.per.acre)

write_csv(regen_alpha, "regen_per_acre_2020_alpha_point_untidy.csv")





```


### code 06



### making data tidy - comm/regen
```{r, making data tidy commercial species}
# commercial regeneration species
# 316 318 372 541 621 762  802 832 833 951 
# 129, 261, 371,407, 402
# maple, yeb, swb, wa, yp, oaks, ba , bh, sh
# no populus 740, 743, 746, no pab 375


comm <- regen_alpha %>% 
  
  filter(spp %in% c("wp", "he", "rm", "sm","yb", "sb", "wa", "yp",
                     "bc", "wo", "co", "ro", "ba", "bl", "rp", "cu", "sh", "bh"
                    )) %>% 
  mutate(
    type = as.factor("commercial")
  ) %>% 
  pivot_longer(c(sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total, winners),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))



```

###making data tidy - diverse/regen
```{r, making data tidy diverse}

# desirable diversity species (no interfering, no commercial, but includes Populus, paper birch, not pin cherry)
# 129 261 316 318 356 372 375 500 541 621 743 746 762 802 832 833  90 920 951, 407, 931

diverse <- regen_alpha %>% 
  
  filter(spp %in% c("svb", "pb", "gb", "haw", "asp", "bta", "qa",
                    "wi", "sh", "wh","sa","ec"
                    )) %>% 
  #

  mutate(
    type = as.factor("diversity")
  ) %>% 
  pivot_longer(c(sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total, winners),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))


```

###making data tidy - interfering species/regen
```{r, tidy for interfering species}

# interfering species

interfere <- regen_alpha %>% 
  
  filter(spp %in% c("ab", "stm", "eh", "pc", "au")) %>% 
  mutate(
    type = as.factor("interfering")
  ) %>% 
  pivot_longer(c(sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total, winners),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))


```

###making data tidy - other species
```{r}
# other species

other <- regen_alpha %>% 
  
  filter(spp %in% c("empty", "ah", "la", "sp", "unk", "cc", "ns"
                    )) %>% 
  mutate(
    type = as.factor("other"),
    year = as.numeric(year)
  ) %>% 
  pivot_longer(c(sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total, winners),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))


```

 

###Write tidy data for each point
```{r, tidy regen and plot data, assign stocking label to each point}

tidy_regen_point <- 
  bind_rows(comm, diverse, interfere, other)
 
 
write_csv(tidy_regen_point, "regen_point_per_acre_2020_alpha_tidy.csv")

```




###Make Plot Data Tidy



```{r}

# not ready to run this code
# 
# 
plot_data_tidy <- #read the csv file having basal area calculations %>% 

# (need to input the calculated values of plot variables 
#  such as basal area per acre.  
  
  select(!c(baf,
            low_spp, inv_spp, mid_spp, upper_spp, vine_spp)) %>% 
  
  pivot_longer(c(disturb, slash, fern, herb, 
                 low_cov, inv_cov, mid_cov, upper_cov, vine_cov,
                  comm_stock, pre_comm_stock, diver_stock,
                  pre_diver_stock,  interf_woody_stock, fern_stock, all_intf,
                 winner_stock
                 ),
               names_to = "plot_attrib",
               values_to = "attrib_val") %>% 
  
  replace_na(list("attrib_val" = "no", comm_per_acre = 0, pre_comm_per_acre = 0,
                  diver_per_acre = 0, pre_diver_per_acre = 0, interfere_per_acre = 0,
                  pre_interfere_per_acre = 0))

write_csv(plot.data_tidy, "tidy_plot_2019.csv")




```



<!-- 
 includes factor and number variables...can't blend via "pivot_longer"
 pivot_longer(c(disturb, slash, fern, herb, 
                 low_cov, inv_cov, mid_cov, upper_cov, vine_cov,
                 ba, seedling_numb, comm_per_acre, comm_stock,
                 pre_comm_per_acre, pre_comm_stock,diver_per_acre, diver_stock,
                 pre_diver_per_acre, pre_diver_stock, interfere_per_acre, interfere_stock,
                 pre_interfere_per_acre, pre_interfere_stock
                 ),
               names_to = "plot_attrib",
               values_to = "attrib_val")
-->
