str---
title: "Regen and Plot Input of 2020 Data and Cleaning/Tidy"
author: "Peter Smallidge"
date: "10/17/2020"
output: github_document
```

<!-- README

File Description: The purpose of this file is to:
- input data from the 2021 growing season regeneration plots (code 01)
- input the plot site characteristics (code 02)
- A separate file will read/input the seedling height data. 
- standardize the names of the harvests (code 02)
- count the number of sample plots and compare to number of veg plots (code 03)
- insert "empty" veg plots where no veg occurred, using spp code "001" (code 04)
- convert stem counts to stems per acre (code 05)
- convert harvest names to consistent usage across years (code 05)
- convert species numeric to alphabetic abbrev (code 05)
- convert untidy to tidy data, alphabetic species (code 06)


exported data files include:
plot_data_2022_untidy.csv (line 237)
regen_per_acre_2022_numeric_point_untidy.csv (line 353)
regen_point_per_acre_2022_alpha_tidy.csv (line 530)

 -->


```{r}
# code from one of Maria's "notes" pages
# TOC may only work with html output

# title: "Lesson 13: Factors"
# output: 
#  html_document:
#    keep_md: yes 
#    toc: true
#    toc_depth: 2 
```


```{r setup, include=FALSE}
```

### Analysis of slash wall regeneation plot data - 2022

###Initialize Libraries
```{r, include=FALSE}

#library(ggplot2)
#library(ggpubr)
#library(readr)
library(readxl)
#library(skimr)
library(tidyr)
library(tidyverse)

#
# ALT plus hypen = <- (within chunk)
# ctrl plus shift plus m = %>% 
# ctrl + ALT + I = insert chunk
# ALT + - = insert assignment operator
# filter rows, select columns
# RMD comment = <!-- comment -->
# 
# 

```


### read data from csv file, check variables and data structure

<!-- note to self: in the "read_csv" code, after the file location the code
for 'col..." etc ensures that the specified column is adjusted to correct data type
the file "cheat sheet data-import" has details -->

<!--     -->

### code 01
###Input Regen Data (code 01) 

```{r, input regen data}
# Seedling data, located in the same folder as this 
# the initial read of the data revealed that some columns were incorrectly
# recognized when inputted. the code for "col_types=" was added in a second pass
# of "read_csv" to clarify the desired data type.

#new raw date at "slash.wall.data.2022.7february2023_incomplete.xlsx"

# the 2022 growing season data file still needs data input from DECKER plot and 
# plot data to be collected from red pine and camp ridge



regen2022 <-  read_excel("slash.wall.data.2022.16may2023_final.xlsx", 
    sheet = "regen", col_types = c("text", 
        "numeric", "text", "numeric", "numeric", 
        "text", "text", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "skip")) %>% # replace missing values before changing character variables to factor variables

    replace_na(list(date = "other",  origin = "u", location = 0,
                  sup.seed01 = 0, sup.seed02 = 0, sup.seed03 = 0, sup.sapl = 0,
                  exp.seed01 = 0, exp.seed02 = 0, exp.seed03 = 0, exp.sapl = 0,
                  top4 = 0
                  )) %>%   #need to change missing values to real values while variable as character
    mutate(
    location = point
      ) %>% 
mutate(
  point = as.character(point),
  location = as.character(location)
) %>% 
   mutate_at(
    vars(harvest),
    list(factor))    %>% 
   
    mutate(harvest = fct_recode(harvest,
        "circle1" = "cricle1")) %>% 
        
           mutate(harvest = fct_recode(harvest,
                             "gas_line" = "gasline",
                              "red_pine" = "redpine",
                              "sta_rd" = "stationrd",
                              "recknagel" = "rectangle1",
                              "recknagel" = "rectangle2",
                              "recknagel" = "rectangle3",
                              "recknagel" = "rectangle4",
                              "circle" = "circle1",
                              "circle" = "circle2",
                              "circle" = "circle3",
                              "sta_rd" = "boot",
                              "sta_rd" = "wedge"

   )) %>% 
   arrange(harvest, point, spp)
   
# RUN table(regen2022$harvest) to look for NA values
table(regen2022$harvest)


 # EX = mutate(mid_spp = recode(mid_spp, "swm" = "rhus")) %>%  #old = new


 # change the value of "location" of the regen data,which was set equal "point", to one of the values: control, interior or perimeter
 # point and location correspondance found in file "regen.data.2021_cleaned.xlsx" in worksheet "point_by_location"

 
point_location <- read_excel("~/R/slash-wall-vegetation/location_by_point.xlsx")%>% 
   mutate(harvest = fct_recode(harvest,
                             "gas_line" = "gasline",
                              "red_pine" = "redpine",
                              "sta_rd" = "stationrd",
                              "recknagel" = "rectangle1",
                              "recknagel" = "rectangle2",
                              "recknagel" = "rectangle3",
                              "recknagel" = "rectangle4",
                              "circle" = "circle1",
                              "circle" = "circle2",
                              "circle" = "circle3",
                              "sta_rd" = "boot",
                              "sta_rd" = "wedge"

   )) %>% 
   mutate(
   point = as.character(point)
    ) %>% 
   arrange(harvest, point)

table(point_location$harvest)

        

```


    # full join allowed matching of the location=treatment with the location which was temporarily (see above) set equal
    # to point#.  Variables harvest and point as characters. Full join created duplicate variables of "location.x" and "location.y"


```{r}
veg_data1 <- full_join(regen2022, point_location, by = (c("harvest", "point"))) %>% 
    select(!"location.x") %>%  
    mutate(
    season = 2022
    ) %>% 
        # this code selects all variables except "location.x" which was set to the point number, 
        # retaining "location.y" which is i, p, or c
        
    rename("location" = "location.y") %>% 
    replace_na(list(date = "other",  origin = "s", location = "u", spp = "001",
           sup.seed01 = 0, sup.seed02 = 0, sup.seed03 = 0, sup.sapl = 0,
           exp.seed01 = 0, exp.seed02 = 0, exp.seed03 = 0, exp.sapl = 0,
           top4 = 0
           )) %>%   
    arrange(harvest, point, spp) %>% 
    select(harvest, season, date, point, location, wall, everything())
   
# for "rename", "newname" = "old name"
# for fct_recode, newname = oldname
  
```

```{r}

plot_data01  <- read_excel("~/R/slash-wall-vegetation/slash.wall.data.2022.16may2023_final.xlsx", 
    sheet = "site-plot", col_types = c("text", 
        "text", "text", "text", "numeric", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "numeric", "skip"), skip = 6) %>% 
        select(harvest, season, date, location, point, comm_sdl) %>% 
  replace_na(list(comm_sdl = 0)) %>%
  mutate(spp = "004", origin = "s") %>% 
  mutate(
    location = as.character(location),
    point = as.character(point)
) %>% 
  mutate(location = point) %>% 
  mutate(harvest = fct_recode(harvest,
                              "gas_line" = "gasline",
                              "recknagel" = "rectangle1",
                              "recknagel" = "rectangle2",
                              "recknagel" = "rectangle3",
                              "recknagel" = "rectangle4",
                              "circle" = "circle1",
                              "circle" = "circle2",
                              "circle" = "circle3",
                              "sta_rd" = "boot",
                              "sta_rd" = "wedge"
  )) %>% 
  arrange(harvest, point, spp)
  
  
  plot_data <- full_join(plot_data01, point_location, by = (c("harvest", "point"))) %>% 
      select(!"location.x") %>% 
      rename("location" = "location.y") %>% 
      arrange(harvest, point, spp) %>% 
      mutate(
      season = as.numeric(season)
)
      # select(harvest, date, season, point, location, spp, origin, comm_sdl)

# the missing values in df=plot_data are created because some plot data haven't been entered ("plot_data")
# but there are data for the location assignment of the point ("point_location"), In full_join
# all entries are included.

write_csv(plot_data, "plot_data_2022_untidy.csv")

```



<!--
The following code creates two new variables. One to identify the year the slash wall was finished, and the other calculate the
number of growing seasons represented by the data as "numb_yrs_grow"


-->
```{r}

# the code for "filter_at" looks for missing values in spp or (?) origin and if missing that observation is removed via "!(is.na)"

   harvested2017 <- full_join(veg_data1, plot_data,  
        by = (c("harvest", "point", "location", "origin", "spp", "season", "date"))) %>%
        select(-wall.y) %>% 
        rename("wall" = "wall.x") %>% 
      filter(harvest %in% c("boot", "gas_line", "red_pine", "wedge", "sta_rd")) %>% 
      filter_at(vars(spp, origin), all_vars(!(is.na(.)))) %>% 
      replace_na(list(origin = "s", 
                  sup.seed01 = 0, sup.seed02 = 0, sup.seed03 = 0, sup.sapl = 0,
                  exp.seed01 = 0, exp.seed02 = 0, exp.seed03 = 0, exp.sapl = 0,
                  top4 = 0, comm_sdl = 0
                  )) %>% 
      mutate(
          cut = 2017,
          year = season,
          numb_yrs_grow = year - cut
          ) %>% 
      select(harvest, year, date, location, wall, point, spp, origin, season, cut, numb_yrs_grow,
        sup.seed01, sup.seed02, sup.seed03, sup.sapl, exp.seed01,exp.seed02, exp.seed03, exp.sapl,
        top4, comm_sdl) %>% 
      arrange(harvest, point, spp)


  harvested2019 <- full_join(veg_data1, plot_data,  
      by = (c("harvest", "point", "location", "origin", "spp", "season", "date"))) %>% 
         select(-wall.y) %>% 
        rename("wall" = "wall.x") %>% 
      filter_at(vars(spp, origin), all_vars(!(is.na(.)))) %>% 
      filter(harvest %in% c("campridge", "circle", "decker","recknagel")) %>% 
      replace_na(list(origin = "s", 
                  sup.seed01 = 0, sup.seed02 = 0, sup.seed03 = 0, sup.sapl = 0,
                  exp.seed01 = 0, exp.seed02 = 0, exp.seed03 = 0, exp.sapl = 0,
                  top4 = 0, comm_sdl = 0
                  )) %>% 
      mutate(
          cut = 2019,
          year = season,
          numb_yrs_grow = year - cut
          ) %>% 
      select(harvest, year, date, location, wall, point, spp, origin, season, cut, numb_yrs_grow,
        sup.seed01, sup.seed02, sup.seed03, sup.sapl, exp.seed01,exp.seed02, exp.seed03, exp.sapl,
        top4, comm_sdl) %>% 
      arrange(harvest, point, spp)
# 
# plot_data (486 obs) primarily contains the counts of commercial seedlings at each plot, coded as spp=004 other hdwd
# veg_data1 (1416 obs) has the stem counts by species and height class.
# veg_data increases the number of rows for points that had both comm_sdl and other stem data (e.g., #299, 300), 
# but not for points that were otherwise empty (e.g., #334 - 344)

veg_data <- 
  bind_rows(harvested2017, harvested2019) %>% 
    mutate(location = recode(location,
    "c" = "control", "i" = "interior", "p" = "perimeter"
    )) %>% 
         select(harvest, location, point, season, cut, year, numb_yrs_grow,
         origin, spp, comm_sdl, sup.seed01, sup.seed02, sup.seed03, sup.sapl, 
          exp.seed01, exp.seed02, exp.seed03, exp.sapl,
           top4, date)


```


#str(veg_data)  - this shows data structure
#table(veg_data$harvest) - this shows table of count of each level of the variable "harvest"
# use str and table code in the console to verify mis coded data, counts of points by harvest, etc.
#
# compare to file "count_of_points_harvest_x_location in R>slash-wall-vegetation


### code 02
###Input Plot Data (code 02)



```{r, input plot data}


```


### code 03
###Count Number of Sample Points (code 03)
```{r, count number of sample points}

point_count_01 <-  point_location %>% # create variable for total plots for denominator of average density
  group_by(harvest, location, point) %>% 
  summarize(
    n =n()
    ) %>% 
  mutate(count = n/n) %>% 
  arrange(point) %>% 
  select(harvest, location, point, count)
  
  
# counts the number of points by each location: interior, perimeter, control
point_count_location <- point_count_01 %>%    #denominator for harvest x location, number of plots
  group_by(harvest, location) %>% 
  summarize(
    location_denom = sum(count))

# counts the number of points by inside the wall ("protected") vs. outside the wall ("control")
point_count_inside <- point_count_location %>% 
      filter(location %in% c("i", "p")) %>% 
      mutate(wall = "protected")%>% 
      group_by(harvest, wall) %>% 
      summarize(
        wall_denom = sum(location_denom))

point_count_outside <- point_count_location %>% 
      filter(location %in% c("c")) %>% 
      mutate(wall = "control")%>% 
      group_by(harvest, wall) %>% 
      summarize(
        wall_denom = sum(location_denom))

point_count_wall <- bind_rows(point_count_inside, point_count_outside) %>% 
arrange(harvest, desc(wall))


write_csv(point_count_location, "count_of_points_harvest_x_location.csv")
write_csv(point_count_wall, "count_of_points_harvest_x_treatment.csv")

# for mutate/recode which changes values, oldname=newname

# for each point, the value of "count" = 1
```


### code 04
###Create plots for regen with no observations (code 04)

```{r, filling out empty sample points}

#There are 486 plots locations in 2020. 

#The missing points would cause an error for calculation
#of the number of stems per acre as an average of points.
#
#  IGNORE IGNORE
#  code 04 not needed. Expansion of veg_data occurred at prior step. 486 unique plots
#  there are now no missing point
#  assess using console "unique(veg_data$point)
#  IGNORE dataframe = site_data, rather use dataframe = "veg_data"

site_data01 <- full_join(point_location, veg_data, by = "point") %>% 
  select(!c(harvest.y, location.y)) %>% 
  rename(harvest = harvest.x, location = location.x)   

site_data <- site_data01 %>% 
  # select(!c(season.y, date.y, location.y, spp.y, origin.y, comm_sdl.y)) %>% 
  # rename( location = location.x, date = date.x, season = season.x, spp = spp.x, comm_sdl = comm_sdl.x, origin = origin.x) %>% 
  replace_na(list(date = "other",  origin = "u", spp = "001",
                  sup.seed01 = 0, sup.seed02 = 0, sup.seed03 = 0, sup.sapl = 0,
                  exp.seed01 = 0, exp.seed02 = 0, exp.seed03 = 0, exp.sapl = 0,
                  top4 = 0, comm_sdl = 0
                  )) %>%    #need to change missing values to real values with variable as character
    mutate(location = recode(location,
    "c" = "control", "i" = "interior", "p" = "perimeter"
    ))  #"old value" = "new value"


#renames a variable label from new = old

# 2022 data
# confirmed that all plot numbers are unique
#date = character
#spp, origin, location and point are all character
#all other = numeric


```

```{r}

### this code is not needed, do not run
inside <- veg_data %>% 
    filter(location %in% c("interior", "perimeter")) %>% 
    droplevels %>% 
    mutate(
      wall = "protected"
    )

outside <- veg_data %>% 
    filter(location %in% c("control")) %>% 
    droplevels %>% 
    mutate(
      wall = "control"
    )

```


### code 05
###data manipulations - per acre calculations & spp names
### output csv files of untidy data with spp numeric and spp alphabetic
```{r, regen per acre at points - calculations}
#
#height classes: 
#   00 = 4" to 12", within 3.7 ft radius ("less than" "LT" seedlings)
#   01 = 12" to 4.5ft, within 3.7 ft radius
#   02 = 4.5 ft to 9ft, within 6 ft radius
#   03 = 9 ft. to 1" dbh, within 6 ft radius
#   winners = top4, identification of the 4 stems deemed winners, within 6 ft radius
  #following code calculates the number of seedlings per acre
  #by species for each height class. Not yet tidy.
        #exclude winners, which would double count


regen_per_acre <- site_data %>% 
    mutate(  #convert plot count of stems to stems/acre
    LT_seedling = comm_sdl * (43560 /((3.14156) * (3.7 * 3.7))),
    sup_short = sup.seed01 * (43560 /((3.14156) * (3.7 * 3.7))),
    sup_med = sup.seed02 * (43560 /((3.14156) * (6 * 6))),
    sup_tall = sup.seed03 * (43560 /((3.14156) * (6 * 6))),
    sup_sapl = sup.sapl * (43560 /((3.14156) * (6 * 6))),
    exp_short = exp.seed01 * (43560 /((3.14156) * (3.7 * 3.7))),
    exp_med = exp.seed02 * (43560 /((3.14156) * (6 * 6))),
    exp_tall = exp.seed03 * (43560 /((3.14156) * (6 * 6))),
    exp_sapl = exp.sapl * (43560 /((3.14156) * (6 * 6))),
    ) %>% 
  mutate(
    suppressed = sup_short + sup_med + sup_tall + sup_sapl,
    exposed = exp_short + exp_med + exp_tall + exp_sapl,
    total = suppressed + exposed, 
    winners = top4 * (43560 /((3.14156) * (6 * 6))) 
    ) %>% 
    select(!c(sup.seed01, sup.seed02, sup.seed03, sup.sapl,
            exp.seed01, exp.seed02, exp.seed03, exp.sapl,
            date, top4, comm_sdl)) %>% 
  
#unify/aggregate values of a factor, "new" = "old"

  mutate(origin = fct_recode(origin,
                             "veg" = "v",
                             "singl/unk" = "s",
                             "singl/unk" = "u"
    ))
    
write_csv(regen_per_acre, "regen_per_acre_2022_numeric_point_untidy.csv")

  
  regen_alpha <- regen_per_acre %>% 
   mutate(spp = fct_recode(spp,   #comment out species not present in data
                         "empty" = "1",
                          "empty" = "001",
                          "ohw" = "004",
                          "ab" = "531",
                          "stm" = "315",
                          "sm" = "318",
                          "rm" = "316",
                          "rp" = "125",
                          "wp" = "129",
                          "he" = "261",
                          #"svb" = "356",
                          "svb" = "355",
                          #"yb" = "371",
                          "sb" = "372",
                          "pb" = "375",
                          #"gb" = "379",
                          "ah" = "391",
                          "haw" = "500",
                          "wa" = "541",
                          "yp" = "621",
                          #"cu" = "651",
                          "app" = "660",
                          "la" = "70",
                          "eh" = "701",
                          #"asp" = "740",
                          "bta" = "743",
                          "qa" = "746",
                          "pc" = "761",
                          "bc" = "762",
                          "cc" = "763",
                          "wo" = "802",
                          "co" = "832",
                          "ro" = "833",
                          #"bl" = "901",
                          "wi" = "920",
                          "ba" = "951",
                          "sp" = "90", 
                          "sa" = "931",
                          "wh" = "585",
                          "sh" = "407",
                          #"ns" = "91",
                          "bh" = "402",
                          #"au" = "535",
                          #"ec" = "742",
                          #"unk" = "999",
                          
#recode data entry errors (avoids repeated fixes with data file updates)
# These numeric codes were entered, and because the raw data file is repeatedly updated, the
# errors would have to be correct each time. This lacks finesse but is more effecient.
                          "sb" = "273",
                          "ab" = "331",
                          "ab" = "351"
                          ))

table(regen_alpha$spp)
table(regen_alpha$harvest)
#view(regen.per.acre)

write_csv(regen_alpha, "regen_per_acre_2022_alpha_point_untidy.csv")





```


### code 06



### making data tidy - comm/regen
```{r, making data tidy commercial species}
# commercial regeneration species
# 316 318 372 541 621 762  802 832 833 951 
# 129, 261, 371,407, 402
# maple, yeb, swb, wa, yp, oaks, ba , bh, sh
# no populus 740, 743, 746, no pab 375


comm <- regen_alpha %>% 
  
  filter(spp %in% c("wp", "he", "rm", "sm","yb", "sb", "wa", "yp",
                     "bc", "wo", "co", "ro", "ba", "bl", "rp", "cu", "sh", "bh", "ohw"
                    )) %>% 
  mutate(
    type = as.factor("commercial")
  ) %>% 
  pivot_longer(c(LT_seedling, sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total, winners),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))


```

###making data tidy - diverse/regen
```{r, making data tidy diverse}

# desirable diversity species (no interfering, no commercial, but includes Populus, paper birch, not pin cherry)
# 129 261 316 318 356 372 375 500 541 621 743 746 762 802 832 833  90 920 951, 407, 931

diverse <- regen_alpha %>% 
  
  filter(spp %in% c("svb", "pb", "gb", "haw", "asp", "bta", "qa",
                    "wi", "sh", "wh","sa","ec"
                    )) %>% 
  #

  mutate(
    type = as.factor("diversity")
  ) %>% 
  pivot_longer(c(LT_seedling, sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total, winners),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))


```

###making data tidy - interfering species/regen
```{r, tidy for interfering species}

# interfering species

interfere <- regen_alpha %>% 
  
  filter(spp %in% c("ab", "stm", "eh", "pc", "au")) %>% 
  mutate(
    type = as.factor("interfering")
  ) %>% 
  pivot_longer(c(LT_seedling, sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total, winners),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))


```

###making data tidy - other species
```{r}
# other species

other <- regen_alpha %>% 
  
  filter(spp %in% c("empty", "ah", "la", "sp", "unk", "cc", "ns"
                    )) %>% 
  mutate(
    type = as.factor("other")
  ) %>% 
  pivot_longer(c(LT_seedling, sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total, winners),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))


```

 

###Write tidy data for each point
```{r, tidy regen and plot data, assign stocking label to each point}

tidy_regen_point <- 
  bind_rows(comm, diverse, interfere, other)
 
 
write_csv(tidy_regen_point, "regen_point_per_acre_2022_alpha_tidy.csv")

```




###Make Plot Data Tidy



```{r}

# not ready to run this code
# not ready to run this code
# not ready to run this code
# not ready to run this code

# 
plot_data_tidy <- #read the csv file having basal area calculations %>% 

# (need to input the calculated values of plot variables 
#  such as basal area per acre.  
  
  select(!c(baf,
            low_spp, inv_spp, mid_spp, upper_spp, vine_spp)) %>% 
  
  pivot_longer(c(disturb, slash, fern, herb, 
                 low_cov, inv_cov, mid_cov, upper_cov, vine_cov,
                  comm_stock, pre_comm_stock, diver_stock,
                  pre_diver_stock,  interf_woody_stock, fern_stock, all_intf,
                 winner_stock
                 ),
               names_to = "plot_attrib",
               values_to = "attrib_val") %>% 
  
  replace_na(list("attrib_val" = "no", comm_per_acre = 0, pre_comm_per_acre = 0,
                  diver_per_acre = 0, pre_diver_per_acre = 0, interfere_per_acre = 0,
                  pre_interfere_per_acre = 0))

write_csv(plot.data_tidy, "tidy_plot_2019.csv")




```



<!-- 
 includes factor and number variables...can't blend via "pivot_longer"
 pivot_longer(c(disturb, slash, fern, herb, 
                 low_cov, inv_cov, mid_cov, upper_cov, vine_cov,
                 ba, seedling_numb, comm_per_acre, comm_stock,
                 pre_comm_per_acre, pre_comm_stock,diver_per_acre, diver_stock,
                 pre_diver_per_acre, pre_diver_stock, interfere_per_acre, interfere_stock,
                 pre_interfere_per_acre, pre_interfere_stock
                 ),
               names_to = "plot_attrib",
               values_to = "attrib_val")
-->
