---
title: "Regen and Plot Input of Data and Cleaning/Tidy"
author: "Peter Smallidge"
date: "10/17/2020"
output: github_document
---


```{r}
# code from one of Maria's "notes" pages
# TOC may only work with html output

# title: "Lesson 13: Factors"
# output: 
#  html_document:
#    keep_md: yes 
#    toc: true
#    toc_depth: 2 
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Analysis of slash wall regeneation plot data - 2019

###Initialize Libraries
```{r, include=FALSE}

library(tidyverse)
library(skimr)        ## install.packages("skimr")
library(readxl)
#
# ALT plus hypen = <- (within chunk)
# ctrl plus shift plus m = %>% 
# ctrl + ALT + I = insert chunk
# ALT + - = insert assignment operator
# filter rows, select columns
# RMD comment = <!-- comment -->
# 
# 

```


### read data from csv file, check variables and data structure

<!-- note to self: in the "read_csv" code, after the file location the code
for 'col..." etc ensures that the specified column is adjusted to correct data type
the file "cheat sheet data-import" has details -->

###Input Regen Data
```{r, input regen data}
# Seedling data, located in the same folder as this 
# the initial read of the data revealed that some columns were incorrectly
# recognized when inputted. the code for "col_types=" was added in a second pass
# of "read_csv" to clarify the desired data type.

input.data <- read_csv("~/R/slash-wall-vegetation/regen-data-2019.csv",
                      col_types=cols(spp=col_character(), 
                                     sup.seed03=col_number(),
                                     sup.sapl=col_number(),
                                     point=col_character())) %>% 
  filter(spp != "ush") %>% 
  mutate_at(vars(harvest, point, location, spp, origin),
            list(factor)) %>% 
  filter(!point %in% c("864", "866", "810"))

  


  
#str(input.data)


```


###Input Plot Data 
```{r, input plot data}

plot.data <- read_excel("plot-data-2019.xlsx", 
    col_types = c("text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "numeric", 
        "text", "text", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric")) %>% 
  mutate_at(
    vars(harvest, season, location, point, disturb, slash, fern,
         herb, low_cov, low_spp, inv_cov, inv_spp, mid_cov, mid_spp,
         upper_cov, upper_spp, vine_cov, vine_spp),
    list(factor)
           ) %>% 
   filter(!point %in% c("864", "866", "810")) %>% 
   rename("year" = "season") %>% 
  mutate(mid_spp = recode(mid_spp, "swm" = "rhus")) %>%  #old = new

   mutate(harvest = fct_recode(harvest,
                             "gas_line" = "cgl",
                              "red_pine" = "crp",
                              "sta_rd" = "cwb",
                              "red_pine" = "rp",
                              "gas_line" = "gl",
                              "boot" = "boot",
                              "wedge" = "wedge"
   ))


#view(plot.data)


```


###Count Number of Sample Points
```{r, count number of sample points}

point.count <-  plot.data %>% # create variable for total plots for denominator of average density
  group_by(harvest, location, point) %>% 
  summarize(n =n()) %>% 
  mutate(count = n/n) %>% 
  arrange(point) %>% 
  select(harvest, location, point, count) 

# for each point, the value of "count" = 1
```


###Create plots for regen with no observations
```{r, filling out empty sample points}

#There are 248 (245, excluding gas line points) plots, 
#but 242 regen points that had observations in 2019. 6 are excluded because
#of no data. The missing points would cause an error for calculation
#of the number of stems per acre as an average of points.
#
#

regen.data <- full_join(point.count, input.data, by = "point") %>% 
  select(!c(harvest.y, location.y, other)) %>% 
#this assumes that all point numbers are unique. Using "unique(point.count$point)"
#in the console assures there are 248 unique point numbers

  replace_na(list(date = "other", spp = "001", origin = "u", radius.01 = 3.7, 
                  radius.other = 6,
                  sup.seed01 = 0, sup.seed02 = 0, sup.seed03 = 0, sup.sapl = 0,
                  exp.seed01 = 0, exp.seed02 = 0, exp.seed03 = 0, exp.sapl = 0,
                  top4 = 0
                  )) 
  
#date = character
#spp, origin = factor
#all other = numeric


#view(regen.data)

```


###data manipulations - per acre calculations & spp names
```{r, regen per acre at points - calculations}
#
#height classes: 
#   01 = 12" to 4.5ft, 
#   02 = 4.5 ft to 9ft
#   03 = 9 ft. to 1" dbh

regen.per.acre <- regen.data %>% 
    rename("harvest" = "harvest.x", "location" = "location.x") %>% 
  
    mutate(
    year = "2019",
    sup_short = sup.seed01 * (43560 /((3.14156) * (3.7 * 3.7))),
    sup_med = sup.seed02 * (43560 /((3.14156) * (6 * 6))),
    sup_tall = sup.seed03 * (43560 /((3.14156) * (6 * 6))),
    sup_sapl = sup.sapl * (43560 /((3.14156) * (6 * 6))),
    exp_short = exp.seed01 * (43560 /((3.14156) * (3.7 * 3.7))),
    exp_med = exp.seed02 * (43560 /((3.14156) * (6 * 6))),
    exp_tall = exp.seed03 * (43560 /((3.14156) * (6 * 6))),
    exp_sapl = exp.sapl * (43560 /((3.14156) * (6 * 6)))
  ) %>% 
  
   mutate_at(vars(year),
            list(factor)) %>% 
  #following code calculates the number of seedlings per acre
  #by species for each height class. Not yet tidy.
  mutate(
    suppressed = sup_short + sup_med + sup_tall,
    exposed = exp_short + exp_med + exp_tall,
    total = suppressed + exposed, 
        #exclude winners, which would double count
        #excludes saplings
    winners = top4 * (43560 /((3.14156) * (6 * 6))) 
  ) %>% 
  
  select(!c(sup.seed01, sup.seed02, sup.seed03, sup.sapl,
            exp.seed01, exp.seed02, exp.seed03, exp.sapl,
            date, radius.01, radius.other, top4)) %>% 
  
   mutate(harvest = fct_recode(harvest,
                             "gas_line" = "cgl",
                              "red_pine" = "crp",
                              "sta_rd" = "cwb",
                              "red_pine" = "rp",
                              "gas_line" = "gl",
                              "boot" = "boot",
                              "wedge" = "wedge"
)) %>% 
  mutate(origin = fct_recode(origin,
                             "root" = "r",
                             "stump" = "t",
                             "singl/unk" = "s",
                             "none" = "u"
)) %>% 
  
   mutate(spp = fct_recode(spp,
                          "mia" = "001",
                          "ab" = "531",
                          "stm" = "315",
                          "sm" = "318",
                          "rm" = "316",
                          "rp" = "125",
                          "wp" = "129",
                          "he" = "261",
                          "svb" = "356",
                          "yb" = "371",
                          "sb" = "372",
                          "pb" = "375",
                          "gb" = "379",
                          "ah" = "391",
                          "haw" = "500",
                          "wa" = "541",
                          "yp" = "621",
                          "cu" = "651",
                          "la" = "70",
                          "eh" = "701",
                          "asp" = "740",
                          "bta" = "743",
                          "qa" = "746",
                          "pc" = "761",
                          "bc" = "762",
                          "cc" = "763",
                          "wo" = "802",
                          "co" = "832",
                          "ro" = "833",
                          "bl" = "901",
                          "wi" = "920",
                          "ba" = "951",
                          "sp" = "90", 
                          "unk" = "999"
                          ))

#view(regen.per.acre)

write_csv(regen.per.acre, "untidy_regen_point_2019.csv")


```


###making data tidy - comm/regen
```{r, making data tidy commercial species}
# commercial regeneration species
# 316 318 372 541 621 762  802 832 833 951 
# 129, 261, 371,
# maple, yeb, swb, wa, yp, oaks, ba 
# no populus 740, 743, 746, no pab 375


comm <- regen.per.acre %>% 
  
  filter(spp %in% c("wp", "he", "rm", "sm","yb", "sb", "wa", "yp",
                     "bc", "wo", "co", "ro", "ba", "bl", "rp", "cu"
                    )) %>% 
  mutate(
    type = as.factor("commercial")
  ) %>% 
  pivot_longer(c(sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total, winners),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))

#view(comm)
  
```

###making data tidy - diverse/regen
```{r, making data tidy diverse}

# desirable diversity species (no interfering, but includes Populus, paper birch, not pin cherry)
# 129 261 316 318 356 372 375 500 541 621 743 746 762 802 832 833  90 920 951

diverse <- regen.per.acre %>% 
  
  filter(spp %in% c("svb", "pb", "gb", "haw", "asp", "bta", "qa",
                    "wi",
                    "wp", "he", "rm", "sm","yb", "sb", "wa", "yp",
                     "bc", "wo", "co", "ro", "ba", "bl", "rp", "cu"
                    )) %>% 
  # includes commercial spp.  avoid summation that would overestimat total stems
  # per acre
  
  mutate(
    type = as.factor("diversity")
  ) %>% 
  pivot_longer(c(sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total, winners),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))

#view(diverse)

```

###making data tidy - interfering species/regen
```{r, tidy for interfering species}

# interfering species

interfere <- regen.per.acre %>% 
  
  filter(spp %in% c("ab", "stm", "eh", "pc")) %>% 
  mutate(
    type = as.factor("interfering")
  ) %>% 
  pivot_longer(c(sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total, winners),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))

#view(interfere)
```

###making data tidy - other species
```{r}
# other species

other <- regen.per.acre %>% 
  
  filter(spp %in% c("mia", "ah", "la", "sp", "unk", "cc"
                    )) %>% 
  mutate(
    type = as.factor("interfering")
  ) %>% 
  pivot_longer(c(sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total, winners),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))

#view(other)
```

 

###Write tidy data assign stocking label to each point
```{r, tidy regen and plot data, assign stocking label to each point}

tidy_regen_point <- 
  bind_rows(comm, diverse, interfere, other)
 
 
write_csv(tidy_regen_point, "tidy_regen_point_2019.csv")


#for stocking breaks, assess heights assuming deer impact = 1 or 2
#for deer impact=1, there should be 15 stems less than 5 ft per 6 ft plot = 5775/ac
#for deer impact=2, there should be 30 stems less than 5 ft per 6 ft plot = 11550/ac
#sapling stocking is 2 per plot = 770
#

winner_stocking <- tidy_regen_point %>% 
  filter(type %in% c("commercial")) %>% 
  filter(size %in% c("winners")) %>% 
  group_by(harvest, location, point) %>% 
  summarize(
    win_comm_per_acre = sum(stems.per.acre)
  ) %>% 
  mutate(winner_stock = cut(win_comm_per_acre, breaks=c(0,385,770,1000000), 
                            labels=c("fail", "win_01", "win_02")
                            ))

comm_stocking <- tidy_regen_point %>% 
  filter(type %in% c("commercial")) %>% 
  filter(size %in% c("exp_med", "exp_tall", "exp_sapl")) %>% 
  group_by(harvest, location, point) %>% 
  summarize(
    comm_per_acre = sum(stems.per.acre)
  ) %>% 
  filter(comm_per_acre >769) %>% 
  mutate(comm_stock = cut(comm_per_acre, breaks=c(0, 770, 1000000),labels=c("comm_not","comm_stkd")))

pre_comm_stocking <- tidy_regen_point %>% 
  filter(type %in% c("commercial")) %>% 
  filter(size %in% c("exp_short")) %>% 
  group_by(harvest, location, point) %>% 
  summarize(
    pre_comm_per_acre = sum(stems.per.acre)
  ) %>% 
  filter(pre_comm_per_acre>5774) %>% 
  mutate(pre_comm_stock =  cut(pre_comm_per_acre, 
                               breaks=c(0, 5775, 11550, 1000000),
                               labels=c("pre_comm_not","pre_comm_deer1", "pre_comm_deer2")))


diver_stocking <- tidy_regen_point %>% 
  filter(type %in% c("diversity")) %>% 
  filter(size %in% c("exp_med", "exp_tall", "exp_sapl")) %>% 
  group_by(harvest, location, point) %>% 
  summarize(
    diver_per_acre = sum(stems.per.acre)
  ) %>% 
  filter(diver_per_acre > 769) %>% 
  #mutate(diver_stock = "div_stkd")
  mutate(diver_stock = cut(diver_per_acre, breaks=c(0, 770, 1000000),labels=c("div_not","div_stkd")))

pre_diver_stocking <- tidy_regen_point %>% 
  filter(type %in% c("diversity")) %>% 
  filter(size %in% c("exp_short")) %>% 
  group_by(harvest, location, point) %>% 
  summarize(
    pre_diver_per_acre = sum(stems.per.acre)
  ) %>% 
  filter(pre_diver_per_acre > 5774) %>% 
  mutate(pre_diver_stock = 
           cut(pre_diver_per_acre, breaks=c(0, 5775, 11550, 1000000),labels=c("pre_div_not","pre_diver_deer1", "pre_diver_deer2")))


#according to SILVAH ne_gtr96, p.46,table 5
#a plot (radius 6 ft) has interference if number of stems less than 12" is 12 or > 12" is 6
#or fern covers more than 30% of the plot ("m" or "h")
#12*385 = 4620 (these weren't recorded); 6 *385 = 2310

interfere_stocking <- tidy_regen_point %>% 
  filter(type %in% c("interfering")) %>% 
  filter(size %in% c("total")) %>%  #recovers sum of stem count for all exposed height classes per plot
  group_by(harvest, location, point) %>% 
  summarize(
    interfere_per_acre = sum(stems.per.acre)
  ) %>% 
  filter(interfere_per_acre > 2310) %>% 
  mutate(interfere_stock = "inft_stkd") %>% 
  mutate_at(vars(interfere_stock), list (factor))
#  mutate(interfere_stock = cut(interfere_per_acre, breaks=c(0, 2310, #1000000),labels=c("intf_not","intf_stkd")))

fern_stocking <- plot.data %>%
  select(harvest, location, point, fern) %>% 
  filter(fern %in% c("m", "h")) %>% 
  group_by(harvest, location, point) %>% 
  mutate(fern_stock = "intf_stkd"
          ) %>% 
  mutate_at(vars(fern_stock),
            list(factor)
  )


all_interference <- 
  full_join(interfere_stocking, fern_stocking, by = "point") %>% 
  select(!c(harvest.y, location.y)) %>% 
  mutate(
      all_intf = "all_intf"
  ) %>% 
  mutate_at(vars(all_intf), list(factor)) %>% 
  rename("harvest" = "harvest.x", "location" = "location.x") %>% 
  arrange(point) %>% 
  select(c(point, interfere_per_acre, interfere_stock, fern_stock, all_intf))
#
#join the plots
#
#

plot.data_01 <- 
  left_join(plot.data, comm_stocking,     by = "point") %>% 
  select(!c(harvest.y, location.y)) %>% 
  rename("harvest" = "harvest.x", "location" = "location.x") 

plot.data_02 <- 
  left_join(plot.data_01, pre_comm_stocking,     by = "point") %>% 
  select(!c(harvest.y, location.y)) %>% 
  rename("harvest" = "harvest.x", "location" = "location.x") 

plot.data_03 <- 
  left_join(plot.data_02, diver_stocking,     by = "point") %>% 
  rename("harvest" = "harvest.x", "location" = "location.x") 

plot.data_04 <- 
  left_join(plot.data_03, pre_diver_stocking,     by = "point") %>% 
  rename("harvest" = "harvest.x", "location" = "location.x") %>% 
  select(!c("harvest.y", "harvest.y.y", "location.y", "location.y.y"))
 
plot.data_05 <- 
  left_join(plot.data_04, interfere_stocking,     by = "point") %>% 
  select(!c(harvest.y, location.y)) %>%  
  rename("harvest" = "harvest.x", "location" = "location.x") 

plot.data_partial <- 
  left_join(plot.data_05, all_interference,     by = "point") %>% 
  rename( "seedling_numb" = "seedling-numb", "interf_woody_stock"="interfere_stock.x", 
          "interf_woody_per_acre" = "interfere_per_acre.x") %>% 
select(!c(interfere_stock.y, interfere_per_acre.y)) %>% 
  select(!c(oakhick, beech, maple, birch, hdwd, conifer)) 

plot.data_full <- 
  left_join(plot.data_partial, winner_stocking,   by = "point" )%>% 
  select(!c(harvest.y, location.y)) %>% 
  rename("harvest" = "harvest.x", "location" = "location.x") %>% 
  arrange(harvest, location, point) %>% 
select(harvest, location, point,everything() )




  
write_csv(plot.data_full, "untidy_plot_2019.csv")
```




###Make Plot Data Tidy
```{r}

plot.data_tidy <- plot.data_full %>% 
  
  select(!c(baf,
            low_spp, inv_spp, mid_spp, upper_spp, vine_spp)) %>% 
  
  pivot_longer(c(disturb, slash, fern, herb, 
                 low_cov, inv_cov, mid_cov, upper_cov, vine_cov,
                  comm_stock, pre_comm_stock, diver_stock,
                  pre_diver_stock,  interf_woody_stock, fern_stock, all_intf,
                 winner_stock
                 ),
               names_to = "plot_attrib",
               values_to = "attrib_val") %>% 
  
  replace_na(list("attrib_val" = "no", comm_per_acre = 0, pre_comm_per_acre = 0,
                  diver_per_acre = 0, pre_diver_per_acre = 0, interfere_per_acre = 0,
                  pre_interfere_per_acre = 0))

write_csv(plot.data_tidy, "tidy_plot_2019.csv")




```



<!-- 
 includes factor and number variables...can't blend via "pivot_longer"
 pivot_longer(c(disturb, slash, fern, herb, 
                 low_cov, inv_cov, mid_cov, upper_cov, vine_cov,
                 ba, seedling_numb, comm_per_acre, comm_stock,
                 pre_comm_per_acre, pre_comm_stock,diver_per_acre, diver_stock,
                 pre_diver_per_acre, pre_diver_stock, interfere_per_acre, interfere_stock,
                 pre_interfere_per_acre, pre_interfere_stock
                 ),
               names_to = "plot_attrib",
               values_to = "attrib_val")
-->
