---
title: "Regneration Analysis"
author: "Peter Smallidge"
date: "10/17/2020"
output: github_document
---


```{r}
# code from one of Maria's "notes" pages
# TOC may only work with html output

# title: "Lesson 13: Factors"
# output: 
#  html_document:
#    keep_md: yes 
#    toc: true
#    toc_depth: 2 
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Analysis of slash wall regeneation plot data - 2019


```{r, include=FALSE}

library(tidyverse)
library(skimr)        ## install.packages("skimr")
#
# ALT plus hypen = <- (within chunk)
# ctrl plus shift plus m = %>% 
# ctrl + ALT + I = insert chunk
# ALT + - = insert assignment operator
# filter rows, select columns
# RMD comment = <!-- comment -->
# 
# 



```




### read data from csv file, check variables and data structure

<!-- note to self: in the "read_csv" code, after the file location the code
for 'col..." etc ensures that the specified column is adjusted to correct data type
the file "cheat sheet data-import" has details -->

###Input Regen Data
```{r, input regen data}
# Seedling data, located in the same folder as this 
# the initial read of the data revealed that some columns were incorrectly
# recognized when inputted. the code for "col_types=" was added in a second pass
# of "read_csv" to clarify the desired data type.

input.data <- read_csv("~/R/slash-wall-vegetation/regen-data-2019.csv",
                      col_types=cols(spp=col_character(), 
                                     sup.seed03=col_number(),
                                     sup.sapl=col_number(),
                                     point=col_character())) %>% 
  filter(spp != "ush") %>% 
  mutate_at(vars(harvest, point, location, spp, origin),
            list(factor)
  )


  
#str(input.data)


```


###Input Plot Data
```{r, input plot data}

plot.data <- read_csv("~/R/slash-wall-vegetation/plot-data-2019.csv",
                              col_types = cols(point=col_character(),
                                              upper_spp=col_character(),
                                               baf=col_number(),
                                              oakhick=col_number(),
                                              beech=col_number(),
                                              maple=col_number(),
                                              birch=col_number(),
                                              hdwd=col_number(),
                                              conifer=col_number(),
                                              ba=col_number()
                                              )) %>% 
  mutate_at(
    vars(harvest, location:vine_spp),
    list(factor)
  ) %>% 
  mutate(mid_spp = recode(mid_spp, "swm" = "rhus")) #old = new

view(plot.data)


```


###Count Number of Points
```{r, count number of points}

point.count <-  plot.data %>% # create variable for total plots for denominator of average density
  group_by(harvest, location, point) %>% 
  summarize(n =n()) %>% 
  mutate(count = n/n) %>% 
  arrange(point) %>% 
  select(harvest, location, point, count) 


```


###Create plots for regen with no observations
```{r}

#there are 248 plots, but 242 regen points. 6 are excluded because
#of no data. The missing points would cause an error for calculation
#of the number of stems per acre as an average of points.
#
#

regen.data <- full_join(point.count, input.data, by = "point") %>% 
  select(!c(harvest.y, location.y, other)) %>% 
#this assumes that all point numbers are unique. Using "unique(point.count$point)"
#in the console assures there are 248 unique point numbers

  replace_na(list(date = "other", spp = "001", origin = "u", radius.01 = 3.7, 
                  radius.other = 6,
                  sup.seed01 = 0, sup.seed02 = 0, sup.seed03 = 0, sup.sapl = 0,
                  exp.seed01 = 0, exp.seed02 = 0, exp.seed03 = 0, exp.sapl = 0,
                  top4 = 0
                  )) 
  
#date = character
#spp, origin = factor
#all other = numeric


view(input.data2)

```


###data manipulations - per acre calculations, spp names
```{r}
regen.per.acre <- regen.data %>% 
    rename("harvest" = "harvest.x", "location" = "location.x") %>% 
  
    mutate(
    year = "2019",
    sup_short = sup.seed01 * (43560 /((3.14156) * (3.7 * 3.7))),
    sup_med = sup.seed02 * (43560 /((3.14156) * (6 * 6))),
    sup_tall = sup.seed03 * (43560 /((3.14156) * (6 * 6))),
    sup_sapl = sup.sapl * (43560 /((3.14156) * (6 * 6))),
    exp_short = exp.seed01 * (43560 /((3.14156) * (3.7 * 3.7))),
    exp_med = exp.seed02 * (43560 /((3.14156) * (6 * 6))),
    exp_tall = exp.seed03 * (43560 /((3.14156) * (6 * 6))),
    exp_sapl = exp.sapl * (43560 /((3.14156) * (6 * 6)))
  ) %>% 
  
   mutate_at(vars(year),
            list(factor)) %>% 
  
  mutate(
    suppressed = sup_short + sup_med + sup_tall,
    exposed = exp_short + exp_med + exp_tall,
    total = suppressed + exposed, 
        #exclude winners, which would double count
        #excludes saplings
    winners = top4 * (43560 /((3.14156) * (6 * 6))) 
  ) %>% 
  
  select(!c(sup.seed01, sup.seed02, sup.seed03, sup.sapl,
            exp.seed01, exp.seed02, exp.seed03, exp.sapl,
            date, radius.01, radius.other, top4)) %>% 
  
   mutate(harvest = fct_recode(harvest,
                             "gas_line" = "cgl",
                              "red_pine" = "crp",
                              "sta_rd" = "cwb",
                              "red_pine" = "rp",
                              "gas_line" = "gl",
                              "boot" = "boot",
                              "wedge" = "wedge"
)) %>% 
  mutate(origin = fct_recode(origin,
                             "root" = "r",
                             "stump" = "t",
                             "singl/unk" = "s",
                             "none" = "u"
)) %>% 
  
   mutate(spp = fct_recode(spp,
                          "mia" = "001",
                          "ab" = "531",
                          "stm" = "315",
                          "sm" = "318",
                          "rm" = "316",
                          "rp" = "125",
                          "wp" = "129",
                          "he" = "261",
                          "svb" = "356",
                          "yb" = "371",
                          "sb" = "372",
                          "pb" = "375",
                          "gb" = "379",
                          "ah" = "391",
                          "haw" = "500",
                          "wa" = "541",
                          "yp" = "621",
                          "cu" = "651",
                          "la" = "70",
                          "eh" = "701",
                          "asp" = "740",
                          "bta" = "743",
                          "qa" = "746",
                          "pc" = "761",
                          "bc" = "762",
                          "cc" = "763",
                          "wo" = "802",
                          "co" = "832",
                          "ro" = "833",
                          "bl" = "901",
                          "wi" = "920",
                          "ba" = "951",
                          "sp" = "90", 
                          "unk" = "999"
                          ))

view(regen.per.acre)


```


###making data tidy - comm
```{r, making data tidy}
# commercial regeneration species
# 316 318 372 541 621 762  802 832 833 951 
# 129, 261, 371,
# maple, yeb, swb, wa, yp, oaks, ba 
# no populus 740, 743, 746, no pab 375


comm <- regen.per.acre %>% 
  
  filter(spp %in% c("wp", "he", "rm", "sm","yb", "sb", "wa", "yp",
                     "bc", "wo", "co", "ro", "ba", "bl", "rp", "cu"
                    )) %>% 
  mutate(
    type = as.factor("commercial")
  ) %>% 
  pivot_longer(c(sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total, winners),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))

view(comm)
  
```

###making data tidy - diverse
```{r, making data tidy diverse}

# desirable diversity species (no interfering, but includes Populus, paper birch, not pin cherry)
# 129 261 316 318 356 372 375 500 541 621 743 746 762 802 832 833  90 920 951

diverse <- regen.per.acre %>% 
  
  filter(spp %in% c("svb", "pb", "gb", "haw", "asp", "bta", "qa",
                    "wi")) %>% 
  mutate(
    type = as.factor("diversity")
  ) %>% 
  pivot_longer(c(sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total, winners),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))

view(diverse)

```

###making data tidy - interfering species
```{r, tidy for interfering species}

# interfering species

interfere <- regen.per.acre %>% 
  
  filter(spp %in% c("ab", "stm", "eh", "pc")) %>% 
  mutate(
    type = as.factor("interfering")
  ) %>% 
  pivot_longer(c(sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total, winners),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))

view(interfere)
```

###making data tidy - other species
```{r}
# other species

other <- regen.per.acre %>% 
  
  filter(spp %in% c("mia", "ah", "la", "sp", "unk", "cc"
                    )) %>% 
  mutate(
    type = as.factor("interfering")
  ) %>% 
  pivot_longer(c(sup_short, sup_med, sup_tall, sup_sapl,
                 exp_short, exp_med, exp_tall, exp_sapl,
                 suppressed, exposed, total, winners),
               names_to = "size",
               values_to = "stems.per.acre") %>% 
  
   mutate_at(vars(size),
            list(factor))

view(other)
```


```

###Write tidy data
```{r}

```






write_csv(avid_tidy, "tidy_avid_2018-2020_gasline.csv")