---
title: "seedling height input and formatting"
author: "Peter Smallidge"
date: "10/16/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Analysis of seedling heights 2018 & 2022

```{r, include=FALSE}

library(tidyverse)
library(skimr)        ## install.packages("skimr")
library(readxl)
library(readr)


#
# ALT plus hypen = <- (within chunk)
# ctrl plus shift plus m = %>% 
# ctrl + ALT + I = insert chunk
# filter rows, select columns
# 

```

### read data from csv file, check variables and data structure

<!-- note to self: in the "read_csv" code, after the file location the code
for 'col..." etc ensures that the specified column is adjusted
the file "cheat sheet data-import" has details 

It is also possible to right-click a file name and import excel


-->






```{r}
input.data.2022 <- read_excel("slash.wall.data.2022.16may2023_final.xlsx", 
    sheet = "growth-height") %>% 
    mutate(
    point = as.character(point),
    location = as.character(location),
    location = point
    ) %>% 
    mutate(
      wall = "neutral"
    ) %>% 
  select(harvest, season,date,wall, location, point, spp, origin, basal, height) %>% 
    mutate(
      point = as.character(point)
    )
```


#import point x location 

```{r}
point_location_seedling <- read_excel("location_by_point.xlsx") %>% 
    mutate(
      point = as.character(point)
    )

```


```{r}
height_with_location <- full_join(input.data.2022, point_location_seedling, by = (c("harvest", "point"))) %>% 
    select(!c(location.x, wall.x)) %>%  
        drop_na %>% 
#        droplevels() %>% 
    rename(
      location = location.y,
      wall = wall.y
      ) %>%  # new variable name = old variable name
    replace_na(list(date = "other",  origin = "u", location = "u", spp = 001
           )) %>%   #need to change missing values to real values while variable as character
    arrange(harvest, point, spp) %>% 
    mutate(
      year = season
    ) %>% 
    mutate(harvest = fct_recode(harvest,
                              "gas_line" = "gasline",
                              "red_pine" = "redpine",
                              "sta_rd" = "stationrd",
                              "recknagel" = "rectangle1",
                              "recknagel" = "rectangle2",
                              "recknagel" = "rectangle3",
                              "recknagel" = "rectangle4",
                              "circle" = "circle1",
                              "circle" = "circle2",
                              "circle" = "circle3",
                              "sta_rd" = "boot",
                              "sta_rd" = "wedge"

    )) %>% 
    # mutate(
    #   cut = if_else((harvest %in% c("gas_line", "red_pine")), 2017, 2018),
    #   cut = if_else((harvest %in% c("boot", "wedge", "sta_rd")), 2018, 2020),
    #   cut = if_else((harvest %in% c("recknagel", "circle", "decker")), 2020, 9999),
    # )

    mutate(
      cut = case_when(
        (harvest %in% c("gas_line", "red_pine")) ~ 2017,
        (harvest %in% c("boot", "wedge", "sta_rd")) ~ 2018,
        (harvest %in% c("campridge")) ~ 2019,
        (harvest %in% c("recknagel", "circle", "decker")) ~ 2020
      ),
        number_yrs_grow = year - cut + 1
       )
```
  The last mutate code changes the value of "cut" to 2017 if the harvest name is list. otherwise the year of cut is 2019
  
  (from chatGPT, for two conditions)
  mutate(
    value = if_else(value > 20, value * 2, value),
  )

(from chatGPT, for multiple conditions)
  mutate(z = case_when(
    x > 2 & y > 8 ~ "high_x_and_high_y",
    x > 2 & y <= 8 ~ "high_x_and_low_y",
    x <= 2 & y > 8 ~ "low_x_and_high_y",
    x <= 2 & y <= 8 ~ "low_x_and_low_y"
  ))

```{r}
  seedling_ht_2022_alpha <- height_with_location %>% 
    mutate (
      spp = as.character(spp)
      )%>% 
   mutate(spp = fct_recode(spp,   #comment out species not present in data
                           "ab" = "531",
                         # "stm" = "315",
                          "sm" = "318",
                          "rm" = "316",
                          "rp" = "125",
                          "wp" = "129",
                         "he" = "261",
                          #"svb" = "356",
                          #"svb" = "355",
                          #"yb" = "371",
                          "sb" = "372",
                          #"pb" = "375",
                         #"gb" = "379",
                         # "ah" = "391",
                          #"haw" = "500",
                          "wa" = "541",
                          #"yp" = "621",
                          "cu" = "651",
                          #"la" = "70",
                          "eh" = "701",
                          #"asp" = "740",
                          #"bta" = "743",
                          #"qa" = "746",
                          #"pc" = "761",
                          "bc" = "762",
                          #"cc" = "763",
                          #"wo" = "802",
                          "co" = "832",
                          "ro" = "833",
                          #"bl" = "901",
                          #"wi" = "920",
                          "ba" = "951",
                         # "sp" = "90", 
                         # "sa" = "931",
                          #"wh" = "585",
                          "sh" = "407",
                         "spr" = "90",
                          #"ns" = "91",
                          "bh" = "402",
                          #"au" = "535",
                          #"ec" = "742",
                          #"unk" = "999"
                          ))

#view(regen.per.acre)

write_csv(seedling_ht_2022_alpha, "seedling_height_2022_alpha_point_untidy.csv")




```

