---
title: "determine_silvah_stocking_levels"
author: "Peter Smallidge"
date: "5/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


###Initialize Libraries
```{r, include=FALSE}

library(tidyverse)
library(skimr)        ## install.packages("skimr")
library(readxl)
library(readr)
#
# ALT plus hypen = <- (within chunk)
# ctrl plus shift plus m = %>% 
# ctrl + ALT + I = insert chunk
# ALT + - = insert assignment operator
# filter rows, select columns
# RMD comment = <!-- comment -->
# 
# 

```



###Write tidy data assign stocking label to each point
```{r, tidy regen and plot data, assign stocking label to each point}

stem_counts <- read_csv("tidy_regen_point_2019.csv") %>% 
  mutate_at(vars(harvest, point, location, spp, origin, type, size),
            list(factor)) %>% 
  mutate(harvest = fct_recode(harvest,
                              "sta_rd" = "boot",
                              "sta_rd" = "wedge")) %>% 
  droplevels() #drops unused levels of a factor
  
plot_data <- read_csv("untidy_plot_2019.csv") %>% 
  select("harvest", "location", "point", "year", "disturb", "slash", "fern", "herb",
         "low_cov", "inv_cov", "mid_cov", "upper_cov","vine_cov", "ba", "seedling_numb") %>% 
  mutate_at(vars(harvest, location, point, disturb, slash, fern, herb, 
                 low_cov, inv_cov, mid_cov, upper_cov, vine_cov),
            list(factor))


  
     
```

### set stocking levels for silvah thresholds
```{r SILVAH stocking thresholds}
# guide for pre-harvest assessment and prescription
#
#for stocking breaks, assess heights assuming deer impact = 1 or 2
#for deer impact=1, there should be 15 stems less than 5 ft per 6 ft plot = 5775/ac
#for deer impact=2, there should be 30 stems less than 5 ft per 6 ft plot = 11550/ac
#sapling stocking is 2 per plot = 770
#
#The basis for stocking thresholds is from SILVAH ne-gtr-96.
#prescription stocking thresholds are described on print-page-40, document page 46.
#post-harvest assessment of stocking is described on print-page-74, document page 80
#the SILVAH guidelines do not differentiate between overtopped and exposed (free to grow) seedlings.
#
# p74  post-harvest assessment at 2 years.
# on 6ft radius plots ~"stocked successfully if 
# (i) "silvah01" 25 stems/plot > 12 inches [=9625/ac, 23774/ha],
# (ii) "silvah02" 5 stem > 3 ft tall [1925/ac, 4755/ha], -or- 
# (iii) "silvah03" 2 stems > 5 ft [770/ac, 1901/ha].
# Once 70% of plots have (iii) stocking, the stand is considered regenerated.


silvah01 <- stem_counts %>% 
  filter(type %in% c("commercial")) %>% 
  filter(size %in% c("exp_short", "exp_med", "exp_tall")) %>% 
  group_by(harvest, location, point) %>% 
  summarize(
    silvah01_per_acre = (sum(stems.per.acre))
  ) %>% 
  filter(silvah01_per_acre > 9624) %>% 
   mutate(silvah25 = cut(silvah01_per_acre, breaks=c(0,9625,9000000000),
    labels=c("not", "stocked25"))) 

silvah02 <- stem_counts %>% 
  filter(type %in% c("commercial")) %>% 
  filter(size %in% c("sup_med", "exp_med", "exp_tall")) %>% 
  group_by(harvest, location, point) %>% 
  summarize(
    silvah02_per_acre = (sum(stems.per.acre))
  ) %>% 
  mutate(silvah5 = cut(silvah02_per_acre, breaks=c(0,1925,9000000000),
    labels=c("not", "stocked5")))

silvah03 <- stem_counts %>% 
  filter(type %in% c("commercial")) %>% 
  filter(size %in% c("sup_tall", "exp_med", "exp_tall")) %>% 
  group_by(harvest, location, point) %>% 
  summarize(
    silvah03_per_acre = (sum(stems.per.acre))
  ) %>% 
  mutate(silvah2 = cut(silvah03_per_acre, breaks=c(0,770,9000000000),
    labels=c("not", "stocked2")))
```

### create df for each threshold and diversity species
#### edge join each to allow a factor to capture a positive in any of the three indicies

```{r SILVAH diversity stocking thresholds}
silvah01div <- stem_counts %>% 
  filter(type %in% c("commercial", "diversity")) %>% 
  filter(size %in% c("exp_short", "exp_med", "exp_tall")) %>% 
  group_by(harvest, location, point) %>% 
  summarize(
    silvah01div_per_acre = (sum(stems.per.acre))
  ) %>% 
  mutate(div_stk25 = cut(silvah01div_per_acre, breaks=c(0,9625,9000000000),
    labels=c("not", "yes")))

silvah02div <- stem_counts %>% 
  filter(type %in% c("commercial", "diversity")) %>% 
  filter(size %in% c("exp_med", "exp_tall")) %>% 
  group_by(harvest, location, point) %>% 
  summarize(
    silvah02div_per_acre = (sum(stems.per.acre))
  ) %>% 
  mutate(div_stk5 = cut(silvah02div_per_acre, breaks=c(0,1925,9000000000),
    labels=c("not", "yes")))

silvah03div <- stem_counts %>% 
  filter(type %in% c("commercial", "diversity")) %>% 
  filter(size %in% c("sup_tall", "exp_med", "exp_tall")) %>% 
  group_by(harvest, location, point) %>% 
  summarize(
    silvah03div_per_acre = (sum(stems.per.acre))
  ) %>% 
  mutate(div_stk2 = cut(silvah03div_per_acre, breaks=c(0,770,9000000000),
    labels=c("not", "yes")))

div_join1 <- full_join(silvah01div, silvah02div, by = c("harvest", "location", "point"))

div_join2 <- full_join(div_join1, silvah03div, by = c("harvest", "location", "point"))

diver_stocking <- div_join2 %>% 
  filter(
    div_stk2 %in% c("yes") |  # if started with div_stk25, no additional obs were added
    div_stk5 %in% c("yes") |
    div_stk25 %in% c("yes")
  ) %>% 
  mutate(
    div_stock = "stocked_div"
  ) %>% 
  select(harvest, location, point, silvah01div_per_acre, div_stock)


```

### Interference Stocking Thresholds for Woody, fern, All
```{r Interference Stocking Thresholds for Woody, fern, All}
#according to SILVAH ne_gtr96, p.46,table 5
# (this is for prescriptions, pre-harvest. No guidance on post harvest INTF thresholds)
# 
#a plot (radius 6 ft) has interference if:
# (i) number of stems less than 12" is 12, [12 stems * 385 = 4620/ac and 11411/ha] 
# (ii) number of stems > 12" is 6 [6 stems * 385 = 2310/ac and 5706/ha],
# or
# (iii) fern covers more than 30% of the plot ("m" or "h")


interfere_stocking <- stem_counts %>% 
  filter(type %in% c("interfering")) %>% 
  filter(size %in% c("total")) %>%  
      #variable total is the sum of all suppressed and exposed stems
      #in this case of type = interfering
  group_by(harvest, location, point) %>% 
  summarize(
    interfere_per_acre = sum(stems.per.acre)
  ) %>% 
  filter(interfere_per_acre > 2309) %>% 
  mutate(woody_stock = "inft_stkd") %>% 
  mutate_at(vars(woody_stock), list (factor))



fern_stocking <- plot.data %>%
  select(harvest, location, point, fern) %>% 
 # filter(fern %in% c("m", "h")) %>% 
  filter(fern %in% c("m")| fern %in% c("h")) %>% 
  group_by(harvest, location, point) %>% 
  mutate(fern_stock = "intf_stkd"
          ) %>% 
  mutate_at(vars(fern_stock),
            list(factor)
  )


all_interference <- 
  full_join(interfere_stocking, fern_stocking,  by = c("harvest", "location", "point")) %>% 
  mutate(
      all_intf = "all_intf"
  ) %>% 
  mutate_at(vars(all_intf), list(factor)) %>% 
  arrange(harvest, location, point) %>% 
  select(c(harvest, location, point, interfere_per_acre, woody_stock, fern_stock, all_intf))

```


### make a wide data frame with each of the stocking variables and stocking amounts
```{r edge join the stocking data frames}

#
#join the plots
#


plot.data_01 <- 
  left_join(plot_data, silvah01,     by = c("harvest", "location", "point"))

plot.data_02 <- 
  left_join(plot.data_01, silvah02,     by = c("harvest", "location", "point"))

plot.data_03 <- 
  left_join(plot.data_02, silvah03,     by = c("harvest", "location", "point"))

plot.data_04 <- 
  left_join(plot.data_03, diver_stocking,     by = c("harvest", "location", "point"))

plot.data_partial <- 
  left_join(plot.data_04, all_interference,      by = c("harvest", "location", "point")) %>%  
  arrange(harvest, location, point) %>% 
  select(harvest, location, point, everything() )

plot.data_full <- plot.data_partial %>% 
  #remove missing from character variables  
  mutate(  
    silvah25=replace_na(silvah25, "not"),
    silvah5 = replace_na(silvah5, "not"),
    silvah2 = replace_na(silvah2, "not"),
    div_stock = replace_na(div_stock, "not"),
    woody_stock = replace_na(woody_stock, "not"),
    fern_stock = replace_na(fern_stock, "not"),
    all_intf = replace_na(all_intf, "not")
    ) %>% 
  
  # remove missing from numeric variables
  mutate(
    silvah01_per_acre = replace_na(silvah01_per_acre, 0),
    silvah02_per_acre = replace_na(silvah02_per_acre, 0),
    silvah03_per_acre = replace_na(silvah03_per_acre, 0),
    silvah01div_per_acre = replace_na(silvah01div_per_acre, 0),
    interfere_per_acre = replace_na(interfere_per_acre, 0),
 ) %>% 

  #change character to factor
  mutate_at(vars(harvest, year, point, location, 
                 disturb, slash, fern, herb, 
                 low_cov, inv_cov, mid_cov, upper_cov, vine_cov,
                 silvah25, silvah5, silvah2, div_stock, woody_stock,
                 fern_stock, all_intf),
            list(factor)) 


write_csv(plot.data_full, "untidy_plot_2019_stocking.csv")
```




###Make Plot Data Tidy
```{r}

plot.data_tidy <-  read_csv("untidy_plot_2019_stocking.csv") %>% 
  
  pivot_longer(c(disturb, slash, fern, herb, 
                 low_cov, inv_cov, mid_cov, upper_cov, vine_cov,
                 silvah25, silvah5, silvah2, div_stock, woody_stock, fern_stock, all_intf,
                ),
               names_to = "plot_attrib",
               values_to = "attrib_val") %>% 
  
  replace_na(list("attrib_val" = "not"
                  ))

write_csv(plot.data_tidy, "tidy_plot_2019.csv")



```

