---
title: "slash wall analyses 2025"
author: "Peter Smalldige"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# activate library functions
```{r message=FALSE}

library(ggplot2)
library(ggpattern)
#library(ggpubr)
library(expss)
library(readr)
library(readxl)
library(writexl)
library(skimr)
library(tidyverse)  # includes: tidyr, dplyr, haven, readr, readxl, forcats, stringer, ggplot2. See all with tidyverse_packages()
library(knitr)
library(janitor)
#install.packages("openxlsx")
library(openxlsx)
# install.packages("shinyaddins")
# library(shinyaddins)

```




### code 01
### Create file path input and the output folder (code 01) 

file paths, input and output
```{r}
#complete data
file_path_veg_complete <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\veg_data_2019to2023_complete(19feb2025)(read_me).xlsx"
file_path_denom <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\count_of_points_stand_by_protection.xlsx"


output_folder <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\"
```


# read denominators, but they mostly aren't used
```{r}
denom <- read_excel(file_path_denom, sheet = "Sheet 1") %>% 
    rename(
        harvest = stand,
        trmt = status
        ) %>%   
    mutate(
      harvest = case_when(
      harvest == "boot" ~ "sta_rd",
      harvest == "circle1" ~ "circle",
      harvest == "circle2" ~ "circle",
      harvest == "circle3" ~ "circle",
      harvest == "cricle1" ~ "circle",
      harvest == "gasline" ~ "gas_line",
      harvest == "recksouth" ~ "recknagel",
      harvest == "redpine" ~ "red_pine",
      harvest == "stationrd" ~ "sta_rd",
      harvest == "wedge" ~ "sta_rd",
      harvest == "old" ~ "new",
      TRUE ~ harvest  # Keeps other values unchanged
    )
      )

denom_stand_trmt <- denom %>% 
    group_by(harvest, trmt) %>% 
    summarize(
        denom = sum(wall_denom)
    ) %>% 
    mutate(
        filename = "slash_wall_analyses_2025.Rmd"
    ) %>% 
    select(filename, everything())

output_file <- file.path(output_folder, paste0("sample_size_table(16feb2025)", ".xlsx"))
write_xlsx(denom_stand_trmt, output_file)


```


#Notes
# Input raw data - data are complete at the level of the wall
```{r}
d0 <- read_excel(file_path_veg_complete, sheet = "data", skip = 2) %>% 
    left_join(denom_stand_trmt, by = c("harvest", "trmt")) %>% 
        mutate(
        GTR5ft = supD + supP + expD + expP,
#        total = supB + supC1 + supC2 + supP + expB + expC1 + expC2 + expP, variable already exists
        all_sizes = seedA + total
    ) %>% 
        mutate(spp = case_when(
        spp %in% c("004") ~ "ohw",
        TRUE ~ spp,  # Keeps the existing value of "type" for all other cases
    )) %>%
    mutate(type = case_when(
            type %in% c("blank") ~ "commercial",
            TRUE ~ type,  # Keeps the existing value of "type" for all other cases
        ),
        origin = case_when(
            origin %in% c("unknown") ~ "single",
            TRUE ~ origin,  # Keeps the existing value of "origin" for all other cases
        )
    ) %>% 
   mutate(type = case_when(
        spp %in% c("sb") ~ "sb",
        TRUE ~ type,  # Keeps the existing value of "type" for all other cases
    )) %>%     
    arrange(harvest, season, trmt, wall, point, spp) %>% 
    select( -upperexposed, -stemcount, -total, -GTR5ft, -all_sizes)

d0_sub <- d0 %>% 
    filter(harvest == "6_6" & season == "2023") 


#
#change date in file name as needed
#
output_file <- file.path(output_folder, paste0("input_data_subset(d0)(5march2025)", ".xlsx"))
write_xlsx(d0_sub, output_file)

```




#Pre-process df for subsequent manipulations. Sum within species
```{r}
d1.1 <- d0 %>% #aggregate or average to collapse "origin"
# use d1a    filter(harvest == "6_6" & season == "2023") %>% 
    group_by(harvest, season, wall, point, spp) %>%  #combines origin into spp
    summarize(
        filename = first(filename),
        point_count = n_distinct(point),
        n = n(),
        trmt = first(trmt),
        type = first(type),
        firstgrow = first(firstgrow),
        year = first(year),
        denom = first(denom),
        growingseasons = first(growingseasons),
        height = first(height),
        clumps.acre = sum(clumps.acre),  # Sum all specified numeric clumps.acre and stems.ramet calc confirmed)
        stems.ramet = sum(stems.ramet),
        seedA = sum(seedA),
        supB = sum(supB),
#        supC1 = sum(supC1), C! and C2 don't exist in 2023 data
#        supC2 = sum(supC2),
        supD = sum(supD),
        supP = sum(supP),
        expB = sum(expB),
#        expC1 = sum(expC1),
#        expC2 = sum(expC2),
        expD = sum(expD),
        expP = sum(expP),
        # across(all_of(numeric_vars1), sum, na.rm = TRUE),  
        # across(all_of(numeric_vars2), mean, na.rm = TRUE),  # average of all specified numeric vars2
        .groups = "drop"
    ) %>% 
    mutate(
        GTR5ft = supD + supP + expD + expP,
        total = supB + supD + supP + expB + expD + expP, 
        all_sizes = seedA + total
    )

#
# go to slash_veg_complete and search for "cafri_point_labels" and where 6_6, ohw seedA wall 11 didn't get the correct height (likely others)
# 

d1 <- d1.1 %>%  #d1 = d1.1, just arranged differently 
    arrange(desc(season), harvest, wall, type, spp, point) %>% 
    select(filename, harvest, wall, point, type, spp, season, trmt, everything())

d1a <- d1 %>%   #d1a is useful to test calculation from the mostly raw data
    filter(season == "2023") %>% 
    filter(harvest == "6_6") %>% 
    filter(all_sizes > 0) %>% 
    arrange(desc(season), harvest, wall, type, spp, point) 

    arrange(filename, (harvest), desc(season), wall, type, spp) %>% 

```



#define numeric variables
```{r}
#numeric vars to sum by origin
numeric_vars1 <- c("total", "all_sizes", "GTR5ft", "seedA", "supB", "supD", "supP", "expB",
                  "expD", "expP")

#numeric_vars2 = average by origin
numeric_vars2 <- c("firstgrow", "year", "denom", "growingseasons", "height")
```



# count # spp within type present at each point. Summarize tpa by point
```{r}
type_summary <- d1 %>%
#    filter(season == "2023") %>%  # filter for season while checking code, then run for all seasons
    group_by(harvest, season, wall, point, type) %>%
    summarize(
        trmt = first(trmt),
        point_count = n_distinct(point),
        denominator = n(), #denominator = # spp x type
        type_count = n(),
        any_present = sum(all_sizes > 0, na.rm = TRUE),  # Record presence at points by type where "all_sizes" > 0, for stems > 4" tall
        GTR1ft_present = sum(total > 0, na.rm = TRUE),  # Record presence at points by type where total > 0, for stems > 12" tall
        GTR5ft_present = sum(GTR5ft > 0, na.rm = TRUE), # Record presence at points by type where "GTR5" >0
        across(all_of(numeric_vars1), sum, na.rm = TRUE),  # Sum all specified numeric vars1 (logic for sum confirmed)
        across(all_of(numeric_vars2), mean, na.rm = TRUE),  # Mean of all specified numeric vars2

          # Calculate mean for clumps.acre and stems.ramet based on non-zero count
       clumps.acre = if_else(sum(clumps.acre > 0, na.rm = TRUE) > 0,  #counts the number of non-zero values of clump.acre
                                  sum(clumps.acre, na.rm = TRUE) / sum(clumps.acre > 0, na.rm = TRUE), #if there are non-zero values, completes the calc
                                  NA_real_), #calculations for clumps.acre and stems.ramet confirmed
       
       stems.ramet = if_else(sum(stems.ramet > 0, na.rm = TRUE) > 0,
                                  sum(stems.ramet, na.rm = TRUE) / sum(stems.ramet > 0, na.rm = TRUE),
                                  NA_real_),

                .groups = "drop"
        
        #need to average clumps.acre and stems.ramet for the non-zero values within a type
        ) %>%

        mutate(
            across(where(is.numeric), ~ round(., 1)),
            filename = "slash_wall_analyses_2025.Rmd"
        ) %>% 

        select(filename, harvest, season, trmt, wall, point, type,point_count, denominator, type_count, 
           any_present, all_sizes, GTR1ft_present, total, growingseasons, clumps.acre, 
           stems.ramet, everything()) %>% 
        arrange(desc(season), harvest, wall, type, point)

```

# occupancy by wall for type. Summ of stem counts by type
# aggregate into average by wall x type at the harvest scale
```{r}
wall_type_occupancy <- type_summary %>%
  group_by(harvest, season, trmt, wall, type) %>%
  summarize(
    point_count = n_distinct(point),# Total plots in the wall

    plots_with_any = sum(any_present > 0, na.rm = TRUE), # Count of plots where type is represented by a spp
    occup_any_pct = (plots_with_any / point_count) * 100, #by type, the percentage of plots within a wall with spp any size of that type

    plots_with_GTR1ft = sum(total > 0, na.rm = TRUE), # Count of plots where type is represented by a spp >12" tall
    GTR1ft_pct = (plots_with_GTR1ft / point_count) * 100,

    plots_with_GTR5ft = sum(GTR5ft_present > 0, na.rm = TRUE), # Count of plots where type is represented by a spp
    GTR5ft_pct = (plots_with_GTR5ft / point_count) * 100,
    
    clumps.acre = sum(clumps.acre / point_count, na.rm = TRUE), # this will calc clumps.acre based on the number of points per wall. points with zero included
    stems.ramet = if_else(sum(stems.ramet > 0, na.rm = TRUE) > 0,
                                  sum(stems.ramet, na.rm = TRUE) / sum(stems.ramet > 0, na.rm = TRUE),
                                  NA_real_), # this will calc stems.ramet based on the number of points w/i wall that have clumps

    across(all_of(numeric_vars1), ~ sum(.x, na.rm = TRUE) / point_count),  # Normalize by point_count
    across(all_of(numeric_vars2), ~ sum(.x, na.rm = TRUE) / point_count),  # Normalize by point_count
    .groups = "drop"
  )  %>%   
    mutate(
        across(where(is.numeric), ~ round(., 1))
  ) %>% 
  mutate(
        filename = "slash_wall_analyses_2025.Rmd"
  ) %>% 
    select(filename, harvest, season, trmt, wall, growingseasons, type, plots_with_any, point_count, occup_any_pct, plots_with_GTR1ft, 
           GTR1ft_pct, plots_with_GTR5ft, GTR5ft_pct, clumps.acre, stems.ramet, everything())

output_file <- file.path(output_folder, paste0("type_occupancy_x_wall_x_trmt_x_season(4march2025)", ".xlsx"))
write_xlsx(wall_type_occupancy, output_file)


```


# calcuate the values of metrics at wall x spp
```{r}
species_by_wall <- d1 %>%
#  filter(season == "2023") %>%  # filter for season while checking code, then run for all seasons
  group_by(harvest, season, trmt, wall, spp) %>%
   summarize(
       filename = first(filename), 
       type = first(type),
       point_count = n_distinct(point),
            #calculate the numb points 
       numb_pts_any = sum(all_sizes > 0, na.rm = TRUE),  # Count species where all_sizes > 0
       numb_pts_GTR1ft = sum(total > 0, na.rm = TRUE),  # Count species where "total" 12"+ > 0
       numb_pts_GTR5ft = sum(GTR5ft > 0, na.rm = TRUE), # count of species where GTR5>0
       any_size_pct = numb_pts_any / point_count *100,
       GTR1ft_pct = numb_pts_GTR1ft / point_count * 100,
       GTR5ft_pct = numb_pts_GTR5ft / point_count * 100,
       
       across(all_of(numeric_vars1), ~ sum(.x, na.rm = TRUE) / point_count),  # Normalize by point_count
       across(all_of(numeric_vars2), ~ sum(.x, na.rm = TRUE) / point_count),  # Normalize by point_count
       
       x1 = sum(stems.ramet, na.rm = TRUE),
       x2 = sum(stems.ramet > 0, na.rm = TRUE), 
       clumps.acre = sum(clumps.acre),
       stems.ramet = sum(stems.ramet),
       clumps.acre = sum(clumps.acre / point_count, na.rm = TRUE), 
            # this will calc clumps.acre based on the number of points per wall. points with zero included
       stems.ramet = if_else(x2 > 0, x1 / x2, NA_real_),
       # stems.ramet.2 = case_when(
       #      sum(stems.ramet > 0, na.rm = TRUE) > 0 ~ sum(stems.ramet, na.rm = TRUE) / sum(stems.ramet > 0, na.rm = TRUE),
       #      TRUE ~ NA_real_), # calc confirmed, stems.ramet based on # spp that have ramets

           .groups = "drop"
  )  %>%   
  mutate(
        across(where(is.numeric), ~ round(., 1)),
        df = "species_by_wall"
  ) %>%
    filter(all_sizes > 0) %>% 
    arrange(filename, (harvest), desc(season), trmt, type, spp) %>% 
    select(filename, df, harvest, season, trmt, wall, growingseasons, spp, type, point_count, numb_pts_any, any_size_pct, numb_pts_GTR1ft, GTR1ft_pct, 
           numb_pts_GTR5ft, GTR5ft_pct, clumps.acre, stems.ramet,  everything())

output_file <- file.path(output_folder, paste0("spp_x_wall_x_trmt_x_season(5march2025)", ".xlsx"))
write_xlsx(species_by_wall, output_file)


#rerun spp x wall as type x harvest

```

# species averages across walls by harvest
```{r}
point_counts <- species_by_wall %>% 
        group_by(harvest, season, trmt, wall) %>% 
        summarize(
            point_count = first(point_count), # the number of points in a wall
            .groups = "drop"
        ) %>% 
        arrange((harvest), desc(season), trmt, wall) 

    
wall_counts <- point_counts %>%
    group_by(harvest, season, trmt) %>%
    summarize(
        wall_count = n_distinct(wall), # number of walls in a harvest
        points_per_wall = sum(point_count), # could also be called points_per_trmt
        .groups = "drop")

species_by_harvest <- species_by_wall %>% 
#        filter(harvest == "6_6" & season == "2023") %>% 
        left_join(wall_counts, by = c("harvest", "season", "trmt")) %>%  # Join to get the correct wall_count
        group_by(harvest, season, trmt, spp) %>% 
        summarize(
            filename = first(filename),
            points_per_harvest = first(points_per_wall),
            # numb_pts_any = first(numb_pts_any), # I don't think this is needed
            # numb_pts_seedl = first(numb_pts_seedl), # same
            # numb_pts_GTR5ft = first(numb_pts_GTR5ft), # same
            type = first(type),
            freq_by_grouping = n(),
            wall_count = first(wall_count),
#            avg_pts_wall_any = sum(numb_pts_any) / wall_count,
#            avg_pts_wall_seedl = sum(numb_pts_seedl) / wall_count,
#            avg_pts_wall_GTR5ft = sum(numb_pts_GTR5ft) / wall_count,
            pct_points_any = (sum(numb_pts_any)) / points_per_harvest *100,
            pct_points_GTR1ft = (sum(numb_pts_GTR1ft)) / points_per_harvest *100, # for stems > 12" tall
            pct_points_GTR5ft = (sum(numb_pts_GTR5ft)) / points_per_harvest * 100,
            
           across(all_of(numeric_vars1), ~ sum(.x, na.rm = TRUE) / wall_count),  # summ numeric from each wall, divide by number of walls = avg/harvest
           across(all_of(numeric_vars2), mean, na.rm = TRUE),
           #across(all_of(numeric_vars2), ~ sum(.x, na.rm = TRUE) / wall_count)  # Normalize by wall_count
           #
       x1 = sum(stems.ramet, na.rm = TRUE),
       x2 = sum(stems.ramet > 0, na.rm = TRUE), 
       clumps.acre = sum(clumps.acre),
       stems.ramet = sum(stems.ramet),
       clumps.acre = sum(clumps.acre / points_per_wall, na.rm = TRUE), 
            # this will calc clumps.acre based on the number of points per wall. points with zero included
       stems.ramet = if_else(x2 > 0, x1 / x2, NA_real_),

       .groups = "drop"
        
    ) %>% 
    arrange(filename, (harvest), desc(season), trmt, type, spp) %>% 
    mutate(df = "species_by_harvest") %>% 
#    select(-x1, -x2, -supC1, -supC2, -expC1, -expC2, -denom) %>% 
        select(filename, df, harvest, season, trmt,  wall_count, points_per_harvest, growingseasons, spp, type, pct_points_any, pct_points_GTR1ft,
           pct_points_GTR5ft, clumps.acre, stems.ramet, everything()) %>% 
    mutate(
        across(where(is.numeric), ~ round(., 1))
    )

output_file <- file.path(output_folder, paste0("spp_x_harvest_x_trmt_x_season(5march2025)", ".xlsx"))
write_xlsx(species_by_harvest, output_file)


```


# TYPE averages across walls by harvest
```{r}
point_counts <- species_by_wall %>% 
        group_by(harvest, season, trmt, wall) %>% 
        summarize(
            point_count = first(point_count),
            .groups = "drop"
        )
    
wall_counts <- point_counts %>%
    group_by(harvest, season, trmt) %>%
    summarize(
        wall_count = n_distinct(wall),
        points_per_wall = sum(point_count),
        .groups = "drop")

type_by_harvest <- species_by_harvest %>% 
#        filter(harvest == "6_6" & season == "2023") %>% 
        left_join(wall_counts, by = c("harvest", "season", "trmt", "points_per_wall", "wall_count")) %>%  # Join to get the correct wall_count
        group_by(harvest, season, trmt, type) %>% 
        summarize(
            filename = first(filename),
            points_per_wall = first(points_per_wall),
            type = first(type),
            freq_by_grouping = n(),
            wall_count = first(wall_count),
            pts_wall_any = sum(numb_pts_any), #nonsensical.  Go to sum@point, flag a point if any comm, if any interf, etc. then summ
            pts_wall_seedl = sum(numb_pts_seedl),
            pts_wall_GTR5ft = sum(numb_pts_GTR5ft),
            pct_points_any = sum(numb_pts_any) / points_per_wall *100,
            pct_points_seedl = sum(numb_pts_seedl) / points_per_wall *100,
            pct_points_GTR5ft = sum(numb_pts_GTR5ft) / points_per_wall * 100,
           type_diversity = n_distinct(spp), # the number of species per type. NEED to standardize this for area
           across(all_of(numeric_vars1), ~ sum(.x, na.rm = TRUE) / wall_count),  # Normalize by wall_count
           across(all_of(numeric_vars2), mean, na.rm = TRUE),
           #across(all_of(numeric_vars2), ~ sum(.x, na.rm = TRUE) / wall_count)  # Normalize by wall_count
           #
       x1 = sum(stems.ramet, na.rm = TRUE), #the total of the stems per ramet
       x2 = sum(stems.ramet > 0, na.rm = TRUE), # the number of spp with clumps
       clumps.acre = sum(clumps.acre),
       clumps.acre = sum(clumps.acre),
       stems.ramet = if_else(x2 > 0, x1 / x2, NA_real_), #average stems.ramet

       .groups = "drop"
        
    ) %>% 
    arrange(filename, (harvest), desc(season), trmt, type) %>% 
    select(!c(pts_wall_any, pct_points_any,pts_wall_seedl,pct_points_seedl,pts_wall_GTR5ft, pct_points_GTR5ft,
              supC1, supC2, expC1, expC2, denom)) %>% 
        select(filename, harvest, season, growingseasons, trmt,  wall_count, points_per_wall, type, 
           clumps.acre, stems.ramet, type_diversity, everything()) %>% 
    mutate(
        across(where(is.numeric), ~ round(., 1))
    )
output_file <- file.path(output_folder, paste0("type_x_harvest_x_trmt_x_season(20feb2025)", ".xlsx"))
write_xlsx(type_by_harvest, output_file)


```



# graphs of df = type x harvest
# data preparation
```{r}
prep_type <- type_by_harvest %>% 
    group_by(harvest) %>% 
    filter(year == max(year)) %>% # retain the most recent growing season within each harvest
    ungroup() %>% 
    filter(!harvest %in% c("decker", "recknagel_north")) %>%  # remove decker and recknagel_north
    mutate(
        cluster = "recknagel",
        cluster = case_when(
            harvest %in% c("6_6") ~ "6_6",
            harvest %in% c("6_9") ~ "6_9",
            harvest %in% c("gas_line") ~ "gas_line",
            harvest %in% c("red_pine") ~ "red_pine",
            harvest %in% c("sta_rd") ~ "sta_rd",
            TRUE ~ cluster,
    )) %>% 
  mutate(trmt = case_when(
    harvest == "recknagel" & is.na(trmt) ~ "control",
    TRUE ~ trmt  # Retain existing values in 'trmt' for all other cases
  ))


desired_order <- c("6_6", "6_9", "recknagel", "red_pine", "sta_rd", "gas_line") #places cluster elements in desired order, young to old
```


# Diversity, total, figure_D1
```{r}
diversity <- prep_type %>% 
  select(harvest, cluster, trmt, type_diversity, season) %>% 
  group_by(harvest, season, trmt) %>% 
  summarize(
    cluster = first(cluster),                       # Keeps the first occurrence of the cluster
    Number_of_Species = sum(type_diversity, na.rm = TRUE),  # Calculate the mean while ignoring NA values
    .groups = "drop"                                # Avoids grouping in the result
  ) %>% 
  mutate(cluster = factor(cluster, levels = desired_order)) %>%  # Set the desired order for clusters
  arrange(cluster, harvest, trmt)                        # Arrange by cluster, harvest, and treatment


figure_d1 <- ggplot(diversity, aes(x = cluster, y = Number_of_Species, fill = trmt)) +
  geom_bar(stat = "identity", position = "dodge")+ 
  scale_fill_manual(values = c("control" = "blue", "protected" = "green")) +
  scale_pattern_manual(values = c("control" = "stripe", "protected" = "crosshatch")) +  # Define patterns
  labs(
       x = "Cluster",
       y = "Number of Species",
      title = "Number of Species by Cluster and Treatment",
      subtitle = "Growing Season 2023, slash_wall_analyses_2025.Rmd. Figure_D1",
       fill = "trmt") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


  ggsave("figures/spp_diversity_cluster_trmt_Figure_D1.jpg", plot = figure_d1, width = 8, height = 6, dpi = 300)
```

# Stem Density, commercial and diversity. GTR5ft, sapl, commercial only
# Figure_S1, Figure_S2
```{r}
gtr5 <- prep_type %>% 
  select(harvest, cluster, trmt, GTR5ft, type, season) %>% 
  filter(type == "commercial" | type == "diversity") %>% 
      group_by(harvest, season, trmt) %>% 
  mutate(GTR5ft = case_when(
    harvest == "recknagel" & trmt == "control" ~ 10,
    TRUE ~ GTR5ft  # Insert small values (10) to allow graphing visualization. Retain existing values in 'trmt' for all other cases
  )) %>% 
  summarize(
      cluster = first(cluster),
      Comm_DiversityTPA = sum(GTR5ft),
      .groups = "drop"
      ) %>% 
  mutate(cluster = factor(cluster, levels = desired_order)) %>% 
    arrange(cluster, harvest, trmt)

figure_S1 <- ggplot(gtr5, aes(x = cluster, y = Comm_DiversityTPA, fill = trmt)) +
  geom_bar(stat = "identity", position = "dodge")+ 
  scale_fill_manual(values = c("control" = "blue", "protected" = "green")) +
  scale_pattern_manual(values = c("control" = "stripe", "protected" = "crosshatch")) +  # Define patterns
  labs(
       x = "Cluster",
       y = "Commercial and Diversity TPA",
      title = "Commercial and Diversity Species GTR 5 ft (TPA) by Cluster and Treatment",
      subtitle = "Growing Season 2023, slash_wall_analyses_2025.Rmd. Figure_S1",
       fill = "trmt") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


  ggsave("figures/comm_diversity_tpa_cluster_trmt_Figure_S1.jpg", plot = figure_S1, width = 8, height = 6, dpi = 300)


sapl <- prep_type %>% 
  select(harvest, cluster, trmt, expP, type, season) %>% 
  filter(type == "commercial" | type == "diversity") %>% 
  # mutate(GTR5ft = case_when(
  #   harvest == "recknagel" & trmt == "control" ~ 10,
  #   TRUE ~ GTR5ft  # Retain existing values in 'trmt' for all other cases
  # )) %>% 
  mutate(expP = case_when(
    harvest == "recknagel" & trmt == "control" ~ 3,
    TRUE ~ expP  # Recknagel controls had 0 stems, so plug data for graph visualization. Retain existing values in 'expP' for all other cases
  )) %>% 
    group_by(harvest, season, trmt) %>% 
    summarize(
      cluster = first(cluster),
      Exposed_Sapl_CommDiv = sum(expP),
      .groups = "drop"
      ) %>%
    group_by(cluster, trmt) %>% 
        summarize(
      cluster = first(cluster),
      Exposed_Sapl_CommDiv = mean(Exposed_Sapl_CommDiv),
      .groups = "drop"
      ) %>%
  mutate(cluster = factor(cluster, levels = desired_order))


figure_S2 <- ggplot(sapl, aes(x = cluster, y = Exposed_Sapl_CommDiv, fill = trmt)) +
  geom_bar(stat = "identity", position = "dodge")+ 
  scale_fill_manual(values = c("control" = "blue", "protected" = "green")) +
  scale_pattern_manual(values = c("control" = "stripe", "protected" = "crosshatch")) +  # Define patterns
  labs(
       x = "Cluster",
       y = "Exposed Saplings Commercial and Diversity TPA",
      title = "Exposed Saplings (Comm & Div Species) (TPA) by Cluster and Treatment",
      subtitle = "Growing Season 2023, slash_wall_analyses_2025.Rmd. Figure_S2",
       fill = "trmt") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


  ggsave("figures/exposed_sapling_CommDivSpp_tpa_cluster_trmt_Figure_S2.jpg", plot = figure_S2, width = 8, height = 6, dpi = 300)

  interf <- prep_type %>% 
  select(harvest, cluster, trmt, GTR5ft, type, season) %>% 
  filter(type == "interfere") %>% 
      group_by(harvest, season, trmt) %>% 
  # mutate(GTR5ft = case_when(
  #   harvest == "recknagel" & trmt == "control" ~ 10,
  #   TRUE ~ GTR5ft  # Retain existing values in 'trmt' for all other cases
  # )) %>% 
    group_by(harvest, season, trmt) %>% 
    summarize(
      cluster = first(cluster),
      InterGTR5 = sum(GTR5ft),
      .groups = "drop"
      ) %>%
    group_by(cluster, trmt) %>% 
    summarize(
      cluster = first(cluster),
      InterGTR5 = mean(InterGTR5),
      .groups = "drop"
      ) %>%
    mutate(cluster = factor(cluster, levels = desired_order))

figure_F1 <- ggplot(interf, aes(x = cluster, y = InterGTR5, fill = trmt)) +
  geom_bar(stat = "identity", position = "dodge")+ 
  scale_fill_manual(values = c("control" = "blue", "protected" = "green")) +
  scale_pattern_manual(values = c("control" = "stripe", "protected" = "crosshatch")) +  # Define patterns
  labs(
       x = "Cluster",
       y = "Interfering Species GTR 5 ft (TPA)",
      title = "Interfering Species GTR 5 ft (TPA) by Cluster and Treatment",
      subtitle = "Growing Season 2023, slash_wall_analyses_2025.Rmd. Figure_F1",
       fill = "trmt") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


  ggsave("figures/interfering_GTR5ft_tpa_cluster_trmt_Figure_F1.jpg", plot = figure_F1, width = 8, height = 6, dpi = 300)


```





# Stem Density, commercial only GTR5ft, sapl
# Figure_C1, Figure_S2
```{r}
comm01 <- prep_type %>% 
  select(harvest, cluster, trmt, GTR5ft, supB, expB, supD, expD, supP, total, expP, type, season) %>% 
  filter(type == "commercial") %>% 
  mutate(
      seedl_sml = supB + expB,
      seedl_lgr = supD + expD,
      sapl = supP + expP,
      GTR5ft = case_when(
    harvest == "recknagel" & trmt == "control" ~ 10,
    TRUE ~ GTR5ft  # Insert small values (10) to allow graphing visualization. Retain existing values in 'trmt' for all other cases
  )) %>% 
  group_by(cluster, trmt) %>% 
  summarize(
 #     harvest = first(harvest),
      seedl_sml = mean(seedl_sml),
      seedl_lgr = mean(seedl_lgr),
      sapl = mean(sapl),
#      total = mean(total),
      .groups = "drop"
      ) %>% 
  pivot_longer(cols = c(seedl_sml, seedl_lgr, sapl),
               names_to = "height_class",
               values_to = "tpa") %>%
  mutate(tpa = case_when(
    tpa == 0 ~ 100,  # Replace 0 with 25
    TRUE ~ tpa      # Keep all other values unchanged
  )) %>% 
  mutate(
    cluster = factor(cluster, levels = desired_order),
    height_class = factor(height_class, levels = c("seedl_sml", "seedl_lgr", "sapl", "total")) # Adjust order
  ) %>%
    arrange(cluster, trmt, height_class)


comm01 <- comm01 %>%
  mutate(fill_color = case_when(
    trmt == "control" ~ control_colors[height_class],
    trmt == "protected" ~ protected_colors[height_class]
  ))




figure_c1 <- ggplot(comm01, aes(x = cluster, y = tpa, fill = fill_color, pattern_fill = height_class)) +
  geom_bar_pattern(stat = "identity", position = "dodge", 
                   pattern = "stripe", pattern_density = 0.1, pattern_spacing = 0.05) +
  scale_fill_identity() +  # Uses preassigned colors
  scale_pattern_fill_manual(values = c("seedl_sml" = "stripe", "seedl_lgr" = "crosshatch", "sapl" = "dots")) +
  labs(
    x = "Cluster",
    y = "Commercial Trees Per Acre (TPA)",
    title = "Commercial Species 12-60in, 60in to 1in, >1in (TPA) by Cluster and Treatment",
    subtitle = "Growing Season 2023, slash_wall_analyses_2025.Rmd. Figure_C1",
    fill = "Treatment",
    pattern_fill = "Height Class"
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

#
#these patterns didn't quite work
#
# figure_c1 <- ggplot(comm01, aes(x = cluster, y = tpa, fill = trmt, pattern = height_class)) +
#   geom_bar_pattern(stat = "identity", position = "dodge",
#                    aes(pattern_fill = height_class),  # Ensure pattern is mapped to height_class
#                    pattern_density = 0.5, pattern_spacing = 0.02) +  
#   scale_fill_manual(values = c("control" = "blue", "protected" = "green")) +  # Treatment = Color
#   scale_pattern_manual(values = c("seedl_sml" = "stripe", 
#                                   "seedl_lgr" = "crosshatch", 
#                                   "sapl" = "circle")) +   # Ensure sapl has a pattern
# #                                  "total" = "none")) +  
#   labs(
#     x = "Cluster",
#     y = "Commercial Trees Per Acre (TPA)",
#     title = "Commercial Species 12-60in, 60in to 1in, total (TPA) by Cluster and Treatment",
#     subtitle = "Growing Season 2023, slash_wall_analyses_2025.Rmd. Figure_C1",
#     fill = "Treatment",   
#     pattern = "Height Class"  
#   ) +
#   theme_classic() +
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


  ggsave("figures/comm_tpa_height-class_cluster_trmt_Figure_C1.jpg", plot = figure_c1, width = 8, height = 6, dpi = 300)


```


# Merge the TPA data for species x harvest x trmt, assemble into an output file that can be used to make tables.
```{r}

## Step 1: Select top 5 species within each group
## there are no calculations here, just sorting the top five spp for each group_by
top_spp <- species_by_harvest %>%
  group_by(season, harvest, trmt) %>%
  arrange(desc(total), .by_group = TRUE) %>%
  slice_head(n = 5) %>%
  ungroup()

# Step 2: Calculate remainder for all other species
remainder <- species_by_harvest %>%
  group_by(season, harvest, trmt) %>%
  filter(!spp %in% top_spp$spp) %>%  # Exclude top 5 spp
  summarize(
    spp = "Remainder",  # Label for remaining species
    across(where(is.numeric), sum, na.rm = TRUE),  # Sum numeric variables
    #(average?)  stems.ramet = mean(stems.ramet)
    .groups = "drop"
  )

# Step 3: Calculate total (sum of top 5 + remainder)
total_row <- bind_rows(top_spp, remainder) %>%
  group_by(season, harvest, trmt) %>%
  summarize(
    spp = "Total",  # Label for total
    across(where(is.numeric), sum, na.rm = TRUE),
    .groups = "drop"
  )

# Step 4: Combine top 5 species, remainder,  total and add a blank row.
# use a dummy variable to force the sort given there is no existing variable that
# would sort ascending or descending.

# Combine and sort Remainder and TOTAL first
remain_total <- bind_rows(remainder, total_row) %>%
  arrange(season, harvest, trmt, spp) %>%   # Sort by spp to ensure Remainder comes before TOTAL
  mutate(spp = case_when(
        spp %in% c("Total") ~ "TOTAL", # see step 3 above where spp defined as "Total" but could change to "TOTAL"
        TRUE ~ spp,  # Keeps the existing value of "spp" for all other cases
    )) %>% 
  mutate(dummy = case_when(
        spp == "Remainder" ~ 6,
        spp == "TOTAL" ~ 7,
        TRUE ~ NA_real_  # Assign NA for other cases
  ))

# Assign dummy values to top_spp
top_spp <- top_spp %>%
  group_by(season, harvest, trmt) %>%
  mutate(dummy = row_number()) %>%  # Assign 1 to 5
  ungroup()

blank_rows <- top_spp %>%
  group_by(season, harvest, trmt) %>%
  summarize(spp = "", total = NA, dummy = 8, .groups = "drop")

# Now combine with top_spp
final_result <- bind_rows(top_spp, remain_total, blank_rows) %>%
  arrange(harvest, season, trmt, dummy) %>%  # Sort by dummy for correct order
  mutate(
    date = "19 feb 2025",
    across(
      c(wall_count, growingseasons, stems.ramet, firstgrow, year, denom, height, 
        pct_points_any, pct_points_seedl, pct_points_GTR5ft, freq_by_grouping),
      ~ if_else(dummy %in% c(6, 7), NA_real_, .x)  # Correct syntax
    )
  ) %>%  
  select(-x1, -x2, -denom) %>%  
  select(filename, date, harvest, trmt, season, wall_count, growingseasons, spp, everything())





# final_result <- bind_rows(top_spp, remain_total, blank_rows) %>%
#   arrange(harvest, season, trmt, dummy) %>%    # Sort by dummy for correct order
#   mutate(
#     date = "19 feb 2025",
#     across(
#       c(wall_count, growingseasons, 
#         stems.ramet, firstgrow, year, denom, height, pct_points_any, pct_points_seedl,
#            pct_points_GTR5ft, freq_by_grouping),
#       ~ if_else(dummy %in% c(6, 7), NA_real_, .)# this code removes the values for these variables # for the Remainder and Total rows, for which they are nonsensical
#     ) %>% 
#   select(-x1, -x2, -denom)%>% 
#   select(filename, date, harvest, trmt, season, wall_count, growingseasons, spp, everything())
 


output_file <- file.path(output_folder, paste0("top_spp_by_harvest_x_trmt_x_season(19feb2025)", ".xlsx"))
write_xlsx(final_result, output_file)







```





```
