---
title: "seedling height analysis"
author: "Peter Smallidge"
date: "10/16/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Analysis of seedling heights 2018 & 2019


```{r, include=FALSE}

library(tidyverse)
library(skimr)        ## install.packages("skimr")
library(dplyr)
#
# ALT plus hypen = <- (within chunk)
# ctrl plus shift plus m = %>% 
# ctrl + ALT + I = insert chunk
# filter rows, select columns
# 

```


#### code to replace one character value with another character value
```{r}
library(dplyr)
char_vec <- sample(c("a", "b", "c"), 10, replace = TRUE)
print(char_vec)
recode(char_vec, a = "Apple")
char_vec <- recode(char_vec, a = "Apple", b = "Banana") %>% as.data.frame()
print(char_vec)
```



### read data from csv file, check variables and data structure

<!-- note to self: in the "read_csv" code, after the file location the code
for 'col..." etc ensures that the specified column is adjusted
the file "cheat sheet data-import" has details -->


```{r}
input.data = read_csv("~/R/slash-wall-vegetation/height-data-2019.csv",
                      col_types=cols(spp=col_character()))
str(input.data)

ht.data <- input.data %>% 
  rename(ht.2019 = tot.ht)

#ht.data %>% recode(spp, "318" = "SM", "531", "AB", "372" = "SB")

#recode(y , "pear" = "apple", "banana" = "pineapple" )

# in an earlier csv file there was an "S" and all should have been lower case.  The code for 
# mutate that follows did not work. I changed the value in the original data file
#ht.data %>% 
#  mutate_if(spp=recode(spp, '531' = 'AB'))

table(ht.data$spp)

#df %>% mutate(result=recode(result, 'Win'='1', 'Loss'='0'))

# learn code to convert numeric values to character values , 316 = "RM", 318 = "SM", 356 = "SV" 
#  recode('531' = "AB")
glimpse(ht.data)



skim(ht.data)
#str(ht.data)  there are 1158 rows in original data frame
head(ht.data)
#dim(ht.data)

```

### Determine average total height at end of 2019

```{r}
# determine average total height and annual increment by harvest, origin, and spp
# 1. filter to retain priority species
# 2. select numeric variable
# 3. group by harvest, origin, and spp
# 4. summarize
ht.avg <- ht.data %>% 
  filter(spp == "531" |spp == "316" | spp == "318" | spp == "356" | spp == "372" |
          spp == "375" |spp ==  "541" | spp ==  "621" | spp ==  "743" | spp ==  "761" |
            spp == "762" | spp == "833" | spp == "951") %>% 
  select(harvest, origin, spp, ht.2019, inc.2019, inc.2018, basal.dia) %>% 
  group_by(harvest, origin, spp) %>% 
  add_count(spp) %>% 
  summarise(
    ht.2019 = mean(ht.2019),
    inc.2019 = mean(inc.2019),
    n=n()
  )
  
  

(ht.avg)


dim(ht.avg)
  

```

### Calcuate Species Heights for Boot and Wedge Combined VS Control Plots

```{r}
# mean height values for select species in boot or wedge harvests vs. controls of boot/wedge
# 1. import ht.data, all values
# 2. filter beech plus desired species
# 3. filter rows of data from boot, wedge or their control areas
# 4. filter to retain interior or control plots
# 5. select to retain focal variables
# 6. group by location (interior vs. control) and by species
# 7. count and summarize total height (ht.2019)

glimpse(ht.data)
tail(ht.data)
table(ht.data$harvest) # display the values in the variable "harvest"

ht.avg.boot.wedge <- ht.data %>% 
  filter(spp == "531" |spp == "316" | spp == "318" | spp == "356" | spp == "372" |
          spp == "375" |spp ==  "541" | spp ==  "621" | spp ==  "743" | spp ==  "761" |
            spp == "762" | spp == "833" | spp == "951") %>% 
  filter(harvest == "Boot" | harvest == "W" | harvest == "CWB") %>% 
  #(ht.avg.boot.wedge)
  #
  
  filter(location == "interior" | location == "control") %>% 
  #(ht.avg.boot.wedge)
  #

  select(harvest, location, origin, spp, ht.2019, inc.2019, inc.2018, basal.dia) %>% 
  
  group_by( location, spp) %>% 
  add_count(spp) %>% 
  summarise(
    ht.2019 = mean(ht.2019),
    inc.2019 = mean(inc.2019),
    n=n()
  )

(ht.avg.boot.wedge)

write.table(ht.avg.boot.wedge, "boot_wedge_ht.txt", sep="\t")

```



### Calcuate Species Heights for Gas Line Interior VS Control Plots

```{r}

# mean height values for select species in gas line harvests vs. controls of gas line

# 1. import ht.data, all values
# 2. filter beech plus desired species
# 3. filter rows of data from gas line or their control areas
# 4. filter to retain interior or control plots
# 5. select to retain focal variables
# 6. group by location (interior vs. control) and by species
# 7. count and summarize total height (ht.2019)

#glimpse(ht.data)
#tail(ht.data)
table(ht.data$harvest) # display the values in the variable "harvest"

ht.avg.gasline <- ht.data %>% 
  filter(spp == "531" |spp == "316" | spp == "318" | spp == "356" | spp == "372" |
          spp == "375" |spp ==  "541" | spp ==  "621" | spp ==  "743" | spp ==  "761" |
            spp == "701" | spp == "762" | spp == "833" | spp == "951") %>% 
 #glimpse(ht.avg.gasline), spp 701 = Ostrya
  
  
  filter(harvest == "GL" | harvest == "CGL") %>% 
 #glimpse(ht.avg.gasline)
  
  
  filter(location == "interior" | location == "control") %>% 
# glimpse(ht.avg.gasline)
  

  select(harvest, location, origin, spp, ht.2019, inc.2019, inc.2018, basal.dia) %>% 
  
  group_by( location, spp) %>% 
  add_count(spp) %>% 
  summarise(
    ht.2019 = mean(ht.2019),
    inc.2019 = mean(inc.2019),
    n=n()
  )

#glimpse(ht.avg.gasline)

write.table(ht.avg.gasline, "gasline_ht.txt", sep="\t")

```








### Visualize data relationships

```{r}

library(ggplot2)

ht.data.com <- ht.data %>% 
  filter(spp == "531" |spp == "316" | spp == "318"  | spp == "372" |
           spp ==  "541"  |
            spp == "762" | spp == "833" | spp == "951") 

ggplot(data=ht.data.com, mapping = aes(x=basal.dia, y = ht.2019, size = spp)) +
  geom_point(mapping = aes(size = spp), shape = 1) +
  geom_smooth(method="lm", se=F) +
  facet_wrap(~origin, nrow=1) +
  theme_bw()


```

```{r}
ht.data.com %>% 
  filter(spp==531) %>% 
  group_by(origin) %>% 
  ggplot()+
    geom_boxplot(mapping=aes(x=spp, y=ht.2019))
```

