---
title: "Regneration Analysis"
author: "Peter Smallidge"
date: "10/17/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Analysis of regeneation plot data - 2019


```{r, include=FALSE}

library(tidyverse)
library(skimr)        ## install.packages("skimr")
library(dplyr)
#
# ALT plus hypen = <- (within chunk)
# ctrl plus shift plus m = %>% 
# ctrl + ALT + I = insert chunk
# filter rows, select columns
# RMD comment = <!-- comment -->

```



### read data from csv file, check variables and data structure

<!-- note to self: in the "read_csv" code, after the file location the code
for 'col..." etc ensures that the specified column is adjusted to correct data type
the file "cheat sheet data-import" has details -->


```{r}
input.data <- read_csv("~/R/slash-wall-vegetation/regen-data-2019.csv",
                      col_types=cols(spp=col_character(), 
                                     sup.seed03=col_number(),
                                     sup.sapl=col_number(),
                                     point=col_character()))
str(input.data)

plot.data <- read_csv("~/R/slash-wall-vegetation/plot-data-2019.csv",
                              col_types = cols(point=col_character(),
                                              upper_spp=col_character(),
                                               baf=col_number(),
                                              oakhick=col_number(),
                                              beech=col_number(),
                                              maple=col_number(),
                                              birch=col_number(),
                                              hdwd=col_number(),
                                              conifer=col_number(),
                                              ba=col_number()
                                              ))
str(plot.data) # there are 248 plots in 2019 data

 

regen.data <- input.data %>% 
  select(-other) %>%  # removes the column for "other"
  filter(!spp == "ush") %>%   #removes the row where species = "ush"
  replace(is.na(.), 0) %>% 
  arrange(harvest, location, point, spp) %>% 
  print(n=15)

#str(regen.data)
unique(regen.data$point)  # n=242 points with seedlings from 2019 field data

#check to ensure no unexpected values
  table(regen.data$spp)
  table(regen.data$origin) # origin = "u" = spp = 001 = empty seedling plot
  table(regen.data$radius.01)
  table(regen.data$radius.other)

#glimpse(regen.data)

skim(regen.data)
#str(regen.data)  there are xxxx rows in original data frame
head(regen.data)
#dim(regen.data)

```

### Create tibble of the count of points

```{r}

point.count <-  plot.data %>% # create variable for total plots for denominator of average density
  group_by(harvest, location, point) %>% 
  summarize(n =n()) %>% 
  mutate(count = n/n) %>% 
  select(harvest, location, point, count) 

dim(point.count) # n=248 unique points
head(point.count,n=10L) 


```



### Convert plot seedling counts to per acre values by plot (var=point)

```{r}
# convert seedling counts to per acre values, data resolution for species by plot
# 1. convert missing (NA) to 0 else density values inflated (completed in previous chunk)
# 2. multiple *.seed01 x 1000 (radius 3.7 ft). use mutate
# 3. multiple *. x (radius = 6) use mutate
# 4. aggregate into "all suppressed" and "all exposed" and top4 per acre for each point
# 5. subsequently, need to average points for final values per acre

regen.per.acre <- regen.data %>% 
     mutate(
    sup.seed01A = sup.seed01 * (43560 /((3.14156) * (3.7 * 3.7))),
    sup.seed02A = sup.seed02 * (43560 /((3.14156) * (6 * 6))),
    sup.seed03A = sup.seed03 * (43560 /((3.14156) * (6 * 6))),
    sup.saplA = sup.sapl * (43560 /((3.14156) * (6 * 6))),
    exp.seed01A = exp.seed01 * (43560 /((3.14156) * (3.7 * 3.7))),
    exp.seed02A = exp.seed02 * (43560 /((3.14156) * (6 * 6))),
    exp.seed03A = exp.seed03 * (43560 /((3.14156) * (6 * 6))),
    exp.saplA = exp.sapl * (43560 /((3.14156) * (6 * 6)))
  ) %>% 
  mutate(
    suppressed = sup.seed01A + sup.seed02A + sup.seed03A + sup.saplA,
    exposed = exp.seed01A + exp.seed02A + exp.seed03A + exp.saplA,
    top4A = top4 * (43560 /((3.14156) * (6 * 6))) 
  ) %>% 
  select(harvest, location, point, spp, origin, sup.seed01A, sup.seed02A, sup.seed03A, sup.saplA,
         exp.seed01A, exp.seed02A, exp.seed03A, exp.saplA, suppressed, exposed, top4A)

regen.per.acre %>% 
  arrange(harvest, location, point, spp) %>% 
  head(n=15L)


dim(regen.per.acre)

table(regen.per.acre$spp)
#view(regen.per.acre)
#write.table(regen.per.acre, "regen-acre-expanded.txt", sep="\t")

regen.per.acre %>% 
  write_csv(path = "regen-acre-expanded.csv") 
#this path places file into the current directory
#the data include per acre density by species by point by location by harvest

###   calculations were confirmed for boot>interior>#350 and #355


# commercial regeneration species
# 316 318 372 541 621 762  802 832 833 951 
# 
# desirable diversity species (no interfering, but includes Populus, paper birch, not pin cherry)
# 129 261 316 318 356 372 375 500 541 621 743 746 762 802 832 833  90 920 951
```

## DIVERSITY SPECIES
### calculate total seedling density for "diversity" species
#### diversity species include: spr, ewp, hem, rem, sum, serv, swb, pab, hawt, wha, yep, bta, qua, blc, oaks, willow, bassw 

```{r}
# Calculate total seedling desity of desirable non-interfering species
# 1. filter undesired or interfering species to retain species of "diversity value"
# 2. calculate density
diversity.per.acre_point <- regen.per.acre %>% 
  filter(spp %in% c("001", "29", "261", "316", "318", "356", "372", "375", "500", "541", "621", "743",
                    "746", "762", "802", "832", "833", "90", "920", "951")) %>% 
  group_by(harvest, location, point) %>% 
  summarise(
    per.acre_point.supp = sum(suppressed), #need to total all species within a point
    per.acre_point.exp = sum(exposed),
    per.acre_point.top4 = sum(top4A)
  )%>% 
  full_join(point.count, by = c("harvest",  "location",  "point")) %>% 
  replace(is.na(.), 0)

# use "full_join" because plots with no diversity species have been removed from df = diversity.per.acre_point. Replace na=0 ensures accurate calculation of mean stem count per acre.

dim(diversity.per.acre_point)
head(diversity.per.acre_point, n=10L)

diversity.per.acre_point %>% 
write_csv(path = "diversity-acre-point.csv") # n=248 points

###   calculations at point-level were confirmed 10/29/2020 for boot>interior>#350, #355 and #361



diversity_per_acre <- diversity.per.acre_point %>% 
  group_by(harvest, location) %>% 
  summarize(
    div.per.acre.supp = sum(per.acre_point.supp) / sum(count),
    div.per.acre.exp = sum(per.acre_point.exp) / sum(count),
    div.per.acre.top4 = sum(per.acre_point.top4) / sum(count)
  ) 
head(diversity_per_acre, n=15L)

diversity_per_acre %>% 
  write_csv(path = "diversity-per-acre.csv")
#calculations confirmed 10/30/2020 for boot interior and wedge interior


```


## COMMERCIAL SPECIES
### calculate total seedling density for commercial species
#### diversity species include: rem, sum, swb, wha, yep,blc, oaks,bassw 

```{r}
# Calculate total seedling desity of desirable non-interfering species
# 1. filter undesired or interfering species
# 2. 

commercial.per.acre_point <- regen.per.acre %>% 
  filter(spp %in% c("316", "318",  "372", "541", "621",
                     "762", "802", "832", "833", "951")) %>% 
  group_by(harvest, location, point) %>% 
  summarise(
    comm.per.acre_point.supp = sum(suppressed), #total all commercial species within a point
    comm.per.acre_point.exp01 = sum(exp.seed01A),
    comm.per.acre_point.exp02 = sum(exp.seed02A),
    comm.per.acre_point.exp03 = sum(exp.seed03A),
    comm.per.acre_point.expsapl = sum(exp.saplA),
    comm.per.acre_point.exp = sum(exposed),
    comm.per.acre_point.top4 = sum(top4A)
  )%>% 
  full_join(point.count, by = c("harvest",  "location",  "point")) %>% 
  replace(is.na(.), 0)

commercial.per.acre_point %>% 
  arrange(harvest, location, point) %>% 
  head(n=10L)
# points without commercial species have value = "0"
# 

commercial.per.acre_point %>% 
  write_csv(path = "commercial-per-acre-point.csv")  # allows for confirmation of calculations

# OLD format ...write.table(commercial.per.acre_point, "commercial-acre-point.txt", sep="\t")


commercial.per.acre <- commercial.per.acre_point %>% 
  group_by(harvest, location) %>% 
  summarize(
    comm.per.acre.supp = mean(comm.per.acre_point.supp),
    comm.per.acre.exp01 = mean(comm.per.acre_point.exp01),
    comm.per.acre.exp02 = mean(comm.per.acre_point.exp02),
    comm.per.acre.exp03 = mean(comm.per.acre_point.exp03),
    comm.per.acre.expsapl = mean(comm.per.acre_point.expsapl),
    comm.per.acre.exp = mean(comm.per.acre_point.exp),
    comm.per.acre.top4 = mean(comm.per.acre_point.top4)
  )
head(commercial.per.acre, n=15L)

commercial.per.acre %>% 
  write_csv(path = "commercial-per-acre.csv") # data desired for description of treatments

# OLD format...write.table(commercial.per.acre, "commercial-acre.txt", sep="\t")


```



