---
title: "Regneration Analysis"
author: "Peter Smallidge"
date: "10/17/2020"
output: github_document
---


```{r}
# code from one of Maria's "notes" pages
# TOC may only work with html output

# title: "Lesson 13: Factors"
# output: 
#  html_document:
#    keep_md: yes 
#    toc: true
#    toc_depth: 2 
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Analysis of slash wall regeneation plot data - 2019


```{r, include=FALSE}

library(tidyverse)
library(skimr)        ## install.packages("skimr")
library(dplyr)
#
# ALT plus hypen = <- (within chunk)
# ctrl plus shift plus m = %>% 
# ctrl + ALT + I = insert chunk
# filter rows, select columns
# RMD comment = <!-- comment -->

```



### read data from csv file, check variables and data structure

<!-- note to self: in the "read_csv" code, after the file location the code
for 'col..." etc ensures that the specified column is adjusted to correct data type
the file "cheat sheet data-import" has details -->


```{r}
# Seedling data

input.data <- read_csv("~/R/slash-wall-vegetation/regen-data-2019.csv",
                      col_types=cols(spp=col_character(), 
                                     sup.seed03=col_number(),
                                     sup.sapl=col_number(),
                                     point=col_character()))
str(input.data)

# Plot data
# 
plot.data <- read_csv("~/R/slash-wall-vegetation/plot-data-2019.csv",
                              col_types = cols(point=col_character(),
                                              upper_spp=col_character(),
                                               baf=col_number(),
                                              oakhick=col_number(),
                                              beech=col_number(),
                                              maple=col_number(),
                                              birch=col_number(),
                                              hdwd=col_number(),
                                              conifer=col_number(),
                                              ba=col_number()
                                              ))
str(plot.data) # there are 248 plots in 2019 data

 

regen.data <- input.data %>% 
  select(-other) %>%  # removes the column for "other"
  filter(!spp == "ush") %>%   #removes the row where species = "ush"
  replace(is.na(.), 0) %>% 
  arrange(harvest, location, point, spp) %>% 
  print(n=15)

#str(regen.data)
unique(regen.data$point)  # n=242 points with seedlings from 2019 field data

#check to ensure no unexpected values
  table(regen.data$spp)
  table(regen.data$origin) # origin = "u" = spp = 001 = empty seedling plot
  table(regen.data$radius.01)
  table(regen.data$radius.other)

#glimpse(regen.data)

skim(regen.data)
#str(regen.data)  there are xxxx rows in original data frame
head(regen.data)
#dim(regen.data)

```

### Create tibble of the count of points

```{r}

point.count <-  plot.data %>% # create variable for total plots for denominator of average density
  group_by(harvest, location, point) %>% 
  summarize(n =n()) %>% 
  mutate(count = n/n) %>% 
  select(harvest, location, point, count) 

dim(point.count) # n=248 unique points
head(point.count,n=10L) 


```



### Convert plot seedling counts to per acre values by plot (var=point)

```{r}
# convert seedling counts to per acre values, data resolution = species x plot
# 1. convert missing (NA) to 0 else density values inflated (completed in previous chunk)
# 2. multiple *.seed01 x 1000 (radius 3.7 ft). use mutate
# 3. multiple *. x (radius = 6) use mutate
# 4. aggregate into "all suppressed" and "all exposed" and top4 per acre for each point
# 5. subsequently, need to average points for final values per acre

regen.per.acre <- regen.data %>% 
     mutate(
    sup.seed01A = sup.seed01 * (43560 /((3.14156) * (3.7 * 3.7))),
    sup.seed02A = sup.seed02 * (43560 /((3.14156) * (6 * 6))),
    sup.seed03A = sup.seed03 * (43560 /((3.14156) * (6 * 6))),
    sup.saplA = sup.sapl * (43560 /((3.14156) * (6 * 6))),
    exp.seed01A = exp.seed01 * (43560 /((3.14156) * (3.7 * 3.7))),
    exp.seed02A = exp.seed02 * (43560 /((3.14156) * (6 * 6))),
    exp.seed03A = exp.seed03 * (43560 /((3.14156) * (6 * 6))),
    exp.saplA = exp.sapl * (43560 /((3.14156) * (6 * 6)))
  ) %>% 
  mutate(
    suppressed = sup.seed01A + sup.seed02A + sup.seed03A,
    exposed = exp.seed01A + exp.seed02A + exp.seed03A,
    total = sup.seed01A + sup.seed02A + sup.seed03A + 
      exp.seed01A + exp.seed02A + exp.seed03A,
    top4A = top4 * (43560 /((3.14156) * (6 * 6))) 
  ) %>% 
  select(harvest, location, point, spp, origin, sup.seed01A, sup.seed02A, sup.seed03A, sup.saplA,
         exp.seed01A, exp.seed02A, exp.seed03A, exp.saplA, suppressed, exposed, total, top4A)

regen.per.acre %>% 
  arrange(harvest, location, point, spp) %>% 
  head(n=15L)


dim(regen.per.acre)

table(regen.per.acre$spp)
#view(regen.per.acre)
#write.table(regen.per.acre, "regen-acre-expanded.txt", sep="\t")

regen.per.acre %>% 
  write_csv(path = "regen-acre-expanded.csv") 
#this path places file into the current directory
#the data include per acre density by species by point by location by harvest

###   calculations were confirmed for boot>interior>#350 and #355


# commercial regeneration species
# 316 318 372 541 621 762  802 832 833 951 
# 
# desirable diversity species (no interfering, but includes Populus, paper birch, not pin cherry)
# 129 261 316 318 356 372 375 500 541 621 743 746 762 802 832 833  90 920 951
```

## DIVERSITY SPECIES
### calculate total seedling density for "diversity" species
#### diversity species include: spr, ewp, hem, rem, sum, serv, swb, pab, hawt, wha, yep, bta, qua, blc, oaks, willow, bassw 

```{r}

# Calculate total seedling density of desirable non-interfering species
# 1. filter undesired or interfering species to retain species of "diversity value"
# 2. calculate density

diversity.per.acre_point <- regen.per.acre %>% 
  filter(spp %in% c("001", "29", "261", "316", "318", "356", "372", "375", "500", "541", "621", "743",
                    "746", "762", "802", "832", "833", "90", "920", "951")) %>% 

    group_by(harvest, location, point) %>% 
  summarise(
    
    per.acre_point.sup.seed01A = sum(sup.seed01A), #need to total all species within a point
    per.acre_point.sup.seed02A = sum(sup.seed02A), 
    per.acre_point.sup.seed03A = sum(sup.seed03A), 
    per.acre_point.sup.saplA = sum(sup.saplA), 
    
    per.acre_point.exp.seed01A = sum(exp.seed01A),
    per.acre_point.exp.seed02A = sum(exp.seed02A),
    per.acre_point.exp.seed03A = sum(exp.seed03A),
    per.acre_point.exp.saplA = sum(exp.saplA),
    
    per.acre_point.supp = sum(suppressed), 
    per.acre_point.exp = sum(exposed),
    per.acre_point.total = sum(total),
    per.acre_point.top4 = sum(top4A)
  )%>% 
  full_join(point.count, by = c("harvest",  "location",  "point")) %>% 
  replace(is.na(.), 0)


# use "full_join" because plots with no diversity species have been removed from 
# dataframe = diversity.per.acre_point. 
# Replace na=0 ensures accurate calculation of mean stem count per acre.

dim(diversity.per.acre_point)

# Print "divesity.per.acre_point"
head(diversity.per.acre_point, n=10L)

diversity.per.acre_point %>% 
write_csv(path = "diversity-acre-point.csv") # n=248 points




###   calculations at point-level were confirmed 10/29/2020 for boot>interior>#350, #355 and #361



diversity_per_acre <- diversity.per.acre_point %>% 
  group_by(harvest, location) %>% 
  summarize(
    div.per.acre.sup01 = mean(per.acre_point.sup.seed01A),
    div.per.acre.sup02 = mean(per.acre_point.sup.seed02A),
    div.per.acre.sup03 = mean(per.acre_point.sup.seed03A),
    
    div.per.acre.exp01 = mean(per.acre_point.exp.seed01A),
    div.per.acre.exp02 = mean(per.acre_point.exp.seed02A),
    div.per.acre.exp03 = mean(per.acre_point.exp.seed03A),
    
    div.per.acre.supsapl = mean(per.acre_point.sup.saplA),
    div.per.acre.expsapl = mean(per.acre_point.exp.saplA),

    #different way to calculate the mean
    div.per.acre.supp = sum(per.acre_point.supp) / sum(count),
    div.per.acre.exp = sum(per.acre_point.exp) / sum(count),
    div.per.acre.total = sum(per.acre_point.total) / sum(count),
    div.per.acre.top4 = sum(per.acre_point.top4) / sum(count)
  ) 

# Print "diversity_per_acre"
head(diversity_per_acre, n=15L)

diversity_per_acre %>% 
  write_csv(path = "diversity-per-acre.csv")
#calculations confirmed 10/30/2020 for boot interior and wedge interior


```

#### calculate stocking of diversity species

```{r}
# post harvest stocking thresholds inside fences
# 
# set a "stocking" variable for each height class = 0
# increment through data and if stocking is > threshold then change stocking = 1
# calculate percentage of "stocked plots"  as sum(stocking.x)/sum(count)*100
# 
# for stems > 5 ft requires 540/acre
# 
# stocking thresholds from SILVAH
# my filename is "2013-silvah-quick-reference-guide-allegheny-hardwoods"
# for "other desirable" seedling need 15 per 6 ft radius = 15 * 385 = 5777 for deer impact = 1
# for saplings need 2 per plot = 770 per acre

div_stocking_per_acre <- diversity.per.acre_point %>% 
  mutate(stk_supp = 0,
         stk_exp01 = 0,
         stk_exp02 = 0,
         stk_exp03 = 0,
         stk_exp.01.02 = 0,
         stk_exp = 0)

head(div_stocking_per_acre)
    
 # if(div_stocking_per_acre$per.acre_point.sup.seed01A > 550) {  
#    mutate(stk_supp = 1)
 # }



```



## COMMERCIAL SPECIES
### calculate total seedling density for commercial species
#### diversity species include: rem, sum, swb, wha, yep,blc, oaks,bassw 

```{r}
# Calculate total seedling desity of desirable non-interfering species
# 1. filter undesired or interfering species
# 2. 

commercial.per.acre_point <- regen.per.acre %>% 
  filter(spp %in% c("316", "318",  "372", "541", "621",
                     "762", "802", "832", "833", "951")) %>% 
  group_by(harvest, location, point) %>% 
  summarise(
    comm.per.acre_point.sup01 = sum(sup.seed01A), 
    comm.per.acre_point.sup02 = sum(sup.seed02A), 
    comm.per.acre_point.sup03 = sum(sup.seed03A), 
    comm.per.acre_point.supsapl = sum(sup.saplA),
    comm.per.acre_point.supp = sum(suppressed), #total all commercial species within a point
    comm.per.acre_point.exp01 = sum(exp.seed01A),
    comm.per.acre_point.exp02 = sum(exp.seed02A),
    comm.per.acre_point.exp03 = sum(exp.seed03A),
    comm.per.acre_point.expsapl = sum(exp.saplA),
    comm.per.acre_point.exp = sum(exposed),
    comm.per.acre_point.total = sum(total), # total includes 01. 02. and 03. not sapl
    comm.per.acre_point.top4 = sum(top4A)
  )%>% 
  full_join(point.count, by = c("harvest",  "location",  "point")) %>% 
  replace(is.na(.), 0)




# Sort and Print "commercial.per.acre_point"
commercial.per.acre_point %>% 
  arrange(harvest, location, point) %>% 
  head(n=10L)
# points without commercial species have value = "0"
# 

commercial.per.acre_point %>% 
  write_csv(path = "commercial-per-acre-point.csv")  # allows for confirmation of calculations

# OLD "write" format ...write.table(commercial.per.acre_point, "commercial-acre-point.txt", sep="\t")


commercial.per.acre <- commercial.per.acre_point %>% 
  group_by(harvest, location) %>% 
  summarize(
    comm.per.acre.sup01 = mean(comm.per.acre_point.sup01),
    comm.per.acre.sup02 = mean(comm.per.acre_point.sup02),
    comm.per.acre.sup03 = mean(comm.per.acre_point.sup03),
    comm.per.acre.supp = mean(comm.per.acre_point.supp),
    comm.per.acre.exp01 = mean(comm.per.acre_point.exp01),
    comm.per.acre.exp02 = mean(comm.per.acre_point.exp02),
    comm.per.acre.exp03 = mean(comm.per.acre_point.exp03),
    comm.per.acre.expsapl = mean(comm.per.acre_point.expsapl),
    comm.per.acre.exp = mean(comm.per.acre_point.exp),
    comm.per.acre.total = mean(comm.per.acre_point.total),
    comm.per.acre.top4 = mean(comm.per.acre_point.top4)
  )
head(commercial.per.acre, n=15L)

commercial.per.acre %>% 
  write_csv(path = "commercial-per-acre.csv") # data desired for description of treatments

# OLD format...write.table(commercial.per.acre, "commercial-acre.txt", sep="\t")


```



