---
title: "slash wall analyses 2025"
author: "Peter Smalldige"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# activate library functions
```{r message=FALSE}

library(ggplot2)
library(ggpattern)
#library(ggpubr)
library(expss)
library(readr)
library(readxl)
library(writexl)
library(skimr)
library(tidyverse)  # includes: tidyr, dplyr, haven, readr, readxl, forcats, stringer, ggplot2. See all with tidyverse_packages()
library(knitr)
library(janitor)
#install.packages("openxlsx")
library(openxlsx)
# install.packages("shinyaddins")
# library %>% (shinyaddins)

```


##
##
##  NOTE - unbrushed plot data 



```{r eval=FALSE}

# the purpose of this code is to take the tabulated and "complete" data from 2019 through 2024 and produce output tables and graphs
# that evaluate the effectiveness of slash walls.
# 
# The source file is from slash_veg_complete.Rmd df = "yay2_harvest" which includes data from the 2019 to 2024 growing seasons.
# This data is located in R/slash-wall-vegetation/summary output. These data are the stem counts/acre of hardwood regeneration (some conifers).
# Additional data are within R/slash-wall-vegetation/summary out/site_data_2019_2024(22july2025)(readme) which include % cover of several attributes
# and basal area / acre of the residual overstory. BA was not measured in every year.
# 
# The input data are complete at the season x harvest level
```



### code 01
### Create file path input and the output folder (code 01) 

file paths, input and output
```{r}
#complete data
file_path_veg_complete <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\veg_data_2019to2024_complete(12sept2025)(readme).xlsx"


file_path_site <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\site_data_2019_2024(22july2025)(readme).xlsx"


file_path_points <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\point_labels_all(readme).xlsx"

output_folder <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\"
```

#read veg data
```{r}
veg_data <- read_excel(file_path_veg_complete, sheet = "Sheet1", skip = 0) %>% 
#search for "stand =" and "date =" and change based on filters and today's date
     filter(harvest %in% c("gas_line", "red_pine", "sta_rd", "recknagel", "campridge", "circle", "decker",
                           "danby_block_2", "goundry_a17", "fairfield_a5")) %>% 
     filter(!(harvest == "decker" & trmt == "protected")) %>% 
#    filter(harvest %in% c("6_6")) %>%   
     filter(growingseasons > 2)
 
# gas_line,goundry_a17,danby_block_2    
# use this table command to refine the above filters     
table(veg_data$harvest, veg_data$growingseasons)
table(veg_data$harvest, veg_data$trmt)
table(veg_data$harvest, veg_data$brush)


table(veg_data$harvest, veg_data$season, veg_data$trmt)
table(veg_data$harvest, veg_data$forest_type, veg_data$trmt)
```


#calc denominators for averaging
```{r}
forest_type_df <- veg_data %>% # no longer "forest type" but leave this nomenclature to simplify subsequent code
    mutate(point_id = paste0(harvest, "_", point)) %>%  #danby and goundry are both point #1-14 and 1-20
    group_by(forest_type, growingseasons, trmt, brush ) %>% #sequence doesn't matter
    summarise(
      n_forest_type = n_distinct(point_id), 
      .groups = "drop"
    )

 trmt_df <- veg_data %>% 
     mutate(point_id = paste0(harvest, "_", point)) %>%  #danby and goundry are both point #1-14 and 1-20
     group_by( growingseasons, trmt ) %>% #sequence doesn't matter
     summarise(
       n_trmt_df = n_distinct(point_id), 
       .groups = "drop"
     )
```






#read site data
```{r}
site_data = read_excel(file_path_site, sheet = "Sheet1", skip = 0)

```


#don't run this use 
# read denominators, but they mostly aren't used
```{r eval=FALSE}

point_labels <- read_excel(file_path_points, sheet = "all_sites", skip = 0)%>% 
    mutate(
        cluster = "drop"
    )   %>%  
    mutate(cluster = case_when(
        harvest %in% c("gas_line", "red_pine", "sta_rd", "fairfield_a5", "danby_block_2", "goundry_a17", "fairfield_a6") ~ "early",
        harvest %in% c("campridge", "recknagel", "circle") ~ "second",
        harvest %in% c("6_6", "6_9") ~ "cafri",
    TRUE ~ cluster
))


point_denom <- point_labels %>%
  count(harvest, wall, height, location, brush, trmt) %>%
  filter(n > 1) %>% 
    rename(
        point_count = n
    )


denom_stand_trmt <- point_labels %>% 
    group_by(forest_type, trmt, brush,harvest) %>% 
    count(forest_type, harvest, trmt, brush) %>% 
    filter(n>1) %>% 
    rename(
        plot_count = n
    )

cluster_point_count <- point_labels %>%  #calculation verified 16sept2025
    group_by(cluster, trmt, brush) %>% 
    count(cluster, trmt, brush) %>% 
        rename(
        point_count = n
    )



output_file <- file.path(output_folder, paste0("sample_size_table(31july2025)", ".xlsx"))
write_xlsx(denom_stand_trmt, output_file)


```


#Notes
# DF = FT_calc1
# Input raw data - data are "complete" at the level of the wall
# Calcuations to summarize by forest type and season
```{r}
FT_calc1 <- veg_data %>%  # df = FT = forest type calculations
    left_join(forest_type_df, by = c("forest_type", "growingseasons", "trmt", "brush")) %>%  
#    left_join(trmt_df, by = c("growingseasons", "trmt")) %>% #left_join at FT_calc4
    mutate(
        filename = "slash_wall_analyses_2025(by stand)"
    ) %>% 

group_by(forest_type, growingseasons, brush, trmt, spp, point) %>%  #summary by species and point to remove origin
    summarize(
        filename = first(filename),
        type = first(type),
        firstgrow = first(firstgrow),
        year = first(year),
#        n_trmt_df = first(n_trmt_df),
        n_forest_type = first(n_forest_type),
        clumps_per_acre = sum(clumps_per_acre),  # Sum all specified numeric clumps.acre and stems.ramet calc confirmed)
        stems_per_ramet = sum(stems_per_ramet),
        seed_4_12 = sum(seedA + sup_sd4_12 + exp_sd4_12),
        sup_sd12_5 = sum(sup_sd12_5),    #suppressed seedling 12" to 5 ft
        sup_sd5_1 = sum(sup_sd5_1),   # suppressed seedling 5 ft to 1" dbh
        sup_sp_1_5 = sum(sup_sp_1_5),  # suppressed sapling 1" to 5"
        exp_sd12_5 = sum(exp_sd12_5), # exposed seedling 12" to 5 ft
        exp_sd5_1 = sum(exp_sd5_1),
        exp_sp_1_5 = sum(exp_sp_1_5),   #expP = sum(expP),
        upperexposed = sum(exp_sd5_1+ exp_sp_1_5),
        stemcount = sum(seed_4_12+ sup_sd12_5+ sup_sd5_1+ sup_sp_1_5+ exp_sd12_5+ exp_sd5_1+ exp_sp_1_5), # the total number of stems
        # across(all_of(numeric_vars1), sum, na.rm = TRUE),  
        # across(all_of(numeric_vars2), mean, na.rm = TRUE),  # average of all specified numeric vars2
        .groups = "drop"
    ) %>% 
    mutate(
        GTR5ft = exp_sd5_1 + exp_sp_1_5 + sup_sd5_1 + sup_sp_1_5,
        GTR1ft = sup_sd12_5 + sup_sd5_1 + sup_sp_1_5 + exp_sd12_5 + exp_sd5_1 + exp_sp_1_5, 
        all_sizes = seed_4_12 + GTR1ft, 
        df = "FT_calc1"
    ) %>% 
      mutate(
    brush = case_when(
      brush == "b" ~ "brushed",
      brush == "u" ~ "unbrushed",
      TRUE ~ brush  # Keep any other values unchanged
    )
  )

# calc1_df <- FT_calc1 %>% # no longer "forest type" but leave this nomenclature to simplify subsequent code
#     group_by(growingseasons, trmt) %>% #sequence doesn't matter (forest_type, brush)
#     summarise(
#       n_forest_type = n_distinct(point), 
#       .groups = "drop"
#     )
# 
# rm(calc1_df)

table(veg_data$harvest,veg_data$trmt, veg_data$growingseasons)

```



# test of richness in fairfield_a6
```{r}
old <- veg_data %>% 
    filter(harvest == "fairfield_a6")%>%
#    summarise(distinct_species = n_distinct(spp))
    distinct(spp) %>%
    pull(spp)

print(old)
```


#checking data structure
```{r}
FT_calc2_check <- FT_calc1 %>% 
  group_by(forest_type, growingseasons, brush, trmt, spp) %>%
  summarise(
    unique_n_forest_type = n_distinct(n_forest_type),
    min_n = min(n_forest_type),
    max_n = max(n_forest_type),
    .groups = "drop"
  ) %>%
  filter(unique_n_forest_type > 1)

FT_calc2_check2 <- FT_calc1 %>% 
  group_by(forest_type, growingseasons, brush, trmt, spp) %>%
  summarise(
    count_rows = n(),
    unique_n_forest_type = n_distinct(n_forest_type),
    n_forest_values = paste(unique(n_forest_type), collapse = ", "),
    .groups = "drop"
  ) %>%
  filter(unique_n_forest_type > 1) %>%
  head(10)

group_check <- FT_calc1 %>% 
  group_by(forest_type, growingseasons, brush, trmt, spp) %>%
  summarise(rows_per_group = n(), .groups = "drop") %>%
  arrange(desc(rows_per_group))

print(group_check)

FT_calc1a <- FT_calc1 %>% 
  group_by(forest_type, growingseasons, brush, trmt, spp) %>%
  summarise(
    unique_filename = n_distinct(filename),
    unique_type = n_distinct(type),
    unique_n_forest_type = n_distinct(n_forest_type),
    .groups = "drop"
  ) %>%
  filter(unique_filename > 1 | unique_type > 1 | unique_n_forest_type > 1)

test_summary <- FT_calc1 %>% 
  group_by(forest_type, growingseasons, brush, trmt, spp) %>%  
  summarize(
    count_rows = n(),
    total_clumps = sum(clumps_per_acre),
    .groups = "drop"
  )

nrow(test_summary)

temp_check <- FT_calc1 %>%
  select(forest_type, growingseasons, brush, trmt, spp, n_forest_type) %>%
  group_by(forest_type, growingseasons, brush, trmt, spp) %>%
  summarise(
    n_forest_values = paste(sort(unique(n_forest_type)), collapse = ", "),
    count_unique = n_distinct(n_forest_type),
    .groups = "drop"
  )

print(temp_check)  # Look at results
rm(temp_check)     # Remove it when done

# Test without any division by n_forest_type to isolate the issue
test_no_division <- FT_calc1 %>% 
  group_by(forest_type, growingseasons, brush, trmt, spp) %>%  
  summarize(
    filename = first(filename),
    type = first(type),
    n_forest_type = first(n_forest_type),
    clumps_per_acre = sum(clumps_per_acre),  # No division
    stems_per_ramet = sum(stems_per_ramet),  # No division
    .groups = "drop"
  ) %>% 
    mutate(
        clumps_per_acre = clumps_per_acre / n_forest_type,
        stems_per_ramet = stems_per_ramet / n_forest_type
    # ... do division after summarizing
  )

test_groups <- FT_calc1 %>% 
  group_by(forest_type, growingseasons, brush, trmt, spp) %>%
  summarise(
    n_points = n_distinct(point),
    n_rows = n(),
    .groups = "drop"
  )

cat("Number of unique combinations:", nrow(test_groups), "\n")
cat("Points per combination (first few):", head(test_groups$n_points), "\n")

FT_calc2_test <- FT_calc1 %>% 
  group_by(forest_type, growingseasons, brush, trmt, spp) %>%  
  summarize(
    n_forest_type = first(n_forest_type),
    clumps_per_acre = sum(clumps_per_acre) / first(n_forest_type),
    .groups = "drop"
  )

nrow(test_no_division)  # Should be ~900
rm(test_groups)
rm(test_no_division)
rm(FT_calc1a)
rm(FT_calc2_check)
rm(FT_calc2_check2)
rm(FT_calc2_test)
rm(test_summary)
rm(group_check)
```






# group by forest type and spp
#FT_calc2 dataframe
```{r}
FT_calc2 <- FT_calc1 %>% 

group_by(forest_type, growingseasons, brush, trmt, spp) %>%  #summary by species by forest type and treatment labels
    summarize(
        filename = first(filename),
        type = first(type),
        n_forest_type = first(n_forest_type),
#        growingseasons = first(growingseasons),
        clumps_per_acre = sum(clumps_per_acre) / first(n_forest_type),  # Sum all specified numeric clumps.acre and stems.ramet calc confirmed)
        stems_per_ramet = sum(stems_per_ramet) / first(n_forest_type),
        seed_4_12 = sum(seed_4_12) / first(n_forest_type),
        sup_sd12_5 = sum(sup_sd12_5) / first(n_forest_type),    #suppressed seedling 12" to 5 ft
        sup_sd5_1 = sum(sup_sd5_1) / first(n_forest_type),   # suppressed seedling 5 ft to 1" dbh
        sup_sp_1_5 = sum(sup_sp_1_5) / first(n_forest_type),  # suppressed sapling 1" to 5"
        exp_sd12_5 = sum(exp_sd12_5) / first(n_forest_type), # exposed seedling 12" to 5 ft
        exp_sd5_1 = sum(exp_sd5_1) / first(n_forest_type),
        exp_sp_1_5 = sum(exp_sp_1_5) / first(n_forest_type),   #expP = sum(expP),
        upperexposed = sum(upperexposed) / first(n_forest_type),
        stemcount = sum(stemcount) / first(n_forest_type), # the total number of stems
        GTR5ft = sum(coalesce(GTR5ft, 0), na.rm = TRUE) / first(n_forest_type),
        GTR1ft = sum(GTR1ft) / first(n_forest_type),
        all_sizes = sum(all_sizes)/ first(n_forest_type),
        .groups = "drop"
    ) %>% 
    mutate(
        df = "FT_calc2",
        stand = "all_not_CAFRI"
    ) %>% 
    filter(growingseasons > 2) %>%  # retain the oldest growing season
    filter(all_sizes >0) %>% 
    arrange(
        forest_type, growingseasons, brush, trmt, type,spp) %>% 
    select(
        filename, df, stand, forest_type,n_forest_type,  growingseasons, brush, trmt, spp, type,  upperexposed, GTR5ft, GTR1ft, all_sizes,
        clumps_per_acre, seed_4_12, sup_sd12_5, sup_sd5_1, sup_sp_1_5, exp_sd12_5, exp_sd5_1, exp_sp_1_5, stems_per_ramet
    ) %>% 
    mutate_if(is.numeric, round, 0)
   
table(FT_calc2$growingseasons) # result shows the growing season and the #obs

 

```



#test calc of FT_cal2
# test calc validated 16oct2025
```{r}
FT_calc2_test <- FT_calc1 %>% 
    filter(growingseasons == "3" & spp %in%c("bc", "rm") & trmt == "control") %>% 
    select(forest_type,n_forest_type,  growingseasons, brush, trmt, spp, type,   GTR5ft, GTR1ft, all_sizes) %>%
    group_by(forest_type, growingseasons, brush, trmt, spp) %>%  #summary by species by forest type and treatment labels
    summarize(
        type = first(type),
        n_forest_type = first(n_forest_type),
        GTR5ft = sum(coalesce(GTR5ft, 0), na.rm = TRUE) / first(n_forest_type),
        GTR1ft = sum(GTR1ft) / first(n_forest_type),
        all_sizes = sum(all_sizes)/ first(n_forest_type),
        .groups = "drop"
    ) %>% 

    filter(growingseasons > 2) %>%  # retain the oldest growing season
    filter(all_sizes >0) %>% 
    arrange(
        forest_type, growingseasons, brush, trmt, type,spp) %>% 
    select(
        forest_type,n_forest_type,  growingseasons, brush, trmt, spp, type,   GTR5ft, GTR1ft, all_sizes
    ) %>% 
    mutate_if(is.numeric, round, 0)

```



#FT_calc3
#summary of stem density by forest_type, growingseason, trmt and type
```{r}
FT_calc3 <- FT_calc1 %>% 

group_by(forest_type, growingseasons, brush, trmt, type) %>%  #summary by type (not spp) by forest type and treatment labels
    summarize(
        filename = first(filename),
        n_forest_type = first(n_forest_type),
#        growingseasons = first(growingseasons),
        clumps_per_acre = sum(clumps_per_acre) / first(n_forest_type),  # Sum all specified numeric clumps.acre and stems.ramet calc confirmed)
        stems_per_ramet = sum(stems_per_ramet) / first(n_forest_type),
        seed_4_12 = sum(seed_4_12) / first(n_forest_type),
        sup_sd12_5 = sum(sup_sd12_5) / first(n_forest_type),    #suppressed seedling 12" to 5 ft
        sup_sd5_1 = sum(sup_sd5_1) / first(n_forest_type),   # suppressed seedling 5 ft to 1" dbh
        sup_sp_1_5 = sum(sup_sp_1_5) / first(n_forest_type),  # suppressed sapling 1" to 5"
        exp_sd12_5 = sum(exp_sd12_5) / first(n_forest_type), # exposed seedling 12" to 5 ft
        exp_sd5_1 = sum(exp_sd5_1) / first(n_forest_type),
        exp_sp_1_5 = sum(exp_sp_1_5) / first(n_forest_type),   #expP = sum(expP),
        upperexposed = sum(upperexposed) / first(n_forest_type),
        stemcount = sum(stemcount) / first(n_forest_type), # the total number of stems
        GTR5ft = sum(coalesce(GTR5ft, 0), na.rm = TRUE) / first(n_forest_type),
        GTR1ft = sum(GTR1ft) / first(n_forest_type),
        all_sizes = sum(all_sizes)/ first(n_forest_type),
        .groups = "drop"
    ) %>% 
    mutate(
        df = "FT_calc3",
        stand = "all_not_CAFRI"
    ) %>% 
#   filter(growingseasons > 2) %>%  # growing season filtered in df=veg_data
    filter(all_sizes >0) %>% 
    arrange(
        forest_type, growingseasons, brush, trmt, type) %>% 
    select(
        filename, df, stand, forest_type, growingseasons, brush, trmt,  type,  upperexposed, GTR5ft, GTR1ft, all_sizes,
        clumps_per_acre, seed_4_12, sup_sd12_5, sup_sd5_1, sup_sp_1_5, exp_sd12_5, exp_sd5_1, exp_sp_1_5, n_forest_type, stems_per_ramet
    ) %>% 
    mutate_if(is.numeric, round, 0)
   
table(FT_calc3$growingseasons) # result shows the growing season and the #obs

```




#FT_calc4 - summary by trmt and growing season
```{r}


FT_calc4 <- FT_calc1 %>% 
left_join(trmt_df, by = c("growingseasons", "trmt")) %>% 
filter(!brush == "unbrush") %>% #unbrushed removed in the "top 10" code at the bottom
group_by(growingseasons, trmt, type) %>%  #summary by type (not spp) and treatment labels
    summarize(
        filename = first(filename),
        n_trmt_df = first(n_trmt_df),
        clumps_per_acre = sum(clumps_per_acre) / first(n_trmt_df),  # Sum all specified numeric clumps.acre and stems.ramet calc confirmed)
        stems_per_ramet = sum(stems_per_ramet) / first(n_trmt_df),
        seed_4_12 = sum(seed_4_12) / first(n_trmt_df),
        sup_sd12_5 = sum(sup_sd12_5) / first(n_trmt_df),    #suppressed seedling 12" to 5 ft
        sup_sd5_1 = sum(sup_sd5_1) / first(n_trmt_df),   # suppressed seedling 5 ft to 1" dbh
        sup_sp_1_5 = sum(sup_sp_1_5) / first(n_trmt_df),  # suppressed sapling 1" to 5"
        exp_sd12_5 = sum(exp_sd12_5) / first(n_trmt_df), # exposed seedling 12" to 5 ft
        exp_sd5_1 = sum(exp_sd5_1) / first(n_trmt_df),
        exp_sp_1_5 = sum(exp_sp_1_5) / first(n_trmt_df),   #expP = sum(expP),
        upperexposed = sum(upperexposed) / first(n_trmt_df),
        stemcount = sum(stemcount) / first(n_trmt_df), # the total number of stems
        GTR5ft = sum(coalesce(GTR5ft, 0), na.rm = TRUE) / first(n_trmt_df),
        GTR1ft = sum(GTR1ft) / first(n_trmt_df),
        all_sizes = sum(all_sizes)/ first(n_trmt_df),
        .groups = "drop"
    ) %>% 
    mutate(
        df = "FT_calc4",
        stand = "all_not_CAFRI"
    ) %>% 
#   filter(growingseasons > 2) %>%  # growing season filtered in df=veg_data
    filter(all_sizes >0) %>% 
    arrange(
        growingseasons, trmt, type) %>% 
    select(
        filename, df, stand,  growingseasons, trmt,  type,  upperexposed, GTR5ft, GTR1ft, all_sizes,
        clumps_per_acre, seed_4_12, sup_sd12_5, sup_sd5_1, sup_sp_1_5, exp_sd12_5, exp_sd5_1, exp_sp_1_5, n_trmt_df, stems_per_ramet
    ) %>% 
    mutate_if(is.numeric, round, 0)
   
table(FT_calc4$growingseasons) # result shows the growing season and the #obs

```



## summary output from FT_calc2 doesn't agree with FT_calc4 (14october2025)
# this chunk was to rectify the miscalc of FT_ calc2 and calc4. Due to point counts where danby and goundry had same point numbers
# as of 16oct2025 there is still misalignment of calculations
```{r eval=TRUE}
d2 <- FT_calc2 %>% 
    filter(growingseasons == 3 | growingseasons == 5) %>%  #retain if growingseasons equal 3 or 5
    filter(brush == "brushed") %>% 
    select(forest_type, n_forest_type, growingseasons, brush, trmt, spp, type, GTR1ft, GTR5ft, all_sizes) %>% 
    group_by(trmt, growingseasons, type) %>% 
    summarize(
        n = sum(n_forest_type),
        n_forest_type_d2 = first(n_forest_type),
        all_sizes_d2 = sum(all_sizes),
        GTR1ft_d2 = sum(GTR1ft),
        GTR5ft_d2 = sum(GTR5ft)
    )

d2_total <- d2 %>% 
    group_by(trmt, growingseasons) %>% 
  summarize(
    n_d2 = first(n_forest_type_d2),
    type = "total",  # Label for remaining species
    across(where(is.numeric), sum, na.rm = TRUE),  # Sum numeric variables
    .groups = "drop"
  ) %>% 
    bind_rows(d2) %>% 
    arrange(growingseasons, trmt, type)

d4 <- FT_calc4 %>% 
     filter(growingseasons == 3 | growingseasons == 5) %>%  #retain if growingseasons equal 3 or 5
     select(growingseasons, n_trmt_df, trmt, type, GTR1ft, GTR5ft, all_sizes) %>% 
     group_by(trmt, growingseasons, type) %>% 
     summarize(
        n_trmt_df_d4 = first(n_trmt_df),
        all_sizes_d4 = sum(all_sizes),
        GTR1ft_d4 = sum(GTR1ft),
        GTR5ft_d4 = sum(GTR5ft)
     )

d4_total <- d4 %>% 
    group_by(trmt, growingseasons) %>% 
    summarize(
    n_d4 = first(n_trmt_df_d4),
        type = "total",  # Label for remaining species
    across(where(is.numeric), sum, na.rm = TRUE),  # Sum numeric variables
    .groups = "drop"
  ) %>% 
    bind_rows(d4) %>% 
    arrange(growingseasons, trmt, type)
   
mesh <- d2_total %>% 
    full_join(d4_total, by = c("trmt", "growingseasons", "type")) %>% 
    select(trmt,type, growingseasons,n_forest_type_d2,n_trmt_df_d4,all_sizes_d2,all_sizes_d4,
           GTR1ft_d2,GTR1ft_d4,GTR5ft_d2,GTR5ft_d4)
```



#recommended test of coding provided by claude AI on 17 oct 2025
```{r}
# Check the calculation method
d4_check <- FT_calc1 %>%
    left_join(trmt_df, by = c("growingseasons", "trmt")) %>% 
    group_by(growingseasons, trmt, type) %>%
    summarise(
        total_stems = sum(stemcount, na.rm = TRUE),
        trmt_df = first(n_trmt_df),
        avg_stems_per_acre2 = total_stems / trmt_df,
        .groups = "drop"
    )

# Verify by spot-checking one combination manually
d4_check %>% filter(growingseasons == 3, trmt == "control", type == "commercial")
# the value from this checks with the output from the TYPE BY TRMT code at the bottom


d4_by_forest <- FT_calc1 %>%
#  left_join(forest_type_df, by = c("growingseasons", "trmt", "forest_type")) %>%
  group_by(growingseasons, trmt, forest_type, brush, type, spp) %>%
  summarise(
    total_stems = sum(stemcount, na.rm = TRUE),
    forest_n = first(n_forest_type),
    avg_stems_per_acre = total_stems / forest_n,
    .groups = "drop"
  )

# Then sum across forest types
d4_from_forests <- d4_by_forest %>%
  group_by(growingseasons, trmt, type) %>%
  summarise(
    combined_total_stems = sum(total_stems),
    combined_n_points = sum(forest_n),
    combined_avg = combined_total_stems / combined_n_points,
    .groups = "drop"
  )

# Compare with your d4_check
comparison <- d4_check %>%
  left_join(d4_from_forests, by = c("growingseasons", "trmt", "type")) %>%
  mutate(diff = round(avg_stems_per_acre2 - combined_avg, 4))

comparison

```



#ignore output_file code.  All files written to the same excel.  See line 1460
```{r}
# output_file <- file.path(output_folder, paste0("forest_type_spp_summary_cafri6.6(23sept2025)", ".xlsx"))
# write_xlsx(FT_calc2, output_file)

```



#IGNORE
#Pre-process df for subsequent manipulations. Sum within species
```{r eval=FALSE}
d1.1 <- d0 %>% #aggregate or average to collapse "origin"
# use d1a    filter(harvest == "6_6" & season == "2023") %>% 
    group_by(cluster, harvest, season, wall, brush, point, spp) %>%  #combines origin into spp for simple summary tables
    summarize(
        filename = first(filename),
        point_count = n_distinct(point),
        n = n(),
        trmt = first(trmt),
        forest_type = first(forest_type),
        type = first(type),
        firstgrow = first(firstgrow),
        year = first(year),
        denom = first(denom),
        growingseasons = first(growingseasons),
        height = first(height),
        clumps_per_acre = sum(clumps_per_acre),  # Sum all specified numeric clumps.acre and stems.ramet calc confirmed)
        stems_per_ramet = sum(stems_per_ramet),
        seedA = sum(seedA),
        sup_sd12_5 = sum(sup_sd12_5),    #suppressed seedling 12" to 5 ft
#        supC1 = sum(supC1), C! and C2 don't exist in 2023 data
#        supC2 = sum(supC2),
        sup_sd5_1 = sum(sup_sd5_1),   # suppressed seedling 5 ft to 1" dbh
        sup_sp_1_5 = sum(sup_sp_1_5),  # suppressed sapling 1" to 5"
        exp_sd12_5 = sum(exp_sd12_5), # exposed seedling 12" to 5 ft
#        expC1 = sum(expC1),
#        expC2 = sum(expC2),
        exp_sd5_1 = sum(exp_sd5_1),
        exp_sp_1_5 = sum(exp_sp_1_5),   #expP = sum(expP),
        # across(all_of(numeric_vars1), sum, na.rm = TRUE),  
        # across(all_of(numeric_vars2), mean, na.rm = TRUE),  # average of all specified numeric vars2
        .groups = "drop"
    ) %>% 
    mutate(
        GTR5ft = exp_sd5_1 + exp_sp_1_5,
        GTR1ft = sup_sd12_5 + sup_sd5_1 + sup_sp_1_5 + exp_sd12_5 + exp_sd5_1 + exp_sp_1_5, 
            # the old variable names supB + supD + supP + expB + expD + expP 
            # this used to be "total"
        all_sizes = seedA + GTR1ft, 
        df = "d1"
    )


d1 <- d1.1 %>%  #d1 = d1.1, just arranged differently 
    arrange(desc(season), harvest, wall, type, spp, point) %>% 
    select(filename, df, harvest, wall, point, type, spp, season, trmt, brush, everything())

d1_stub <- d1 %>%   #d1a is useful to test calculation from the mostly raw data
    filter(season == "2024") %>% 
    filter(harvest == "6_6") %>% 
    filter(all_sizes > 0) %>% 
    arrange(desc(season), harvest, trmt, type, brush, spp, wall,  point)

#    arrange(filename, (harvest), desc(season), wall, type, spp) %>% 

```



#define numeric variables
```{r}
#nu vars to sum by origin
numeric_vars1 <- c("GTR1ft", "all_sizes", "GTR5ft", "upperexposed", "seed_4_12", "sup_sd12_5", 
                   "sup_sd5_1", "sup_sp_1_5", "exp_sd12_5", "exp_sd5_1", "exp_sp_1_5")

```



# count # spp within type present at each point. Summarize tpa by point
#df = richness from FT_calc2
```{r}
richness <- FT_calc2 %>%
#    filter(season == "2023") %>%  # filter for season while checking code, then run for all seasons
    group_by(forest_type, growingseasons, trmt, brush, type) %>%  # include "brush"
    summarize(
        filename = first(filename),
        type = first(type),
        growingseasons = first(growingseasons),
        n_forest_type = first(n_forest_type),
        denominator = n(), #denominator = # spp x type
        type_count = n(),
        richness_type_all_sizes = sum(all_sizes > 0, na.rm = TRUE),  # Record presence at points by type where "all_sizes" > 0, for stems > 4" tall
        richness_type_GTR1ft = sum(GTR1ft > 0, na.rm = TRUE),  # Record presence at points by type where total > 0, for stems > 12" tall
        richness_type_GTR5ft = sum(GTR5ft > 0, na.rm = TRUE), # Record presence at points by type where "GTR5" >0
        across(all_of(numeric_vars1), sum, na.rm = TRUE),  # Sum all specified numeric vars1 (logic for sum confirmed)

       # Calculate mean for clumps.acre and stems.ramet based on non-zero count
       clumps_per_acre = if_else(sum(clumps_per_acre > 0, na.rm = TRUE) > 0,  #counts the number of non-zero values of clump.acre
                                  sum(clumps_per_acre, na.rm = TRUE) / sum(clumps_per_acre > 0, na.rm = TRUE),
                                 #if there are non-zero values, completes the calc
                                 #
                                  NA_real_), #calculations for clumps_per_acre and stems.ramet confirmed

       stems_per_ramet = if_else(sum(stems_per_ramet > 0, na.rm = TRUE) > 0,
                                  sum(stems_per_ramet, na.rm = TRUE) / sum(stems_per_ramet > 0, na.rm = TRUE),
                                  NA_real_),

                .groups = "drop"
        
        #need to average clumps.acre and stems.ramet for the non-zero values within a type
        ) %>%

        mutate(
            across(where(is.numeric), ~ round(., 1)),
            df = "richness"
        ) %>% 

        select(filename, df,  forest_type, growingseasons, brush, trmt,type, denominator, type_count, 
           richness_type_all_sizes, richness_type_GTR1ft, richness_type_GTR5ft, all_sizes,  GTR1ft, GTR5ft,  
           everything()) %>% 
        arrange(forest_type, growingseasons, brush, trmt, type)
    #arrange(desc(season), harvest, trmt, type, brush, spp, wall,  point)
    #

spp_list <- FT_calc1 %>% 
    select (forest_type, growingseasons, brush, trmt, spp, GTR5ft) %>% 
    filter(GTR5ft >0) %>% 
    group_by(forest_type, growingseasons, brush, trmt) %>%
    summarise(
        species_present = toString(unique(spp)),   # comma-separated list
        reported = "GTR5ft",
        .groups = "drop"
    ) %>% 
    select(forest_type, growingseasons, trmt, brush, reported, species_present) %>% 
    arrange(forest_type, growingseasons, trmt, brush)

      
  

richness_summ <- richness %>% 
    group_by(forest_type, growingseasons, trmt, brush) %>% 
        summarize(
        filename = first(filename),
        richness_all_sizes = sum(richness_type_all_sizes, na.rm = TRUE),  # Sum the actual values
        richness_GTR1ft = sum(richness_type_GTR1ft, na.rm = TRUE),        # Sum the actual values  
        richness_GTR5ft = sum(richness_type_GTR5ft, na.rm = TRUE),        # Sum the actual values
        .groups = "drop"
        )

# this summarize code sums the values.  To count the values, include "sum(variable > 0, na.rm=TRUE)" as in the previous df
```




#graph richnes
```{r}
figure_rich_gtr5 <- richness_summ %>%
  filter(forest_type == "n_hdwd") %>%  # Filter for both brush and forest_type
  mutate(growingseasons = as.numeric(growingseasons)) %>%  # Convert back to numeric
  ggplot(aes(x = factor(growingseasons), 
             y = richness_GTR5ft, 
             fill = trmt)) +
  geom_col(position = "dodge", width = 0.7) +  # Side-by-side bars
  facet_wrap(~brush, scales = "fixed") +
#  facet_wrap(~brush, scales = "free_y") +
  labs(
    title = "Forest Type Woody Richness by Growing Season (N. Hardwood)",
    subtitle = "Number of Species >5ft tall fig=rich_gtr5_NH slash_wall_analyses2025(by cluster).Rmd",
    x = "Growing Seasons",
    y = "Number of Species",
    fill = "Treatment"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  ) +
  scale_fill_manual(values = c("control" = "blue", "protected" = "green"))

# Display the plot
print(figure_rich_gtr5)

# Save the plot
ggsave("figures/forest_type_richness_Figure_rich__NH_gtr5.jpg", 
       plot = figure_rich_gtr5, width = 8, height = 6, dpi = 300)

```



# occupancy by forest type for type. 
# 
```{r}
forest_type_occupancy <- FT_calc1 %>%
#    filter(growingseasons > 2) %>% 
  group_by(forest_type, trmt, brush, growingseasons, type) %>%
  summarize(
    filename = first(filename),
    n_forest_type = first(n_forest_type),
    # Count distinct points where ANY species of this type has all_sizes > 0
    occup_any_count = n_distinct(point[all_sizes > 0], na.rm = TRUE),
    occup_any_pct = occup_any_count / first(n_forest_type) * 100,
    
    occup_GTR1ft_count = n_distinct(point[GTR1ft > 0], na.rm = TRUE),
    occup_GTR1ft_pct = occup_GTR1ft_count / first(n_forest_type) * 100,
    
    occup_GTR5ft_count = n_distinct(point[GTR5ft > 0], na.rm = TRUE),
    occup_GTR5ft_pct = occup_GTR5ft_count / first(n_forest_type) * 100,
    .groups = "drop"
  )  %>%   
    mutate(
        across(where(is.numeric), ~ round(., 1))
  ) %>% 
  mutate(
        df = "forest_type_occupancy PCT of points",
        stand = "all_not_CAFRI"
  ) %>% 
    arrange(forest_type, growingseasons, trmt, brush, type) %>% 
    select(filename, df, stand, forest_type, trmt, brush, growingseasons, type, n_forest_type, occup_any_count, 
           occup_any_pct, occup_GTR1ft_pct, occup_GTR5ft_pct, everything())
```

#something wrong with filter into ggplot, next chunk
```{r}
forest_type_occupancy %>%
  count(brush, forest_type) %>%
  arrange(brush, forest_type)

step1 <- forest_type_occupancy %>% filter(brush == "b")
step2 <- forest_type_occupancy %>% filter(forest_type == "n_hdwd") 
step3 <- forest_type_occupancy %>% filter(brush == "b", forest_type == "n_hdwd")

cat("brush == 'b':", nrow(step1), "\n")
cat("forest_type == 'n_hdwd':", nrow(step2), "\n") 
cat("Both filters:", nrow(step3), "\n")
```

#graph forest type occupancy
# figure = occ_1
```{r}
# Create bar graph of forest type occupancy
library(ggplot2)

# Filter for brush = "u" only and create clean plot
figure_occ_B_bars <- forest_type_occupancy %>%
  mutate(growingseasons = as.numeric(growingseasons)) %>%  # Convert back to numeric
  filter(brush == "unbrushed", forest_type == "n_hdwd", growingseasons < 7) %>%  # Filter for both brush and forest_type
  filter(!(type == "other")) %>% 

  ggplot(aes(x = factor(growingseasons), 
             y = occup_GTR5ft, 
             fill = trmt)) +

  geom_col(position = "dodge", width = 0.7) +  # Side-by-side bars
  facet_wrap(~type, scales = "fixed") +
#  facet_wrap(~type, scales = "free_y") +
  labs(
    title = "Forest Type Occupancy by Growing Season (N. Hardwood, Unbrushed)",
    subtitle = "% of points stems >5ft tall fig=occ_U_NH slash_wall_analyses2025(by cluster).Rmd",
    x = "Growing Seasons",
    y = "Occupancy (%)",
    fill = "Treatment"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  ) +
  scale_fill_manual(values = c("control" = "blue", "protected" = "green"))

# Display the plot
print(figure_occ_B_bars)

#notes
#1. oak-hdwd unbrushed, and U NH suggest there is zero occupancy comm in 7th growing season protected.
#   confirm there are dat for this and why the pattern isn't present. The only site with 7 growing
#   seasons and oak pine unbrushed is Danby. I don't recall that had 75% comm occupancy > 5ft.
#   Run those data through a filtered df to test the methods.


# Save the plot
ggsave("figures/forest_type_occupancy_Figure_occ_U_NH_bars.jpg", 
       plot = figure_occ_B_bars, width = 8, height = 6, dpi = 300)
```


cluster occupancy
```{r}
#data output not accurate - need to check calculations (16sept2025)
#
cluster_type_occupancy <- type_summary %>%
    select(-point_count) %>%   #this version of "point_count" won't work within the cluster summary

#group by harvest to include the last measurement season. then compare within the cluster
  group_by(harvest, trmt, brush) %>%
    slice_max(order_by = as.numeric(season), n = 1, with_ties = TRUE) %>% # "with_ties" if multiple rows have the same latest season, you keep them all 
  ungroup() %>% 

    
        left_join(cluster_point_count, by = c("cluster", "trmt", "brush")) %>% 

    group_by(cluster, season, trmt, brush, type) %>%  #include "brush"
    summarize(
    # point_count = first(point_count), #n_distinct(point),# Total points in the wall
    # growingseasons = first(growingseasons),
    # year = first(year),
    point_count = first(point_count),
    plots_with_any = sum(all_sizes > 0, na.rm = TRUE), # Count of plots where type is represented by a spp
    occup_any_pct = (plots_with_any / point_count) * 100, 
    # should this be outside the summarize code?    
    # by type, the percentage of plots within a wall with spp any size of that type

    plots_with_GTR1ft = sum(GTR1ft > 0, na.rm = TRUE), # Count of plots where type is represented by a spp >12" tall
    GTR1ft_pct = (plots_with_GTR1ft / point_count) * 100,

    plots_with_GTR5ft = sum(GTR5ft > 0, na.rm = TRUE), # Count of plots where type is represented by a spp
    GTR5ft_pct = (plots_with_GTR5ft / point_count) * 100,
    
    clumps_per_acre = sum(clumps_per_acre / point_count, na.rm = TRUE), 
        # this will calc clumps.acre based on the number of points per wall. points with zero included
    stems_per_ramet = if_else(sum(stems_per_ramet > 0, na.rm = TRUE) > 0,
                                  sum(stems_per_ramet, na.rm = TRUE) / sum(stems_per_ramet > 0, na.rm = TRUE),
                                  NA_real_), # this will calc stems.ramet based on the number of points w/i wall that have clumps

    across(all_of(numeric_vars1), ~ sum(.x, na.rm = TRUE) / point_count),  # Normalize by point_count
#    across(all_of(numeric_vars2), ~ sum(.x, na.rm = TRUE) / point_count),  # Normalize by point_count
    .groups = "drop"
  )  %>%   
    mutate(
        across(where(is.numeric), ~ round(., 1))
  ) %>% 
  
  mutate(
        filename = "slash_wall_analyses_2025.Rmd",
        df = "cluster_type_occupancy"
  ) %>% 
    select(filename, df, cluster, season, trmt, type, point_count,  plots_with_any, occup_any_pct, plots_with_GTR1ft, 
           GTR1ft_pct, plots_with_GTR5ft, GTR5ft_pct, clumps_per_acre, stems_per_ramet, everything())

table(wall_type_occupancy$harvest, wall_type_occupancy$growingseasons)

```



```{r}
# output_file <- file.path(output_folder, paste0("forest_type_occupancy(16sept2025)", ".xlsx"))
# write_xlsx(forest_type_occupancy, output_file)


```


# calcuate the values of richness (?) metrics at wall x spp
```{r}
species_by_wall <- d1 %>%
#  filter(season == "2023") %>%  # filter for season while checking code, then run for all seasons
  group_by(cluster, harvest, season, trmt, wall, spp) %>%
   summarize(
       filename = first(filename), 
       type = first(type),
       point_count = n_distinct(point),
            #calculate the numb points 
       numb_pts_any = sum(all_sizes > 0, na.rm = TRUE),  # Count species where all_sizes > 0
       numb_pts_GTR1ft = sum(GTR1ft > 0, na.rm = TRUE),  # Count species where "total" 12"+ > 0
       numb_pts_GTR5ft = sum(GTR5ft > 0, na.rm = TRUE), # count of species where GTR5>0
       any_size_pct = numb_pts_any / point_count *100,
       GTR1ft_pct = numb_pts_GTR1ft / point_count * 100,
       GTR5ft_pct = numb_pts_GTR5ft / point_count * 100,
       
       across(all_of(numeric_vars1), ~ sum(.x, na.rm = TRUE) / point_count),  # Normalize by point_count
       across(all_of(numeric_vars2), ~ sum(.x, na.rm = TRUE) / point_count),  # Normalize by point_count
       
       x1 = sum(stems_per_ramet, na.rm = TRUE),
       x2 = sum(stems_per_ramet > 0, na.rm = TRUE), 
       clumps_per_acre = sum(clumps_per_acre),
       stems_per_ramet = sum(stems_per_ramet),
       clumps_per_acre = sum(clumps_per_acre / point_count, na.rm = TRUE), 
            # this will calc clumps.acre based on the number of points per wall. points with zero included
       stems_per_ramet = if_else(x2 > 0, x1 / x2, NA_real_),
       # stems.ramet.2 = case_when(
       #      sum(stems.ramet > 0, na.rm = TRUE) > 0 ~ sum(stems.ramet, na.rm = TRUE) / sum(stems.ramet > 0, na.rm = TRUE),
       #      TRUE ~ NA_real_), # calc confirmed, stems.ramet based on # spp that have ramets

           .groups = "drop"
  )  %>%   
  mutate(
        across(where(is.numeric), ~ round(., 1)),
        df = "species_by_wall"
  ) %>%
    filter(all_sizes > 0) %>% 
    arrange(filename, (harvest), desc(season), trmt, type, spp) %>% 
    select(filename, df, cluster, harvest, season, trmt, wall, growingseasons, spp, type, 
           point_count, numb_pts_any, any_size_pct, numb_pts_GTR1ft, GTR1ft_pct, 
           numb_pts_GTR5ft, GTR5ft_pct, clumps_per_acre, stems_per_ramet,  everything())

output_file <- file.path(output_folder, paste0("spp_x_wall_x_trmt_x_season(15sept2025)", ".xlsx"))
write_xlsx(species_by_wall, output_file)


#rerun spp x wall as type x harvest

```

# species averages across walls by harvest (changed to "by cluster" on 15sept2025)
```{r}
point_counts <- species_by_wall %>% 
        group_by(cluster, harvest, season, trmt, wall) %>% 
        summarize(
            point_count = first(point_count), # the number of points in a wall
            .groups = "drop"
        ) %>% 
        arrange((harvest), desc(season), trmt, wall) 

    
wall_counts <- point_counts %>%
    group_by(cluster, harvest, season, trmt) %>%
    summarize(
        wall_count = n_distinct(wall), # number of walls in a harvest
        points_per_wall = sum(point_count), # could also be called points_per_trmt
        .groups = "drop")

species_by_cluster <- species_by_wall %>% 
#        filter(harvest == "6_6" & season == "2023") %>% 
        left_join(wall_counts, by = c("cluster", "harvest", "season", "trmt")) %>%  # Join to get the correct wall_count
        group_by(cluster, harvest, season, trmt, spp) %>% 
        summarize(
            filename = first(filename),
            points_per_harvest = first(points_per_wall),
            # numb_pts_any = first(numb_pts_any), # I don't think this is needed
            # numb_pts_seedl = first(numb_pts_seedl), # same
            # numb_pts_GTR5ft = first(numb_pts_GTR5ft), # same
            type = first(type),
            freq_by_grouping = n(),
            wall_count = first(wall_count),
#            avg_pts_wall_any = sum(numb_pts_any) / wall_count,
#            avg_pts_wall_seedl = sum(numb_pts_seedl) / wall_count,
#            avg_pts_wall_GTR5ft = sum(numb_pts_GTR5ft) / wall_count,
            pct_points_any = (sum(numb_pts_any)) / points_per_harvest *100,
            pct_points_GTR1ft = (sum(numb_pts_GTR1ft)) / points_per_harvest *100, # for stems > 12" tall
            pct_points_GTR5ft = (sum(numb_pts_GTR5ft)) / points_per_harvest * 100,
            
           across(all_of(numeric_vars1), ~ sum(.x, na.rm = TRUE) / wall_count),  
            # summ numeric from each wall, divide by number of walls = avg/harvest
           across(all_of(numeric_vars2), mean, na.rm = TRUE),
           #across(all_of(numeric_vars2), ~ sum(.x, na.rm = TRUE) / wall_count)  # Normalize by wall_count
           #
       x1 = sum(stems_per_ramet, na.rm = TRUE),
       x2 = sum(stems_per_ramet > 0, na.rm = TRUE), 
       clumps_per_acre = sum(clumps_per_acre),
       stems_per_ramet = sum(stems_per_ramet),
       clumps_per_acre = sum(clumps_per_acre / points_per_wall, na.rm = TRUE), 
            # this will calc clumps.acre based on the number of points per wall. points with zero included
       stems_per_ramet = if_else(x2 > 0, x1 / x2, NA_real_),

       .groups = "drop"
        
    ) %>% 
    arrange(filename, (harvest), desc(season), trmt, type, spp) %>% 
    mutate(df = "species_by_cluster") %>% 
#    select(-x1, -x2, -supC1, -supC2, -expC1, -expC2, -denom) %>% 
        select(filename, df, cluster, harvest, season, trmt,  wall_count, points_per_harvest, growingseasons, 
               spp, type, pct_points_any, pct_points_GTR1ft,
               pct_points_GTR5ft, clumps_per_acre, stems_per_ramet, everything()) %>% 
    mutate(
        across(where(is.numeric), ~ round(., 1))
    )

output_file <- file.path(output_folder, paste0("spp_x_cluster_x_trmt_x_season(15sept2025)", ".xlsx"))
write_xlsx(species_by_cluster, output_file)


```


# TYPE occupancy averages across walls by harvest (by cluster as of 15sept2025)
```{r}
point_counts <- species_by_wall %>% 
        group_by(cluster, harvest, season, trmt, wall) %>% 
        summarize(
            point_count = first(point_count),
            .groups = "drop"
        )
    
wall_counts <- point_counts %>%
    group_by(cluster, harvest, season, trmt) %>%
    summarize(
        wall_count = n_distinct(wall),
        points_per_wall = sum(point_count),
        .groups = "drop")

type_by_cluster <- species_by_wall %>%  #code copied from df=species_by_cluster
        left_join(wall_counts, by = c("cluster", "harvest", "season", "trmt")) %>%  # Join to get the correct wall_count
        group_by(cluster, harvest, season, trmt, type) %>% 
        summarize(
            filename = first(filename),
            growingseasons = first(growingseasons),
            points_per_cluster = sum(points_per_wall), #check this calculation as of 9/15/2025 seems high (early, danby control commercial)
            type = first(type),
            freq_by_grouping = n(),

            pct_points_any = (sum(numb_pts_any)) / points_per_cluster *100,
            pct_points_GTR1ft = (sum(numb_pts_GTR1ft)) / points_per_cluster *100, # for stems > 12" tall
            pct_points_GTR5ft = (sum(numb_pts_GTR5ft)) / points_per_cluster * 100,
            
           across(all_of(numeric_vars1), ~ sum(.x, na.rm = TRUE) / points_per_cluster),  # summ numeric from each wall, divide by number of walls = avg/harvest
           across(all_of(numeric_vars2), mean, na.rm = TRUE),

       x1 = sum(stems_per_ramet, na.rm = TRUE),
       x2 = sum(stems_per_ramet > 0, na.rm = TRUE), 

       clumps_per_acre = sum(clumps_per_acre),
       stems_per_ramet = sum(stems_per_ramet),
       clumps_per_acre = sum(clumps_per_acre / points_per_cluster, na.rm = TRUE), 
            # this will calc clumps.acre based on the number of points per wall. points with zero included
       stems_per_ramet = if_else(x2 > 0, x1 / x2, NA_real_),

       .groups = "drop"
    ) %>% 

    arrange(filename, (harvest), desc(season), trmt, type) %>% 
    mutate(df = "type_by_cluster") %>% 
#    select(-x1, -x2, -supC1, -supC2, -expC1, -expC2, -denom) %>% 
        select(filename, df, cluster, harvest, season, trmt, points_per_cluster, growingseasons, 
               type, pct_points_any, pct_points_GTR1ft,
               pct_points_GTR5ft, clumps_per_acre, stems_per_ramet, everything()) %>% 
    mutate(
        across(where(is.numeric), ~ round(., 1))
    )

    
    
    
    
#####    
#         left_join(wall_counts, by = c("cluster", "harvest", "season", "trmt", "wall_count")) %>%  # Join to get the correct wall_count
#         group_by(cluster, harvest, season, trmt, type) %>% 
#         summarize(
#             filename = first(filename),
#             points_per_wall = first(points_per_wall),
#             type = first(type),
#             freq_by_grouping = n(),
#             wall_count = first(wall_count),
# #             pts_wall_any = sum(numb_pts_any), #nonsensical.  Go to sum@point, flag a point if any comm, if any interf, etc. then summ
# #             pts_wall_seedl = sum(numb_pts_seedl),
# #             pts_wall_GTR5ft = sum(numb_pts_GTR5ft),
#             pct_points_any = sum(numb_pts_any) / points_per_wall *100,
#             pct_points_seedl = sum(numb_pts_seedl) / points_per_wall *100,
#             pct_points_GTR5ft = sum(numb_pts_GTR5ft) / points_per_wall * 100,
#            type_diversity = n_distinct(spp), # the number of species per type. NEED to standardize this for area
#            across(all_of(numeric_vars1), ~ sum(.x, na.rm = TRUE) / wall_count),  # Normalize by wall_count
#            across(all_of(numeric_vars2), mean, na.rm = TRUE),
#            #across(all_of(numeric_vars2), ~ sum(.x, na.rm = TRUE) / wall_count)  # Normalize by wall_count
#            #
#        x1 = sum(stems_per_ramet, na.rm = TRUE), #the total of the stems per ramet
#        x2 = sum(stems_per_ramet > 0, na.rm = TRUE), # the number of spp with clumps
#        clumps_per_acre = sum(clumps_per_acre),
#        clumps_per_acre = sum(clumps_per_acre),
#        stems_per_ramet = if_else(x2 > 0, x1 / x2, NA_real_), #average stems.ramet
# 
#        .groups = "drop"
#         
#     ) %>% 
#     mutate(
#         df = "type_by_cluster"
#     ) %>% 
#     arrange(filename, (harvest), desc(season), trmt, type) %>% 
#     select(!c(pts_wall_any, pct_points_any,pts_wall_seedl,pct_points_seedl,pts_wall_GTR5ft, pct_points_GTR5ft,
#               supC1, supC2, expC1, expC2, denom)) %>% 
#         select(filename, df, cluster, harvest, season, growingseasons, trmt,  wall_count, points_per_wall, type, 
#            clumps.acre, stems.ramet, type_diversity, everything()) %>% 
#     mutate(
#         across(where(is.numeric), ~ round(., 1))
#     )

output_file <- file.path(output_folder, paste0("type_x_harvest_x_trmt_x_season(20feb2025)", ".xlsx"))
write_xlsx(type_by_harvest, output_file)


```



#species richness by wall
```{r eval=TRUE}

# Step 1 — collapse to harvest/season/trmt/spp and record presence (across all walls)
species_presence <- species_by_wall %>%
  group_by(harvest) %>%
  filter(year == max(year)) %>%             # keep most recent growing season for each harvest
  ungroup() %>%
  group_by(harvest, season, trmt, growingseasons, spp) %>%
  summarize(
    filename          = first(filename),
    present_any       = any(numb_pts_any    > 0, na.rm = TRUE),
    present_GTR1ft    = any(numb_pts_GTR1ft > 0, na.rm = TRUE),
    present_GTR5ft    = any(numb_pts_GTR5ft > 0, na.rm = TRUE),
    present_any_pct   = any(any_size_pct    > 0, na.rm = TRUE),
    present_GTR1ft_pct= any(GTR1ft_pct      > 0, na.rm = TRUE),
    present_GTR5ft_pct= any(GTR5ft_pct      > 0, na.rm = TRUE),
    present_all_sizes = any(all_sizes       > 0, na.rm = TRUE),
    present_GTR1ft_all= any(GTR1ft          > 0, na.rm = TRUE),
    present_GTR5ft_all= any(GTR5ft          > 0, na.rm = TRUE),
    .groups = "drop"
  )

# Step 2 — count species and list them
richness <- species_presence %>%
  group_by(harvest, season, trmt, growingseasons) %>%
  summarize(
    filename = first(filename),
    df       = "richness",

    richness_any        = sum(present_any,       na.rm = TRUE),
    spp_any             = paste(sort(unique(spp[present_any])), collapse = ", "),

    richness_GTR1ft     = sum(present_GTR1ft,    na.rm = TRUE),
    spp_GTR1ft          = paste(sort(unique(spp[present_GTR1ft])), collapse = ", "),

    richness_GTR5ft     = sum(present_GTR5ft,    na.rm = TRUE),
    spp_GTR5ft          = paste(sort(unique(spp[present_GTR5ft])), collapse = ", "),

    # richness_any_pct    = sum(present_any_pct,   na.rm = TRUE),
    # spp_any_pct         = paste(sort(unique(spp[present_any_pct])), collapse = ", "),
    # 
    # richness_GTR1ft_pct = sum(present_GTR1ft_pct,na.rm = TRUE),
    # spp_GTR1ft_pct      = paste(sort(unique(spp[present_GTR1ft_pct])), collapse = ", "),
    # 
    # richness_GTR5ft_pct = sum(present_GTR5ft_pct,na.rm = TRUE),
    # spp_GTR5ft_pct      = paste(sort(unique(spp[present_GTR5ft_pct])), collapse = ", "),
    # 
    # richness_all_sizes  = sum(present_all_sizes, na.rm = TRUE),
    # spp_all_sizes       = paste(sort(unique(spp[present_all_sizes])), collapse = ", "),
    # 
    # richness_GTR1ft_all = sum(present_GTR1ft_all,na.rm = TRUE),
    # spp_GTR1ft_all      = paste(sort(unique(spp[present_GTR1ft_all])), collapse = ", "),
    # 
    # richness_GTR5ft_all = sum(present_GTR5ft_all,na.rm = TRUE),
    # spp_GTR5ft_all      = paste(sort(unique(spp[present_GTR5ft_all])), collapse = ", "),

    .groups = "drop"
  ) %>%
  arrange(harvest, season, trmt, growingseasons)


```

```{r}

output_file <- file.path(output_folder, paste0("species_richness_species_list)(8august20252025)", ".xlsx"))
write_xlsx(richness, output_file)


```

# graphs of df = type x harvest
# data preparation
```{r}

##
## need to change use of the term "cluster"  maybe change to "zone"
## 
prep_type <- type_by_harvest %>% 
    group_by(harvest) %>% 
    filter(year == max(year)) %>% # retain the most recent growing season within each harvest
    ungroup() %>% 
    filter(!harvest %in% c("decker", "recknagel_north")) %>%  # remove decker and recknagel_north
    mutate(
        cluster = "recknagel",
        cluster = case_when(
            harvest %in% c("6_6") ~ "6_6",
            harvest %in% c("6_9") ~ "6_9",
            harvest %in% c("gas_line") ~ "gas_line",
            harvest %in% c("red_pine") ~ "red_pine",
            harvest %in% c("sta_rd") ~ "sta_rd",
            TRUE ~ cluster,
    )) %>% 
  mutate(trmt = case_when(
    harvest == "recknagel" & is.na(trmt) ~ "control",
    TRUE ~ trmt  # Retain existing values in 'trmt' for all other cases
  ))


desired_order <- c("6_6", "6_9", "recknagel", "red_pine", "sta_rd", "gas_line") #places cluster elements in desired order, young to old
```


# Diversity, total, figure_D1
```{r}
diversity <- prep_type %>% 
  select(harvest, cluster, trmt, type_diversity, season) %>% 
  group_by(harvest, season, trmt) %>% 
  summarize(
    cluster = first(cluster),                       # Keeps the first occurrence of the cluster
    Number_of_Species = sum(type_diversity, na.rm = TRUE),  # Calculate the mean while ignoring NA values
    .groups = "drop"                                # Avoids grouping in the result
  ) %>% 
  mutate(cluster = factor(cluster, levels = desired_order)) %>%  # Set the desired order for clusters
  arrange(cluster, harvest, trmt)                        # Arrange by cluster, harvest, and treatment


figure_d1 <- ggplot(diversity, aes(x = cluster, y = Number_of_Species, fill = trmt)) +
  geom_bar(stat = "identity", position = "dodge")+ 
  scale_fill_manual(values = c("control" = "blue", "protected" = "green")) +
  scale_pattern_manual(values = c("control" = "stripe", "protected" = "crosshatch")) +  # Define patterns
  labs(
       x = "Cluster",
       y = "Number of Species",
      title = "Number of Species by Cluster and Treatment",
      subtitle = "Growing Season 2023, slash_wall_analyses_2025.Rmd. Figure_D1",
       fill = "trmt") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


  ggsave("figures/spp_diversity_cluster_trmt_Figure_D1.jpg", plot = figure_d1, width = 8, height = 6, dpi = 300)
```



# Stem Density, commercial and diversity. GTR5ft, sapl, commercial only
# Figure_S1, Figure_S2
```{r}
gtr5 <- prep_type %>% 
  select(harvest, cluster, trmt, GTR5ft, type, season) %>% 
  filter(type == "commercial" | type == "diversity") %>% 
      group_by(harvest, season, trmt) %>% 
  mutate(GTR5ft = case_when(
    harvest == "recknagel" & trmt == "control" ~ 10,
    TRUE ~ GTR5ft  # Insert small values (10) to allow graphing visualization. Retain existing values in 'trmt' for all other cases
  )) %>% 
  summarize(
      cluster = first(cluster),
      Comm_DiversityTPA = sum(GTR5ft),
      .groups = "drop"
      ) %>% 
  mutate(cluster = factor(cluster, levels = desired_order)) %>% 
    arrange(cluster, harvest, trmt)

figure_S1 <- ggplot(gtr5, aes(x = cluster, y = Comm_DiversityTPA, fill = trmt)) +
  geom_bar(stat = "identity", position = "dodge")+ 
  scale_fill_manual(values = c("control" = "blue", "protected" = "green")) +
  scale_pattern_manual(values = c("control" = "stripe", "protected" = "crosshatch")) +  # Define patterns
  labs(
       x = "Cluster",
       y = "Commercial and Diversity TPA",
      title = "Commercial and Diversity Species GTR 5 ft (TPA) by Cluster and Treatment",
      subtitle = "Growing Season 2023, slash_wall_analyses_2025.Rmd. Figure_S1",
       fill = "trmt") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


  ggsave("figures/comm_diversity_tpa_cluster_trmt_Figure_S1.jpg", plot = figure_S1, width = 8, height = 6, dpi = 300)


sapl <- prep_type %>% 
  select(harvest, cluster, trmt, expP, type, season) %>% 
  filter(type == "commercial" | type == "diversity") %>% 
  # mutate(GTR5ft = case_when(
  #   harvest == "recknagel" & trmt == "control" ~ 10,
  #   TRUE ~ GTR5ft  # Retain existing values in 'trmt' for all other cases
  # )) %>% 
  mutate(expP = case_when(
    harvest == "recknagel" & trmt == "control" ~ 3,
    TRUE ~ expP  # Recknagel controls had 0 stems, so plug data for graph visualization. Retain existing values in 'expP' for all other cases
  )) %>% 
    group_by(harvest, season, trmt) %>% 
    summarize(
      cluster = first(cluster),
      Exposed_Sapl_CommDiv = sum(expP),
      .groups = "drop"
      ) %>%
    group_by(cluster, trmt) %>% 
        summarize(
      cluster = first(cluster),
      Exposed_Sapl_CommDiv = mean(Exposed_Sapl_CommDiv),
      .groups = "drop"
      ) %>%
  mutate(cluster = factor(cluster, levels = desired_order))


figure_S2 <- ggplot(sapl, aes(x = cluster, y = Exposed_Sapl_CommDiv, fill = trmt)) +
  geom_bar(stat = "identity", position = "dodge")+ 
  scale_fill_manual(values = c("control" = "blue", "protected" = "green")) +
  scale_pattern_manual(values = c("control" = "stripe", "protected" = "crosshatch")) +  # Define patterns
  labs(
       x = "Cluster",
       y = "Exposed Saplings Commercial and Diversity TPA",
      title = "Exposed Saplings (Comm & Div Species) (TPA) by Cluster and Treatment",
      subtitle = "Growing Season 2023, slash_wall_analyses_2025.Rmd. Figure_S2",
       fill = "trmt") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


  ggsave("figures/exposed_sapling_CommDivSpp_tpa_cluster_trmt_Figure_S2.jpg", plot = figure_S2, width = 8, height = 6, dpi = 300)

  interf <- prep_type %>% 
  select(harvest, cluster, trmt, GTR5ft, type, season) %>% 
  filter(type == "interfere") %>% 
      group_by(harvest, season, trmt) %>% 
  # mutate(GTR5ft = case_when(
  #   harvest == "recknagel" & trmt == "control" ~ 10,
  #   TRUE ~ GTR5ft  # Retain existing values in 'trmt' for all other cases
  # )) %>% 
    group_by(harvest, season, trmt) %>% 
    summarize(
      cluster = first(cluster),
      InterGTR5 = sum(GTR5ft),
      .groups = "drop"
      ) %>%
    group_by(cluster, trmt) %>% 
    summarize(
      cluster = first(cluster),
      InterGTR5 = mean(InterGTR5),
      .groups = "drop"
      ) %>%
    mutate(cluster = factor(cluster, levels = desired_order))

figure_F1 <- ggplot(interf, aes(x = cluster, y = InterGTR5, fill = trmt)) +
  geom_bar(stat = "identity", position = "dodge")+ 
  scale_fill_manual(values = c("control" = "blue", "protected" = "green")) +
  scale_pattern_manual(values = c("control" = "stripe", "protected" = "crosshatch")) +  # Define patterns
  labs(
       x = "Cluster",
       y = "Interfering Species GTR 5 ft (TPA)",
      title = "Interfering Species GTR 5 ft (TPA) by Cluster and Treatment",
      subtitle = "Growing Season 2023, slash_wall_analyses_2025.Rmd. Figure_F1",
       fill = "trmt") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


  ggsave("figures/interfering_GTR5ft_tpa_cluster_trmt_Figure_F1.jpg", plot = figure_F1, width = 8, height = 6, dpi = 300)


```





# Stem Density, commercial only GTR5ft, sapl
# Figure_C1, Figure_S2
```{r}
comm01 <- prep_type %>% 
  select(harvest, cluster, trmt, GTR5ft, supB, expB, supD, expD, supP, total, expP, type, season) %>% 
  filter(type == "commercial") %>% 
  mutate(
      seedl_sml = supB + expB,
      seedl_lgr = supD + expD,
      sapl = supP + expP,
      GTR5ft = case_when(
    harvest == "recknagel" & trmt == "control" ~ 10,
    TRUE ~ GTR5ft  # Insert small values (10) to allow graphing visualization. Retain existing values in 'trmt' for all other cases
  )) %>% 
  group_by(cluster, trmt) %>% 
  summarize(
 #     harvest = first(harvest),
      seedl_sml = mean(seedl_sml),
      seedl_lgr = mean(seedl_lgr),
      sapl = mean(sapl),
#      total = mean(total),
      .groups = "drop"
      ) %>% 
  pivot_longer(cols = c(seedl_sml, seedl_lgr, sapl),
               names_to = "height_class",
               values_to = "tpa") %>%
  mutate(tpa = case_when(
    tpa == 0 ~ 100,  # Replace 0 with 25
    TRUE ~ tpa      # Keep all other values unchanged
  )) %>% 
  mutate(
    cluster = factor(cluster, levels = desired_order),
    height_class = factor(height_class, levels = c("seedl_sml", "seedl_lgr", "sapl", "total")) # Adjust order
  ) %>%
    arrange(cluster, trmt, height_class)


comm01 <- comm01 %>%
  mutate(fill_color = case_when(
    trmt == "control" ~ control_colors[height_class],
    trmt == "protected" ~ protected_colors[height_class]
  ))




figure_c1 <- ggplot(comm01, aes(x = cluster, y = tpa, fill = fill_color, pattern_fill = height_class)) +
  geom_bar_pattern(stat = "identity", position = "dodge", 
                   pattern = "stripe", pattern_density = 0.1, pattern_spacing = 0.05) +
  scale_fill_identity() +  # Uses preassigned colors
  scale_pattern_fill_manual(values = c("seedl_sml" = "stripe", "seedl_lgr" = "crosshatch", "sapl" = "dots")) +
  labs(
    x = "Cluster",
    y = "Commercial Trees Per Acre (TPA)",
    title = "Commercial Species 12-60in, 60in to 1in, >1in (TPA) by Cluster and Treatment",
    subtitle = "Growing Season 2023, slash_wall_analyses_2025.Rmd. Figure_C1",
    fill = "Treatment",
    pattern_fill = "Height Class"
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

#
#these patterns didn't quite work
#
# figure_c1 <- ggplot(comm01, aes(x = cluster, y = tpa, fill = trmt, pattern = height_class)) +
#   geom_bar_pattern(stat = "identity", position = "dodge",
#                    aes(pattern_fill = height_class),  # Ensure pattern is mapped to height_class
#                    pattern_density = 0.5, pattern_spacing = 0.02) +  
#   scale_fill_manual(values = c("control" = "blue", "protected" = "green")) +  # Treatment = Color
#   scale_pattern_manual(values = c("seedl_sml" = "stripe", 
#                                   "seedl_lgr" = "crosshatch", 
#                                   "sapl" = "circle")) +   # Ensure sapl has a pattern
# #                                  "total" = "none")) +  
#   labs(
#     x = "Cluster",
#     y = "Commercial Trees Per Acre (TPA)",
#     title = "Commercial Species 12-60in, 60in to 1in, total (TPA) by Cluster and Treatment",
#     subtitle = "Growing Season 2023, slash_wall_analyses_2025.Rmd. Figure_C1",
#     fill = "Treatment",   
#     pattern = "Height Class"  
#   ) +
#   theme_classic() +
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


  ggsave("figures/comm_tpa_height-class_cluster_trmt_Figure_C1.jpg", plot = figure_c1, width = 8, height = 6, dpi = 300)


```



#  SPECIES BY FOREST TYPE, top 10
# Merge the  data for species x forest type x trmt, assemble into an output file that can be used to make tables.
# 
```{r}

## Step 1: Select top 5 species within each group
## there are no calculations here, just sorting the top five spp for each group_by
top_spp <- FT_calc2 %>%
  filter(!brush == "unbrush") %>% 
  group_by(growingseasons, forest_type, trmt, brush) %>%
  arrange(desc(GTR5ft), .by_group = TRUE) %>%
  slice_head(n = 10) %>%  #this was originally n= 5, now n=-10 for publication
  ungroup()

# Step 2: Calculate remainder for all other species
remainder <- FT_calc2 %>%
  filter(!brush == "unbrush") %>% 
  group_by(growingseasons, forest_type, trmt, brush) %>%
  filter(!spp %in% top_spp$spp) %>%  # Exclude top 10 (previously 5) spp
  summarize(
    spp = "Remainder",  # Label for remaining species
    across(where(is.numeric), sum, na.rm = TRUE),  # Sum numeric variables
    #(average?)  stems.ramet = mean(stems.ramet)
    .groups = "drop"
  )

# Step 3: Calculate total (sum of top 5 + remainder)
total_row <- bind_rows(top_spp, remainder) %>%
  group_by(growingseasons, forest_type, trmt, brush) %>%
  summarize(
    spp = "Total",  # Label for total
    across(where(is.numeric), sum, na.rm = TRUE),
    .groups = "drop"
  )

# Step 4: Combine top 5 species, remainder,  total and add a blank row.
# use a dummy variable to force the sort given there is no existing variable that
# would sort ascending or descending.

# Combine and sort Remainder and TOTAL first
remain_total <- bind_rows(remainder, total_row) %>%
  arrange(growingseasons, forest_type, trmt, brush, spp) %>%   # Sort by spp to ensure Remainder comes before TOTAL
  mutate(spp = case_when(
        spp %in% c("Total") ~ "TOTAL", # see step 3 above where spp defined as "Total" but could change to "TOTAL"
        TRUE ~ spp,  # Keeps the existing value of "spp" for all other cases
    )) %>% 
  mutate(dummy = case_when(
        spp == "Remainder" ~ 11, #this was 6, 5 + 1, now 11 = 10+1
        spp == "TOTAL" ~ 12, # this was 7, one greater than the last line of code
        TRUE ~ NA_real_  # Assign NA for other cases
  ))

# Assign dummy values to top_spp
top_spp <- top_spp %>%
  group_by(growingseasons, forest_type, trmt, brush) %>%
  mutate(dummy = row_number()) %>%  # Assign 1 to 5
  ungroup()

blank_rows <- top_spp %>%
  group_by(growingseasons, forest_type, trmt, brush) %>%
  summarize(
      spp = "", 
      total = NA, 
      dummy = 13, # this was 8 
      .groups = "drop"
      )

# Now combine with top_spp
final_result <- bind_rows(top_spp, remain_total, blank_rows) %>%
  arrange(growingseasons, forest_type, trmt, brush, dummy) %>%  # Sort by dummy for correct order

  mutate(
    date = "14oct2025",
    across(
      !dummy,
      ~ if_else(dummy == 13, NA, .x)  # Set to NA when dummy equals 13 (previously 8)
    )) %>%
    
    mutate(
        across(where(is.numeric), ~ round(.x, 1))
    ) %>% 
    
  select(filename, df, date, stand, forest_type, trmt, brush, growingseasons, spp, type, 
         n_forest_type, all_sizes, GTR1ft, GTR5ft, upperexposed, everything())





#see code below to write all dataframe output to one excel file
# output_file <- file.path(output_folder, paste0("top_spp_by_forest_type_and_growingseason(16sept2025)", ".xlsx"))
# write_xlsx(final_result, output_file)




```

# summary by "type" for type x forest type x trmt
```{r}

# Merge the  data for type x forest type x trmt, assemble into an output file that can be used to make tables.
# 


## Step 1: Select top 5 species within each group
## there are no calculations here, just pulling in the type summary 
type_summ <- FT_calc3 %>%
  group_by(growingseasons, forest_type, trmt, brush, type) %>%
  arrange(growingseasons, forest_type, trmt, brush, type, .by_group = TRUE) %>%
  slice_head(n = 5) %>%
  ungroup()

# Step 2: Calculate remainder for all other species
# this step isn't needed
remainder_type <- FT_calc2 %>%
  group_by(growingseasons, forest_type, trmt, brush) %>%
  filter(!type %in% type_summ$type) %>%  # Exclude top 5 spp
  summarize(
    type = "Remainder",  # Label for remaining species
    across(where(is.numeric), sum, na.rm = TRUE),  # Sum numeric variables
    #(average?)  stems.ramet = mean(stems.ramet)
    .groups = "drop"
  )

# Step 3: Calculate total (sum of top 5 + remainder)
total_row_type <- bind_rows(top_spp, remainder) %>%
  group_by(growingseasons, forest_type, trmt, brush) %>%
  summarize(
    type = "TOTAL",  # Label for total
    across(where(is.numeric), sum, na.rm = TRUE),
    .groups = "drop"
  )

# Step 4: Combine top 5 species, remainder,  total and add a blank row.
# use a dummy variable to force the sort given there is no existing variable that
# would sort ascending or descending.

# Combine and sort Remainder and TOTAL first
remain_total_type <- bind_rows(remainder_type, total_row_type) %>%
  arrange(growingseasons, forest_type, trmt, brush, type) %>%   # Sort by spp to ensure Remainder comes before TOTAL
  
    # mutate(type = case_when(
    #     type %in% c("Total") ~ "TOTAL", # see step 3 above where spp defined as "Total" but could change to "TOTAL"
    #     TRUE ~ type,  # Keeps the existing value of "spp" for all other cases
    # )) %>% 
  mutate(dummy = case_when(
        type == "Remainder" ~ 6,
        type == "TOTAL" ~ 7,
        TRUE ~ NA_real_  # Assign NA for other cases
  ))

# Assign dummy values to top_spp
type_summ <- type_summ %>%
  group_by(growingseasons, forest_type, trmt, brush, type) %>%
  mutate(dummy = row_number()) %>%  # Assign 1 to 5
  ungroup()

blank_rows_type <- type_summ %>%
  group_by(growingseasons, forest_type, trmt, brush) %>%
  summarize(
      type = "", 
      TOTAL = NA, 
      dummy = 8, 
      .groups = "drop"
      )

# Now combine with top_spp
final_result_type <- bind_rows(type_summ, remain_total_type, blank_rows_type) %>%
  arrange(growingseasons, forest_type, trmt, brush, dummy) %>%  # Sort by dummy for correct order

  mutate(
    date = "9oct2025",
    across(
      !dummy,
      ~ if_else(dummy == 8, NA, .x)  # Set to NA when dummy equals 8
    )) %>%
    
    mutate(
        across(where(is.numeric), ~ round(.x, 1))
    ) %>% 
    
#  select(-x1, -x2, -denom) %>%  
  select(filename, df, date, stand, forest_type, trmt, brush, growingseasons, type, 
         n_forest_type, upperexposed, GTR5ft, GTR1ft, all_sizes, everything())



```

# TYPE BY TRMT
#summary by type x trmt x growing season
```{r}
# Merge the  data for type x forest type x trmt, assemble into an output file that can be used to make tables.
# 


## Step 1: Select top 5 species within each group
## there are no calculations here, just pulling in the type summary 
type_summ <- FT_calc4 %>%
  group_by(growingseasons, trmt, type) %>% #group_by(growingseasons, forest_type, trmt, brush, type) %>%
  arrange(growingseasons,  trmt, type, .by_group = TRUE) %>%
  slice_head(n = 5) %>%
  ungroup()

# Step 2: Calculate remainder for all other species
# this step isn't needed
remainder_type <- FT_calc4 %>%
  group_by(growingseasons, trmt, type) %>%   #group_by(growingseasons, forest_type, trmt, brush) %>%
  filter(!type %in% type_summ$type) %>%  # Exclude top 5 spp
  summarize(
    type = "Remainder",  # Label for remaining species
    across(where(is.numeric), sum, na.rm = TRUE),  # Sum numeric variables
    #(average?)  stems.ramet = mean(stems.ramet)
    .groups = "drop"
  )

# Step 3: Calculate total (sum of top 5 + remainder)
total_row_type <- bind_rows(type_summ, remainder_type) %>%
  group_by(growingseasons, trmt) %>%   #group_by(growingseasons, forest_type, trmt, brush) %>%
  summarize(
    type = "TOTAL",  # Label for total
    across(where(is.numeric), sum, na.rm = TRUE),
    .groups = "drop"
  )

# Step 4: Combine top 5 species, remainder,  total and add a blank row.
# use a dummy variable to force the sort given there is no existing variable that
# would sort ascending or descending.

# Combine and sort Remainder and TOTAL first
remain_total_type <- bind_rows(remainder_type, total_row_type) %>%
  arrange(growingseasons, trmt, type) %>%   # Sort by spp to ensure Remainder comes before TOTAL
  
    # mutate(type = case_when(
    #     type %in% c("Total") ~ "TOTAL", # see step 3 above where spp defined as "Total" but could change to "TOTAL"
    #     TRUE ~ type,  # Keeps the existing value of "spp" for all other cases
    # )) %>% 
  mutate(dummy = case_when(
        type == "Remainder" ~ 6,
        type == "TOTAL" ~ 7,
        TRUE ~ NA_real_  # Assign NA for other cases
  ))

# Assign dummy values to top_spp
type_summ <- type_summ %>%
  group_by(growingseasons,  trmt, type) %>%
  mutate(dummy = row_number()) %>%  # Assign 1 to 5
  ungroup()

blank_rows_type <- type_summ %>%
  group_by(growingseasons,  trmt) %>%
  summarize(
      type = "", 
      TOTAL = NA, 
      dummy = 8, 
      .groups = "drop"
      )

# Now combine with top_spp
final_result_type <- bind_rows(type_summ, remain_total_type, blank_rows_type) %>%
  arrange(growingseasons, trmt  ) %>%  # Sort by dummy for correct order

  mutate(
    date = "14oct2025",
    across(
      !dummy,
      ~ if_else(dummy == 8, NA, .x)  # Set to NA when dummy equals 8
    )) %>%
    
    mutate(
        across(where(is.numeric), ~ round(.x, 1))
    ) %>% 
    
#  select(-x1, -x2, -denom) %>%  
  select(filename, df, date, stand, trmt, growingseasons, type, 
         n_trmt_df, all_sizes, GTR1ft, GTR5ft, upperexposed, everything())

```





# graphs of stem counts per acre
```{r}
figure_tpa_B <- FT_calc2 %>%
  mutate(growingseasons = as.numeric(growingseasons)) %>%
  filter(forest_type == "n_hdwd", type == "interfere", growingseasons < 7) %>%
  filter(!(growingseasons > 6 & trmt == "control")) %>% 
  ggplot(aes(x = factor(growingseasons), 
             y = upperexposed, 
             fill = trmt)) +
  geom_col(position = "dodge", width = 0.7) +
  facet_wrap(~brush, scales = "fixed") +
#  facet_wrap(~brush, scales = "free_y") +
  labs(
    title = "Interference Stems Per Acre by Growing Season (N. Hardwood)",
    subtitle = "Stems Exposed and >5ft tall fig=tpa_B slash_wall_analyses2025(by cluster).Rmd",
    x = "Growing Seasons",
    y = "Stems Per Acre",
    fill = "Treatment"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  ) +
  scale_fill_manual(values = c("control" = "blue", "protected" = "green"))

# Display the plot
print(figure_tpa_B)

#note the code above excludes one of the state forests
#   either danby or goundry with a 7 growing seasons and apparently high commercial 
#   stem density. Verify those calculations and consider if the comparison is valid.
#   the NH brushed suggests there are no commercial stems outside the wall at year 7.
#   This also doesn't seem correct.

# Save the plot
ggsave("figures/forest_type_stem_density_interfere_NH_Figure_tpa_B.jpg", 
       plot = figure_tpa_B, width = 8, height = 6, dpi = 300)

```



#write all dataframes to a single excel file
```{r}
# The Old Way to Write to Worksheets
# dfs2xlsx(withNames("full_spp_list", FT_calc2,
#                    "summary of richness", richness_summ,
#                    "percent point occupancy", forest_type_occupancy,
#                    "top species", final_result
#                         ), "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\summary cafri 6_6 23sept2025")

#the new way.  Need to run library for openxlsx
wb <- createWorkbook()

# add each worksheet + its data
addWorksheet(wb, "top species")
writeData(wb, "top species", final_result)

addWorksheet(wb, "type summ")
writeData(wb, "type summ", final_result_type)

# addWorksheet(wb, "summary of richness")
# writeData(wb, "summary of richness", richness_summ)

addWorksheet(wb, "list of species")
writeData(wb, "list of species", spp_list)

addWorksheet(wb, "percent point occupancy")
writeData(wb, "percent point occupancy", forest_type_occupancy)

addWorksheet(wb, "full_spp_list")
writeData(wb, "full_spp_list", FT_calc2)

addWorksheet(wb, "type_list")
writeData(wb, "type_list", FT_calc3)


# save workbook (note .xlsx at the end!)
saveWorkbook(
  wb,
  "C:/Users/pjs23/Documents/R/slash-wall-vegetation/summary output/summary all_but_cafri  16oct2025.xlsx",
  overwrite = TRUE)
                   

```
