---
title: "slash veg complete"
author: "Peter Smalldige"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- README

File Description: The purpose of this file is to:
1. read the raw vegetation data and make it complete
2. convert the field measurements to per/acre values
3. make data complete

exported data files include:
xxx (line 237)
xxx (line 353)
xxx (line 530)
xxx

 -->


###Initialize Libraries

```{r message=FALSE}

#library(ggplot2)
#library(ggpubr)
library(expss)
library(readr)
library(readxl)
library(writexl)
library(skimr)
library(tidyverse)  # includes: tidyr, dplyr, haven, readr, readxl, forcats, stringer, ggplot2. See all with tidyverse_packages()
library(knitr)
library(janitor)
#install.packages("openxlsx")
library(openxlsx)
# install.packages("shinyaddins")
# library(shinyaddins)

```

#
# ALT plus hypen = <- (within chunk)
# ctrl plus shift plus m = %>% 
# ctrl + ALT + I = insert chunk
# ALT + - = insert assignment operator
# filter rows, select columns
# RMD comment = <!-- comment -->
# 
# 

<!--
-->

<!--     -->

### code 01
### Create file path input and the output folder (code 01) 

file paths, input and output
```{r}
#file_path_input01 <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\regen_per_acre_all_alpha_tidy_trmt_coded 2019-2022 (with read me).xlsx"
file_path_2023_veg <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\complete_full_2023(readme).xlsx"

file_path_2023_veg_22july2025 <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\raw_data\\veg2023_raw(22july2025).xlsx"

file_path_2019 <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\regen.data.2019.xlsx"
    # line 93
file_path_2020 <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\regen.data.2020_cleaned.xlsx"
    # line 143
file_path_2021 <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\regen.data.2021_cleaned.xlsx"
    # line 179
file_path_2022 <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\slash.wall.data.2022.15february2025.xlsx"
  # this includes cafri and non-cafri data
 
file_path_2024 <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\raw_data\\slash.wall.data_2024_FINAL_1june2025.xlsx"
    # the FINAL 2024 growing season vegetation state forest data added to main folder on 1june2025
    # This file path  includes worksheets for "regen" and "plot" (line 734)
    # 
    # NOTE - for the state forest stands (danby_block_2, fairfied_a5, fairfield_a6, goundry_a17, and decker
    # set location = "control")
     

file_path_points <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\point_labels_all(readme).xlsx"


#plot data for the 4" to 12" seedling data
file_path_2019_plot <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\plot-data-2019.csv"
   
file_path_2020_plot <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\plot_calc_2020_untidy.csv"
    
file_path_2021_plot <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\slash.wall.data.2021.30october2022.xlsx"
    
file_path_2022_plot <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\plot_data_2022_untidy.csv"
#  does not include any CAFRI data

file_path_2023_plot <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\slash.wall.data.2023.23july2024(maybe FINAL).xlsx"

file_path_2023_cafri_overstory <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\overstory dbh data for 6-9 and 6-6 CAFRI_2023_oct_30.xlsx"
# file includes the original dbh data with distances on marginal trees. Retain distance for future assessment
# of when a tree is considered "in"


# 2024 plot data is within "file_path_2024"

output_folder <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\"


```
In 2019 - 2022, seedling regen data were recorded in the following strata:
Site/plot sheet: counts of seedlings 4" to 12", sometimes as categorical
Regen Tally sheet: 
    12" to 4.5ft (B) = formerly seed01
    4.5ft - 9 ft (C1) = formerly seed02
    9ft to 1"(C2)   [C1 + C2 = D], formerly seed03
    1 - 4", formerly sapl

in 2023 and beyond, seedling regen data were recorded in the following strata:
Regen tally sheet: 
    (A = 01) 4" to 12"
    (B = 02) 12" to 60"  (the break point at 60" vs. 54" was to align with SILVAH protocols)
    (D = 03) 60" tall to 1" dbh
    (P = 04) 1" to 4" dbh [in 2024, the upper boundary was 5"...bad idea]

To standardize labels:
A = 4" to 12"           radius = 3.7 ft; count in 2019, 2023, and 2024 otherwise categorical 0 = 0; 1 = 1to5; 2 = 6 to 15; 3 = 16 to 30; 4 = 31+
B = 12" to 4.5/5 ft     radius = 6 ft
C1 = 4.5 ft to 9 ft     radius = 6 ft
C2 = 9ft to 1" dbh      radius = 6 ft
D = 4.5/5 ft to 1"      radius = 6 ft
P = 1" to 4" (saplings) radius thru 2022 = 6 ft; then 16 ft


Radii
4" - 12" = 3.7 ft
12" to 4.5/5ft = 6 ft


plot data for seedlings 4" to 12" (species recorded as 01 in 2023)
folder = R\\slash-wall-vegetation\\old files but do not discard
2019 file = plot-data-2019.csv
2020 file = plot_calc_2020_untidy.csv
2021 file = plot_data_2021_untidy.csv
2022 file = plot_data_2022_untidy.csv
2023 files =
        slash.wall.data.2023.23july2024(maybe FINAL)
        slash.wall.data.2023_1_feb_2024_regen(north, south, rectangle, gasline)_final
2024 files sheet = regen, filename = "slash.wall.data_2024_FINAL"



input point attributes, "point_label_all" this is the new data file of attributes by point

cafri overstory data and cafri point labels
```{r}
# data of the target residual basal area from the 2022 harvest in CAFRI 6-9 and 6-6
cafri_point_residual_overstory <- read_excel("CAFRI Treatment & Plot Location 8-2024.xlsx", sheet = "merged", skip = 6) %>% 
    select(stand, wall, overstory, point) %>% 
    mutate(
        wall = as.character(wall),
        point = as.character(point)
    ) %>% 
    rename(
        harvest = stand,
        residual = overstory
    )

# df includes the trmt, residual basal area target, and wall height (brush trmt not included)
cafri_point_labels <- read_excel("CAFRI Treatment & Plot Location 8-2024.xlsx", sheet = "point counts", skip = 4) %>% 
    rename(
        harvest = stand,
        trmt = status
        ) %>% 
    mutate(
        wall = as.character(wall),
        point = as.character(point),
        location = trmt
    ) %>% 
    mutate(
    location = case_when(
      location == "control" ~ "c",
      location == "protected" ~ "p",
      TRUE ~ location  # Keeps other values unchanged
    )
  ) %>% 
    left_join(cafri_point_residual_overstory, by = c("harvest", "wall", "point" ))

# the point_labels are updated to include state forest sites on about line 1712, df=new_sites
# the excel file of point_labels was updated for sheet = "all_sites". that is imported here and
# subsequent rewriting of the excel file is halted.


#includes height, target BA, brushing, forest type, trmt, point number, wall number
point_labels_revision <- read_excel(file_path_points, sheet = "all_sites", skip = 0) # %>% 
    # arrange(harvest, wall, trmt, point) %>% 
    # left_join(cafri_point_labels, by = c("harvest", "wall", "point", "trmt", "height", "location"))
```



2019 vegetation data
```{r eval=T}

raw2019 <- read_excel(file_path_2019, sheet = "seedling", skip = 0 ) |>
    select(!c(top4, other, radius.01, radius.other)) |> 
    mutate(
        season = "2019",
        sup.seed03 = as.numeric(sup.seed03),
        sup.sapl = as.numeric(sup.sapl),
        point = as.character(point)
    ) |>
    mutate(origin = case_when(
       origin %in% c("s","u") ~ "single",
       origin %in% c("t", "r") ~ "clump",
       TRUE ~ origin  # Keeps other values unchanged
  )) |> 
    select(harvest, season, point, location, spp, origin, everything()) %>% 
    mutate(
    harvest = case_when(
      harvest == "cgl" ~ "gasline",
      harvest == "crp" ~ "redpine",
      harvest == "cwb" ~ "stationrd",
      harvest == "gl" ~ "gasline",
      harvest == "rp" ~ "redpine",
      TRUE ~ harvest  # Keeps other values unchanged
    )
  ) %>% 
    mutate(
    location = case_when(
      location == "control" ~ "c",
      location == "interior" ~ "i",
      location == "perimeter" ~ "p",
      TRUE ~ harvest  # Keeps other values unchanged
    )
  )

  
# clumps were not separated on the tally sheet. 
# It isn't possible to  the number of clumps per acre
# It would be possible to represent the % of total stems present in clumps 
# (e.g.) 40% of stem density is of clump origin
# 

regen2019 <- raw2019 %>% 
    mutate(across(where(is.numeric), ~replace_na(., 0))) %>% 
    mutate(across(starts_with("su"), ~replace_na(., 0))) %>%    # Replace NA with 0 for "su" variables, the "supressed" variables
    mutate(across(starts_with("ex"), ~replace_na(., 0))) %>%       # Replace NA with 0 for "ex" variables, the "exposed" variables
    mutate(harvest = case_when(
        harvest %in% c("boot", "wedge", "stationrd") ~ "sta_rd",
        harvest %in% c("gasline") ~ "gas_line",
        harvest %in% c("redpine") ~ "red_pine",
        TRUE ~ harvest, # keeps the existing value of "harvest" if none of the cases are true
    )) %>% 
        left_join(point_labels_revision, by = c("harvest", "point", "location")) %>% 
    select(-wall, -trmt, -height, -harvest_original, -forest_type, -brush) %>% 
      mutate(across(
        where(is.numeric),  # Select all numeric columns
        ~ if_else(is.na(spp), 0, .x)  # Set to 0 when spp is NA, otherwise keep original values
  ))


 
```
 

2020 data
```{r}
raw2020 <- read_excel(file_path_2020, sheet = "regen", skip = 0 ) |>
    select(! c(top4, Comments)) |> 
    mutate(
        season = as.character(season),
        spp = as.character(spp),
        sup.seed03 = as.numeric(sup.seed03),
        sup.sapl = as.numeric(sup.sapl)
    ) |>
    mutate(origin = case_when(
       origin %in% c("s","u") ~ "single",
       origin %in% c("t", "r") ~ "clump",
       TRUE ~ origin  # Keeps other values unchanged
  )) |> 
    select(harvest, season, point, location, spp, origin, everything())

  
# clumps were not separated on the tally sheet. 
# It isn't possible to  the number of clumps per acre
# It would be possible to represent the % of total stems present in clumps 
# (e.g.) 40% of stem density is of clump origin
# 

regen2020 <- raw2020 %>% 
    mutate(point = as.character(point)) %>% 
    mutate(across(where(is.numeric), ~replace_na(., 0))) %>% 
    mutate(across(starts_with("su"), ~replace_na(., 0))) %>%    # Replace NA with 0 for "su" variables
    mutate(across(starts_with("ex"), ~replace_na(., 0)))      # Replace NA with 0 for "ex" variables


```

2021 data
```{r}
raw2021 <- read_excel(file_path_2021, sheet = "regen", skip = 2 ) |>
    select(! c(top4)) |> 
    mutate(
        season = as.character(season),
        location = as.character(location),
        spp = as.character(spp)
   ) |>
    mutate(origin = case_when(
       origin %in% c("s","u") ~ "single",
       origin %in% c("v", "r") ~ "clump",
       TRUE ~ origin  # Keeps other values unchanged
  )) |> 
    mutate(
        sup.seed03 = as.numeric(sup.seed03),
        sup.sapl = as.numeric(sup.sapl)
    ) %>% 
    select(harvest, season, point, location, spp, origin, everything()) 
  
  
# clumps were not separated on the tally sheet. 
# It isn't possible to  the number of clumps per acre
# It would be possible to represent the % of total stems present in clumps 
# (e.g.) 40% of stem density is of clump origin
# 

regen2021 <- raw2021 %>% 
    mutate(
        location = as.character(location),
        point = as.character(point)
    ) %>% 
     mutate(across(where(is.numeric), ~replace_na(., 0))) %>% 
   mutate(across(starts_with("su"), ~replace_na(., 0))) %>%    # Replace NA with 0 for "su" variables
    mutate(across(starts_with("ex"), ~replace_na(., 0)))      # Replace NA with 0 for "ex" variables


```




2022 data
```{r message=FALSE, warning=FALSE}
raw2022 <- read_excel(file_path_2022, sheet = "regen", skip = 0 ) |>
    select(! c(top4, comments)) |> 
    select(-wall) %>% 
    mutate(
        season = as.character(season),
        location = as.character(location),
        spp = as.character(spp),
        point = as.character(point)
    ) |>
    mutate(origin = case_when(
       origin %in% c("s","u") ~ "single",
       origin %in% c("v", "r") ~ "clump",
       TRUE ~ origin  # Keeps other values unchanged
  )) |> 
    mutate(        
        sup.seed02 = as.numeric(sup.seed03),
        sup.seed03 = as.numeric(sup.seed03),
        sup.sapl = as.numeric(sup.sapl)
    ) %>% 
    select(harvest, season, point, location, spp, origin, everything()) 
  
# clumps were not separated on the tally sheet. 
# It isn't possible to  the number of clumps per acre
# It would be possible to represent the % of total stems present in clumps 
# (e.g.) 40% of stem density is of clump origin
# 

regen2022 <- raw2022 %>% 
    mutate(
        location = as.character(location)
    ) %>% 
    mutate(across(where(is.numeric), ~replace_na(., 0))) %>% 
    mutate(across(starts_with("su"), ~replace_na(., 0))) %>%    # Replace NA with 0 for "su" variables
    mutate(across(starts_with("ex"), ~replace_na(., 0)))      # Replace NA with 0 for "ex" variables


```


veg2023_peracre_calc <- data01 %>% 
    mutate(
        multip01 = 43560 / (pi * (3.7^2)),  # = 1012.79
        multip02_03 = 43560 / (pi * (6^2)), # = 43560/113.1 = 385.15
        multip_sapl  = 43560 / (pi * (16^2)) # = 43560 / 256 = 170.1
    ) |> 


2023 veg data
```{r warning=FALSE}
# input data are for counts of stems per point, not per acre
# 
#regen2023 <- read_excel(file_path_2023_veg, sheet = "Sheet1", skip = 0 ) %>% 
regen2023 <- read_excel(file_path_2023_veg_22july2025, sheet = "Sheet1", skip = 0) %>% 
    rename(
        supA = sup.seed01, #confirmed 7/24/2025 01 = 4" to 12"
        supB = sup.seed02, # 12" to 5 ft
        supD = sup.seed03, # 5 ft to 1"
        supP = sup.sapl, #1" to 5"
        expA = exp.seed01,
        expB = exp.seed02,
        expD = exp.seed03,
        expP = exp.sapl
        ) %>% 
    
#TODO need to double check this calculation #DONE, calculation confirmed 2/13/2025    
    mutate(
    numb.per.clump = if_else(origin == "clump",  
                     rowSums(across(starts_with("sup") | starts_with("exp"), 
                            ~ if_else(. > 0, ., NA_real_)), na.rm = TRUE) * count, #within a row, assess if >0, then calculates mean of clump size 
                          NA_real_)
   ) %>% 
    
    #Next, convert stem counts to "per acre" values
    mutate(
        multip01 = 43560 / (pi * (3.7^2)),  # = 1012.79
        multip02_03 = 43560 / (pi * (6^2)), # = 43560/113.1 = 385.15
        multip_sapl  = 43560 / (pi * (16^2)), # = 43560 / 256 = 170.1
        wall = as.character(wall)
    ) %>% 
    
    mutate(
        supA = supA * count * multip01,
        supB = supB * count * multip02_03,
        supD = supD * count * multip02_03,
        supP = supP * count * multip_sapl,
        expA = expA * count * multip01,
        expB = expB * count * multip02_03,
        expD = expD * count * multip02_03,
        expP = expP * count * multip_sapl
    ) %>% 
    mutate(
        type = "blank"
    ) %>% 
    mutate(type = case_when(
    spp %in% c("ba", "bc", "bih", "bl", "co", "cu", "he", "ohw", "rm", "ro", "sa", 
                "sh", "sm", "wa", "wo", "wp", "yb", "yp",
               "951", "762", "402", "901", "832", "651", "261", "004", "4", "316", "833", "931", #004 same as 4
                "407", "318", "541", "802", "129", "371", "621") ~ "commercial",
    
    spp %in% c("sb", "372") ~ "sb",

    spp %in% c("ab", "eh", "pc", "stm", "toh", "au",
               "531", "701", "761", "315", "998", "535") ~ "interfere",

    spp %in% c("app", "bta", "haw", "brs", "dws", "ec", "la", "ns",
               "pb", "gb", "qa", "rp", "scp", "smc", "svb", "wh", "wi",
               "660", "743", "500", "brs", "dws", "742", "70", "71", "91",  # 70 same as 71
               "375", "379", "746", "125", "130", "866", "355", "356", "585", "920") ~ "diversity", # 355 same as 356

    spp %in% c("mia", "ah", "cc", "unk", "sp",
                "001", "1", "391", "763", "999", "90") ~ "other",  # 001 same as 1
    TRUE ~ type,  # Keeps the existing value of "type" for all other cases
    )) %>%  
    select(-df, -filename) %>% 
    select(harvest, wall, season, date, trmt, location, point, spp, type, origin, count, everything())



regen2023seedl <- regen2023 %>%  # extract the data for the 4" to 12" seedlings to merge into the site data permanent file
# data are counts of stems per acre by point
    filter(type == "commercial") %>% 
    mutate(
        comm_sdl = supA + expA
    ) %>% 
    select(harvest, wall, point, season, type, spp, comm_sdl) %>% 
    group_by(harvest, point) %>% 
    summarize(
       season = first(season),
       wall = first(wall),
       comm_sdl = sum(comm_sdl),
       seedA = sum(comm_sdl),
       richness = n_distinct(spp),
       .groups = "drop"
    )

# (july 24 2025) in the field data read into df=regen2023, the 4" to 12" class included all species
# in df = regen2023seedl, the data are sorted first for commercial species and
# seedA is the representation of comm_sdl on a per acre basis. The same as for regen 2019 throu regen2022

```


2024 vegetation
```{r warning=FALSE}

# input data values in the input file are count of stems in each plot (not per acre)

regen2024 <- read_excel(
  path = file_path_2024,
  sheet = "regen",
  skip = 2,
  col_types = c("text", "text", "text", "text", "text","text",
                "numeric", "numeric", "numeric", "numeric","numeric", "numeric", "numeric", "numeric","numeric",
                "text")           # etc.
) %>% 
    select(-comments) %>% 
    filter(!season < 2024) %>% 
    rename(
        harvest = stand
    ) %>% 
    mutate(
        count = as.numeric(gsub("[^0-9]", "", count))
    ) %>%  #removes any non-numeric characters and changes count from "char" to "num"
      mutate(point = str_replace(point, "^0+", "")) %>%  #from the start (^), match one or more zeros (0+) and replace with ""

    mutate(
        spp = as.character(spp),
        season = as.character(season),
        point = as.character(point),
        sup.sapl = as.numeric(sup.sapl)
    ) %>% 

    replace_na(list(date = "other", count = 1,
        sup.seed01 = 0, sup.seed02 = 0, sup.seed03 = 0, sup.sapl = 0,
        exp.seed01 = 0, exp.seed02 = 0, exp.seed03 = 0, exp.sapl = 0
    ))  %>% 

    rename(
        supA = sup.seed01, # 4 to 12", radius = 3.7 ft
        supB = sup.seed02, # 12 to 60", radius = 6 ft
        supD = sup.seed03, # 60" to 1"dbh, radius = 6 ft  #elsewhere, label "C" is for stems 60" to 9 ft
        supP = sup.sapl, # 1" to 5" dbh, radius = 16 ft
        expA = exp.seed01,
        expB = exp.seed02,
        expD = exp.seed03,
        expP = exp.sapl
    ) %>%   #  need to check this code, added on 4/16/2025 but not tested

    mutate(
      origin = case_when(
      count < 2 ~ "single",
      count >= 2 ~ "clump"
      )) %>% 
    
#number.per.clump calculation confirmed 2/13/2025    
    mutate(
    numb.per.clump = if_else(origin == "clump",  
                          rowSums(across(starts_with("sup") | starts_with("exp"), 
                                          ~ if_else(. > 0, ., NA_real_)), na.rm = TRUE) * count, #within a row, assess if >0, then calculates mean of clump size 
                          NA_real_)
   ) %>%
    
#Next, convert stem counts to "per acre" values
    mutate(
        multip01 = 43560 / (pi * (3.7^2)),  # = 1012.79
        multip02_03 = 43560 / (pi * (6^2)), # = 43560/113.1 = 385.15
        multip_sapl  = 43560 / (pi * (16^2)), # = 43560 / 256 = 170.1
        wall = as.character(wall)
    ) %>% 
    
    mutate(
        supA = supA * count * multip01,
        supB = supB * count * multip02_03,
        supD = supD * count * multip02_03,
        supP = supP * count * multip_sapl,
        expA = expA * count * multip01,
        expB = expB * count * multip02_03,
        expD = expD * count * multip02_03,
        expP = expP * count * multip_sapl
    ) %>% 
    mutate(
        type = "blank"
    ) %>% 
    mutate(type = case_when(
    spp %in% c("ba", "bc", "bih", "bl", "co", "cu", "he", "ohw", "rm", "ro", "sa", 
                "sh", "sm", "wa", "wo", "wp", "yb", "yp",
               "951", "762", "402", "901", "832", "651", "261", "004", "4", "316", "833", "931", #004 same as 4
                "407", "318", "541", "802", "129", "371", "621") ~ "commercial",
    
    spp %in% c("sb", "372") ~ "sb",

    spp %in% c("ab", "eh", "pc", "stm", "toh", "au",
               "531", "701", "761", "315", "998", "535") ~ "interfere",

    spp %in% c("app", "bta", "haw", "brs", "dws", "ec", "la", "ns",
               "pb", "gb", "qa", "rp", "scp", "smc", "svb", "wh", "wi",
               "660", "743", "500", "brs", "dws", "742", "70", "71", "91",  # 70 same as 71
               "375", "379", "746", "125", "130", "866", "355", "356", "585", "920") ~ "diversity", # 355 same as 356

    spp %in% c("mia", "ah", "cc", "unk", "sp",
                "001", "1", "391", "763", "999", "90") ~ "other",  # 001 same as 1 = empty plot = "mia"
    TRUE ~ type,  # Keeps the existing value of "type" for all other cases
    )) %>%  

      mutate(harvest = case_when(
      harvest == "boot" ~ "sta_rd",
      harvest == "decker control" ~ "decker",
      harvest == "recknagel south" ~ "recknagel",
      harvest == "wedge" ~ "sta_rd",
      harvest == "wedge control" ~ "sta_rd",
      harvest == "stand_5.1" ~ "5_1",
      TRUE ~ harvest  # Keeps other values unchanged
    )
  )  

# create this df to merge its seedling data (i.e., stems 4"to 12") with the plot attribute data. See line 785
# or search for the df name = site_data
regen2024seedl <- regen2024 %>% 
    filter(type == "commercial") %>%  #retains only commercial species to calculate "seedA"
    mutate(
        comm_sdl = supA + expA
    ) %>% 
    select(harvest, wall, point, season, type, spp, comm_sdl) %>% 
    group_by(harvest, point) %>% 
    summarize(
       season = first(season),
       wall = first(wall),
       comm_sdl = sum(comm_sdl),
       seedA = sum(comm_sdl),
       richness = n_distinct(spp),
       .groups = "drop"
    ) %>% 
mutate(
        wall = case_when(
        harvest %in% c("danby_block_2", "fairfield_a5", "fairfield_a6", "goundry_a17", "5_1") ~ "99",
        TRUE ~ wall
  )
)
```


###  site/plot data read and manipulate

plot 2019
```{r}
plot2019 <- read.csv(file_path_2019_plot) %>% 
   mutate(
    location = case_when(
      location == "control" ~ "c",
      location == "interior" ~ "i",
      location == "perimeter" ~ "p",
      TRUE ~ location  # Keeps other values unchanged
    )
  ) %>% 
    mutate(
        comm_sdl = as.numeric(comm_sdl),
        seedA = comm_sdl,
        seedA = if_else(is.na(seedA), 0, seedA ),
        rhizom = fern,
        season = as.character(season),
        upper.spp = as.character(upper.spp),
        maple = as.integer(maple),
        point = as.character(point)
    ) %>% 
    mutate(across(c(comm_sdl), ~ replace_na(., 0))) %>% 
    rename(
        native_cov = low.cov,
        native_spp = low.spp,
        invasive_spp = invasive.spp,
        invasive_cov = invasive.cov,
        mid_cov = mid.cov,
        mid_spp = mid.spp,
        upper_cov = upper.cov,
        upper_spp = upper.spp,
        vine_cov = vine.cov,
        vine_spp = vine.spp
    ) %>% 
   mutate(seedA = seedA *(43560 / (pi * 3.7 * 3.7))) %>% 
    select(harvest,  season,  date,  location,  point,  disturb,  slash,  fern,  herb,  native_cov,  
           native_spp,  invasive_cov,  invasive_spp,  mid_cov,  mid_spp,  upper_cov,  upper_spp,  vine_cov,  
           vine_spp,  baf,  oakhick,  beech,  maple,  birch,  hdwd,  conifer,  comm_sdl,    seedA )    # ba_acre, cut,  year,  numb_yrs_grow,
 #   filter( point == "338" & season == "2019")
```


plot 2020
```{r}
plot2020 <- read.csv(file_path_2020_plot) %>%  # this has data for boot, wedge and sta_rd separately presented
    mutate(across(c(comm_sdl), ~ replace_na(., 0))) %>% 
    mutate(
        point = str_remove(point, "^0+"),
        point = as.numeric(point)
        ) %>% 
    mutate(
        season = as.character(season),
        rhizom = fern,
        mid_spp = as.character(mid_spp),
        upper_spp = as.character(upper_spp),
        point = as.numeric(point),
        point = as.character(point),
        seedA = case_when(
          comm_sdl == 0 ~ 0,  # If category is 0, the random value is 0
          comm_sdl == 1 ~ sample(1:5, size = n(), replace = TRUE),  # Random number from 1 to 5 for each row
          comm_sdl == 2 ~ sample(6:15, size = n(), replace = TRUE), # Random number from 6 to 15 for each row
          comm_sdl == 3 ~ sample(16:30, size = n(), replace = TRUE), # Random number from 16 to 30 for each row
          comm_sdl == 4 ~ sample(31:50, size = n(), replace = TRUE) # Random number from 31 to 100 for each row
    )
  ) %>% 
   mutate(seedA = seedA *(43560 / (pi * 3.7 * 3.7))) %>% 
        select(harvest,  season,  date,  location,  point,  disturb,  slash,  fern,  herb,  native_cov,  
           native_spp,  invasive_cov,  invasive_spp,  mid_cov,  mid_spp,  upper_cov,  upper_spp,  vine_cov,  
           vine_spp,  baf,  oakhick,  beech,  maple,  birch,  hdwd,  conifer,  comm_sdl,    seedA ) # ba_acre, cut,  year,  numb_yrs_grow,

# for 2020, comm_sdl is a categorical value from 0 to 4.
# seedA is initially a value based on random assignment of comm_sdl within a bracket,
# and subsequently expanded from milacre to acre. 

```


plot 2021
```{r}
plot2021 <- read_excel(file_path_2021_plot, sheet = "site-plot", skip = 6) %>%  
    mutate(across(c(comm_sdl), ~ replace_na(., 0))) %>% 
    mutate(
    season = as.character(season),
    rhizom = fern,
    mid_spp = as.character(mid_spp),
    upper_spp = as.character(upper_spp),
    invasive_spp = as.character(invasive_spp),
    point = as.character(point),
    seedA = case_when(
      comm_sdl == 0 ~ 0,  # If category is 0, the random value is 0
      comm_sdl == 1 ~ sample(1:5, size = n(), replace = TRUE),  # Random number from 1 to 5 for each row
      comm_sdl == 2 ~ sample(6:15, size = n(), replace = TRUE), # Random number from 6 to 15 for each row
      comm_sdl == 3 ~ sample(16:30, size = n(), replace = TRUE), # Random number from 16 to 30 for each row
      comm_sdl == 4 ~ sample(31:50, size = n(), replace = TRUE) # Random number from 31 to 100 for each row
    )
  ) %>% 
   mutate(seedA = seedA *(43560 / (pi * 3.7 * 3.7))) %>% 
        select(harvest,  season,  date,  location,  point,  disturb,  slash,  fern,  herb,  native_cov,  
           native_spp,  invasive_cov,  invasive_spp,  mid_cov,  mid_spp,  upper_cov,  upper_spp,  vine_cov,  
           vine_spp,  baf,  oakhick,  beech,  maple,  birch,  hdwd,  conifer,  comm_sdl,    seedA ) # ba_acre, cut,  year,  numb_yrs_grow,

# for 2021, comm_sdl is a categorical value from 0 to 4.
# seedA is initially a value based on random assignment of comm_sdl within a bracket,
# and subsequently expanded from milacre to acre. 

```


plot 2022
```{r}
plot2022_original <- read_excel(file_path_2022, sheet = "site-plot", skip = 6) %>% 
    mutate(
        point = as.character(point)
    )

plot2022  <- read_excel(file_path_2022, sheet = "site-traits", skip = 1) %>% #read in the cafri plot traits
    mutate(
        point = as.character(point)
    ) %>% 
    bind_rows(plot2022_original) %>% 
    mutate(across(c(comm_sdl), ~ replace_na(., 0))) %>% 
    mutate(
    season = as.character(season),
    rhizom = fern,
    point = as.character(point),
    seedA = case_when(
      comm_sdl == 0 ~ 0,  # If category is 0, the random value is 0
      comm_sdl == 1 ~ sample(1:5, size = n(), replace = TRUE),  # Random number from 1 to 5 for each row
      comm_sdl == 2 ~ sample(6:15, size = n(), replace = TRUE), # Random number from 6 to 15 for each row
      comm_sdl == 3 ~ sample(16:30, size = n(), replace = TRUE), # Random number from 16 to 30 for each row
      comm_sdl == 4 ~ sample(31:50, size = n(), replace = TRUE) # Random number from 31 to 100 for each row
    )
# for 2022, comm_sdl is a categorical value from 0 to 4.
# seedA is initially a value based on random assignment of comm_sdl within a bracket,
# and subsequently expanded from milacre to acre. 

  ) %>% 
   mutate(seedA = seedA * (43560 / (pi * 3.7 * 3.7))) %>% 
   select(harvest,  season,  date,  location,  point,  comm_sdl,    seedA ) # ba_acre, cut,  year,  numb_yrs_grow,
```


2023 plot and site CAFRI and other (includes CAFRI 2023 dbh with distances)
```{r warning=FALSE}

# note 5may2025 - the data from the 2023 overstory sampling including distance to tree and status as
# marginal needs to be amended to the worksheet "2024 overstory data."  Once amended, the overstory data needs to be joined into the df=plot2024
# before it can be joined into df=site_data.
# 
# UPDATE 17july2025 the dbh data for cafri 2023 were adjusted to include distance, a calculation of limiting distance, and status as "in" or "out"

# see also plot_basal_area_calc_summ2025.Rmd
# and output file = plot_basal_area_calc_summ2025.Rmd



plot2023cafri <- read_excel(file_path_2023_plot, sheet = "cafri_site", skip = 3) %>% 
   select(-comments) %>% 
    rename(harvest = stand) %>% 
    mutate(
        wall = as.character(wall),
        season = as.character(season),
        invasive_spp = as.character(invasive_spp),
        mid_spp = as.character(mid_spp),
        point = as.character(point)
    ) %>% 
   full_join(cafri_point_labels, by = c("harvest", "wall", "point"), suffix = c("_cafri", "_labels")) %>% 
    arrange(harvest, point)


cafri2023dbh <- read_excel(file_path_2023_cafri_overstory, sheet = "dbh2023cafri", skip = 5) %>% 
    mutate(
        baf = 9.1,
        limit_distance_2023 = 2.75 * dbh2023,
        status = "in",
        basalarea = baf,
        stem_count = 43560 / (pi * (limit_distance_2023^2)), 
        wall = as.character(wall),
        point = as.character(point)
    ) %>% 
   mutate(
    status = case_when(
      distance > limit_distance_2023 ~ "out", # "distance" is the measured distance in feet from plot center to tree
      TRUE ~ status  # Keeps other values unchanged
    )
  ) %>% 
    arrange(harvest, point, tag) %>% 
    mutate(
        file = "slash_veg_complete.Rmd",
        df = "cafri2023dbh"
    ) %>% 
    select(file, df, harvest, baf, wall, point, spp, tag, status, dbh2023, distance, limit_distance_2023, basalarea, stem_count )

cafri2023_structure_point_summary <- cafri2023dbh %>% 
    filter(
        status == "in"
    ) %>% 
    group_by(harvest, wall, point) %>% 
    summarize(
        ba_acre = sum(basalarea),
        stems_acre = sum(stem_count),
        x_dbh = mean(dbh2023),
        .groups = "drop"
    ) %>% 
    left_join(point_labels_revision, by = c("harvest", "wall", "point")) %>% 
    mutate(
        file = "slash_veg_complete.Rmd",
        df = "cafri2023_structure_point_summary"
    ) %>% 
    select(file, df, harvest, wall, point, ba_acre, stems_acre, x_dbh, trmt, residual, height, brush ) %>% 
    mutate_if(is.numeric, round, 0)

# a list of wall-level attributes
point_labels_unique <- point_labels_revision %>%
  group_by(harvest, wall, brush) %>%
  summarize(
      brush = first(brush),
      trmt = first(trmt),
      height = first(height),
      residual = first(residual),
      .groups = "drop"
      )

cafri2023_structure_wall_summary <- cafri2023_structure_point_summary %>% # use the df = point summary which is filtered for "in" and then average by point
    left_join(point_labels_unique, by = c("harvest", "wall", "trmt", "residual", "height", "brush")) %>% 
    select(harvest, wall, trmt, height, x_dbh, brush, stems_acre, ba_acre, residual) %>% 
    group_by(harvest, wall,brush) %>% 
    summarize(
        ba_acre_2023 = mean(ba_acre),
        stems_acre_2023 = mean(stems_acre),
        x_dbh = mean(x_dbh),
        trmt = first(trmt),
        height = first(height),
        residual = first(residual),
        .groups = "drop"
    ) %>% 
    rename(
        residual_target = residual,
        overstory_TPA = stems_acre_2023
    ) %>% 
    mutate(
        file = "slash_veg_complete.Rmd",
        df = "cafri2023_structure_wall_summary"
    ) %>% 
    select(file, df, harvest, wall, trmt, height, brush, overstory_TPA, ba_acre_2023, x_dbh, residual_target) %>% 
    mutate_if(is.numeric, round, 0)

# I'm not sure why I used "bind rows". that appends the dbh data below the wall_summary data. Maybe for summary printout?
cafri2023_summary <- cafri2023_structure_wall_summary %>% 
    bind_rows(cafri2023dbh) #cafri2023_structure_point_summary, 

output_file <- file.path(output_folder, paste0("cafri2023_dbh_basalarea_raw_summ", ".xlsx"))
write_xlsx(cafri2023_summary, output_file)



plot2023other <- read_excel(file_path_2023_plot, sheet = "other_site_baf10", skip = 2) %>% 
    select(-comments) %>% 
    rename(harvest = stand) %>% 
    mutate(
        spp = as.character(spp),
        wall = as.character(wall),
        season = as.character(season),
        invasive_spp = as.character(invasive_spp),
        mid_spp = as.character(mid_spp),
        point = as.character(point)
    ) %>% 
    mutate(wall = if_else(is.na(wall), "99", wall)    )

plot2023_overstory <- plot2023other %>% 
    select(harvest, wall, location, season, date, point, spp, dbh, in16ft, status,crown, logs)

output_file <- file.path(output_folder, paste0("campridge_overstory_by_point_2023data(29may2025)", ".xlsx"))
write_xlsx(plot2023_overstory, output_file)


plot2023_site <- plot2023other %>% 
    select(harvest, wall, location, season, date, point, rhizom,	clump,	grass,
           herb,	herb_spp,	native_cov,	native_spp,	mid_cov,	mid_spp,	invasive_cov,	invasive_spp,	vine_cov,	vine_spp) %>% 
    #the data are read in twice, so the code "distinct" retains one unique (distinct) version of each point.
    #I'm not sure what decision criteria would be used if the rows varied.
    distinct(point, .keep_all = TRUE)



basalarea2023 <- plot2023_overstory %>% 
    select(harvest, wall, point, spp, dbh, season) %>% 
    mutate(
        ba = (dbh/dbh) * 9.1,  #note the angle gauge BAF = 9.1
        stem_count = 43560 / (pi * ((2.75 * dbh)^2))
    ) %>% 
    group_by(harvest, wall, point) %>% 
    summarize(
        season = first(season),
        ba_acre = sum(ba),
        overstory_stems_acre = sum(stem_count),
        .groups = "drop"
    ) %>% 
    mutate(across(c(ba_acre, overstory_stems_acre), ~ replace_na(.,0))) %>% 
    bind_rows(cafri2023_structure_point_summary) %>% 
    select(-file, -df, -trmt, -x_dbh, -residual, -height, -brush) %>% 
    mutate(
        season = "2023"        
    )

    
#create plot2023 to include seedlings less than 12" and the basal area per acre
  
plot2023 <- bind_rows(plot2023cafri, plot2023_site) %>% 
    mutate(
        point = as.character(point)
    ) %>% 
    left_join(regen2023seedl, by = c("harvest", "season", "wall", "point")) %>% 
        #regen2023seedl has the seedling 4" - 12" data as "seedA" from the vegetation file.
        #
    left_join(basalarea2023, by = c("harvest", "wall", "point", "season")) %>% 
    rename(
        basalarea = ba_acre
    ) %>% 
    select(harvest, wall, season, point, basalarea, everything()) %>% 
    select(-date)
```

2024 plot (includes state forest plots)
```{r}
plot2024_orig <- read_excel(file_path_2024, sheet = "site data", skip = 3 ) %>% 
    select(-location, -wall, -comments) %>% 
    mutate(
        season = "2024"
    ) %>% 
      mutate(point = str_replace(point, "^0+", "")) %>%  
        # string replacement (str_replace) from the start (^), match one or more zeros (0+) and replace with ""

# harvest_original = harvest, forest_type=n_hdwd or oak_hdwd
# this df only has basal area values (not original data) for the state forest land sites
       
    mutate(harvest = case_when(
      harvest == "boot" ~ "sta_rd",  #this changes "harvest" from boot to sta_rd
      harvest == "decker control" ~ "decker",
      harvest == "recknagel south" ~ "recknagel",
      harvest == "wedge" ~ "sta_rd",
      harvest == "wedge control" ~ "sta_rd",
      harvest == "rectangle" ~ "recknagel",
    TRUE ~ harvest  # Keeps other values unchanged
    )
  ) %>% 

left_join(point_labels_revision, by = c("harvest", "point")) %>% 
    mutate(
        height = case_when(
        harvest %in% c("danby_block_2", "fairfield_a5", "fairfield_a6", "goundry_a17", "5_1") ~ 0,
        TRUE ~ height
  )
) %>% 
    mutate(
        trmt = case_when(
        harvest %in% c("danby_block_2", "fairfield_a5", "fairfield_a6", "goundry_a17", "decker", "rectangle", "5_1") ~ "control",
        TRUE ~ trmt
  )
) %>% 
    mutate(
        location = case_when(
        harvest %in% c("danby_block_2", "fairfield_a5", "fairfield_a6", "goundry_a17", "5_1") ~ "c",
        TRUE ~ location
  )
) %>% 
    
    mutate(
        wall = case_when(
        harvest %in% c("danby_block_2", "fairfield_a5", "fairfield_a6", "goundry_a17", "5_1") ~ "99",
        TRUE ~ wall
  )
) %>% 

    mutate(
        brush = case_when(
        harvest %in% c("fairfield_a5") ~ "b",
        harvest %in% c("danby_block_2",  "fairfield_a6", "goundry_a17", "5_1") ~ "u",
        TRUE ~ brush
  )
) %>% 
        
    mutate(
        forest_type = case_when(
        harvest %in% c("danby_block_2", "fairfield_a5", "5_1") ~ "oak_hdwd",
        harvest %in% c("fairfield_a6", "goundry_a17") ~ "n_hdwd",
        TRUE ~ forest_type
  )
) %>% 

    mutate(
        harvest_original = case_when(
        harvest %in% c("danby_block_2", "fairfield_a5", "fairfield_a6", "goundry_a17", "5_1") ~ harvest,
        TRUE ~ harvest_original
  )
)



#merge the seedlings 4" to 12" within df=regen2024seedl with site attribute data to create df=plot2024
# on line number 505, only type = "commercial" were retained
plot2024 <- left_join(plot2024_orig, regen2024seedl, by = c("harvest", "point", "wall", "season")) %>%  
#    filter(!harvest == "stand_5.1") %>% 
    mutate(harvest = case_when(
      harvest == "boot" ~ "sta_rd",  # old ~ new
      harvest == "recknagel south" ~ "recknagel",
      harvest == "wedge" ~ "sta_rd",
      harvest == "wedge control" ~ "sta_rd",
      harvest =="stand_5.1" ~ "5_1",
    TRUE ~ harvest  # Keeps other values unchanged
    )
  ) %>% 
    mutate(seedA = comm_sdl)


```


To standardize labels the strata will be denoted as:
A = 4" to 12" (these are not sorted as suppressed vs. exposed and spp is "004" = other hardwood)
B = 12" to 4.5/5 ft
C1 = 4.5 ft to 9 ft
C2 = 9ft to 1" dbh
D = 4.5/5 ft to 1" (the aggregate of C1 and C2)
P = 1" to 4" (saplings)


#Merge Site/Plot Data
df=site_data_early (2019 - 2022)
```{r}
site_data_early <- bind_rows(plot2019, plot2020, plot2021, plot2022) %>% 
    select(-c(location)) %>% 
    mutate( 
      baf = 9.1,
      
      harvest = case_when(
      harvest == "b" ~ "sta_rd",
      harvest == "boot" ~ "sta_rd",
      harvest == "cgl" ~ "gas_line",
      harvest == "circle1" ~ "circle",
      harvest == "circle2" ~ "circle",
      harvest == "circle3" ~ "circle",
      harvest == "cricle1" ~ "circle",
      harvest == "decker control" ~ "decker",
      harvest == "crp" ~ "red_pine",
      harvest == "cwb" ~ "sta_rd",
      harvest == "gasline" ~ "gas_line",
      harvest == "gl" ~ "gas_line",
      harvest == "recksouth" ~ "recknagel",
      harvest == "rectangle" ~ "recknagel",
      harvest == "rect1" ~ "recknagel",
      harvest == "rect2" ~ "recknagel",
      harvest == "rect3" ~ "recknagel",
      harvest == "rect4" ~ "recknagel",
      harvest == "rectangle1" ~ "recknagel",
      harvest == "rectangle2" ~ "recknagel",
      harvest == "rectangle3" ~ "recknagel",
      harvest == "rectangle4" ~ "recknagel",
      harvest == "rp" ~ "red_pine",
      harvest == "redpine" ~ "red_pine",
      harvest == "stationrd" ~ "sta_rd",
      harvest == "w" ~ "sta_rd",
      harvest == "wb" ~ "sta_rd",
      harvest == "wedge" ~ "sta_rd",
      harvest == "old" ~ "new",
      TRUE ~ harvest  # Keeps other values unchanged
    )
  ) %>% 
    left_join(point_labels_revision, by = c("harvest", "point")) %>% 
    mutate(across(c(oakhick, beech, maple, birch, hdwd, conifer), ~ replace_na(., 0))) %>% 
    mutate(across(c(baf), ~ replace_na(., 9.1))) %>% 
    mutate(
        basalarea = baf * (oakhick + beech + maple + birch + hdwd + conifer)
    ) %>% 
    select(-oakhick, -beech, -maple, -birch, -hdwd, -conifer, -baf)
    
site_data <- bind_rows(site_data_early, plot2023, plot2024) %>% 
    #
    #plot2019, plot2023 and plot2024 have actual counts of 4" to 12" seedlings. Variable "seedA"
    #is the value expanded to a "stems per acre" basis. plot2020 - plot2022 use a categorical variable
    #that is randomly expanded to a "stems per acre" basis as variable "seedA"
    #
    
    filter(!(harvest == "gas_line" & point %in% c(866, 864))) %>%   #corrected code selects for either point at the 
        # intersection of harvest and point
    
       select(harvest, wall, season, date, point, location, trmt, height, disturb, slash, fern, rhizom, clump,
           grass, herb, herb_spp, native_cov, native_spp, mid_cov, mid_spp, vine_cov, vine_spp, invasive_cov,
           invasive_spp, upper_cov, upper_spp, basalarea,
           comm_sdl, seedA) %>%
    arrange(harvest, season, wall, point) %>% 
    mutate(
            rhiz_fern = coalesce(fern, rhizom)
        ) %>% 
        select( -fern, -rhizom, ) %>% 
    mutate(
        filename = "slash_veg_complete.Rmd",
        df = "site_data"
        ) %>% 

    select(filename, df, harvest, wall, season, date, point, location, trmt, height, everything()) 


#   filter(season == "2023")
#    filter(wall == "98")

#write the aggregation of all site data to the output file
##change the name of the output file to today's date

output_file <- file.path(output_folder, paste0("site_data_2019_2024(22july2025)", ".xlsx"))

write_xlsx(site_data, output_file)


```


To standardize labels:
A = 4" to 12"           radius = 3.7 ft; count in 2019, 2023, and 2024 otherwise categorical 0 = 0; 1 = 1to5; 2 = 6 to 15; 3 = 16 to 30; 4 = 31+
B = 12" to 4.5/5 ft     radius = 6 ft  (size 01)
C1 = 4.5 ft to 9 ft     radius = 6 ft  (size 02)
C2 = 9ft to 1" dbh      radius = 6 ft  (size 03)
D = 4.5/5 ft to 1"      radius = 6 ft  (size 02 and 03)
P = 1" to 4" (saplings) radius thru 2022 = 6 ft; then 16 ft  (size sapl)



#merging regeneration data across years, select 4" to 12" seedlings from the site/plot data
create df =seedlings2019_2024

```{r}
seedlingspre2023 <- site_data %>% #from counts of seedlings onto site tally sheets
    filter (!season == "2023", !season == "2024" #remove season >= 2023
    ) %>% 
    select(harvest, wall, point, season, seedA) %>% 
    mutate(
        spp ="004"
    ) # %>% 
#    filter(harvest == "red_pine" & season == "2019")

seedlings2019_2024 <- bind_rows(regen2023seedl, regen2024seedl) %>%  # these df include only commercial species stem counts < 12" tall
    #this is the count of commercial seedlings 4" to 12" from the 2023 and 2024 veg data
    bind_rows(seedlingspre2023) %>% #seedlings before 2023
    mutate(
        spp = "004",
        origin = "single"
    )%>% 
    select(harvest, wall, point, season, spp, seedA, richness, origin) # %>% 
#   filter(seedA >0)
    
#    filter(harvest == "red_pine"  & season == "2019")
#    
#    (24july2025) for 2023 and 2024, seedA was collected in veg plots as "seed01". in 2023, only commercial species and sweet birch
#    were included. in 2024, all species were included.
```


#df=merge2019_2022_a
this df standarizes the labels for seedling height classes for years 2019 thru 2022
```{r}
#convert regen data to "per acre" for 2019 - 2022 before merging with 2023 and 2024
#regen2019-regen2022 data are stems counts per point
# regen2023 and regen2024 are stem counts per acre at each point.

merge2019_2022_a <- bind_rows(regen2019, regen2020, regen2021, regen2022) %>% 
    filter(!(harvest == "gas_line" & point %in% c(866, 864)))  %>%  
        #correct - code selects for either point at the intersection of harvest and point
    select(harvest, season, location, point, origin, spp, date, sup.seed01, sup.seed02, sup.seed03, 
           sup.sapl, exp.seed01, exp.seed02, exp.seed03, exp.sapl) %>% 

    bind_rows(seedlings2019_2024) %>%  #, by = c("harvest", "point", "season", "origin")) %>% 
    select(-wall, -richness, -location) %>% 
    filter(!season %in% c("2023", "2024")) %>% 
    
    #next, standardize the seedling labels for variance across years
        rename(
        supB = sup.seed01, #prior to 2023, seed01 was 12" to 60" and 4" to 12" was collected as site data
        supC1 = sup.seed02, #60" to 9ft
        supC2 = sup.seed03, # 9 ft to 1" dbh
        supP = sup.sapl, # 1" to 5" dbh
        expB = exp.seed01,
        expC1 = exp.seed02,
        expC2 = exp.seed03,
        expP = exp.sapl
    ) %>% 
    mutate(across(where(is.numeric), ~ replace_na(., 0))) %>% 
#next, standardize harvest names
    mutate(
    harvest = case_when(
      harvest == "b" ~ "sta_rd",
      harvest == "boot" ~ "sta_rd",
      harvest == "cgl" ~ "gas_line",
      harvest == "circle1" ~ "circle",
      harvest == "circle2" ~ "circle",
      harvest == "circle3" ~ "circle",
      harvest == "cricle1" ~ "circle",
      harvest == "crp" ~ "red_pine",
      harvest == "cwb" ~ "sta_rd",
      harvest == "gasline" ~ "gas_line",
      harvest == "gl" ~ "gas_line",
      harvest == "recksouth" ~ "recknagel",
      harvest == "rect1" ~ "recknagel",
      harvest == "rect2" ~ "recknagel",
      harvest == "rect3" ~ "recknagel",
      harvest == "rect4" ~ "recknagel",
      harvest == "rectangle1" ~ "recknagel",
      harvest == "rectangle2" ~ "recknagel",
      harvest == "rectangle3" ~ "recknagel",
      harvest == "rectangle4" ~ "recknagel",
      harvest == "rp" ~ "red_pine",
      harvest == "redpine" ~ "red_pine",
      harvest == "stationrd" ~ "sta_rd",
      harvest == "w" ~ "sta_rd",
      harvest == "wb" ~ "sta_rd",
      harvest == "wedge" ~ "sta_rd",
      harvest == "old" ~ "new",
      TRUE ~ harvest  # Keeps other values unchanged
    )
      )
```




this df converts to per acre values at each point (clump data not available for 2019 to 2022)
note that clump data were not distinct in 2019 through 2022, such that no estimates of ramets/clump or clump/acre are possible
#df=merge2019_2022_b
```{r}
#Next, calculate per acre values for regen 2019 through 2022
#

merge2019_2022_b <- merge2019_2022_a %>% 
    select(harvest, season, point, origin, spp, everything()) %>% 

#Next, convert plot counts to "per acre" values
   mutate( #converts plot count to count per acre
        # count = if_else(is.na(count), 1, count), # if count missing (it's a single) then make count = 1
        # stems.ramet = if_else(is.na(stems.ramet), 0, stems.ramet), # if variable is NA, change to 0, otherwise leave as is
        radius01 = 3.7,
        radius02 = 6,
        radius03 = 16,
        radius = if_else(season == "2023", radius03, radius02),  
          # Dynamically assigns radius to radius03 when season is 2023 and sets to radius02 if not 2023 (no 2023 data included as of 2/13/2025)
        supB = (supB * (43560 / (pi * radius02^2))),
        supC1 = (supC1 * (43560 / (pi * radius02^2))),
        supC2 = (supC2 * (43560 / (pi * radius02^2))),
        supP = (supP * (43560 / (pi * radius^2))),  # Uses the new "radius" variable here
        expB = (expB * (43560 / (pi * radius02^2))),
        expC1 = (expC1 * (43560 / (pi * radius02^2))),
        expC2 = (expC2 * (43560 / (pi * radius02^2))),
        expP = (expP * (43560 / (pi * radius^2)))   # Uses the new "radius" variable here
    ) %>%
    ungroup() %>% 
  mutate(  #if there is concern there are blank spaces in the grouping variable that cause grouping problems
    season = str_trim(season),
    harvest = str_trim(harvest),
    point = str_trim(point),
    spp = str_trim(spp),
    origin = str_trim(origin)
  ) %>%  #

#
# it appears the group_by summary happens in the next df.  I'll wait to join in "point_labels_revision until after the calculations 
    group_by(harvest, season, point, spp, origin) %>% 
    summarize(
        n = n(),
        date = first(date),
        seedA = seedA,
        supB = sum(supB),
        supC1 = sum(supC1),
        supC2 = sum(supC2),
        supP = sum(supP),
        expB = sum(expB),
        expC1 = sum(expC1), 
        expC2 = sum(expC2),
        expP = sum(expP),
       .groups = "drop"
    )
```



#Next, aggregate numeric values by harvest, season, point, spp, and origin. 

df=merge2019_2022_c
```{r relationship = "many-to-many}

merge2019_2022_c <- merge2019_2022_b %>% 
    
#this summarize is redundant with the same action from df=merge2019_2022_b
#the df doesn't change, so I'll leave it rather than potentially causing
#problems

        group_by(season, harvest, point, spp, origin) %>% 
        summarize(
            date = first(date),
            seedA = seedA,
            supB = sum(supB),
            supC1 = sum(supC1),
            supC2 = sum(supC2),
            supP = sum(supP),
            expB = sum(expB),
            expC1 = sum(expC1),
            expC2 = sum(expC2),
            expP = sum(expP),
            .groups = "drop"
        ) %>% 
    select(harvest, season,  point, spp, origin, everything()) %>% 
    arrange(harvest, season, point, spp, origin) %>% 
    mutate(
        stemcount = supB + supC1 + supC2 + supP + expB + expC1 + expC2 + expP,
        upperexposed = expC1 + expC2 + expP,
        present = if_else(stemcount > 0, 1, stemcount), # used to calculate diversity if a spp is present at a point
        success = if_else(upperexposed > 0, 1, upperexposed) # used to calculate diversity of spp of successful stems
#        stems.ramet = if_else(is.na(stems.ramet), 0, stems.ramet)
    ) %>%
    
    left_join(point_labels_revision, by= c("harvest", "point")) %>%  #this includes residual BA/ac for cafri walls
    
    left_join(seedlings2019_2024, by = c("harvest", "season", "wall", "spp", "point", "origin")) %>% 

    mutate(
        seedA = coalesce(seedA.x, seedA.y),
        spp = if_else(is.na(spp), "empty", spp)
    ) %>% 
    select(
        -seedA.x, -seedA.y
    ) 
 
 
```


create "numeric_vars"
#create a stand alone reference to the numeric variables that will be in the "complete" datafile
```{r}
numeric_vars <- c( "supB", "supC1", "supC2","height", 
                  "supP", "expB", "expC1", "expC2", "expP", "seedA", "stemcount", 
                  "upperexposed", "present", "success")

numeric_vars.2 <- c("supA", "supB", "supC1", "supC2", "seedA",
                  "supD", "supP", "expA", "expB", "expC1", "expC2","expD", "expP",
                  "stems_per_ramet", "clumps_per_acre")

```




# df=regen2023_updated, seedlings > 4", matching variables to bind rows with 2019-2022 data
```{r}
regen2023_updated <- regen2023 %>% 
  select(-harvest_original, -forest_type, -brush) %>% #add these later, after summarize
  rename( #new = old
        stems_per_ramet = numb.per.clump
    ) %>% 

#create "seedA" by adding supA+expA for commercial species
  mutate(seedA = case_when(
    type == "commercial" ~ supA + expA,
    TRUE ~ NA_real_
  )) %>% 
    
# this "mutate" results in a variable seedA that have values when type = commercial and is NA for other types.

#(july 23 2025) create metric for number of clumps by species
#each clump represents the "multiplier" based on plot size as the clumps per acre
# the sum of supA::expP is because there is only one value among those variables, the others are zero
# dividing the number of stems by count = the plot multiplier
# 
  mutate(clump_per_acre_each = if_else
         (origin == "clump", (supA+supB+supD+supP+expA+expB+expD+expP)/count, NA_real_)) %>% 

#for regen2023 data, aggregate the stem data at the point spp and origin level
group_by(harvest, point, spp, origin) %>% 
    summarize(
       wall = first(wall),
       season = first(season),
       date = first(date),
       trmt = first(trmt),
       location = first(location),
       type = first(type),
       supA = sum(supA),
       supB = sum(supB),
       supD = sum(supD),
       supP = sum(expP),
       expA = sum(expA),
       expB = sum(expB),
       expD = sum(expD),
       expP = sum(expP),
       seedA = sum(seedA),
       stems_per_ramet = mean(stems_per_ramet),
       clumps_per_acre = sum(clump_per_acre_each),
       .groups = "drop"
    ) %>% 

  left_join(point_labels_revision, by = c("harvest", "wall", "location", "trmt", "point"))
    
```


# df = regen2024_updated, same code as regen2023_updated
```{r}
#all this code basically identical to the previous chunck for "df=regen2023_updated"
#
regen2024_updated <- regen2024 %>% 
    select(-wall, -multip01, -multip02_03, -multip_sapl) %>% 
    filter(!harvest == "5_1") %>% 
  rename( #new = old
        stems_per_ramet = numb.per.clump
    ) %>% 

 #create "seedA" by adding supA+expA for commercial species
  mutate(seedA = case_when(
    type == "commercial" ~ supA + expA,
    TRUE ~ NA_real_
  )) %>% 
   
#(july 23 2025) create metric for number of clumps by species
#each clump represents the "multiplier" based on plot size as the clumps per acre
# the sum of supA::expP is because there is only one value, the others are zero
# dividing the number of stems by count = the plot multiplier
# 
  mutate(clump_per_acre_each = if_else
         (origin == "clump", (supA+supB+supD+supP+expA+expB+expD+expP)/count, NA_real_)) %>% 

#for regen2024 data, aggregate the stem data at the point spp and origin level
group_by(harvest, point, spp, origin) %>% 
    summarize(
       season = first(season),
       date = first(date),
       type = first(type),
       supA = sum(supA),
       supB = sum(supB),
       supD = sum(supD),
       supP = sum(expP),
       expA = sum(expA),
       expB = sum(expB),
       expD = sum(expD),
       expP = sum(expP),
       seedA = sum(seedA),
       stems_per_ramet = mean(stems_per_ramet),
       clumps_per_acre = sum(clump_per_acre_each),
       .groups = "drop"
    ) %>% 

      left_join(point_labels_revision, by = c("harvest", "point")) #%>% 


```




bind all data into one df and then make complete for all species in a wall for each season
the df below, yay, lacked completeness for the 2023 and 2024 data.
#df = all_seasons (to be made complete)
```{r}
data_2019_2022 <- merge2019_2022_c %>% 
    select(harvest, season, wall, point, spp, origin, supB, supC1, supC2, supP, expB, expC1, 
           expC2, expP, seedA)

data_2023 <- regen2023_updated %>% 
        select(harvest, season, wall, point, spp, origin, date, supA, supB, supD, supP, expA, expB,  
               expD, expP, seedA, stems_per_ramet, clumps_per_acre)

data_2024 <- regen2024_updated %>% 
        select(harvest, season, wall, point, spp, origin, date, supA, supB, supD, supP, expA, expB,  
               expD, expP, seedA, stems_per_ramet, clumps_per_acre)

spp_lookup <- read_csv("../species table.csv", skip = 1, col_names = c("spp_code", "spp_name"))


all_seasons <- bind_rows(data_2019_2022, data_2023, data_2024) %>% 
  mutate(spp = as.character(spp)) %>%
  left_join(spp_lookup, by = c("spp" = "spp_code")) %>%
  mutate(
    spp = coalesce(spp_name, spp),
    spp = as.character(spp)
  ) %>%
  select(-spp_name) %>% 
  mutate(harvest = case_when(
      harvest == "rectangle" ~ "recknagel",
    TRUE ~ harvest  # Keeps other values unchanged
  )) 
  
all_seasons %>% 
    filter(origin == "unknown")



table(all_seasons$spp)
```

make complete 2019 through 2024
#df= all_seasons that results in df=yay2
# (season x wall)
```{r eval=FALSE}
# step 1 - make a list of all species for each combination of season and wall

species_per_wall <- all_seasons %>%
  group_by(season, wall) %>%
  summarize(
      all_species = list(unique(spp)), # this must make a narrow-long list of species
      .groups = "drop")

# Step 2 - build all combinations

all_combinations <- all_seasons %>%
  select(harvest, season, wall, point) %>%
  distinct() %>%
  left_join(species_per_wall, by = c("season", "wall")) %>% # the list of all possible harvest/season/wall/point is long, and replicated across the list of species
  unnest(all_species) %>%
  rename(spp = all_species)

# Step 3 - combine the df and make the numeric_vars complete

yay2_wall <- all_combinations %>%
  left_join(all_seasons, by = c("harvest", "season", "wall", "point", "spp")) %>%
  mutate(
    across(all_of(numeric_vars.2), ~ replace_na(.x, 0)),
    origin = replace_na(origin, "unknown"),
    date = replace_na(date, "unknown")
  ) %>%
  arrange(harvest, season, wall, point, spp, origin) %>% 
  
     mutate(
        firstgrow = 2035, # default value for the first growing season
        firstgrow = case_when(
          harvest %in% c("gas_line", "red_pine") ~ 2017,
          harvest %in% c("sta_rd") ~ 2018,
          harvest %in% c("campridge", "recknagel", "recknagel_north", "circle", "decker") ~ 2020,
          harvest %in% c("6_6", "6_9") ~ 2022,
          harvest %in% c("danby_block_2") ~ 2018,  #see comments after this chunk about the State Forest "firstgrow"
          harvest %in% c("fairfield_a5") ~ 2022,
          harvest %in% c("fairfield_a6") ~ 2010,
          harvest %in% c("goundry_a17") ~ 2018,
          TRUE ~ firstgrow  # Keeps the existing value of "firstgrow" for all other cases
    ),
        year = as.numeric(season),
        growingseasons = (year - firstgrow) + 1
  ) %>% 
    
    mutate(
        type = "blank"
        ) %>% 
    mutate(type = case_when(
    spp %in% c("ba", "bc", "bih", "bl", "co", "he", "ohw", "rm", "ro", "sa", 
                "ph","sh", "sm", "wa", "wo", "wp", "yb", "yp", "comm_sdl") ~ "commercial",
    
    spp %in% c("sb") ~ "sb",
    
    spp %in% c("ab", "eh", "pc", "stm", "toh", "au" ) ~ "interfere",

    spp %in% c("app", "bta", "haw", "brs", "dws", "ec", "la", "ns", "cu", "erc", "ows","amc",
               "pb", "gb", "qa", "rp", "scp", "smc", "svb", "wh", "wi", "viac", "visp" ) ~ "diversity",

    spp %in% c("ah", "cc", "unk", "sp", "empty" ) ~ "other",
    TRUE ~ type,  # Keeps the existing value of "type" for all other cases
    )) %>% 
    

    mutate(
    stemcount = coalesce(supB, 0) + coalesce(supC1, 0) + coalesce(supC2, 0) + coalesce(supP, 0) + 
                coalesce(expB, 0) + coalesce(expC1, 0) + coalesce(expC2, 0) + coalesce(expP, 0),

    upperexposed = coalesce(expC1, 0) + coalesce(expC2, 0) + coalesce(expP, 0),

    present = if_else(stemcount > 0, 1, 0),  # Ensures output is never NA

    success = if_else(upperexposed > 0, 1, 0),  # Ensures output is never NA


    stems_per_ramet = if_else(is.na(stems_per_ramet), 0, stems_per_ramet)
    ) %>% 
    mutate(
        across(all_of(numeric_vars.2), ~ ifelse(seedA > 0 & is.na(.), 0, .)),
        across(all_of(numeric_vars.2), ~ ifelse(is.na(.), 0, .)), 
        across(all_of(numeric_vars.2), ~ ifelse(seedA <1 & spp == "004", 0, .)),
        seedA = if_else(is.na(seedA), 0, seedA),
        total = (supB + supC1 + supC2 + supP + expB + expC1 + expC2 + expP) #bracket D = C1 + C2
    ) %>% 
    mutate(
        filename = "slash_veg_complete.Rmd",
        df = "yay2_wall"
    )
```


MAKE COMPLETE
make complete 2019 through 2024
#df= all_seasons that results in df=yay2_harvest
# (season x harvest)
```{r}
#Step 1 - Create a list of distinct levels of attributes and group by season x harvest
species_origin_per_harvest <- all_seasons %>%
  distinct(season, harvest, spp, origin) %>%
  group_by(season, harvest) %>%
  summarize(
    spp_origin = list(tibble(spp = spp, origin = origin)),
    .groups = "drop"
  )


# Step 2 - build all combinations
all_combinations_harvest <- all_seasons %>%
  select(harvest, season, wall, point) %>%
  distinct() %>%
  left_join(species_origin_per_harvest, by = c("season", "harvest")) %>%
  unnest(spp_origin)  # this will give you separate spp and origin columns



# Step 3 - combine the df and make the numeric_vars complete

yay2_harvest <- all_combinations_harvest %>%
  select(-wall) %>% 
  left_join(all_seasons, by = c("harvest", "season",  "point", "spp", "origin")) %>%
  select(-wall) %>% 
  mutate(
    across(all_of(numeric_vars.2), ~ replace_na(.x, 0)),
    origin = replace_na(origin, "single"),
    date = replace_na(date, "unknown")
  ) %>%
  arrange(harvest, season, point, spp, origin) %>% 
  
# Fairfield and Danby = Region 7, Petit, Travis D (DEC) <Travis.Petit@dec.ny.gov> 
# fairfield SF A5, contract open sept 2020 - november 2025, first growing season = 2021 (confirmed 7/31/2025)
# fairfield SF a6 (aka A-6.1) cut 2009, first growing season = 2010 confirmed (logging finished 2009, confirmed 7/31/2025)
# Danby SF block 2, contract open november 2016 to october 2019, EWP 5 whorls in 2024, first growing season = 2018 (7/31/2025 confirmed)
# 
# Goundry Hill SF = Region 8, Cicora, Gretchen L (DEC) <gretchen.cicora@dec.ny.gov>
# Goundry hill SF , goundry_a17, contract open december 2016 to january 2020, first growing season = 2018 (confirmed 8/8/2025)

     mutate(
        firstgrow = 2035, # default value for the first growing season
        firstgrow = case_when(
          harvest %in% c("gas_line", "red_pine") ~ 2017,
          harvest %in% c("sta_rd") ~ 2018,
          harvest %in% c("campridge", "recknagel", "recknagel_north", "circle", "decker") ~ 2020,
          harvest %in% c("6_6", "6_9") ~ 2022,
          harvest %in% c("danby_block_2") ~ 2018,  #(confirmed) see comments after this chunk about the State Forest "firstgrow"
          harvest %in% c("fairfield_a5") ~ 2021,   #(confirmed)
          harvest %in% c("fairfield_a6") ~ 2010,   #(confirmed)
          harvest %in% c("goundry_a17") ~ 2018,
          TRUE ~ firstgrow  # Keeps the existing value of "firstgrow" for all other cases
    ),
        year = as.numeric(season),
        growingseasons = (year - firstgrow) + 1
  ) %>% 
    
    mutate(
        type = "blank"
        ) %>% 
    mutate(type = case_when(
    spp %in% c("ba", "bc", "bih", "bl", "co", "he", "ohw", "rm", "ro", "sa", 
                "ph","sh", "sm", "wa", "wo", "wp", "yb", "yp", "comm_sdl") ~ "commercial",
    
    spp %in% c("sb") ~ "sb",
    
    spp %in% c("ab", "eh", "pc", "stm", "toh", "au" ) ~ "interfere",

    spp %in% c("app", "bta", "haw", "brs", "dws", "ec", "la", "ns", "cu", "erc", "ows","amc",
               "pb", "gb", "qa", "rp", "scp", "smc", "svb", "wh", "wi", "viac", "visp" ) ~ "diversity",

    spp %in% c("ah", "cc", "unk", "sp", "empty" ) ~ "other",
    TRUE ~ type,  # Keeps the existing value of "type" for all other cases
    )) %>% 
    

    mutate(
    stemcount = coalesce(supB, 0) + coalesce(supC1, 0) + coalesce(supC2, 0) + coalesce(supP, 0) + 
                coalesce(expB, 0) + coalesce(expC1, 0) + coalesce(expC2, 0) + coalesce(expP, 0),

    upperexposed = coalesce(expC1, 0) + coalesce(expC2, 0) + coalesce(expP, 0),

    present = if_else(stemcount > 0, 1, 0),  # Ensures output is never NA

    success = if_else(upperexposed > 0, 1, 0),  # Ensures output is never NA


    stems_per_ramet = if_else(is.na(stems_per_ramet), 0, stems_per_ramet)
    ) %>% 
    
# there are duplicated rows in the df, often associated with a species that has both origins single and clump.
# The result is there are unbalanced combinations of actual data rows per harvest x season vs. expected (#species_origin x # points)
# the code "distinct" should remove the duplicated row.
# testing for this is below following df = expected_missing
    distinct(season, harvest, point, spp, origin, .keep_all=TRUE) %>% 
    
    mutate(
        across(all_of(numeric_vars.2), ~ ifelse(seedA > 0 & is.na(.), 0, .)),
        across(all_of(numeric_vars.2), ~ ifelse(is.na(.), 0, .)), 
        across(all_of(numeric_vars.2), ~ ifelse(seedA <1 & spp == "004", 0, .)),
        seedA = if_else(is.na(seedA), 0, seedA),
        total = (supB + supC1 + supC2 + supP + expB + expC1 + expC2 + expP) #bracket D = C1 + C2
    ) %>% 

# check for replication of "empty" at points that have other species
# this might have happened when rejoining bracket A seedling data into the larger groups
# prior to this step "table(yay2_harvest$spp)" showed empty = 1619
        
  group_by(harvest, season, point) %>%  
  mutate(
      n_non_empty = sum(spp != "empty" #counts the number of spp at harvest x season x point that are not "empty"
                        )) %>%
  ungroup() %>%
  filter(!(spp == "empty" & n_non_empty > 0)) %>% #removes (reverse filter) when spp = empty and other species >0
  select(-n_non_empty) %>%   # optional cleanup

    mutate(
        filename = "slash_veg_complete.Rmd",
        df = "yay2_harvest"
    ) %>% 
    arrange(season, harvest, point, spp, origin) %>% 
    rename(
        sup_sd4_12 = supA,
        sup_sd12_5 = supB,
        sup_sd5_9 = supC1,
        sup_sd9_1 = supC2,
        sup_sd5_1 = supD, #D = C1 + C2
        sup_sp_1_5 = supP,
        exp_sd4_12 = expA,
        exp_sd12_5 = expB,
        exp_sd5_9 = expC1,
        exp_sd9_1 = expC2,
        exp_sd5_1 = expD, #D = C1 + C2
        exp_sp_1_5 = expP
    ) %>% 
    mutate(
    cluster = "spare"
)   %>%  
    mutate(cluster = case_when(
        harvest == "decker" & location == "c" ~ "early",
        harvest %in% c("gas_line", "red_pine", "sta_rd", "fairfield_a5", "danby_block_2", "goundry_a17", "fairfield_a6") ~ "early",
        harvest %in% c("campridge", "recknagel", "circle") ~ "second",
        harvest %in% c("6_6", "6_9") ~ "cafri",
    TRUE ~ cluster
)) %>% 
    
    
    left_join(point_labels_revision, by = c("harvest", "point")) %>% 
 
      mutate(
    wall = if_else(
      wall == "99" & harvest %in% c("campridge", "danby_block_2", "decker", "fairfield_a5",
                                    "fairfield_a6", "gas_line", "goundry_a17",
                                    "recknagel", "recknagel_north", "red_pine"),
      harvest,   # assign the harvest value to wall
      wall       # otherwise keep wall as is
    )) %>%
  
    
    select(filename, df, cluster, season, harvest, wall, trmt, location, brush, residual, 
           point, growingseasons, type, spp, origin,
           seedA, total, stemcount, upperexposed, present, success,
           sup_sd4_12, sup_sd12_5, sup_sd5_9, sup_sd9_1, sup_sd5_1, sup_sp_1_5,
           exp_sd4_12, exp_sd12_5, exp_sd5_9, exp_sd9_1, exp_sd5_1, exp_sp_1_5,
           clumps_per_acre, stems_per_ramet, firstgrow, year, date, height, harvest_original, forest_type)

#        sup_sd4_12, sup_sd12_5, sup_sd5_9, sup_sd9_1, sup_sd5_1, sup_sp_1_5,
 #       exp_sd4_12, exp_sd12_5, exp_sd5_9, exp_sd9_1, exp_sd5_1, exp_sp_1_5

table(yay2_harvest$harvest, yay2_harvest$wall)
table(yay2_harvest$harvest, yay2_harvest$growingseasons)
table(yay2_harvest$harvest, yay2_harvest$cluster)
table(yay2_harvest$wall)

# Within a harvest there can be multiple walls (e.g., harvest = recknagel includes walls: recknagel, r1, r2, r3 and r4).  
# A higher order variable "cluster" has values of: early, second, cafri and drop. "drop" should be excluded from analysis.
# Cluster aggregates multiple harvests that happen in the same year or same 
# vicinity (e.g., cluster = "second" for the 2019-2020 campridge/recknagel spatial cluster).
# Each level of cluster includes trmt = "protected" (inside) and trmt = "control" (outside)

```

    mutate(
    harvest = case_when(
      harvest == "cgl" ~ "gasline",
      harvest == "crp" ~ "redpine",
      harvest == "cwb" ~ "stationrd",
      harvest == "gl" ~ "gasline",
      harvest == "rp" ~ "redpine",
      TRUE ~ harvest  # Keeps other values unchanged
    )
  ) %>% 



checking for errors

```{r}
 wontjoin <- yay2_harvest %>% 
  anti_join(point_labels_revision, by = c("harvest", "wall", "point")) %>%
  count(harvest, wall) 


error_check <- yay2_harvest %>% 
    filter(spp == "empty")

table(error_check$harvest, error_check$season, error_check$type)

error_check02 <- yay2_harvest %>% 
    filter(is.na(spp))
```



# Fairfield and Danby = Region 7, Petit, Travis D (DEC) <Travis.Petit@dec.ny.gov> 
# fairfield SF A5, contract open sept 2020 - november 2025, first growing season = 2021 (confirmed 7/31/2025)
# fairfield SF a6 (aka A-6.1) cut 2009, first growing season = 2010 confirmed (logging finished 2009, confirmed 7/31/2025)
# Danby SF block 2, contract open november 2016 to october 2019, EWP 5 whorls in 2024, first growing season = 2018 (7/31/2025 confirmed)
# 
# Goundry Hill SF = Region 8, Cicora, Gretchen L (DEC) <gretchen.cicora@dec.ny.gov>
# Goundry hill SF , goundry_a17, contract open december 2016 to january 2020, first growing season = 2018 (confirmed 8/8/2025)




testing yay1 vs yay2
```{r eval=FALSE}
combined_df <- bind_rows(yay1, yay2, .id = "source")  # 'source' will be "1" or "2"

combinded_df <- combined_df %>% 
    mutate(source = recode(source, `1` = "yay1", `2` = "yay2"))
    
summary_df <- combined_df %>%
  group_by(source, season, harvest) %>%
  summarize(mean_seedA = mean(seedA, na.rm = TRUE), .groups = "drop")

library(ggplot2)

ggplot(summary_df, aes(x = season, y = mean_seedA, color = source)) +
  geom_point() +
  geom_line(aes(group = source)) +
  facet_wrap(~harvest, scales = "free_y") +
  labs(
    title = "Mean SeedA by Harvest and Season",
    x = "Season",
    y = "Mean seedA"
  ) +
  theme_minimal()

ggplot(summary_df, aes(x = season, y = mean_seedA, color = source)) +
  geom_point(position = position_dodge(width = 0.3)) +
  geom_line(aes(group = source), position = position_dodge(width = 0.3)) +
  facet_wrap(~harvest, scales = "free_y") +
  labs(title = "Mean SeedA by Harvest and Season", x = "Season", y = "Mean seedA") +
  theme_minimal()

```


testing basic metrics to provide confidence that df=yay2 is complete and ready to go
#tests if the number of species is consistent among points by harvest x wall x season (the result was TRUE, no deviants)
```{r}
#tests if the number of species is consistent among points by harvest x wall x season (the result was TRUE, no deviants)
yay_test1 <- yay2_harvest %>% 
 group_by(harvest, season, wall, point) %>%
  summarize(
      n_spp = n_distinct(spp), 
      .groups = "drop") %>% 
      group_by(harvest, season, wall) %>%
  mutate(max_spp = max(n_spp),
         min_spp = min(n_spp),
         n_points = n(),
         is_deviant = n_spp != max_spp | n_spp != min_spp) %>%
  ungroup() %>%
  filter(is_deviant) %>% 
  arrange(harvest, season, wall)
```


#assess if the actual number of records = # points x # species
```{r}

#one
yay2_expected_points <- yay2_harvest %>% 
  group_by(season, harvest) %>%
  summarize(
    points_distinct = n_distinct(point),
    spp_distinct = n_distinct(spp),
    expected = n_distinct(point) * n_distinct(spp),
    actual = n()
  ) %>%
  filter(actual != expected)

# yay2_harvest_sub <- yay2_harvest %>% 
#     filter(harvest == "circle" & season == "2020")


#there were 41 mis-matched actual vs. expected. The following code may help resolve

#two
# Identify all expected combinations
expected_combos <- yay2_harvest %>%
  select(season, harvest, point, spp, origin) %>%
  distinct()

#three
# Group to find which are missing
expected_missing <- yay2_harvest %>%
  select(season, harvest, point, spp, origin) %>%
  distinct() %>%
  right_join(expected_combos, by = c("season", "harvest", "point", "spp", "origin")) %>%
  filter(is.na(season))  # these were not present in yay2

#four
# another way of coding the two previous pieces
# this produces an output because "origin" isn't included
yay2_harvest_test_expected <- yay2_harvest %>%
  group_by(season, harvest) %>%
  summarize(
    points = n_distinct(point),
    spp = n_distinct(spp),
    expected = points * spp,
    actual = n(),
    difference = actual - expected
  ) %>%
  filter(difference != 0)

#five
#another check
nrow(expected_combos) == nrow(yay2_harvest %>% distinct(season, harvest, point, spp, origin))

#six
#the mis match of actual and expected may be a result of duplicate rows. the test follows.
yay2_harvest %>%
  count(season, harvest, point, spp, origin) %>%
  filter(n > 1)
```

#assess species distribution by season and origin (not sure why this matters)
```{r}
yay2_harvest %>%
  count(season, spp, origin) %>%
  pivot_wider(names_from = spp, values_from = n, values_fill = 0)

yay2_harvest %>%
  count(season, spp, origin) %>%
  pivot_wider(
    names_from = spp,
    values_from = n,
    values_fill = 0
  )

```




```{r}
# Define the output file path
output_file <- file.path(output_folder, paste0("veg_data_2019to2024_complete(12sept2025)", ".xlsx"))
write_xlsx(yay2_harvest, output_file)
```

