---
title: "slash veg complete"
author: "Peter Smalldige"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- README

File Description: The purpose of this file is to:
1. read the raw vegetation data and make it complete
2. convert the field measurements to per/acre values
3. make data complete

exported data files include:
xxx (line 237)
xxx (line 353)
xxx (line 530)
xxx

 -->


###Initialize Libraries

```{r message=FALSE}

#library(ggplot2)
#library(ggpubr)
library(expss)
library(readr)
library(readxl)
library(writexl)
library(skimr)
library(tidyverse)  # includes: tidyr, dplyr, haven, readr, readxl, forcats, stringer, ggplot2. See all with tidyverse_packages()
library(knitr)
library(janitor)
#install.packages("openxlsx")
library(openxlsx)
# install.packages("shinyaddins")
# library(shinyaddins)

```

#
# ALT plus hypen = <- (within chunk)
# ctrl plus shift plus m = %>% 
# ctrl + ALT + I = insert chunk
# ALT + - = insert assignment operator
# filter rows, select columns
# RMD comment = <!-- comment -->
# 
# 

<!--
-->

<!--     -->

### code 01
### Create file path input and the output folder (code 01) 

file paths, input and output
```{r}
#file_path_input01 <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\regen_per_acre_all_alpha_tidy_trmt_coded 2019-2022 (with read me).xlsx"
file_path_2023_veg <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\complete_full_2023(readme).xlsx"

file_path_2023_veg_22july2025 <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\raw_data\\veg2023_raw(22july2025).xlsx"

file_path_2019 <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\regen.data.2019.xlsx"
    # line 93
file_path_2020 <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\regen.data.2020_cleaned.xlsx"
    # line 143
file_path_2021 <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\regen.data.2021_cleaned.xlsx"
    # line 179
file_path_2022 <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\slash.wall.data.2022.15february2025.xlsx"
  # this includes cafri and non-cafri data
 
file_path_2024 <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\raw_data\\slash.wall.data_2024_FINAL_1june2025.xlsx"
    # the FINAL 2024 growing season vegetation state forest data added to main folder on 1june2025
    # This file path  includes worksheets for "regen" and "plot" (line 734)
    # 
    # NOTE - for the state forest stands (danby_block_2, fairfied_a5, fairfield_a6, goundry_a17, and decker
    # set location = "control")
     

file_path_points <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\point_labels_all(readme).xlsx"


#plot data for the 4" to 12" seedling data
file_path_2019_plot <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\plot-data-2019.csv"
   
file_path_2020_plot <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\plot_calc_2020_untidy.csv"
    
file_path_2021_plot <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\slash.wall.data.2021.30october2022.xlsx"
    
file_path_2022_plot <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\plot_data_2022_untidy.csv"
#  does not include any CAFRI data

file_path_2023_plot <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\slash.wall.data.2023.23july2024(maybe FINAL).xlsx"

file_path_2023_cafri_overstory <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\overstory dbh data for 6-9 and 6-6 CAFRI_2023_oct_30.xlsx"
# file includes the original dbh data with distances on marginal trees. Retain distance for future assessment
# of when a tree is considered "in"


# 2024 plot data is within "file_path_2024"

output_folder <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\"


```
In 2019 - 2022, seedling regen data were recorded in the following strata:
Site/plot sheet: counts of seedlings 4" to 12", sometimes as categorical
Regen Tally sheet: 
    12" to 4.5ft (B) = formerly seed01
    4.5ft - 9 ft (C1) = formerly seed02
    9ft to 1"(C2)   [C1 + C2 = D], formerly seed03
    1 - 4", formerly sapl

in 2023 and beyond, seedling regen data were recorded in the following strata:
Regen tally sheet: 
    (A = 01) 4" to 12"
    (B = 02) 12" to 60"  (the break point at 60" vs. 54" was to align with SILVAH protocols)
    (D = 03) 60" tall to 1" dbh
    (P = 04) 1" to 4" dbh [in 2024, the upper boundary was 5"...bad idea]

To standardize labels:
A = 4" to 12"           radius = 3.7 ft; count in 2019, 2023, and 2024 otherwise categorical 0 = 0; 1 = 1to5; 2 = 6 to 15; 3 = 16 to 30; 4 = 31+
B = 12" to 4.5/5 ft     radius = 6 ft
C1 = 4.5 ft to 9 ft     radius = 6 ft
C2 = 9ft to 1" dbh      radius = 6 ft
D = 4.5/5 ft to 1"      radius = 6 ft
P = 1" to 4" (saplings) radius thru 2022 = 6 ft; then 16 ft


Radii
4" - 12" = 3.7 ft
12" to 4.5/5ft = 6 ft


plot data for seedlings 4" to 12" (species recorded as 01 in 2023)
folder = R\\slash-wall-vegetation\\old files but do not discard
2019 file = plot-data-2019.csv
2020 file = plot_calc_2020_untidy.csv
2021 file = plot_data_2021_untidy.csv
2022 file = plot_data_2022_untidy.csv
2023 files =
        slash.wall.data.2023.23july2024(maybe FINAL)
        slash.wall.data.2023_1_feb_2024_regen(north, south, rectangle, gasline)_final
2024 files sheet = regen, filename = "slash.wall.data_2024_FINAL"



input point attributes, "point_label_all" this is the new data file of attributes by point

cafri overstory data and cafri point labels
```{r}
cafri_point_residual_overstory <- read_excel("CAFRI Treatment & Plot Location 8-2024.xlsx", sheet = "merged", skip = 6) %>% 
    select(stand, wall, overstory, point) %>% 
    mutate(
        wall = as.character(wall),
        point = as.character(point)
    ) %>% 
    rename(
        harvest = stand,
        residual = overstory
    )

cafri_point_labels <- read_excel("CAFRI Treatment & Plot Location 8-2024.xlsx", sheet = "point counts", skip = 4) %>% 
    rename(
        harvest = stand,
        trmt = status
        ) %>% 
    mutate(
        wall = as.character(wall),
        point = as.character(point),
        location = trmt
    ) %>% 
    mutate(
    location = case_when(
      location == "control" ~ "c",
      location == "protected" ~ "p",
      TRUE ~ location  # Keeps other values unchanged
    )
  ) %>% 
    left_join(cafri_point_residual_overstory, by = c("harvest", "wall", "point" ))

# the point_labels are updated to include state forest sites on about line 1712, df=new_sites
# the excel file of point_labels was updated for sheet = "all_sites". that is imported here and
# subsequent rewriting of the excel file is halted.

point_labels_revision <- read_excel(file_path_points, sheet = "all_sites", skip = 0) # %>% 
    # arrange(harvest, wall, trmt, point) %>% 
    # left_join(cafri_point_labels, by = c("harvest", "wall", "point", "trmt", "height", "location"))
```


```{r eval=FALSE}


output_file <- file.path(output_folder, paste0("point_labels_all", ".xlsx"))
write_xlsx(point_labels_revision, output_file)



```

import the revised point label data file
```{r}

point_labels <- read_excel(file_path_points, sheet = "all_sites", skip = 0)
# the sheet="all_sites" includes the state forest sites. 
# there is potential within this code for circularity such that the df=point_labels
# here includes residuals, but that point_labels_residuals is created within this code.


```



2019 vegetation data
```{r eval=T}

raw2019 <- read_excel(file_path_2019, sheet = "seedling", skip = 0 ) |>
    select(!c(top4, other, radius.01, radius.other)) |> 
    mutate(
        season = "2019",
        sup.seed03 = as.numeric(sup.seed03),
        sup.sapl = as.numeric(sup.sapl),
        point = as.character(point)
    ) |>
    mutate(origin = case_when(
       origin %in% c("s","u") ~ "single",
       origin %in% c("t", "r") ~ "clump",
       TRUE ~ origin  # Keeps other values unchanged
  )) |> 
    select(harvest, season, point, location, spp, origin, everything()) %>% 
    mutate(
    harvest = case_when(
      harvest == "cgl" ~ "gasline",
      harvest == "crp" ~ "redpine",
      harvest == "cwb" ~ "stationrd",
      harvest == "gl" ~ "gasline",
      harvest == "rp" ~ "redpine",
      TRUE ~ harvest  # Keeps other values unchanged
    )
  ) %>% 
    mutate(
    location = case_when(
      location == "control" ~ "c",
      location == "interior" ~ "i",
      location == "perimeter" ~ "p",
      TRUE ~ harvest  # Keeps other values unchanged
    )
  )

  
# clumps were not separated on the tally sheet. 
# It isn't possible to  the number of clumps per acre
# It would be possible to represent the % of total stems present in clumps 
# (e.g.) 40% of stem density is of clump origin
# 

regen2019 <- raw2019 %>% 
    mutate(across(where(is.numeric), ~replace_na(., 0))) %>% 
    mutate(across(starts_with("su"), ~replace_na(., 0))) %>%    # Replace NA with 0 for "su" variables, the "supressed" variables
    mutate(across(starts_with("ex"), ~replace_na(., 0))) %>%       # Replace NA with 0 for "ex" variables, the "exposed" variables
    mutate(harvest = case_when(
        harvest %in% c("boot", "wedge", "stationrd") ~ "sta_rd",
        harvest %in% c("gasline") ~ "gas_line",
        harvest %in% c("redpine") ~ "red_pine",
        TRUE ~ harvest, # keeps the existing value of "harvest" if none of the cases are true
    )) %>% 
        left_join(point_labels, by = c("harvest", "point", "location")) %>% 
    select(-wall, -trmt, -height, -harvest_original, -forest_type, -brush) %>% 
      mutate(across(
        where(is.numeric),  # Select all numeric columns
        ~ if_else(is.na(spp), 0, .x)  # Set to 0 when spp is NA, otherwise keep original values
  ))


 
```
 

2020 data
```{r}
raw2020 <- read_excel(file_path_2020, sheet = "regen", skip = 0 ) |>
    select(! c(top4, Comments)) |> 
    mutate(
        season = as.character(season),
        spp = as.character(spp),
        sup.seed03 = as.numeric(sup.seed03),
        sup.sapl = as.numeric(sup.sapl)
    ) |>
    mutate(origin = case_when(
       origin %in% c("s","u") ~ "single",
       origin %in% c("t", "r") ~ "clump",
       TRUE ~ origin  # Keeps other values unchanged
  )) |> 
    select(harvest, season, point, location, spp, origin, everything())

  
# clumps were not separated on the tally sheet. 
# It isn't possible to  the number of clumps per acre
# It would be possible to represent the % of total stems present in clumps 
# (e.g.) 40% of stem density is of clump origin
# 

regen2020 <- raw2020 %>% 
    mutate(point = as.character(point)) %>% 
    mutate(across(where(is.numeric), ~replace_na(., 0))) %>% 
    mutate(across(starts_with("su"), ~replace_na(., 0))) %>%    # Replace NA with 0 for "su" variables
    mutate(across(starts_with("ex"), ~replace_na(., 0)))      # Replace NA with 0 for "ex" variables


```

2021 data
```{r}
raw2021 <- read_excel(file_path_2021, sheet = "regen", skip = 2 ) |>
    select(! c(top4)) |> 
    mutate(
        season = as.character(season),
        location = as.character(location),
        spp = as.character(spp)
   ) |>
    mutate(origin = case_when(
       origin %in% c("s","u") ~ "single",
       origin %in% c("v", "r") ~ "clump",
       TRUE ~ origin  # Keeps other values unchanged
  )) |> 
    mutate(
        sup.seed03 = as.numeric(sup.seed03),
        sup.sapl = as.numeric(sup.sapl)
    ) %>% 
    select(harvest, season, point, location, spp, origin, everything()) 
  
  
# clumps were not separated on the tally sheet. 
# It isn't possible to  the number of clumps per acre
# It would be possible to represent the % of total stems present in clumps 
# (e.g.) 40% of stem density is of clump origin
# 

regen2021 <- raw2021 %>% 
    mutate(
        location = as.character(location),
        point = as.character(point)
    ) %>% 
     mutate(across(where(is.numeric), ~replace_na(., 0))) %>% 
   mutate(across(starts_with("su"), ~replace_na(., 0))) %>%    # Replace NA with 0 for "su" variables
    mutate(across(starts_with("ex"), ~replace_na(., 0)))      # Replace NA with 0 for "ex" variables


```




2022 data
```{r message=FALSE, warning=FALSE}
raw2022 <- read_excel(file_path_2022, sheet = "regen", skip = 0 ) |>
    select(! c(top4, comments)) |> 
    select(-wall) %>% 
    mutate(
        season = as.character(season),
        location = as.character(location),
        spp = as.character(spp),
        point = as.character(point)
    ) |>
    mutate(origin = case_when(
       origin %in% c("s","u") ~ "single",
       origin %in% c("v", "r") ~ "clump",
       TRUE ~ origin  # Keeps other values unchanged
  )) |> 
    mutate(        
        sup.seed02 = as.numeric(sup.seed03),
        sup.seed03 = as.numeric(sup.seed03),
        sup.sapl = as.numeric(sup.sapl)
    ) %>% 
    select(harvest, season, point, location, spp, origin, everything()) 
  
# clumps were not separated on the tally sheet. 
# It isn't possible to  the number of clumps per acre
# It would be possible to represent the % of total stems present in clumps 
# (e.g.) 40% of stem density is of clump origin
# 

regen2022 <- raw2022 %>% 
    mutate(
        location = as.character(location)
    ) %>% 
    mutate(across(where(is.numeric), ~replace_na(., 0))) %>% 
    mutate(across(starts_with("su"), ~replace_na(., 0))) %>%    # Replace NA with 0 for "su" variables
    mutate(across(starts_with("ex"), ~replace_na(., 0)))      # Replace NA with 0 for "ex" variables


```


# origin_test <- data01 %>% 
#     group_by(stand, origin) |> 
#     summarize(
#     test_clump = mean(count)  
     
# this code should produce the average number of stems per clump in each stand. 
#Note the older harvests have a lower average number

veg2023_peracre_calc <- data01 %>% 
    mutate(
        multip01 = 43560 / (pi * (3.7^2)),  # = 1012.79
        multip02_03 = 43560 / (pi * (6^2)), # = 43560/113.1 = 385.15
        multip_sapl  = 43560 / (pi * (16^2)) # = 43560 / 256 = 170.1
    ) |> 


2023 veg data
```{r warning=FALSE}
# data are for counts of stems per point, not per acre
# 
#regen2023 <- read_excel(file_path_2023_veg, sheet = "Sheet1", skip = 0 ) %>% 
regen2023 <- read_excel(file_path_2023_veg_22july2025, sheet = "Sheet1", skip = 0) %>% 
    rename(
        supA = sup.seed01, #confirmed 7/24/2025 01 = 4" to 12"
        supB = sup.seed02, # 12" to 5 ft
        supD = sup.seed03, # 5 ft to 1"
        supP = sup.sapl, #1" to 5"
        expA = exp.seed01,
        expB = exp.seed02,
        expD = exp.seed03,
        expP = exp.sapl
        ) %>% 
    
#TODO need to double check this calculation #DONE, calculation confirmed 2/13/2025    
    mutate(
    numb.per.clump = if_else(origin == "clump",  
                     rowSums(across(starts_with("sup") | starts_with("exp"), 
                            ~ if_else(. > 0, ., NA_real_)), na.rm = TRUE) * count, #within a row, assess if >0, then calculates mean of clump size 
                          NA_real_)
   ) %>% 
    
    #Next, convert stem counts to "per acre" values
    mutate(
        multip01 = 43560 / (pi * (3.7^2)),  # = 1012.79
        multip02_03 = 43560 / (pi * (6^2)), # = 43560/113.1 = 385.15
        multip_sapl  = 43560 / (pi * (16^2)), # = 43560 / 256 = 170.1
        wall = as.character(wall)
    ) %>% 
    
    mutate(
        supA = supA * count * multip01,
        supB = supB * count * multip02_03,
        supD = supD * count * multip02_03,
        supP = supP * count * multip_sapl,
        expA = expA * count * multip01,
        expB = expB * count * multip02_03,
        expD = expD * count * multip02_03,
        expP = expP * count * multip_sapl
    ) %>% 
    mutate(
        type = "blank"
    ) %>% 
    mutate(type = case_when(
    spp %in% c("ba", "bc", "bih", "bl", "co", "cu", "he", "ohw", "rm", "ro", "sa", 
                "sh", "sm", "wa", "wo", "wp", "yb", "yp",
               "951", "762", "402", "901", "832", "651", "261", "004", "4", "316", "833", "931", #004 same as 4
                "407", "318", "541", "802", "129", "371", "621") ~ "commercial",
    
    spp %in% c("sb", "372") ~ "sb",

    spp %in% c("ab", "eh", "pc", "stm", "toh", "au",
               "531", "701", "761", "315", "998", "535") ~ "interfere",

    spp %in% c("app", "bta", "haw", "brs", "dws", "ec", "la", "ns",
               "pb", "gb", "qa", "rp", "scp", "smc", "svb", "wh", "wi",
               "660", "743", "500", "brs", "dws", "742", "70", "71", "91",  # 70 same as 71
               "375", "379", "746", "125", "130", "866", "355", "356", "585", "920") ~ "diversity", # 355 same as 356

    spp %in% c("mia", "ah", "cc", "unk", "sp",
                "001", "1", "391", "763", "999", "90") ~ "other",  # 001 same as 1
    TRUE ~ type,  # Keeps the existing value of "type" for all other cases
    )) %>%  
    select(-df, -filename) %>% 
    select(harvest, wall, season, date, trmt, location, point, spp, type, origin, count, everything())



regen2023seedl <- regen2023 %>%  # extract the data for the 4" to 12" seedlings to merge into the site data permanent file
# data are counts of stems per acre by point
    filter(type == "commercial") %>% 
    mutate(
        comm_sdl = supA + expA
    ) %>% 
    select(harvest, wall, point, season, type, spp, comm_sdl) %>% 
    group_by(harvest, point) %>% 
    summarize(
       season = first(season),
       wall = first(wall),
       comm_sdl = sum(comm_sdl),
       seedA = sum(comm_sdl),
       richness = n_distinct(spp),
       .groups = "drop"
    )

# (july 24 2025) in the field data read into df=regen2023, the 4" to 12" class included all species
# in df = regen2023seedl, the data are sorted first for commercial species and
# seedA is the representation of comm_sdl on a per acre basis. The same as for regen 2019 throu regen2022

```


2024 vegetation
```{r warning=FALSE}

# data values i input file are count of stems in each plot (not per acre)
regen2024 <- read_excel(file_path_2024, sheet = "regen", skip = 2 ) %>% 
    select(-comments) %>% 
    filter(!season < 2024) %>% 
    rename(
        harvest = stand
    ) %>% 
    mutate(
        count = as.numeric(gsub("[^0-9]", "", count))
    ) %>%  #removes any non-numeric characters and changes count from "char" to "num"
      mutate(point = str_replace(point, "^0+", "")) %>%  #from the start (^), match one or more zeros (0+) and replace with ""

    mutate(
        spp = as.character(spp),
        season = as.character(season),
        point = as.character(point),
        sup.sapl = as.numeric(sup.sapl)
    ) %>% 

    replace_na(list(date = "other", count = 1,
        sup.seed01 = 0, sup.seed02 = 0, sup.seed03 = 0, sup.sapl = 0,
        exp.seed01 = 0, exp.seed02 = 0, exp.seed03 = 0, exp.sapl = 0
    ))  %>% 

    rename(
        supA = sup.seed01, # 4 to 12", radius = 3.7 ft
        supB = sup.seed02, # 12 to 60", radius = 6 ft
        supD = sup.seed03, # 60" to 1"dbh, radius = 6 ft  #elsewhere, label "C" is for stems 60" to 9 ft
        supP = sup.sapl, # 1" to 5" dbh, radius = 16 ft
        expA = exp.seed01,
        expB = exp.seed02,
        expD = exp.seed03,
        expP = exp.sapl
    ) %>%   #  need to check this code, added on 4/16/2025 but not tested

    mutate(
      origin = case_when(
      count < 2 ~ "single",
      count >= 2 ~ "clump"
      )) %>% 
    
#number.per.clump calculation confirmed 2/13/2025    
    mutate(
    numb.per.clump = if_else(origin == "clump",  
                          rowSums(across(starts_with("sup") | starts_with("exp"), 
                                          ~ if_else(. > 0, ., NA_real_)), na.rm = TRUE) * count, #within a row, assess if >0, then calculates mean of clump size 
                          NA_real_)
   ) %>%
    
#Next, convert stem counts to "per acre" values
    mutate(
        multip01 = 43560 / (pi * (3.7^2)),  # = 1012.79
        multip02_03 = 43560 / (pi * (6^2)), # = 43560/113.1 = 385.15
        multip_sapl  = 43560 / (pi * (16^2)), # = 43560 / 256 = 170.1
        wall = as.character(wall)
    ) %>% 
    
    mutate(
        supA = supA * count * multip01,
        supB = supB * count * multip02_03,
        supD = supD * count * multip02_03,
        supP = supP * count * multip_sapl,
        expA = expA * count * multip01,
        expB = expB * count * multip02_03,
        expD = expD * count * multip02_03,
        expP = expP * count * multip_sapl
    ) %>% 
    mutate(
        type = "blank"
    ) %>% 
    mutate(type = case_when(
    spp %in% c("ba", "bc", "bih", "bl", "co", "cu", "he", "ohw", "rm", "ro", "sa", 
                "sh", "sm", "wa", "wo", "wp", "yb", "yp",
               "951", "762", "402", "901", "832", "651", "261", "004", "4", "316", "833", "931", #004 same as 4
                "407", "318", "541", "802", "129", "371", "621") ~ "commercial",
    
    spp %in% c("sb", "372") ~ "sb",

    spp %in% c("ab", "eh", "pc", "stm", "toh", "au",
               "531", "701", "761", "315", "998", "535") ~ "interfere",

    spp %in% c("app", "bta", "haw", "brs", "dws", "ec", "la", "ns",
               "pb", "gb", "qa", "rp", "scp", "smc", "svb", "wh", "wi",
               "660", "743", "500", "brs", "dws", "742", "70", "71", "91",  # 70 same as 71
               "375", "379", "746", "125", "130", "866", "355", "356", "585", "920") ~ "diversity", # 355 same as 356

    spp %in% c("mia", "ah", "cc", "unk", "sp",
                "001", "1", "391", "763", "999", "90") ~ "other",  # 001 same as 1 = empty plot = "mia"
    TRUE ~ type,  # Keeps the existing value of "type" for all other cases
    )) %>%  

      mutate(harvest = case_when(
      harvest == "boot" ~ "sta_rd",
      harvest == "decker control" ~ "decker",
      harvest == "recknagel south" ~ "recknagel",
      harvest == "wedge" ~ "sta_rd",
      harvest == "wedge control" ~ "sta_rd",
      harvest == "stand_5.1" ~ "5_1",
      TRUE ~ harvest  # Keeps other values unchanged
    )
  )  

# create this df to merge its seedling data (i.e., stems 4"to 12") with the plot attribute data. See line 785
# or search for the df name = site_data
regen2024seedl <- regen2024 %>% 
    filter(type == "commercial") %>% 
    mutate(
        comm_sdl = supA + expA
    ) %>% 
    select(harvest, wall, point, season, type, spp, comm_sdl) %>% 
    group_by(harvest, point) %>% 
    summarize(
       season = first(season),
       wall = first(wall),
       comm_sdl = sum(comm_sdl),
       seedA = sum(comm_sdl),
       richness = n_distinct(spp),
       .groups = "drop"
    ) %>% 
mutate(
        wall = case_when(
        harvest %in% c("danby_block_2", "fairfield_a5", "fairfield_a6", "goundry_a17", "5_1") ~ "99",
        TRUE ~ wall
  )
)
```


###  site/plot data read and manipulate

plot 2019
```{r}
plot2019 <- read.csv(file_path_2019_plot) %>% 
   mutate(
    location = case_when(
      location == "control" ~ "c",
      location == "interior" ~ "i",
      location == "perimeter" ~ "p",
      TRUE ~ location  # Keeps other values unchanged
    )
  ) %>% 
    mutate(
        comm_sdl = as.numeric(comm_sdl),
        seedA = comm_sdl,
        seedA = if_else(is.na(seedA), 0, seedA ),
        rhizom = fern,
        season = as.character(season),
        upper.spp = as.character(upper.spp),
        maple = as.integer(maple),
        point = as.character(point)
    ) %>% 
    mutate(across(c(comm_sdl), ~ replace_na(., 0))) %>% 
    rename(
        native_cov = low.cov,
        native_spp = low.spp,
        invasive_spp = invasive.spp,
        invasive_cov = invasive.cov,
        mid_cov = mid.cov,
        mid_spp = mid.spp,
        upper_cov = upper.cov,
        upper_spp = upper.spp,
        vine_cov = vine.cov,
        vine_spp = vine.spp
    ) %>% 
   mutate(seedA = seedA *(43560 / (pi * 3.7 * 3.7))) %>% 
    select(harvest,  season,  date,  location,  point,  disturb,  slash,  fern,  herb,  native_cov,  
           native_spp,  invasive_cov,  invasive_spp,  mid_cov,  mid_spp,  upper_cov,  upper_spp,  vine_cov,  
           vine_spp,  baf,  oakhick,  beech,  maple,  birch,  hdwd,  conifer,  comm_sdl,    seedA )    # ba_acre, cut,  year,  numb_yrs_grow,
 #   filter( point == "338" & season == "2019")
```


plot 2020
```{r}
plot2020 <- read.csv(file_path_2020_plot) %>%  # this has data for boot, wedge and sta_rd separately presented
    mutate(across(c(comm_sdl), ~ replace_na(., 0))) %>% 
    mutate(
        point = str_remove(point, "^0+"),
        point = as.numeric(point)
        ) %>% 
    mutate(
        season = as.character(season),
        rhizom = fern,
        mid_spp = as.character(mid_spp),
        upper_spp = as.character(upper_spp),
        point = as.numeric(point),
        point = as.character(point),
        seedA = case_when(
          comm_sdl == 0 ~ 0,  # If category is 0, the random value is 0
          comm_sdl == 1 ~ sample(1:5, size = n(), replace = TRUE),  # Random number from 1 to 5 for each row
          comm_sdl == 2 ~ sample(6:15, size = n(), replace = TRUE), # Random number from 6 to 15 for each row
          comm_sdl == 3 ~ sample(16:30, size = n(), replace = TRUE), # Random number from 16 to 30 for each row
          comm_sdl == 4 ~ sample(31:50, size = n(), replace = TRUE) # Random number from 31 to 100 for each row
    )
  ) %>% 
   mutate(seedA = seedA *(43560 / (pi * 3.7 * 3.7))) %>% 
        select(harvest,  season,  date,  location,  point,  disturb,  slash,  fern,  herb,  native_cov,  
           native_spp,  invasive_cov,  invasive_spp,  mid_cov,  mid_spp,  upper_cov,  upper_spp,  vine_cov,  
           vine_spp,  baf,  oakhick,  beech,  maple,  birch,  hdwd,  conifer,  comm_sdl,    seedA ) # ba_acre, cut,  year,  numb_yrs_grow,

# for 2020, comm_sdl is a categorical value from 0 to 4.
# seedA is initially a value based on random assignment of comm_sdl within a bracket,
# and subsequently expanded from milacre to acre. 

```


plot 2021
```{r}
plot2021 <- read_excel(file_path_2021_plot, sheet = "site-plot", skip = 6) %>%  
    mutate(across(c(comm_sdl), ~ replace_na(., 0))) %>% 
    mutate(
    season = as.character(season),
    rhizom = fern,
    mid_spp = as.character(mid_spp),
    upper_spp = as.character(upper_spp),
    invasive_spp = as.character(invasive_spp),
    point = as.character(point),
    seedA = case_when(
      comm_sdl == 0 ~ 0,  # If category is 0, the random value is 0
      comm_sdl == 1 ~ sample(1:5, size = n(), replace = TRUE),  # Random number from 1 to 5 for each row
      comm_sdl == 2 ~ sample(6:15, size = n(), replace = TRUE), # Random number from 6 to 15 for each row
      comm_sdl == 3 ~ sample(16:30, size = n(), replace = TRUE), # Random number from 16 to 30 for each row
      comm_sdl == 4 ~ sample(31:50, size = n(), replace = TRUE) # Random number from 31 to 100 for each row
    )
  ) %>% 
   mutate(seedA = seedA *(43560 / (pi * 3.7 * 3.7))) %>% 
        select(harvest,  season,  date,  location,  point,  disturb,  slash,  fern,  herb,  native_cov,  
           native_spp,  invasive_cov,  invasive_spp,  mid_cov,  mid_spp,  upper_cov,  upper_spp,  vine_cov,  
           vine_spp,  baf,  oakhick,  beech,  maple,  birch,  hdwd,  conifer,  comm_sdl,    seedA ) # ba_acre, cut,  year,  numb_yrs_grow,

# for 2021, comm_sdl is a categorical value from 0 to 4.
# seedA is initially a value based on random assignment of comm_sdl within a bracket,
# and subsequently expanded from milacre to acre. 

```


plot 2022
```{r}
plot2022_original <- read_excel(file_path_2022, sheet = "site-plot", skip = 6) %>% 
    mutate(
        point = as.character(point)
    )

plot2022  <- read_excel(file_path_2022, sheet = "site-traits", skip = 1) %>% #read in the cafri plot traits
    mutate(
        point = as.character(point)
    ) %>% 
    bind_rows(plot2022_original) %>% 
    mutate(across(c(comm_sdl), ~ replace_na(., 0))) %>% 
    mutate(
    season = as.character(season),
    rhizom = fern,
    point = as.character(point),
    seedA = case_when(
      comm_sdl == 0 ~ 0,  # If category is 0, the random value is 0
      comm_sdl == 1 ~ sample(1:5, size = n(), replace = TRUE),  # Random number from 1 to 5 for each row
      comm_sdl == 2 ~ sample(6:15, size = n(), replace = TRUE), # Random number from 6 to 15 for each row
      comm_sdl == 3 ~ sample(16:30, size = n(), replace = TRUE), # Random number from 16 to 30 for each row
      comm_sdl == 4 ~ sample(31:50, size = n(), replace = TRUE) # Random number from 31 to 100 for each row
    )
# for 2022, comm_sdl is a categorical value from 0 to 4.
# seedA is initially a value based on random assignment of comm_sdl within a bracket,
# and subsequently expanded from milacre to acre. 

  ) %>% 
   mutate(seedA = seedA * (43560 / (pi * 3.7 * 3.7))) %>% 
   select(harvest,  season,  date,  location,  point,  comm_sdl,    seedA ) # ba_acre, cut,  year,  numb_yrs_grow,
```


2023 plot and site CAFRI and other (includes CAFRI 2023 dbh with distances)
```{r warning=FALSE}

# note 5may2025 - the data from the 2023 overstory sampling including distance to tree and status as
# marginal needs to be amended to the worksheet "2024 overstory data."  Once amended, the overstory data needs to be joined into the df=plot2024
# before it can be joined into df=site_data.
# 
# UPDATE 17july2025 the dbh data for cafri 2023 were adjusted to include distance, a calculation of limiting distance, and status as "in" or "out"

# see also plot_basal_area_calc_summ2025.Rmd
# and output file = plot_basal_area_calc_summ2025.Rmd



plot2023cafri <- read_excel(file_path_2023_plot, sheet = "cafri_site", skip = 3) %>% 
   select(-comments) %>% 
    rename(harvest = stand) %>% 
    mutate(
        wall = as.character(wall),
        season = as.character(season),
        invasive_spp = as.character(invasive_spp),
        mid_spp = as.character(mid_spp),
        point = as.character(point)
    ) %>% 
   full_join(cafri_point_labels, by = c("harvest", "wall", "point"), suffix = c("_cafri", "_labels")) %>% 
    arrange(harvest, point)


cafri2023dbh <- read_excel(file_path_2023_cafri_overstory, sheet = "dbh2023cafri", skip = 5) %>% 
    mutate(
        baf = 9.1,
        limit_distance_2023 = 2.75 * dbh2023,
        status = "in",
        basalarea = baf,
        stem_count = 43560 / (pi * (limit_distance_2023^2)), 
        wall = as.character(wall),
        point = as.character(point)
    ) %>% 
   mutate(
    status = case_when(
      distance > limit_distance_2023 ~ "out", # "distance" is the measured distance in feet from plot center to tree
      TRUE ~ status  # Keeps other values unchanged
    )
  ) %>% 
    arrange(harvest, point, tag) %>% 
    mutate(
        file = "slash_veg_complete.Rmd",
        df = "cafri2023dbh"
    ) %>% 
    select(file, df, harvest, baf, wall, point, spp, tag, status, dbh2023, distance, limit_distance_2023, basalarea, stem_count )

cafri2023_structure_point_summary <- cafri2023dbh %>% 
    filter(
        status == "in"
    ) %>% 
    group_by(harvest, wall, point) %>% 
    summarize(
        ba_acre = sum(basalarea),
        stems_acre = sum(stem_count),
        x_dbh = mean(dbh2023),
        .groups = "drop"
    ) %>% 
    left_join(point_labels_revision, by = c("harvest", "wall", "point")) %>% 
    mutate(
        file = "slash_veg_complete.Rmd",
        df = "cafri2023_structure_point_summary"
    ) %>% 
    select(file, df, harvest, wall, point, ba_acre, stems_acre, x_dbh, trmt, residual, height, brush ) %>% 
    mutate_if(is.numeric, round, 0)

point_labels_unique <- point_labels_revision %>%
  group_by(harvest, wall, brush) %>%
  summarize(
      brush = first(brush),
      trmt = first(trmt),
      height = first(height),
      residual = first(residual),
      .groups = "drop"
      )

cafri2023_structure_wall_summary <- cafri2023_structure_point_summary %>% # use the df = point summary which is filtered for "in" and then average by point
    left_join(point_labels_unique, by = c("harvest", "wall", "trmt", "residual", "height", "brush")) %>% 
    select(harvest, wall, trmt, height, x_dbh, brush, stems_acre, ba_acre, residual) %>% 
    group_by(harvest, wall,brush) %>% 
    summarize(
        ba_acre_2023 = mean(ba_acre),
        stems_acre_2023 = mean(stems_acre),
        x_dbh = mean(x_dbh),
        trmt = first(trmt),
        height = first(height),
        residual = first(residual),
        .groups = "drop"
    ) %>% 
    rename(
        residual_target = residual,
        overstory_TPA = stems_acre_2023
    ) %>% 
    mutate(
        file = "slash_veg_complete.Rmd",
        df = "cafri2023_structure_wall_summary"
    ) %>% 
    select(file, df, harvest, wall, trmt, height, brush, overstory_TPA, ba_acre_2023, x_dbh, residual_target) %>% 
    mutate_if(is.numeric, round, 0)

cafri2023_summary <- cafri2023_structure_wall_summary %>% 
    bind_rows(cafri2023_structure_point_summary, cafri2023dbh)

output_file <- file.path(output_folder, paste0("cafri2023_dbh_basalarea_raw_summ", ".xlsx"))
write_xlsx(cafri2023_summary, output_file)



plot2023other <- read_excel(file_path_2023_plot, sheet = "other_site_baf10", skip = 2) %>% 
    select(-comments) %>% 
    rename(harvest = stand) %>% 
    mutate(
        spp = as.character(spp),
        wall = as.character(wall),
        season = as.character(season),
        invasive_spp = as.character(invasive_spp),
        mid_spp = as.character(mid_spp),
        point = as.character(point)
    ) %>% 
    mutate(wall = if_else(is.na(wall), "99", wall)    )

plot2023_overstory <- plot2023other %>% 
    select(harvest, wall, location, season, date, point, spp, dbh, in16ft, status,crown, logs)

output_file <- file.path(output_folder, paste0("campridge_overstory_by_point_2023data(29may2025)", ".xlsx"))
write_xlsx(plot2023_overstory, output_file)


plot2023_site <- plot2023other %>% 
    select(harvest, wall, location, season, date, point, rhizom,	clump,	grass,
           herb,	herb_spp,	native_cov,	native_spp,	mid_cov,	mid_spp,	invasive_cov,	invasive_spp,	vine_cov,	vine_spp) %>% 
    #the data are read in twice, so the code "distinct" retains one unique (distinct) version of each point.
    #I'm not sure what decision criteria would be used if the rows varied.
    distinct(point, .keep_all = TRUE)



basalarea2023 <- plot2023_overstory %>% 
    select(harvest, wall, point, spp, dbh, season) %>% 
    mutate(
        ba = (dbh/dbh) * 9.1,  #note the angle gauge BAF = 9.1
        stem_count = 43560 / (pi * ((2.75 * dbh)^2))
    ) %>% 
    group_by(harvest, wall, point) %>% 
    summarize(
        season = first(season),
        ba_acre = sum(ba),
        overstory_stems_acre = sum(stem_count),
        .groups = "drop"
    ) %>% 
    mutate(across(c(ba_acre, overstory_stems_acre), ~ replace_na(.,0))) %>% 
    bind_rows(cafri2023_structure_point_summary) %>% 
    select(-file, -df, -trmt, -x_dbh, -residual, -height, -brush) %>% 
    mutate(
        season = "2023"        
    )

    
#create plot2023 to include seedlings less than 12" and the basal area per acre
  
plot2023 <- bind_rows(plot2023cafri, plot2023_site) %>% 
    mutate(
        point = as.character(point)
    ) %>% 
    left_join(regen2023seedl, by = c("harvest", "season", "wall", "point")) %>% 
        #regen2023seedl has the seedling 4" - 12" data as "seedA" from the vegetation file.
        #
    left_join(basalarea2023, by = c("harvest", "wall", "point", "season")) %>% 
    rename(
        basalarea = ba_acre
    ) %>% 
    select(harvest, wall, season, point, basalarea, everything()) %>% 
    select(-date)
```

2024 plot
```{r}
plot2024_orig <- read_excel(file_path_2024, sheet = "site data", skip = 3 ) %>% 
    select(-location, -wall, -comments) %>% 
    mutate(
        season = "2024"
    ) %>% 
      mutate(point = str_replace(point, "^0+", "")) %>%  
        # string replacement (str_replace) from the start (^), match one or more zeros (0+) and replace with ""

# harvest_original = harvest, forest_type=n_hdwd or oak_hdwd
# this df only has basal area values (not original data) for the state forest land sites
       
    mutate(harvest = case_when(
      harvest == "boot" ~ "sta_rd",  #this changes "harvest" from boot to sta_rd
      harvest == "decker control" ~ "decker",
      harvest == "recknagel south" ~ "recknagel",
      harvest == "wedge" ~ "sta_rd",
      harvest == "wedge control" ~ "sta_rd",
      harvest == "rectangle" ~ "recknagel",
    TRUE ~ harvest  # Keeps other values unchanged
    )
  ) %>% 

left_join(point_labels, by = c("harvest", "point")) %>% 
    mutate(
        height = case_when(
        harvest %in% c("danby_block_2", "fairfield_a5", "fairfield_a6", "goundry_a17", "5_1") ~ 0,
        TRUE ~ height
  )
) %>% 
    mutate(
        trmt = case_when(
        harvest %in% c("danby_block_2", "fairfield_a5", "fairfield_a6", "goundry_a17", "decker", "rectangle", "5_1") ~ "control",
        TRUE ~ trmt
  )
) %>% 
    mutate(
        location = case_when(
        harvest %in% c("danby_block_2", "fairfield_a5", "fairfield_a6", "goundry_a17", "5_1") ~ "c",
        TRUE ~ location
  )
) %>% 
    
    mutate(
        wall = case_when(
        harvest %in% c("danby_block_2", "fairfield_a5", "fairfield_a6", "goundry_a17", "5_1") ~ "99",
        TRUE ~ wall
  )
) %>% 

    mutate(
        brush = case_when(
        harvest %in% c("fairfield_a5") ~ "b",
        harvest %in% c("danby_block_2",  "fairfield_a6", "goundry_a17", "5_1") ~ "u",
        TRUE ~ brush
  )
) %>% 
        
    mutate(
        forest_type = case_when(
        harvest %in% c("danby_block_2", "fairfield_a5", "5_1") ~ "oak_hdwd",
        harvest %in% c("fairfield_a6", "goundry_a17") ~ "n_hdwd",
        TRUE ~ forest_type
  )
) %>% 

    mutate(
        harvest_original = case_when(
        harvest %in% c("danby_block_2", "fairfield_a5", "fairfield_a6", "goundry_a17", "5_1") ~ harvest,
        TRUE ~ harvest_original
  )
)



#merge the seedlings 4" to 12" within df=regen2024seedl with site attribute data to create df=plot2024
# on line number 505, only type = "commercial" were retained
plot2024 <- left_join(plot2024_orig, regen2024seedl, by = c("harvest", "point", "wall", "season")) %>%  
#    filter(!harvest == "stand_5.1") %>% 
    mutate(harvest = case_when(
      harvest == "boot" ~ "sta_rd",  # old ~ new
      harvest == "recknagel south" ~ "recknagel",
      harvest == "wedge" ~ "sta_rd",
      harvest == "wedge control" ~ "sta_rd",
      harvest =="stand_5.1" ~ "5_1",
    TRUE ~ harvest  # Keeps other values unchanged
    )
  ) %>% 
    mutate(seedA = comm_sdl)


```


To standardize labels the strata will be denoted as:
A = 4" to 12" (these are not sorted as suppressed vs. exposed and spp is "004" = other hardwood)
B = 12" to 4.5/5 ft
C1 = 4.5 ft to 9 ft
C2 = 9ft to 1" dbh
D = 4.5/5 ft to 1" (the aggregate of C1 and C2)
P = 1" to 4" (saplings)


#Merge Site/Plot Data
```{r}
site_data_early <- bind_rows(plot2019, plot2020, plot2021, plot2022) %>% 
    select(-c(location)) %>% 
    mutate( 
      baf = 9.1,
      
      harvest = case_when(
      harvest == "b" ~ "sta_rd",
      harvest == "boot" ~ "sta_rd",
      harvest == "cgl" ~ "gas_line",
      harvest == "circle1" ~ "circle",
      harvest == "circle2" ~ "circle",
      harvest == "circle3" ~ "circle",
      harvest == "cricle1" ~ "circle",
      harvest == "decker control" ~ "decker",
      harvest == "crp" ~ "red_pine",
      harvest == "cwb" ~ "sta_rd",
      harvest == "gasline" ~ "gas_line",
      harvest == "gl" ~ "gas_line",
      harvest == "recksouth" ~ "recknagel",
      harvest == "rectangle" ~ "recknagel",
      harvest == "rect1" ~ "recknagel",
      harvest == "rect2" ~ "recknagel",
      harvest == "rect3" ~ "recknagel",
      harvest == "rect4" ~ "recknagel",
      harvest == "rectangle1" ~ "recknagel",
      harvest == "rectangle2" ~ "recknagel",
      harvest == "rectangle3" ~ "recknagel",
      harvest == "rectangle4" ~ "recknagel",
      harvest == "rp" ~ "red_pine",
      harvest == "redpine" ~ "red_pine",
      harvest == "stationrd" ~ "sta_rd",
      harvest == "w" ~ "sta_rd",
      harvest == "wb" ~ "sta_rd",
      harvest == "wedge" ~ "sta_rd",
      harvest == "old" ~ "new",
      TRUE ~ harvest  # Keeps other values unchanged
    )
  ) %>% 
    left_join(point_labels, by = c("harvest", "point")) %>% 
    mutate(across(c(oakhick, beech, maple, birch, hdwd, conifer), ~ replace_na(., 0))) %>% 
    mutate(across(c(baf), ~ replace_na(., 9.1))) %>% 
    mutate(
        basalarea = baf * (oakhick + beech + maple + birch + hdwd + conifer)
    ) %>% 
    select(-oakhick, -beech, -maple, -birch, -hdwd, -conifer, -baf)
    
site_data <- bind_rows(site_data_early, plot2023, plot2024) %>% 
    #
    #plot2019, plot2023 and plot2024 have actual counts of 4" to 12" seedlings. Variable "seedA"
    #is the value expanded to a "stems per acre" basis. plot2020 - plot2022 use a categorical variable
    #that is randomly expanded to a "stems per acre" basis as variable "seedA"
    #
    
    filter(!(harvest == "gas_line" & point %in% c(866, 864))) %>%   #corrected code selects for either point at the 
        # intersection of harvest and point
    
       select(harvest, wall, season, date, point, location, trmt, height, disturb, slash, fern, rhizom, clump,
           grass, herb, herb_spp, native_cov, native_spp, mid_cov, mid_spp, vine_cov, vine_spp, invasive_cov,
           invasive_spp, upper_cov, upper_spp, basalarea,
           comm_sdl, seedA) %>%
    arrange(harvest, season, wall, point) %>% 
    mutate(
            rhiz_fern = coalesce(fern, rhizom)
        ) %>% 
        select( -fern, -rhizom, ) %>% 
    mutate(
        filename = "slash_veg_complete.Rmd",
        df = "site_data"
        ) %>% 

    select(filename, df, harvest, wall, season, date, point, location, trmt, height, everything()) 


#   filter(season == "2023")
#    filter(wall == "98")

#write the aggregation of all site data to the output file
##change the name of the output file to today's date

output_file <- file.path(output_folder, paste0("site_data_2019_2024(22july2025)", ".xlsx"))

write_xlsx(site_data, output_file)


```


To standardize labels:
A = 4" to 12"           radius = 3.7 ft; count in 2019, 2023, and 2024 otherwise categorical 0 = 0; 1 = 1to5; 2 = 6 to 15; 3 = 16 to 30; 4 = 31+
B = 12" to 4.5/5 ft     radius = 6 ft  (size 01)
C1 = 4.5 ft to 9 ft     radius = 6 ft  (size 02)
C2 = 9ft to 1" dbh      radius = 6 ft  (size 03)
D = 4.5/5 ft to 1"      radius = 6 ft  (size 02 and 03)
P = 1" to 4" (saplings) radius thru 2022 = 6 ft; then 16 ft  (size sapl)



#merging regeneration data across years, select 4" to 12" seedlings from the site/plot data
create df =seedlings2019_2024

```{r}
seedlingspre2023 <- site_data %>% #from counts of seedlings onto site tally sheets
    filter (!season == "2023", !season == "2024" #remove season >= 2023
    ) %>% 
    select(harvest, wall, point, season, seedA) %>% 
    mutate(
        spp ="004"
    ) # %>% 
#    filter(harvest == "red_pine" & season == "2019")

seedlings2019_2024 <- bind_rows(regen2023seedl, regen2024seedl) %>% 
    #this is the count of commercial seedlings 4" to 12" from the 2023 and 2024 veg data
    bind_rows(seedlingspre2023) %>% #seedlings before 2023
    mutate(
        spp = "004",
        origin = "single"
    )%>% 
    select(harvest, wall, point, season, spp, seedA, richness, origin)
#    filter(harvest == "red_pine"  & season == "2019")
#    
#    (24july2025) for 2023 and 2024, seedA was collected in veg plots as "seed01". in 2023, only commercial species and sweet birch
#    were included. in 2024, all species were included.
```


#df=merge2019_2022_a

```{r}
#convert regen data to "per acre" for 2019 - 2022 before merging with 2023 and 2024
#regen2019-regen2022 data are stems counts per point
# regen2023 adn regen2024 are stem counts per acre at each point.

merge2019_2022_a <- bind_rows(regen2019, regen2020, regen2021, regen2022) %>% 
    filter(!(harvest == "gas_line" & point %in% c(866, 864)))  %>%  
        #correct - code selects for either point at the intersection of harvest and point
    select(harvest, season, location, point, origin, spp, date, sup.seed01, sup.seed02, sup.seed03, 
           sup.sapl, exp.seed01, exp.seed02, exp.seed03, exp.sapl) %>% 
#next, standardize the seedling labels for variance across years
        rename(
        supB = sup.seed01, #prior to 2023, 01 was 12" to 60"
        supC1 = sup.seed02, #60" to 9ft
        supC2 = sup.seed03, # 9 ft to 1" dbh
        supP = sup.sapl, # 1" to 5" dbh
        expB = exp.seed01,
        expC1 = exp.seed02,
        expC2 = exp.seed03,
        expP = exp.sapl
    ) %>% 
#next, standardize harvest names
    mutate(
    harvest = case_when(
      harvest == "b" ~ "sta_rd",
      harvest == "boot" ~ "sta_rd",
      harvest == "cgl" ~ "gas_line",
      harvest == "circle1" ~ "circle",
      harvest == "circle2" ~ "circle",
      harvest == "circle3" ~ "circle",
      harvest == "cricle1" ~ "circle",
      harvest == "crp" ~ "red_pine",
      harvest == "cwb" ~ "sta_rd",
      harvest == "gasline" ~ "gas_line",
      harvest == "gl" ~ "gas_line",
      harvest == "recksouth" ~ "recknagel",
      harvest == "rect1" ~ "recknagel",
      harvest == "rect2" ~ "recknagel",
      harvest == "rect3" ~ "recknagel",
      harvest == "rect4" ~ "recknagel",
      harvest == "rectangle1" ~ "recknagel",
      harvest == "rectangle2" ~ "recknagel",
      harvest == "rectangle3" ~ "recknagel",
      harvest == "rectangle4" ~ "recknagel",
      harvest == "rp" ~ "red_pine",
      harvest == "redpine" ~ "red_pine",
      harvest == "stationrd" ~ "sta_rd",
      harvest == "w" ~ "sta_rd",
      harvest == "wb" ~ "sta_rd",
      harvest == "wedge" ~ "sta_rd",
      harvest == "old" ~ "new",
      TRUE ~ harvest  # Keeps other values unchanged
    )
      )
```




Merge veg data, assign attributes, and convert to per acre values at each point (clump data not available for 2019 to 2022)
#df=merge2019_2022_b
```{r}
#Next, calculate per acre values for regen 2019 through 2022
#

merge2019_2022_b <- merge2019_2022_a %>% 
    select(harvest, season, point, origin, spp, everything()) %>% 

#Next, calculate the average # stems per ramet. 
#WAIT
#WAIT - clumps were not distinguished within a spp for 2019 - 2022.  
#All code for ramets, count and clumps will be commented out (20june 2025)

   #  mutate(
   #  stems.ramet = if_else(origin == "clump",  #TODO need to double check this calculation #DONE, calculation confirmed 2/13/2025
   #                        rowMeans(across(starts_with("sup") | starts_with("exp"), 
   #                         ~ if_else(. > 0, ., NA_real_)), na.rm = TRUE), #within a row, assess if >0, then calculates mean of clump size 
   #                        NA_real_)
   # ) %>%
    
#Next, convert plot counts to "per acre" values
   mutate( #converts plot count to count per acre
        # count = if_else(is.na(count), 1, count), # if count missing (it's a single) then make count = 1
        # stems.ramet = if_else(is.na(stems.ramet), 0, stems.ramet), # if variable is NA, change to 0, otherwise leave as is
        radius01 = 3.7,
        radius02 = 6,
        radius03 = 16,
        radius = if_else(season == "2023", radius03, radius02),  
          # Dynamically assigns radius to radius03 when season is 2023 and sets to radius02 if not 2023 (no 2023 data included as of 2/13/2025)
        supB = (supB * (43560 / (pi * radius02^2))),
        supC1 = (supC1 * (43560 / (pi * radius02^2))),
        supC2 = (supC2 * (43560 / (pi * radius02^2))),
        supP = (supP * (43560 / (pi * radius^2))),  # Uses the new "radius" variable here
        expB = (expB * (43560 / (pi * radius02^2))),
        expC1 = (expC1 * (43560 / (pi * radius02^2))),
        expC2 = (expC2 * (43560 / (pi * radius02^2))),
        expP = (expP * (43560 / (pi * radius^2)))   # Uses the new "radius" variable here
    ) %>%
    ungroup() %>% 
  mutate(  #if there is concern there are blank spaces in the grouping variable that cause grouping problems
    season = str_trim(season),
    harvest = str_trim(harvest),
    point = str_trim(point),
    spp = str_trim(spp),
    origin = str_trim(origin)
  ) %>%  #

#
# it appears the group_by summary happens in the next df.  I'll wait to join in "point_labels_revision until after the calculations 
    group_by(harvest, season, point, spp, origin) %>% 
    summarize(
        n = n(),
#        height = first(height),
#        wall = first(wall),
        location = first(location),
#        brush = first(brush),
#        trmt = first(trmt),
#        harvest_original = first(harvest_original),
#        forest_type = first(forest_type),
        supB = sum(supB),
        supC1 = sum(supC1),
        supC2 = sum(supC2),
        supP = sum(supP),
        expB = sum(expB),
        expC1 = sum(expC1), 
        expC2 = sum(expC2),
        expP = sum(expP),
       .groups = "drop"
    )
```



#Next, aggregate numeric values by harvest, season, point, spp, and origin. 

df=merge2019_2022_c
```{r}

merge2019_2022_c <- merge2019_2022_b %>% 
    
#this summarize is redundant with the same action from df=merge2019_2022_b
#the df doesn't change, so I'll leave it rather than potentially causing
#problems

        group_by(season, harvest, point, spp, origin) %>% 
        summarize(
            supB = sum(supB),
            supC1 = sum(supC1),
            supC2 = sum(supC2),
            supP = sum(supP),
            expB = sum(expB),
            expC1 = sum(expC1),
            expC2 = sum(expC2),
            expP = sum(expP),
            .groups = "drop"
        ) %>% 
    select(harvest, season,  point, spp, origin, everything()) %>% 
    arrange(harvest, season, point, spp, origin) %>% 
    mutate(
        stemcount = supB + supC1 + supC2 + supP + expB + expC1 + expC2 + expP,
        upperexposed = expC1 + expC2 + expP,
        present = if_else(stemcount > 0, 1, stemcount), # used to calculate diversity if a spp is present at a point
        success = if_else(upperexposed > 0, 1, upperexposed) # used to calculate diversity of spp of successful stems
#        stems.ramet = if_else(is.na(stems.ramet), 0, stems.ramet)
    ) %>%
    
    left_join(point_labels_revision, by= c("harvest", "point")) %>%  #this includes residual BA/ac for cafri walls
    
    left_join(seedlings2019_2024, by = c("harvest", "season", "wall", "spp", "point", "origin")) %>% 

    
# change numeric spp to alpha
mutate(
  spp = as.character(fct_recode(
    factor(spp),"comm_sdl" = "004",
    "rp" = "125", "wp" = "129", "scp" = "130", "he" = "261",
    "svb" = "356", "sm" = "310", "stm" = "315", "rm" = "316",
    "sm" = "318", "ah" = "341", "ab" = "351", "svb" = "355",
    "yb" = "371", "sb" = "372", "pb" = "375", "ah" = "391",
    "bih" = "402", "sh" = "407", "pc" = "461", "unk" = "492",
    "haw" = "500", "ab" = "521", "ab" = "53", "ab" = "531",
    "wa" = "541", "wh" = "585", "yp" = "621", "ab" = "631",
    "app" = "660", "la" = "70", "eh" = "701", "unk" = "706",
    "bta" = "743", "qa" = "746", "pc" = "761", "bc" = "762",
    "cc" = "763", "wo" = "802", "co" = "832", "ro" = "833",
    "smc" = "866", "ns" = "91", "bl" = "901", "wi" = "920",
    "sa" = "931", "ba" = "951", "toh" = "998", "unk" = "999",
    "dws" = "Devil's Walking Stick",
    #data recording errors or previously missed
    "sb" = "273", "gb" = "379", "au" = "535", "ec" = "742",
    "sp" = "90", "ohw" = "001", "ohw" = "1", "ab" = "331",
    "unk" = "ush"
  ))
 ) 
 
 
```



#create a stand alone reference to the numeric variables that will be in the "complete" datafile
```{r}
numeric_vars <- c( "supB", "supC1", "supC2","height", 
                  "supP", "expB", "expC1", "expC2", "expP", "seedA", "stemcount", 
                  "upperexposed", "present", "success")

numeric_vars.2 <- c("supB", "supC1", "supC2", "seedA",
                  "supD", "expD", "supP", "expB", "expC1", "expC2", "expP")

```

attempt #1 - make the data complete (failed)
```{r eval=FALSE}
veg_2019_2022_complete <- merge2019_2022_c %>%
  ungroup() %>%
  select(harvest, season, wall, point, spp, numeric_vars, -location, -origin, -trmt, -type) %>% 
  arrange(harvest, season, wall, point, spp) %>%
  group_by(harvest, season, wall, point) %>%
#    complete(spp, fill = list(!!!setNames(rep(0, length(numeric_vars)), numeric_vars))) %>% 
    complete(spp, fill = list(!!!setNames(rep(NA, length(numeric_vars)), numeric_vars))) %>% 
  ungroup() %>% 
  arrange(harvest, season, wall, spp, origin, point)

```

attempt #2 - make the data complete (failed)
```{r eval=FALSE}
# Step 1: Create a full list of unique species (spp) per wall
species_per_wall <- merge2019_2022_c %>%
  group_by(wall) %>%
  summarize(tableall_species = list(unique(spp)), .groups = "drop")

# Step 2: Use `complete()` to expand every point within a wall to include all species
veg_2019_2022_complete <- merge2019_2022_c %>%
  left_join(species_per_wall, by = "wall") %>% # Add all_species for each wall
  unnest(all_species) %>%                     # Expand species for each wall
  rename(spp_all = all_species) %>%           # Avoid overwriting original spp
  group_by(harvest, season, wall, point) %>%  # Group by the desired levels
  complete(
    spp = spp_all,                            # Ensure all species are present
    fill = list(
      !!!setNames(rep(0, length(numeric_vars)), numeric_vars), # Fill numeric columns with 0
      origin = "unknown"                      # Fill origin with "unknown"
    )
  ) %>%
  ungroup() %>%
  arrange(harvest, season, wall, spp, origin, point)

```


#attempt #3 - make the data complete (this worked!)
```{r}
# Step 1: Create a full list of unique species (spp) per wall
species_per_wall <- merge2019_2022_c %>%
  group_by(wall) %>%
  summarize(all_species = list(unique(spp)), .groups = "drop")


# Step 2: Create all combinations of `harvest`, `season`, `wall`, `point`, and `all_species`
all_combinations <- merge2019_2022_c %>%
  select(harvest, season, wall, point, -trmt, -location) %>%
  distinct() %>%
  left_join(species_per_wall, by = "wall") %>%  # Attach the full species list per wall
  unnest(all_species) %>%                       # Expand the species list
  rename(spp = all_species)                     # Rename to match the original dataset


# Step 3: Complete the dataset by joining the full combinations with the original data
veg_2019_2022_complete <- all_combinations %>%
left_join(merge2019_2022_c, by = c("harvest", "season", "wall", "point", "spp")) %>%
 select(harvest, season, wall, point, spp, origin, all_of(numeric_vars)) %>%
      mutate(
#    height = as.numeric(height),
          # Fill missing numeric variables with 0
    across(all_of(numeric_vars), ~ replace_na(.x, 0)),
    # Fill `origin` with "unknown" if missing
    origin = replace_na(origin, "unknown")
  ) %>%
      arrange(harvest, season, wall, spp, point, origin) 
```



# df=veg01, builds from veg2019-2022, plus seedlings <12"
```{r}
veg01 <- veg_2019_2022_complete %>% 
    filter(!harvest == "5_1") %>% 
    select(-seedA) %>%  #seedA doesn't exist in the input df. 
    full_join(seedlings2019_2024, by = c("harvest", "season", "wall", "point", "spp", "origin")) %>% 
        # I'm not sure why this isn't left_join given the base is 2019-2022
        # in df=veg01, seasons 2023 and 2024 are present only due to the seedA data
    arrange(harvest, season, wall, point, spp,  origin) %>% 
    select(harvest, season, wall, point, spp, origin, seedA, everything()) %>% 
    select(-height) %>% 
#anti_join(veg01, all_labels, by = c("harvest",  "point", "wall"))
    left_join(point_labels_revision, by = c("harvest", "wall", "point")) %>% 
    arrange(harvest, season, wall, trmt, point, location, spp, origin) %>% 
    mutate(
  # #      seedA = coalesce(seedA.x, seedA.y),
  #       height = coalesce(height.x, height.y),
  #       trmt = coalesce(trmt.x, trmt.y),
  #       location = coalesce(location.x, location.y),
        height = as.numeric(height),
        supD = supC1 + supC2,
        expD = expC1 + expC2
    ) %>% 
    select(harvest, season, wall, trmt, point, location, height, spp, origin, everything()) %>%
    select(-stemcount, -upperexposed) %>% 

# Assign "type" category lables to the species
    mutate(
        type = "blank"
        ) %>% 
    mutate(type = case_when(
    spp %in% c("ba", "bc", "bih", "bl", "co", "he", "ohw", "rm", "ro", "sa", 
                "sh", "sm", "wa", "wo", "wp", "yb", "yp") ~ "commercial",
    
    spp %in% c("sb") ~ "sb",
    
    spp %in% c("ab", "eh", "pc", "stm", "toh", "au" ) ~ "interfere",

    spp %in% c("app", "bta", "haw", "brs", "dws", "ec", "la", "ns", "pb", "gb", "qa", "rp", "scp", "smc", "svb", "wh", "wi" ) ~ "diversity",

    spp %in% c("ah", "cc", "unk", "sp" ) ~ "other",
    TRUE ~ type,  # Keeps the existing value of "type" for all other cases
    )) #%>% 
#    filter(harvest == "red_pine" & point == "338" & season == "2019")


```



test of seedling2019_2024
```{r}
seedling_test <- seedlings2019_2024 %>% 
    filter(season == "2023")
```




# df=regen2023_updated, seedlings > 4", matching variables to bind rows with 2019-2022 data
```{r}
regen2023_updated <- regen2023 %>% 
  select(-harvest_original, -forest_type, -brush) %>% #add these later, after summarize
  rename( #new = old
        stems_per_ramet = numb.per.clump
    ) %>% 

#create "seedA" by adding supA+expA for commercial species
  mutate(seedA = case_when(
    type == "commercial" ~ supA + expA,
    TRUE ~ NA_real_
  )) %>% 

#(july 23 2025) create metric for number of clumps by species
#each clump represents the "multiplier" based on plot size as the clumps per acre
# the sum of supA::expP is because there is only one value, the others are zero
# dividing the number of stems by count = the plot multiplier
# 
  mutate(clump_per_acre_each = if_else
         (origin == "clump", (supA+supB+supD+supP+expA+expB+expD+expP)/count, NA_real_)) %>% 

#for regen2023 data, aggregate the stem data at the point spp and origin level
group_by(harvest, point, spp, origin) %>% 
    summarize(
       wall = first(wall),
       season = first(season),
       date = first(date),
       trmt = first(trmt),
       location = first(location),
       type = first(type),
       supA = sum(supA),
       supB = sum(supB),
       supD = sum(supD),
       supP = sum(expP),
       expA = sum(expA),
       expB = sum(expB),
       expD = sum(expD),
       expP = sum(expP),
       seedA = sum(seedA),
       stems_per_ramet = mean(stems_per_ramet),
       clumps_per_acre = sum(clump_per_acre_each),
       .groups = "drop"
    ) %>% 

  left_join(point_labels_revision, by = c("harvest", "wall", "location", "trmt", "point"))
    
```


# df = regen2024_updated, same code as regen2023_updated
```{r}
regen2024_updated <- regen2024 %>% 
    select(-wall, -multip01, -multip02_03, -multip_sapl) %>% 
    filter(!harvest == "5_1") %>% 
  rename( #new = old
        stems_per_ramet = numb.per.clump
    ) %>% 

 #create "seedA" by adding supA+expA for commercial species
  mutate(seedA = case_when(
    type == "commercial" ~ supA + expA,
    TRUE ~ NA_real_
  )) %>% 
   
#(july 23 2025) create metric for number of clumps by species
#each clump represents the "multiplier" based on plot size as the clumps per acre
# the sum of supA::expP is because there is only one value, the others are zero
# dividing the number of stems by count = the plot multiplier
# 
  mutate(clump_per_acre_each = if_else
         (origin == "clump", (supA+supB+supD+supP+expA+expB+expD+expP)/count, NA_real_)) %>% 

#for regen2024 data, aggregate the stem data at the point spp and origin level
group_by(harvest, point, spp, origin) %>% 
    summarize(
       season = first(season),
       date = first(date),
       type = first(type),
       supA = sum(supA),
       supB = sum(supB),
       supD = sum(supD),
       supP = sum(expP),
       expA = sum(expA),
       expB = sum(expB),
       expD = sum(expD),
       expP = sum(expP),
       seedA = sum(seedA),
       stems_per_ramet = mean(stems_per_ramet),
       clumps_per_acre = sum(clump_per_acre_each),
       .groups = "drop"
    ) %>% 

    
# 
# the new df=point_labels_revision includes these sites and attributes.
# the subsequent mutate statements are not needed.

      left_join(point_labels_revision, by = c("harvest", "point")) #%>% 
    
#         mutate(
#         height = case_when(
#         harvest %in% c("danby_block_2", "fairfield_a5", "fairfield_a6", "goundry_a17", "5_1") ~ 0,
#         TRUE ~ height
#   )
# ) %>% 
#     mutate(
#         trmt = case_when(
#         harvest %in% c("danby_block_2", "fairfield_a5", "fairfield_a6", "goundry_a17", "decker", "rectangle", "5_1") ~ "control",
#         TRUE ~ trmt
#   )
# ) %>% 
#     mutate(
#         location = case_when(
#         harvest %in% c("danby_block_2", "fairfield_a5", "fairfield_a6", "goundry_a17", "5_1") ~ "c",
#         TRUE ~ location
#   )
# ) %>% 
#     
#     mutate(
#         wall = case_when(
#         harvest %in% c("danby_block_2", "fairfield_a5", "fairfield_a6", "goundry_a17", "5_1") ~ "99",
#         TRUE ~ wall
#   )
# ) %>% 
# 
#     mutate(
#         brush = case_when(
#         harvest %in% c("fairfield_a5") ~ "b",
#         harvest %in% c("danby_block_2",  "fairfield_a6", "goundry_a17", "5_1") ~ "u",
#         TRUE ~ brush
#   )
# ) %>% 
#         
#     mutate(
#         forest_type = case_when(
#         harvest %in% c("danby_block_2", "fairfield_a5", "5_1") ~ "oak_hdwd",
#         harvest %in% c("fairfield_a6", "goundry_a17") ~ "n_hdwd",
#         TRUE ~ forest_type
#   )
# ) %>% 
# 
#     mutate(
#         harvest_original = case_when(
#         harvest %in% c("danby_block_2", "fairfield_a5", "fairfield_a6", "goundry_a17", "5_1") ~ harvest,
#         TRUE ~ harvest_original
#   )
# )



```

#update point labels to include the state forest sites
```{r eval=FALSE}
basal2024 <- plot2024_orig %>% 
    select(harvest, point, basalarea) %>% 
    filter(! harvest %in% c("5_1", "6_6", "6_9", "recknagel", "sta_rd", "decker")) %>% 
    group_by(harvest) %>% 
    summarize(
        residual = mean(basalarea)
    ) %>% 
 mutate_if(is.numeric, round, 0)
    
    
new_sites <- regen2024_updated %>% 
    select(harvest, point, height, trmt, location, wall, brush, forest_type, harvest_original) %>% 
    filter(! harvest %in% c("5_1", "6_6", "6_9", "recknagel", "rectangle", "sta_rd", "decker")) %>% 
    distinct(harvest, point, .keep_all = TRUE) %>% 
    left_join(basal2024, by = c("harvest"))

point_labels2 <- bind_rows(point_labels_revision, new_sites)
```

#output point_lables to include state forests
```{r eval=FALSE}
output_file <- file.path(output_folder, paste0("point_labels_all(24july2025)", ".xlsx"))

write_xlsx(point_labels2, output_file)

```


#all_labels adapted to give plot count by wall/brush combination
```{r}
# all_labels %>%
#   count(harvest, wall, height, location, trmt) %>%
#   filter(n > 1)

all_labels <- point_labels_revision %>%
  count(harvest, wall, height, location, brush, trmt) %>%
  filter(n > 1) %>% 
    rename(
        plot_count = n
    )


```

create a df that is based on the complete 2019-2022 data, expanded by 2023 and 2024 data, plus adds attributes.
```{r}
# yay1 <- bind_rows(veg01, regen2023_updated, regen2024_updated) %>% 
#     filter(!harvest == "5_1")

# I tested the similarity of yay1 and yay2. they appear identical based on these two table statements, and the subsequent R chunk that graphs the mean of seedA
# table(yay1$season, yay1$harvest)
# table(yay2$season, yay2$harvest)

#yay2 becomes...
yay <- full_join(veg01, regen2023_updated, by = c("harvest", "point", "spp", "origin", "wall",
                                                  "season", "type", "trmt", "location", "height", "harvest_original",
                                                  "forest_type", "brush", "residual", "seedA",
                                                  "supB", "supD", "supP", "expB", "expD", "expP" )) %>% 
    filter(!harvest == "5_1") %>% 
    full_join(regen2024_updated, by = c("harvest", "point", "spp", "origin", "wall",
                                                  "season", "type", "trmt", "location", "height", "harvest_original",
                                                  "forest_type", "brush", "residual", "seedA", "stems_per_ramet", 
                                                  "clumps_per_acre", "date","supB", "supD", "supP", "expB", "expD", "expP",
                                                  "supA", "expA")) %>% 


#the creation of df=yay results in the same number or rows and columns by either code bind and join. 
    
        #check variables of df=regen2024 and add to bind_rows. Check on ramet types of variables and seedA vs. comm_seedl
    arrange(harvest, season, wall, point, spp) %>% 
    select(harvest, season, wall, spp, point, seedA, everything()) %>% 
    filter(!(harvest == "gas_line" & point %in% c(866, 864)))  %>%  #corrected code selects for either point at the intersection of harvest and point
            # these points dropped because they were uncut in 2017 and heavy to beech.

    #anti_join(yay, all_labels, by = c("harvest",  "point", "wall", "height", "location", "trmt", "brush"))
    left_join(all_labels, by = c("harvest",  "wall", "height", "location", "trmt", "brush")) %>% 

# df=all_labels doesn't appear to have any function in df=yay. since df=all_labels
# counts by point, and filters n>1, the df is empty. 
# when "point" is removed, this provides a count of the sample size for each combination of attributes   

# Fairfield and Danby = Region 7, Petit, Travis D (DEC) <Travis.Petit@dec.ny.gov> 
# fairfield SF A5, contract open sept 2020 - november 2025, first growing season = 2022 (need to confirm)
# fairfield SF a6 (aka A-6.1) cut 2009, first growing season = 2010 confirmed (logging finished 2009)
# Danby SF block 2, contract open november 2016 to october 2019, first growing season = 2018 (need to confirm)
# 
# Goundry Hill SF = Region 8, Cicora, Gretchen L (DEC) <gretchen.cicora@dec.ny.gov>
# Goundry hill SF , goundry_a17, contract open december 2016 to january 2020, first growing season = 2018 (need to confirm)
# 
# 
        select(-present, -success) %>% 
     mutate(
        firstgrow = 2035, # default value for the first growing season
        firstgrow = case_when(
          harvest %in% c("gas_line", "red_pine") ~ 2017,
          harvest %in% c("sta_rd") ~ 2018,
          harvest %in% c("campridge", "recknagel", "recknagel_north", "circle", "decker") ~ 2020,
          harvest %in% c("6_6", "6_9") ~ 2022,
          TRUE ~ firstgrow  # Keeps the existing value of "firstgrow" for all other cases
    ),
        year = as.numeric(season),
        growingseasons = (year - firstgrow) + 1
  ) %>% 
    mutate(
    stemcount = coalesce(supB, 0) + coalesce(supC1, 0) + coalesce(supC2, 0) + coalesce(supP, 0) + 
                coalesce(expB, 0) + coalesce(expC1, 0) + coalesce(expC2, 0) + coalesce(expP, 0),

    upperexposed = coalesce(expC1, 0) + coalesce(expC2, 0) + coalesce(expP, 0),

    present = if_else(stemcount > 0, 1, 0),  # Ensures output is never NA

    success = if_else(upperexposed > 0, 1, 0),  # Ensures output is never NA

#    stems.ramet = replace_na(stems.ramet, 0)  # Replaces NA with 0
    
    stems_per_ramet = if_else(is.na(stems_per_ramet), 0, stems_per_ramet)
    ) %>% 
    mutate(
        across(all_of(numeric_vars.2), ~ ifelse(seedA > 0 & is.na(.), 0, .)),
        across(all_of(numeric_vars.2), ~ ifelse(is.na(.), 0, .)), 
        across(all_of(numeric_vars.2), ~ ifelse(seedA <1 & spp == "004", 0, .)),
        seedA = if_else(is.na(seedA), 0, seedA),
        total = (supB + supC1 + supC2 + supP + expB + expC1 + expC2 + expP)
    ) %>% 
    mutate(
        filename = slash_veg_complete.Rmd,
        df = "yay"
    ) %>% 
    select(filename, df, harvest, season, growingseasons, wall, trmt, point, spp, origin, seedA, supB, supC1, supC2, supP,
           expB, expC1, expC2, expP, total, everything())  
 #   filter(if_any(everything(), is.na))
```

testing yay1 vs yay2
```{r eval=FALSE}
combined_df <- bind_rows(yay1, yay2, .id = "source")  # 'source' will be "1" or "2"

combinded_df <- combined_df %>% 
    mutate(source = recode(source, `1` = "yay1", `2` = "yay2"))
    
summary_df <- combined_df %>%
  group_by(source, season, harvest) %>%
  summarize(mean_seedA = mean(seedA, na.rm = TRUE), .groups = "drop")

library(ggplot2)

ggplot(summary_df, aes(x = season, y = mean_seedA, color = source)) +
  geom_point() +
  geom_line(aes(group = source)) +
  facet_wrap(~harvest, scales = "free_y") +
  labs(
    title = "Mean SeedA by Harvest and Season",
    x = "Season",
    y = "Mean seedA"
  ) +
  theme_minimal()

ggplot(summary_df, aes(x = season, y = mean_seedA, color = source)) +
  geom_point(position = position_dodge(width = 0.3)) +
  geom_line(aes(group = source), position = position_dodge(width = 0.3)) +
  facet_wrap(~harvest, scales = "free_y") +
  labs(title = "Mean SeedA by Harvest and Season", x = "Season", y = "Mean seedA") +
  theme_minimal()

```






```{r}
# Define the output file path
output_file <- file.path(output_folder, paste0("veg_data_2019to2024_complete(25july2025)", ".xlsx"))
#write_xlsx(yay, output_file)

# Create a new workbook and add a worksheet
wb <- createWorkbook()
addWorksheet(wb, "Data")

# Get the Rmd filename, with a fallback for interactive use
rmd_filename <- knitr::current_input()
if (is.null(rmd_filename)) {
  rmd_filename <- "slash_veg_complete.Rmd"  # Fallback when running in console
}

# Write the Rmd filename in the first row
writeData(wb, "Data", data.frame(Filename = rmd_filename), startRow = 1, startCol = 1, colNames = FALSE)

# Write the actual data starting in row 3 (to leave a blank row for clarity)
writeData(wb, "Data", yay, startRow = 3, startCol = 1)

# Save the workbook
saveWorkbook(wb, output_file, overwrite = TRUE)
```


```{r}
test01 <- yay %>% 
  ungroup() %>%
  mutate(
      total = (seedA + supB + supC1 + supC2 + supP + expB + expC1 + expC2 + expP)
  ) %>% 
  group_by(harvest, season, wall, trmt) %>%
  summarize(
    season = first(season),
    firstgrow = mean(firstgrow),
    year = mean(year),
    growingseasons = mean(growingseasons),
    total = mean(total),
    richness = n_distinct(spp[present > 0], na.rm = TRUE),
    group_size = n(),
    .groups = "drop"
  )

```
