---
title: "slash wall analyses 2025"
author: "Peter Smalldige"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r }

#library(ggplot2)
#library(ggpubr)
library(expss)
library(readr)
library(readxl)
library(writexl)
library(skimr)
library(tidyverse)  # includes: tidyr, dplyr, haven, readr, readxl, forcats, stringer, ggplot2. See all with tidyverse_packages()
library(knitr)
library(janitor)
#install.packages("openxlsx")
library(openxlsx)
# install.packages("shinyaddins")
# library(shinyaddins)

```




### code 01
### Create file path input and the output folder (code 01) 

file paths, input and output
```{r}
#complete data
file_path_veg_complete <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\veg_data_2019to2023_complete(16feb2025)(read_me).xlsx"
file_path_denom <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\count_of_points_stand_by_protection.xlsx"


output_folder <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\summary output\\"
```



```{r}
denom <- read_excel(file_path_denom, sheet = "Sheet 1") %>% 
    rename(
        harvest = stand,
        trmt = status
        ) %>%   
    mutate(
      harvest = case_when(
      harvest == "boot" ~ "sta_rd",
      harvest == "circle1" ~ "circle",
      harvest == "circle2" ~ "circle",
      harvest == "circle3" ~ "circle",
      harvest == "cricle1" ~ "circle",
      harvest == "gasline" ~ "gas_line",
      harvest == "recksouth" ~ "recknagel",
      harvest == "redpine" ~ "red_pine",
      harvest == "stationrd" ~ "sta_rd",
      harvest == "wedge" ~ "sta_rd",
      harvest == "old" ~ "new",
      TRUE ~ harvest  # Keeps other values unchanged
    )
      )

denom_stand_trmt <- denom %>% 
    group_by(harvest, trmt) %>% 
    summarize(
        denom = sum(wall_denom)
    ) %>% 
    mutate(
        filename = "slash_wall_analyses_2025.Rmd"
    ) %>% 
    select(filename, everything())

output_file <- file.path(output_folder, paste0("sample_size_table(16feb2025)", ".xlsx"))
write_xlsx(denom_stand_trmt, output_file)


```
#Notes
# - data are complete at the level of the wall
```{r}
d0 <- read_excel(file_path_veg_complete, sheet = "Data", skip = 2) %>% 
    left_join(denom_stand_trmt, by = c("harvest", "trmt")) %>% 
        mutate(
        supD = supC1 + supC2,
        expD = expC1 + expC2
    ) %>% 
    arrange(harvest, season, trmt, wall, point, spp) %>% 
    select(-stemcount, -upperexposed) 

#numeric vars to sum by origin
numeric_vars1 <- c("seedA", "supB", "supC1", "supC2", "supD", "supP", "expB", "expC1", "expC2",
                  "expD", "expP", "total", "clumps.acre", "stems.ramet")

#numeric_vars2 = average by origin
numeric_vars2 <- c("firstgrow", "year", "present", "success", "denom", "growingseasons", "height")

d1.1 <- d0 %>% #aggregate or average to collapse "origin"
    group_by(harvest, season, wall, point, spp) %>% 
    mutate(type = case_when(
        spp %in% c("sb") ~ "sb",
        TRUE ~ type,  # Keeps the existing value of "type" for all other cases
    )) %>%     
    summarize(
        filename = first(filename),
        trmt = first(trmt),
        type = first(type),
        numb_points = n(),
        across(all_of(numeric_vars1), sum, na.rm = TRUE),  # Sum all specified numeric vars1
        across(all_of(numeric_vars2), mean, na.rm = TRUE),  # Sum all specified numeric vars2
        .groups = "drop"
    )

d1 <- d1.1 %>% 
    arrange(desc(season), harvest, wall, type, spp, point) %>% 
    select(harvest, wall, point, type, spp, season, trmt, everything())

d1a <- d1 %>% 
    filter(season == "2023") %>% 
    filter(harvest == "6_6" & season == "2023") %>% 
    filter(total > 0) %>% 
    arrange(desc(season), harvest, wall, type, spp, point) 

```

wall summary, mean values by wall and spp - across points
```{r}
wall_summ <- d1 %>% 
#    filter(spp =="004") %>% 
    filter(season == "2023") %>% 
#    arrange(harvest, season, trmt, wall, point, spp) %>% 
#    select(harvest, season, trmt, wall, point, spp, seedA, total, growingseasons)
    group_by(harvest, season, wall, spp) %>% 
    summarize(
        filename = first(filename),
        numbplots = first(denom),
        trmt = first(trmt),
        type = first(type),
        numbpoints = n(),
        total = mean(total), #seedA and all larger stems
        seedlings = mean(seedA),
#        GTR5ft_sum = sum(supD + supP + expD + expP),
        GTR5ft = mean(supD + supP + expD + expP),
        exp5ft = mean(expD + expP),  #average of points within a wall
        established = mean(supB + supD + supP + expB + expD + expP),  #average of points within a wall
        richness_established = sum((supB + supD + supP + expB + expD + expP) > 0, na.rm = TRUE),  # Count plots by spp where stems >12" occur
        richness_GTR5ft = sum((supD + supP + expD + expP) > 0, na.rm = TRUE), # count of plots by spp for stems > 5 ft
        richness_GTR5ft_commdiv = sum((supD + supP + expD + expP) > 0 & (type == "commercial" | type == "diversity") , na.rm = TRUE),  
            #Count escaped comm and divers spp
        across(all_of(numeric_vars1), mean, na.rm = TRUE),    
        .groups = "drop"
    ) %>% 
    arrange(harvest, trmt,wall, type, spp) %>% 
    filter(total > 0) %>% 
    mutate(
        across(where(is.numeric), ~ round(., 0))
        ) %>% 
    select(filename, harvest, season, trmt, type, spp, wall, everything())

output_file <- file.path(output_folder, paste0("wall_summ(16feb2025)", ".xlsx"))
write_xlsx(wall_summ, output_file)

```


#all summary, mean values by wall and spp - across points
# this chunk allows calc of occupancy by point without double counting spp of the same type
# merge this with the richness calcs at the harvest/season/trmt scale
```{r eval=FALSE}
point_wall_summ <- d1 %>% 
#    filter(spp =="004") %>% 
    filter(season == "2023") %>% 
#create dummy variables
    mutate(
    type = case_when(
    type %in% c("commercial") ~ "commercial",
    
    spp %in% c("ab", "eh", "pc", "stm", "toh" ) ~ "interfere",

    )
    group_by(harvest, season, wall, point, type) %>% 
    summarize(
        filename = first(filename),
        numbplots = first(denom),
        trmt = first(trmt),
        type = first(type),
        numb_spp = n(),
        total = mean(total), #seedA and all larger stems
        seedlings = mean(seedA),
#        GTR5ft_sum = sum(supD + supP + expD + expP),
        GTR5ft = mean(supD + supP + expD + expP),
        exp5ft = mean(expD + expP),  #average of points within a wall
        established = mean(supB + supD + supP + expB + expD + expP),  #average of points within a wall
        richness_established = sum((supB + supD + supP + expB + expD + expP) > 0, na.rm = TRUE),  # Count plots by spp where stems >12" occur
        richness_GTR5ft = sum((supD + supP + expD + expP) > 0, na.rm = TRUE), # count of plots by spp for stems > 5 ft
        richness_GTR5ft_commdiv = sum((supD + supP + expD + expP) > 0 & (type == "commercial" | type == "diversity") , na.rm = TRUE),  
            #Count escaped comm and divers spp
        across(all_of(numeric_vars1), mean, na.rm = TRUE),    
        .groups = "drop"
    ) %>% 
    arrange(harvest, trmt, wall, type, point) %>% 
    filter(total > 0) %>% 
    mutate(
        across(where(is.numeric), ~ round(., 1))
        ) %>% 
    select(filename, harvest, season, trmt, wall, type, point, everything()) 
    

output_file <- file.path(output_folder, paste0("point_wall_summ(17feb2025)", ".xlsx"))
write_xlsx(point_wall_summ, output_file)

```


# initial code for richness
```{r eval=FALSE}
trmt_richness_summ <- wall_summ %>% 
    filter(season == "2023") %>% 
    group_by(harvest, type, trmt, wall) %>% 
    summarize(
        filename = first(filename),
        trmtdenom = first(numbplots),
        numbpoints = first(numbpoints),
        n = n(),
        seedlings = sum(seedlings),
        total = sum(total), #seedA and all larger stems

#established
        num_spp_present =  n_distinct(spp[established > 0], na.rm = TRUE),  # Count non-zero values, positioned prior to calculation of "established"
        established = sum(established), #stem count per acre of stems > 12"
        richness_established = sum(richness_established/numbpoints), #proportion of points occupied by stems > 12"
        spp_occ_pct = ifelse(num_spp_present > 0, (richness_established / num_spp_present*100), NA), #for spp present, their average % occupancy of points within a wall
        point_estab_pct = richness_established * 100, # % points within wall with at least one established seedling
        points_occupied_estab = mean(richness_established/numbpoints*100), # number of points occupied by an established seedling

# GTR5ft
        num_spp_GTR5ft =  n_distinct(spp[GTR5ft > 0], na.rm = TRUE),  # Count non-zero values, positioned prior to calculation of "established" (calc confirmed)
        GTR5ft_tpa = sum(GTR5ft), # number of stems per acre per wall > 5ft
        richness_GTR5ft = sum(richness_GTR5ft/ numbpoints), #     
        points_GTR5ft_pct = (sum(num_spp_GTR5ft)),#/(numbpoints*100), #percentage of points in a wall occupied by stems GTR5ft
    
        occupancy_GTR5ft_commdiv = mean(richness_GTR5ft_commdiv/numbpoints*100),
        across(all_of(numeric_vars1), sum, na.rm = TRUE), 
        .groups = "drop"
        ) %>%
     mutate(
        across(where(is.numeric), ~ round(., 1))
        ) %>% 
  
    group_by(harvest, type, trmt) %>% 
    summarize(
        filename = first(filename),
        trmtdenom = first(numbplots),
        n = n(),
        occupancy_exposed = sum(richness_all_exposed) / trmtdenom * 100,
        across(all_of(numeric_vars1), mean, na.rm = TRUE),    
        GTR5ft = sum(GTR5ft),
        all_exposed = sum(all_exposed),  #average of points within a wall
        escaped = sum(escaped),  #average of walls within a treatment
) %>%
    select(filename, harvest, trmt, type, n,...... everything()) %>% 
    arrange(harvest, type, trmt)

```



Use summ by point to calculate occupancy and avoid double counting plots that have multiple spp of the same type
        total = sum(total), #seedA and all larger stems
```{r}

trmt_occupancy_summ <- point_wall_summ %>% 
    filter(season == "2023") %>% 
    group_by(harvest, type, trmt, wall) %>% 
    summarize(
        filename = first(filename),
        trmtdenom = first(numbplots),
        numbpoints = first(numbpoints),
        n = n(),
        seedlings = sum(seedlings),

#established
        num_spp_present =  n_distinct(spp[established > 0], na.rm = TRUE),  # Count non-zero values, positioned prior to calculation of "established"
        established = sum(established), #stem count per acre of stems > 12"
        richness_established = sum(richness_established/numbpoints), #proportion of points occupied by stems > 12"
        spp_occ_pct = ifelse(num_spp_present > 0, (richness_established / num_spp_present*100), NA), #for spp present, their average % occupancy of points within a wall
        point_estab_pct = richness_established * 100, # % points within wall with at least one established seedling
        points_occupied_estab = mean(richness_established/numbpoints*100), # number of points occupied by an established seedling

# GTR5ft
        num_spp_GTR5ft =  n_distinct(spp[GTR5ft > 0], na.rm = TRUE),  # Count non-zero values, positioned prior to calculation of "established" (calc confirmed)
        GTR5ft_tpa = sum(GTR5ft), # number of stems per acre per wall > 5ft
        richness_GTR5ft = sum(richness_GTR5ft/ numbpoints), #     
        points_GTR5ft_pct = (sum(num_spp_GTR5ft)),#/(numbpoints*100), #percentage of points in a wall occupied by stems GTR5ft
    
        occupancy_GTR5ft_commdiv = mean(richness_GTR5ft_commdiv/numbpoints*100),
        across(all_of(numeric_vars1), sum, na.rm = TRUE), 
        .groups = "drop"
        ) %>%
     mutate(
        across(where(is.numeric), ~ round(., 1))
        ) %>% 
  
    group_by(harvest, type, trmt) %>% 
    summarize(
        filename = first(filename),
        trmtdenom = first(numbplots),
        n = n(),
        occupancy_exposed = sum(richness_all_exposed) / trmtdenom * 100,
        across(all_of(numeric_vars1), mean, na.rm = TRUE),    
        GTR5ft = sum(GTR5ft),
        all_exposed = sum(all_exposed),  #average of points within a wall
        escaped = sum(escaped),  #average of walls within a treatment
) %>%
    select(filename, harvest, trmt, type, n,...... everything()) %>% 
    arrange(harvest, type, trmt)

```


    

```{r}
trmt_spp_summ <- wall_summ %>% 
    group_by(harvest, trmt, type) %>% 
    summarize(
        filename = first(filename),
        trmtdenom = first(numbplots),
        occupancy_exposed = sum(richness_all_exposed) / trmtdenom * 100,
        across(all_of(numeric_vars1), mean, na.rm = TRUE),    
        GTR5ft = mean(GTR5ft),
        all_exposed = mean(all_exposed),  #average of points within a wall
        escaped = mean(escaped),  #average of walls within a treatment
) %>%
    select(filename, harvest, trmt, type, trmtdenom, total, occupancy_exposed, GTR5ft, all_exposed, escaped, everything()) %>% 
    arrange(harvest, type, trmt)

```
