---
title: "summary_regen_veg_data.Rmd"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r}
library(readxl)
library(readr)
#library(skimr)
library(tidyr)
library(tidyverse)
#library(tidytable)  ## CAUTION this may conflict with dplyr functions
library(ggplot2)
library(ggpubr)
library(dplyr)

```

<!-- README 

File Description: The purpose of this file is to:
  CODE 01, line 54
  - input data from the 2019 through 2022 growing seasons. 
    Data was created by files "regen_and_plot_input_20xx_cleaning.Rmd" Bind the data sets for each year together. (code 01)  For the 2023 growing season, the input file is "test_2023_regen_plot_input.Rmd"
    
The original raw data is regen_per_acre_all_alpha_tidy_1june2023.xlsx (or 2020, 2021, 2022). 
The alpha_tidy output from the "input and cleaning" files (e.g., regen_and_plot_input_2022_cleaning.Rmd) 
is copy and pasted into the raw data file. This ensures that if the originating input and cleaning files are rerun
they don't automatically write over the data file used for input here.

CODE 02, line 135
  - The data file "regen_all" has all regen data from all growing seasons available. The commercial species and 
    diversity species are extracted to produce summaries of trees per acre. Code in this section creates averages by various clusters, 
    culminating in average by size class, spp, season, location, and harvest. Additional averaging completed by harvest x season x type.
    
    
-  (lines 300 - 800) summary of tpa by various groupings of subsets of the data (e.g., exposed GT 4.5 ft)
-  (lines 800 - 1000) graphs
-  (code 05)
-  (code 06)


Exported data files (csv) include:
summary_regen_veg_data.Rmd which outputs
•	regen_per_acre_all_alpha_tidy_trmt_coded.csv
•	regen_per_acre_winners_spp_composition_tables.csv
•	regen_per_acre_winners_spp_5_dom.csv
•	regen_per_acre_total_spp_composition_tables.csv
•	regen_per_acre_total_spp_5_dom.csv
•	regen_per_acre_exp_GT_4.5ft_spp_composition_tables.csv
•*	regen_per_acre_exp_GT_4.5ft_spp_5_dom.csv
•	regen_per_acre_desired_spp_composition_tables.csv
•	regen_per_acre_desired_spp_5_dom.csv
•	regen_per_acre_dsrd_exp_gt4.5ft_spp_composition_tables.csv
•	regen_per_acre_dsrd_exp_gt4.5ft_spp_5_dom.csv
•	regen_per_acre_type_exp_gt4.5ft_composition_tables.csv
•	regen_per_acre_type_exp_gt4.5ft_5_dom.csv
•	regen_summary_type_exp_gt4.5ft_composition_tables.csv
•	regen_summary_type_exp_gt4.5ft_5_dom.csv

-->

```{r}
file_path_input <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\regen_per_acre_all_alpha_tidy_1june2023.xlsx"

input_count_harvest_x_location <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\count_of_points_harvest_x_location.csv"

input_count_harvest_x_treatment <- "C:\\Users\\pjs23\\Documents\\R\\slash-wall-vegetation\\old files but do not discard\\count_of_points_harvest_x_treatment.csv"


```


## Code 01
```{r}
#    suppressed = sup_short + sup_med + sup_tall + sup_sapl,
#    exposed = exp_short + exp_med + exp_tall + exp_sapl,
#    total = suppressed + exposed, 
#    winners = top4

# 2019 growing season. 
# 
regen_2019  <- read_excel(file_path_input, sheet = "2019", skip = 2)%>% 
    rename(
      tpa = stems.per.acre
    ) %>% 
             mutate(harvest = fct_recode(harvest,
                              "sta_rd" = "boot",
                              "sta_rd" = "wedge"
             )) %>% 
    mutate(
      location = as.factor(location)
      ) %>% 
    select(harvest, wall, location, season, numb_yrs_grow, point, count, spp, origin, type, size, tpa, cut )

table(regen_2019$harvest)
str(regen_2019)
```


```{r}
#2020 growing season
regen_2020  <- read_excel(file_path_input, sheet = "2020", skip = 3) %>% 
    rename(
      tpa = stems.per.acre
    ) %>% 
    mutate(
    location = as.factor(location),
    harvest = fct_recode(harvest,
                         "sta_rd" = "boot",
                         "sta_rd" = "wedge")
  ) %>% 
      select(harvest, wall, location, season, numb_yrs_grow, point, count, spp, origin, type, size, tpa, cut )
table(regen_2020$harvest)
str(regen_2020)
```


```{r}
#2021 growing season
regen_2021 <- read_excel(file_path_input, sheet = "2021", skip = 3) %>% 
  mutate(
    year = 2021,
    season=year,
    tpa = stems.per.acre,
    harvest = as.factor(harvest),
    location = as.factor(location)
  ) %>% 
    select(harvest, wall, location, season, numb_yrs_grow, point, count, spp, origin, type, size, tpa, cut )
table(regen_2021$harvest)
str(regen_2021)
```


```{r}
#2022 growing season
regen_2022 <- read_excel(file_path_input, sheet = "2022", skip = 3) %>% 
  mutate(
    count = 1,
    tpa = stems.per.acre,
    harvest = as.factor(harvest),
    location = as.factor(location)  
    ) %>% 
    select(harvest, wall, location, season, numb_yrs_grow, point, spp, origin, type, size, tpa, cut )
table(regen_2022$harvest)
str(regen_2022)
```


```{r}
# data for testing calculations
# 4 harvests
testdata01 <- read_excel("regen_per_acre_all_alpha_tidy_20feb2023.xlsx", 
    sheet = "testdata01", skip = 0) %>% 
  rename(TRMT = wall) %>% 
  mutate(
    harvest = as.factor(harvest),
    location = as.factor(location)  
    )
```


```{r}
point_count_location <- read_csv(input_count_harvest_x_location, show_col_types = FALSE) %>% 
   mutate(
     location = fct_recode(location,
    "control" = "c", "interior" = "i", "perimeter" = "p"
    )) 
## old_value = new_value

point_count_wall <- read_csv(input_count_harvest_x_treatment,  show_col_types = FALSE)
```


```{r}
regen_all_base <- bind_rows(regen_2019, regen_2020, regen_2021, regen_2022) %>% 
    arrange(harvest, location, season, spp,
          match(size, c("sup_short", "sup_med", "sup_tall", "sup_sapl", 
                         "exp_short", "exp_med", "exp_tall", "exp_sapl",
                         "suppressed", "exposed", "total")), point) %>% 
    mutate(
      tpha = tpa * 2.47
    ) %>% 
  arrange(desc(harvest), wall, season, point, spp, size)




##
#
regen_all_location <- full_join(regen_all_base, point_count_location, by = (c("harvest", "location"))) 


regen_01 <- full_join(regen_all_location, point_count_wall, by = (c("harvest", "wall"))) %>% 
  rename(TRMT = wall) %>% 
  mutate(
    tpa_x_wall = tpa / wall_denom, #this is the average (="x") tpa by wall treatment
    tpa_x_location = tpa / location_denom,
    tpha_x_wall = tpha / wall_denom,
    tpha_x_location = tpha / location_denom,
    count = 1
  ) %>% 
  arrange(harvest, desc(TRMT), season, point, spp, size) %>% 
  select(harvest, TRMT, season, location, cut, numb_yrs_grow, point, count, spp, origin, size, type, tpa, tpha, 
         location_denom, wall_denom, tpa_x_wall, tpa_x_location, tpha_x_wall, tpha_x_location)


df2017 <- regen_01 %>% 
  filter(harvest %in% c("gas_line", "red_pine")) %>% 
      mutate(
          cut = 2017,
          first_grow = 2017,
          year = season,
          numb_yrs_grow = (year - first_grow) + 1   #e.g., 2021 - 2017 + 1
          ) 

df2018 <- regen_01 %>% 
  filter(harvest %in% c("boot", "wedge", "sta_rd")) %>% 
      mutate(
          cut = 2017,
          first_grow = 2018,
          year = season,
          numb_yrs_grow = (year - first_grow) + 1   #e.g., 2021 - 2018 + 1
          ) 

df2019 <- regen_01 %>% 
  filter(harvest %in% c("campridge", "circle")) %>% 
      mutate(
          cut = 2019,
          first_grow = 2020,
          year = season,
          numb_yrs_grow = (year - first_grow) + 1   #e.g., 2021 - 2020 + 1
          ) 

df2020 <- regen_01 %>% 
  filter(harvest %in% c("recknagel", "decker")) %>% 
      mutate(
          cut = 2019,
          first_grow = 2020,
          year = season,
          numb_yrs_grow = (year - first_grow) + 1   #e.g., 2021 - 2020 + 1
          ) 

regen_all <- bind_rows(df2017, df2018, df2019, df2020) %>% 
    select(harvest, TRMT, season, location, cut, first_grow, numb_yrs_grow, point, count, spp, origin, size, type, tpa, tpha, 
         location_denom, wall_denom, tpa_x_wall, tpa_x_location, tpha_x_wall, tpha_x_location) %>% 
    mutate(TRMT = fct_recode(TRMT,
                    "inside" = "protected",
                    "outside" = "control")
    ) 
table(regen_all$spp)
write_csv(regen_all, "regen_per_acre_all_alpha_tidy_trmt_coded.csv")

#
# the *.csv file saved as "regen_per_acre_all_alpha_tidy_trmt_coded.xlxs" in the root directory R>slash-wall-vegetation
#


```


```{r}
regen_all_harvest_summ <- regen_all %>% 
  group_by(harvest, season, TRMT, spp) %>% 
  summarize(
    harvest_tpa = sum(tpa_x_wall),
    harvest_tpa_location = sum(tpa_x_location),
    harvest_tpha = sum(tpha_x_wall),
    harvest_tpha_location = sum(tpha_x_location)
  )
```




## 


# TESTING THE CODE TO SORT SPP TABLES
# TESTING,  this code doesn't create any new df needed in subsequent calculations


```{r}



#
# test for exposed stems by spp in red_pine for 2021
#

testdata01 <- testdata01 %>% 
  filter(size == "exposed") %>% 
  filter(tpa > 0)

regen_all_test_spp <- testdata01 %>% 
  group_by(harvest, season, TRMT, spp) %>% 
    summarize(
      tpa = sum(tpa_x_wall),
      wall_denom = mean(wall_denom),
      points_occupied = n_distinct(point),
      age = mean(numb_yrs_grow)
    ) %>% 
# the combination of "arrange" and "slice_head" retains the n=i highest values of tpa within the grouping structure
   arrange(harvest, season, TRMT, desc(tpa)) %>% 
  slice_head(n=7) %>% 
  mutate(
    spp_freq = (points_occupied / wall_denom )*100
  ) %>% 
  ungroup %>% 
  arrange(harvest, season, spp, TRMT)

# calculations for tpa_x_wall confirmed on 6/5/2023


# the metric "n_distinct" provides a count of the number of points where a species occurs
# "mean(wall_denom) provides the number of points by wall as control or protected
# must filter (tpa > 0) because n_distinct(variable_name) counts "0" as occupied
# freq (by spp) = count/wall_denom * 100

#
# test for exposed stems by type (ignore spp)

regen_all_test_type <- testdata01 %>% 
  group_by(harvest, season, TRMT, type) %>% 
    summarize(
      tpa = sum(tpa_x_wall),
      wall_denom = mean(wall_denom),
      points_occupied = n_distinct(point),
      age = mean(numb_yrs_grow)
    ) %>% 
  mutate(
    spp_freq = (points_occupied / wall_denom )*100
  ) %>% 
  ungroup %>% 
  arrange(harvest, season, type, TRMT)


```


#.
# COMPOSITION TABLES STEM DENSITY AND FREQUENCEY
#.
# STEMS THAT ARE EXPOSED, ALL SIZE CLASSES

```{r}

#
# Composition for stems 
#





```


# calculation of total winners checked and confirmed as correct on 3/1/2023

```{r winner tabluation}

regen_all_winners_spp_initial <- regen_all %>% 
  filter(tpa>0) %>% 
  filter(size == "winners") %>% 
  mutate(
    describ = "winners"  #labeling data to be clear about the variables selected
  ) %>% 
  group_by(describ, harvest, season, TRMT, spp) %>% 
    summarize(
      tpa = sum(tpa_x_wall),  # the average tpa for each spp is aggregated across points
      tpha = sum(tpha_x_wall),
      wall_denom = mean(wall_denom),
      points_occupied = n_distinct(point),
      age = mean(numb_yrs_grow)
    ) %>% 
  mutate(
    PCT_pts_present = (points_occupied / wall_denom )*100
     ) %>% 
      arrange(harvest, season, TRMT, (desc(tpa)))

  regen_winners_head <- regen_all_winners_spp_initial %>% 
    slice_head(n=5) %>% 
    mutate(spp_count = 1)
  
  regen_winners_tail <- regen_all_winners_spp_initial %>% 
    slice_tail(n=-5) %>% 
    filter(tpa >0) %>% 
    mutate(
      spp = "remainder",
      numb = 1
      ) %>% 
    group_by(describ, harvest, season, TRMT, spp) %>% 
    summarize(
      tpa = sum(tpa),  # the average tpa for each spp is aggregated across points
      tpha = sum(tpha),
      wall_denom = mean(wall_denom),
      age = mean(age),
      spp_count = sum(numb)
    )
 
  #this code provides the sum of stem density by harvest, giving totals plus spp summ in next dataframe 
  regen_winners_summ <- bind_rows(regen_winners_head, regen_winners_tail) %>% 
    select(describ, harvest, season, TRMT, spp, tpa, tpha, PCT_pts_present, wall_denom, age, spp_count) %>% 
    group_by(describ, harvest, season, TRMT) %>% 
    summarize(
      tpa = sum(tpa),
      tpha=sum(tpha),
      spp_count = sum(spp_count)
    ) %>% 
    mutate(spp = "TOTAL")

    
regen_winners_spp <- bind_rows(regen_winners_head, regen_winners_tail, regen_winners_summ) %>% 
    select(describ, harvest, season, TRMT, spp, tpa, tpha, PCT_pts_present, wall_denom, age, spp_count) %>% 
    rename(
      TOTAL_tpa = tpa,
      TOTAL_tpha = tpha
    ) %>% 
    #ungroup %>% 
    arrange(describ, harvest, desc(season), TRMT) %>% 
    dplyr::mutate_if(is.numeric, round, 0)  #changes all numeric variable to round to 0 significant digits


#calculations confirmed 2/21/2023
write_csv(regen_all_winners_spp_initial, "regen_per_acre_winners_spp_composition_tables.csv")
write_csv(regen_winners_spp, "regen_per_acre_winners_spp_5_dom.csv")


# the metric "n_distinct" provides a count of the number of points where a species occurs
# "mean(wall_denom) provides the number of points by wall as control or protected
# freq (by spp) = count/wall_denom * 100



```



# species compositon tables for all stems 'TOTAL' by spp

```{r}

# Composition for stems as "TOTAL" by spp
#

regen_total_spp_initial <- regen_all %>% 
  filter(tpa>0) %>% 
  filter(size == "total") %>% 
  mutate(
    describ = "total"  #labeling data to be clear about the variables selected
  ) %>% 
  group_by(describ, harvest, season, TRMT, spp) %>% arrange(harvest, season, TRMT, (desc(spp))) %>% 
    summarize(
      tpa = sum(tpa_x_wall),  # the average tpa by harvest and season for each spp is aggregated across points
      tpha = sum(tpha_x_wall),
      wall_denom = mean(wall_denom),
      points_occupied = n_distinct(point), #(3mar2023 calculation of points_occupied at harvest scale CONFIRMED)
      age = mean(numb_yrs_grow)
    ) %>% 
  mutate(
    PCT_pts_present = (points_occupied / wall_denom )*100
     ) %>% 
      arrange(harvest, season, TRMT, (desc(tpa)))

  regen_total_head <- regen_total_spp_initial %>% 
    slice_head(n=5) %>% 
    mutate(spp_count = 1)
  
  regen_total_tail <- regen_total_spp_initial %>% 
    slice_tail(n=-5) %>% 
    mutate(
      spp = "remainder",
      numb = 1
      ) %>% 
    group_by(describ, harvest, season, TRMT, spp) %>% 
    summarize(
      tpa = sum(tpa),  # the average tpa for each spp is aggregated across points
      tpha = sum(tpha),
      wall_denom = mean(wall_denom),
      age = mean(age),
      spp_count = sum(numb)
    )
 
  #this code provides the sum of stem density by harvest, giving totals plus spp summ in next dataframe 
  regen_summ <- bind_rows(regen_total_head, regen_total_tail) %>% 
    select(describ, harvest, season, TRMT, spp, tpa, tpha, PCT_pts_present, wall_denom, age, spp_count) %>% 
    group_by(describ, harvest, season, TRMT) %>% 
    summarize(
      tpa = sum(tpa),
      tpha=sum(tpha),
      spp_count = sum(spp_count)
    ) %>% 
    mutate(spp = "TOTAL")

    
regen_total_spp <- bind_rows(regen_total_head, regen_total_tail, regen_summ) %>% 
    select(describ, harvest, season, TRMT, spp, tpa, tpha, PCT_pts_present, wall_denom, age, spp_count) %>% 
    rename(
      TOTAL_tpa = tpa,
      TOTAL_tpha = tpha
    ) %>% 
    #ungroup %>% 
    arrange(describ, harvest, desc(season), TRMT) %>% 
    dplyr::mutate_if(is.numeric, round, 0)  #changes all numeric variable to round to 0 significant digits


#calculations confirmed 2/21/2023

write_csv(regen_total_spp_initial, "regen_per_acre_total_spp_composition_tables.csv")
write_csv(regen_total_spp, "regen_per_acre_total_spp_5_dom.csv")


```




#
# Composition for exposed stems >4.5 ft by species
#

```{r}

# for exposed stems greater than 4.5 ft tall

regen_exp_GT_4.5ft_spp_initial <- regen_all %>% 
  filter(tpa>0) %>% 
  filter(size == "exp_med" | size == "exp_tall" | size == "exp_sapl") %>% 
  mutate(
    describ = "exp_GT_4.5ft"  #labeling data to be clear about the variables selected
  ) %>% 
  group_by(describ, harvest, season, TRMT, spp) %>% 
    summarize(
      tpa = sum(tpa_x_wall),  # the average tpa for each spp is aggregated across points
      tpha = sum(tpha_x_wall),
      wall_denom = mean(wall_denom),
      points_occupied = n_distinct(point),
      age = mean(numb_yrs_grow)
    ) %>% 
  mutate(
    PCT_pts_present = (points_occupied / wall_denom )*100
     ) %>% 
      arrange(harvest, season, TRMT, (desc(tpa)))

  regen_exp_GT_4.5ft_head <- regen_exp_GT_4.5ft_spp_initial %>% 
    slice_head(n=5) %>% 
    mutate(spp_count = 1)
  
  regen_exp_GT_4.5ft_tail <- regen_exp_GT_4.5ft_spp_initial %>% 
    slice_tail(n=-5) %>% 
    filter(tpa >0) %>% 
    mutate(
      spp = "remainder",
      numb = 1
      ) %>% 
    group_by(describ, harvest, season, TRMT, spp) %>% 
    summarize(
      tpa = sum(tpa),  # the average tpa for each spp is aggregated across points
      tpha = sum(tpha),
      wall_denom = mean(wall_denom),
      age = mean(age),
      spp_count = sum(numb)
    )
 
  #this code provides the sum of stem density by harvest, giving totals plus spp summ in next dataframe 
  regen_summ_exp_GT_4.5ft <- bind_rows(regen_exp_GT_4.5ft_head, regen_exp_GT_4.5ft_tail) %>% 
    select(describ, harvest, season, TRMT, spp, tpa, tpha, PCT_pts_present, wall_denom, age, spp_count) %>% 
    group_by(describ, harvest, season, TRMT) %>% 
    summarize(
      tpa = sum(tpa),
      tpha=sum(tpha),
      spp_count = sum(spp_count),
      age = mean(age)
    ) %>% 
    mutate(spp = "TOTAL")

    
regen_exp_GT_4.5ft_spp <- bind_rows(regen_exp_GT_4.5ft_head, regen_exp_GT_4.5ft_tail, regen_summ_exp_GT_4.5ft) %>% 
    select(describ, harvest, season, TRMT, spp, tpa, tpha, PCT_pts_present, wall_denom, age, spp_count) %>% 
    rename(
      TOTAL_tpa = tpa,
      TOTAL_tpha = tpha
    ) %>% 
    #ungroup %>% 
    arrange(describ, harvest, desc(season), TRMT) %>% 
    dplyr::mutate_if(is.numeric, round, 0)  #changes all numeric variable to round to 0 significant digits


#calculations confirmed 2/21/2023

write_csv(regen_exp_GT_4.5ft_spp_initial, "regen_per_acre_exp_GT_4.5ft_spp_composition_tables.csv")
write_csv(regen_exp_GT_4.5ft_spp, "regen_per_acre_exp_GT_4.5ft_spp_5_dom.csv")

spp_count_exp_gt4.5ft <- regen_exp_GT_4.5ft_spp %>% 
  select(describ, harvest, season, TRMT, age, spp, spp_count) %>% 
  filter(spp == "TOTAL") %>% 
  group_by(season, TRMT, age) %>% 
  summarize(
    spp_count = mean(spp_count)
  )

  col03 <- spp_count_exp_gt4.5ft %>% 
    ungroup %>% 
    select(age, TRMT, spp_count) %>% 
    mutate(
      age = as.character(age),
      ) %>%
    mutate(TRMT = fct_recode(TRMT,
      "protected" = "inside",
      "control" = "outside"
    )) %>% 
    group_by(age, TRMT) %>% 
    ggplot()+
    geom_col(aes(x=age, y = spp_count, fill = TRMT)) +  #, position = "dodge2" , fill = harvest
    facet_wrap(~TRMT)+
     theme_pubr() +
    labs(
      x = "Number of Growing Seasons Post Harvest (yrs)",
      y = "Number of Species",
      title = "Average Species Richness for Exposed Stems > 4.5 ft",
      subtitle = "Arnot Forest Slash Walls, summary_regen_veg_data.Rmd, fig=col03"
    )

  set_palette(col03, "uchicago")
  ggsave("figures/species_richness_exp_gt4.5.col03.jpg")
  
  
```


#for exposed desired species, mostly hardwoods
# bc, rm, ro, sm, wa, plus others

```{r}


regen_desired_spp_initial <- regen_all %>% 
  filter(tpa>0) %>% 
  filter(spp == "sm" | spp == "rm" | spp == "ro" | spp == "bc" | spp == "wa"
         | spp == "co" | spp =="wo" | spp == "wp" | spp == "yp" 
         | spp == "ba" | spp == "bh" | spp == "bl" | spp == "sh" 
         | spp == "wp" | spp == "yb") %>% 
  filter(size == "exposed") %>% 
  mutate(
    describ = "desired_exp_all"  #labeling data to be clear about the variables selected
  ) %>% 
  group_by(describ, harvest, season, TRMT, spp) %>% 
    summarize(
      tpa = sum(tpa_x_wall),  # the average tpa for each spp is aggregated across points
      tpha = sum(tpha_x_wall),
      wall_denom = mean(wall_denom),
      points_occupied = n_distinct(point),
      age = mean(numb_yrs_grow)
    ) %>% 
  mutate(
    PCT_pts_present = (points_occupied / wall_denom )*100
     ) %>% 
      arrange(harvest, season, TRMT, (desc(tpa)))

  regen_desired_head <- regen_desired_spp_initial %>% 
    slice_head(n=5) %>% 
    mutate(spp_count = 1)
  
  regen_desired_tail <- regen_desired_spp_initial %>% 
    slice_tail(n=-5) %>% 
    filter(tpa >0) %>% 
    mutate(
      spp = "remainder",
      numb = 1
      ) %>% 
    group_by(describ, harvest, season, TRMT, spp) %>% 
    summarize(
      tpa = sum(tpa),  # the average tpa for each spp is aggregated across points
      tpha = sum(tpha),
      wall_denom = mean(wall_denom),
      age = mean(age),
      spp_count = sum(numb)
    )
 
  #this code provides the sum of stem density by harvest, giving totals plus spp summ in next dataframe 
  regen_summ_desired <- bind_rows(regen_desired_head, regen_desired_tail) %>% 
    select(describ, harvest, season, TRMT, spp, tpa, tpha, PCT_pts_present, wall_denom, age, spp_count) %>% 
    group_by(describ, harvest, season, TRMT) %>% 
    summarize(
      tpa = sum(tpa),
      tpha=sum(tpha),
      spp_count = sum(spp_count)
    ) %>% 
    mutate(spp = "desired total")

    
regen_desired_spp <- bind_rows(regen_desired_head, regen_desired_tail, regen_summ_desired) %>% 
    select(describ, harvest, season, TRMT, spp, tpa, tpha, PCT_pts_present, wall_denom, age, spp_count) %>% 
    rename(
      TOTAL_tpa = tpa,
      TOTAL_tpha = tpha
    ) %>% 
    #ungroup %>% 
    arrange(describ, harvest, desc(season), TRMT) %>% 
    dplyr::mutate_if(is.numeric, round, 0)  #changes all numeric variable to round to 0 significant digits


#calculations confirmed 2/21/2023

write_csv(regen_desired_spp_initial, "regen_per_acre_desired_spp_composition_tables.csv")
write_csv(regen_desired_spp, "regen_per_acre_desired_spp_5_dom.csv")

```


# calculate spp composition for desired species GT 4.5 ft tall that are exposed.
# Not sure how this differs from previous datafile (previous has exposed stems of all heights)

```{r}

# for exposed stems greater than 4.5 ft tall

regen_dsrd_exp_gt4.5ft_spp_initial <- regen_all %>% 
  filter(tpa>0) %>% 
  filter(spp == "sm" | spp == "rm" | spp == "ro" | spp == "bc" | spp == "wa"
         | spp == "co" | spp =="wo" | spp == "wp" | spp == "yp" 
         | spp == "ba" | spp == "bh" | spp == "bl" | spp == "sh" 
         | spp == "wp" | spp == "yb") %>% 
  filter(size == "exp_med" | size == "exp_tall" | size == "exp_sapl") %>% 
  mutate(
    describ = "dsrd_exp_gt4.5ft"  #labeling data to be clear about the variables selected
  ) %>% 
  group_by(describ, harvest, season, TRMT, spp) %>% 
    summarize(
      tpa = sum(tpa_x_wall),  # the average tpa for each spp is aggregated across points
      tpha = sum(tpha_x_wall),
      wall_denom = mean(wall_denom),
      points_occupied = n_distinct(point),
      age = mean(numb_yrs_grow)
    ) %>% 
  mutate(
    PCT_pts_present = (points_occupied / wall_denom )*100
     ) %>% 
      arrange(harvest, season, TRMT, (desc(tpa)))

  regen_dsrd_exp_gt4.5ft_head <- regen_dsrd_exp_gt4.5ft_spp_initial %>% 
    slice_head(n=5) %>% 
    mutate(spp_count = 1)
  
  regen_dsrd_exp_gt4.5ft_tail <- regen_dsrd_exp_gt4.5ft_spp_initial %>% 
    slice_tail(n=-5) %>% 
    filter(tpa >0) %>% 
    mutate(
      spp = "remainder",
      numb = 1
      ) %>% 
    group_by(describ, harvest, season, TRMT, spp) %>% 
    summarize(
      tpa = sum(tpa),  # the average tpa for each spp is aggregated across points
      tpha = sum(tpha),
      wall_denom = mean(wall_denom),
      age = mean(age),
      spp_count = sum(numb)
    )
 
  #this code provides the sum of stem density by harvest, giving totals plus spp summ in next dataframe 
  regen_summ_dsrd_exp_gt4.5ft <- bind_rows(regen_dsrd_exp_gt4.5ft_head, regen_dsrd_exp_gt4.5ft_tail) %>% 
    select(describ, harvest, season, TRMT, spp, tpa, tpha, PCT_pts_present, wall_denom, age, spp_count) %>% 
    group_by(describ, harvest, season, TRMT) %>% 
    summarize(
      tpa = sum(tpa),
      tpha=sum(tpha),
      spp_count = sum(spp_count)
    ) %>% 
    mutate(spp = "dsrd_exp_gt4.5ft")

    
regen_dsrd_exp_GT_4.5ft_spp <- bind_rows(regen_dsrd_exp_gt4.5ft_head, regen_dsrd_exp_gt4.5ft_tail, regen_summ_dsrd_exp_gt4.5ft) %>% 
    select(describ, harvest, season, TRMT, spp, tpa, tpha, PCT_pts_present, wall_denom, age, spp_count) %>% 
    rename(
      TOTAL_tpa = tpa,
      TOTAL_tpha = tpha
    ) %>% 
    #ungroup %>% 
    arrange(describ, harvest, desc(season), TRMT) %>% 
    dplyr::mutate_if(is.numeric, round, 0)  #changes all numeric variable to round to 0 significant digits


#calculations confirmed 2/21/2023

write_csv(regen_dsrd_exp_gt4.5ft_spp_initial, "regen_per_acre_dsrd_exp_gt4.5ft_spp_composition_tables.csv")
write_csv(regen_dsrd_exp_GT_4.5ft_spp, "regen_per_acre_dsrd_exp_gt4.5ft_spp_5_dom.csv")


x1_w_ash <- regen_dsrd_exp_GT_4.5ft_spp %>% 
      filter(!spp %in% c("dsrd_exp_gt4.5ft"))  %>%   
  group_by(harvest, season, TRMT) %>%  
      summarize(
        TOTAL_tpa = sum(TOTAL_tpa),
        age = mean(age, na.rm=TRUE)
      )

x1_wo_ash <- regen_dsrd_exp_GT_4.5ft_spp %>% 
      filter(!spp %in% c("dsrd_exp_gt4.5ft", "wa"))  %>%   
  group_by(harvest, season, TRMT) %>%  
      summarize(
        TOTAL_tpa = sum(TOTAL_tpa),
        age = mean(age, na.rm=TRUE)
      )
ggplot((x1_w_ash),
    mapping = aes(x = age, y = TOTAL_tpa, group = TRMT)) + #(this creates a line for each level of the "group" factor (e.g., TRMT))
    geom_point(mapping = aes(color = harvest), shape = 1) +
    geom_smooth(method = lm, se=FALSE, aes(linetype = as.factor(TRMT))) +
      stat_cor(label.x.npc="left",label.y.npc = "top", 
           method="pearson", r.digits=2, p.digits=4, p.accuracy = 0.001)+
      stat_regline_equation(label.x.npc="center",label.y.npc = "top")+  
          #stat_cor and stat_regline within library(ggpubr)
    theme_classic()+
  labs(
    x = "Number of Growing Seasons Post Harvest",
    y = "Stems per acre",
    title = "Annual Change in Abundance of Exposed Desired Species > 4.5 ft",
    subtitle = "Slash Walls (with ash), summary_regen_veg_data.Rmd, fig=x1"
  )

  ggsave("figures/desired_spp_tpa_w_ash_by_age_fig_x1.jpg")

  x2 <- regen_dsrd_exp_GT_4.5ft_spp
  ggplot((x2),
    mapping = aes(x = age, y = PCT_pts_present, group = TRMT)) + #(this creates a line for each level of the "group" factor (e.g., origin))
    geom_point(mapping = aes(shape = harvest), color = 1) +
    geom_smooth(method = lm, se=FALSE, aes(linetype = as.factor(TRMT))) +
      stat_cor(label.x.npc="left",label.y.npc = "top", 
           method="pearson", r.digits=2, p.digits=4, p.accuracy = 0.001)+
      stat_regline_equation(label.x.npc="center",label.y.npc = "top")+  
          #stat_cor and stat_regline within library(ggpubr)
    theme_classic()+
  labs(
    x = "Number of Growing Seasons Post Harvest (yrs)",
    y = "% Points Occupied by Exposed Desired Stems > 4.5 ft",
    title = "Annual Change in Abundance of Exposed Desired Species > 4.5 ft",
    subtitle = "Slash Walls (with ash), summary_regen_veg_data.Rmd, fig=x2"
  )

  ggsave("figures/desired_spp_PCT-occupied_by_age_fig_x2.jpg")

  
  col01 <- x1_w_ash %>% 
    ungroup %>% 
    select(harvest, age, TRMT, TOTAL_tpa) %>% 
    mutate(
      age = as.character(age),
      ) %>%
    mutate(TRMT = fct_recode(TRMT,
      "protected" = "protected",
      "control" = "control"
    )) %>% 
    group_by(harvest, age, TRMT) %>% 
    ggplot()+
    geom_col(aes(x=age, y = TOTAL_tpa, fill = harvest)) +
    facet_wrap(~TRMT)+
     theme_pubr() +
    labs(
      x = "Number of Growing Seasons Post Harvest (yrs)",
      y = "Stems Per Acre (by Harvest)",
      title = "Annual Change in Abundance of Exposed Desired Species > 4.5 ft",
      subtitle = "Slash Walls (with ash), summary_regen_veg_data.Rmd, fig=col01"
    )

  set_palette(col01, "uchicago")
  ggsave("figures/tpa_exposed_desired_gt4.5_with_ash_fig_col01.jpg")

  
 



```


# it appears the following codes changes the type associated with sweet birch
# such that type = "sb"  and then reruns the summary of GT4.5exp sorted by type


```{r}


regen_sb <- regen_all %>% 
  filter(spp == "sb" ) %>% 
  mutate(type = "sb")

regen_sb_gone <- regen_all %>% 
  filter(!spp == "sb")
  
 regen_type_exp_gt4.5ft_initial <- bind_rows(regen_sb, regen_sb_gone) %>% 
  filter(tpa>0) %>% 
  filter(size == "exp_med" | size == "exp_tall" | size == "exp_sapl") %>% 
  mutate(
    describ = "type_exp_gt4.5ft"  #labeling data to be clear about the variables selected
  ) %>% 
  group_by(describ, harvest, season, TRMT, type) %>% arrange(harvest, season, TRMT, (desc(type))) %>% 
    summarize(
      tpa = sum(tpa_x_wall),  # the average tpa for each spp is aggregated across points
      tpha = sum(tpha_x_wall),
      wall_denom = mean(wall_denom),
      points_occupied = n_distinct(point),
      age = mean(numb_yrs_grow)
    ) %>% 
  mutate(
    PCT_pts_present = (points_occupied / wall_denom )*100
     ) %>% 
      arrange(harvest, age, TRMT, (desc(tpa)))

  regen_type_exp_gt4.5ft_head <- regen_type_exp_gt4.5ft_initial %>% 
    slice_head(n=5) %>% 
    mutate(type_count = 1) %>%   # type_count is nonsensical, the value is always 1 or 5
    ungroup %>% 
    arrange(describ, harvest, desc(season), TRMT, type)

  regen_type_exp_gt4.5ft_tail <- regen_type_exp_gt4.5ft_initial %>% 
    slice_tail(n=-5) %>% 
    mutate(
      type = "remainder",
      numb = 1
      ) %>% 
    group_by(describ, harvest, season, TRMT, type) %>% 
    summarize(
      tpa = sum(tpa),  # the average tpa for each spp is aggregated across points
      tpha = sum(tpha),
      wall_denom = mean(wall_denom),
      age = mean(age),
      type_count = sum(numb)
    ) %>% 
    ungroup %>%   
    arrange(describ, harvest, desc(season), TRMT)

 
  #this code provides the sum of stem density by harvest, giving totals plus spp summ in next dataframe 
  regen_summ_type_exp_gt4.5ft <- bind_rows(regen_type_exp_gt4.5ft_head, regen_type_exp_gt4.5ft_tail) %>% 
    select(describ, harvest, season, TRMT, type, tpa, tpha, PCT_pts_present, wall_denom, age, type_count) %>% 
    group_by(describ, harvest, season, TRMT) %>% 
    summarize(
      tpa = sum(tpa),
      tpha=sum(tpha),
      type_count = sum(type_count)
    ) %>% 
    mutate(type = "SUM")

    
regen_type_exp_GT_4.5ft <- bind_rows(regen_type_exp_gt4.5ft_head, regen_type_exp_gt4.5ft_tail, regen_summ_type_exp_gt4.5ft) %>% 
    select(describ, harvest, season, TRMT, type, tpa, tpha, PCT_pts_present, wall_denom, age, type_count) %>% 
    rename(
      TOTAL_tpa = tpa,
      TOTAL_tpha = tpha
    ) %>% 
    ungroup %>% 
    arrange(describ, harvest, desc(season), TRMT) %>% 
    dplyr::mutate_if(is.numeric, round, 0)  #changes all numeric variable to round to 0 significant digits


#calculations confirmed 2/21/2023

write_csv(regen_type_exp_gt4.5ft_initial, "regen_per_acre_type_exp_gt4.5ft_composition_tables.csv")
write_csv(regen_type_exp_GT_4.5ft, "regen_per_acre_type_exp_gt4.5ft_5_dom.csv")





```


# isolate the oldest year of trmt x harvest and plot the difference between treatments of % occupancy
# use this to illustrate the difference among slash walls and need to pay attention to details.

# https://stackoverflow.com/questions/24237399/how-to-select-the-rows-with-maximum-values-in-each-group-with-dplyr
```{r}
##
## this code does not provide first occurrence data file
##

regen_harvest_occ <- regen_type_exp_gt4.5ft_initial %>% 
    filter(type =="commercial") %>% 
    slice_head(1, by = TRMT, harvest, age) %>% 
    group_by(TRMT, harvest,age) %>% 
    filter(age == max(age))  
    
    #slice(which.max(age))
    #slice_head(n=1)
    #slice_max(age)

```



#
# Summary blending all harvests, averaging harvest values by TYPE
#

## boxplot to show variation among harvests
```{r}
data1 <- regen_dsrd_exp_gt4.5ft_spp_initial %>% 
  filter(age == 3)

data2 <- regen_

boxplot(tpa ~ harvest, data = data1)



```


```{r}

x3 <- regen_type_exp_GT_4.5ft %>% 
      filter(type %in% c("commercial", "diversity"))  %>%   #the input file has sweet birch as a distinct type
  group_by(harvest, season, TRMT) %>%  
      summarize(
        TOTAL_tpa = sum(TOTAL_tpa),
        TOTAL_tpha = sum(TOTAL_tpha),
        age = mean(age, na.rm=TRUE)
      )

x3_averaged <- x3 %>% 
  group_by(age, TRMT) %>% 
  summarize(
    avg_tpa = mean(TOTAL_tpa),
    avg_tpha = mean(TOTAL_tpha),
    age = mean(age)
  ) %>% 
      dplyr::mutate_if(is.numeric, round, 0)  #changes all numeric variable to round to 0 significant digits

x3_HT <- regen_type_exp_GT_4.5ft %>%  #HT = "harvest total"
    filter(!type %in% c("SUM"))  %>%   #the input file has sweet birch as a distinct type
    group_by(harvest, age, season, TRMT) %>% 
    summarize(
    TOTAL_tpa = sum(TOTAL_tpa),
    TOTAL_tpha = sum(TOTAL_tpha),
    age = mean(age)
  )
    


ggplot((x3),
    mapping = aes(x = age, y = TOTAL_tpa, group = TRMT)) + #(this creates a line for each level of the "group" factor (e.g., origin))
    geom_point(mapping = aes(color = harvest), shape = 1) +
    geom_smooth(method = lm, se=FALSE, aes(linetype = as.factor(TRMT))) +
      stat_cor(label.x.npc="left",label.y.npc = "top", 
           method="pearson", r.digits=2, p.digits=4, p.accuracy = 0.001)+
      stat_regline_equation(label.x.npc="center",label.y.npc = "top")+  
          #stat_cor and stat_regline within library(ggpubr)
    theme_classic()+
  labs(
    x = "Number of Growing Seasons Post Harvest",
    y = "Stems per acre",
    title = "Annual Change in Abundance of Exposed Comm and Div Species > 4.5 ft",
    subtitle = "Arnot Forest Slash Walls, summary_regen_veg_data.Rmd, fig=x3"
  )

  ggsave("figures/TPA_commercial_diversity_exp_gt4.5.fig_x3.jpg")

  col04 <- x3_averaged %>% 
    ungroup %>% 
    select(age, TRMT, avg_tpa) %>% 
    mutate(
      age = as.character(age),
      ) %>%
    mutate(TRMT = fct_recode(TRMT,
      "protected" = "protected",
      "control" = "control"
    )) %>% 
    group_by(age, TRMT) %>% 
    ggplot()+
    geom_col(aes(x=age, y = avg_tpa, fill = TRMT)) +
    facet_wrap(~TRMT)+
     theme_pubr() +
    labs(
      x = "Number of Growing Seasons Post Harvest (yrs)",
      y = "Stems Per Acre (average of harvests)",
      title = "Abundance of Exposed Commercial & Diversity Species > 4.5 ft",
      subtitle = "Slash Walls (excludes sweet birch), summary_regen_veg_data.Rmd, fig=col04"
    )

  set_palette(col04, "uchicago")
  ggsave("figures/tpa_comm_diversity_exposed_wo_sweetbirch_gt4.5_fig_col04.jpg")
  
 # graph of total TPA by harvest to illustrate differences, importance of site and wall effectiveness
  # calculations need to be verified, average/summation doesn't appear correct
  col04_HT <- x3_HT %>% 
    ungroup %>% 
    select(harvest, age, TRMT, TOTAL_tpa) %>% 
    mutate(
      age = as.character(age),
      ) %>%
    mutate(TRMT = fct_recode(TRMT,
      "protected" = "protected",
      "control" = "control"
    )) %>% 
    group_by(harvest, age, TRMT) %>% 
    ggplot()+
    geom_col(aes(x=age, y = TOTAL_tpa, fill = harvest)) +
    facet_wrap(~TRMT)+
     theme_pubr() +
    labs(
      x = "Number of Growing Seasons Post Harvest (yrs)",
      y = "Stems Per Acre",
      title = "Cumulative Abundance by Harvest of Exposed Stems > 4.5 ft",
      subtitle = "Arnot Forest Slash Walls, summary_regen_veg_data.Rmd, fig=col04_HT"
    )

  set_palette(col04_HT, "uchicago")
  ggsave("figures/tpa_by_harvest_exposed_gt4.5_fig_col04_HT.jpg")
  
  
  
  
  
  x4 <- regen_type_exp_GT_4.5ft %>% 
#  filter(spp == "ab" | spp == "wa" | spp == "sb" | spp == "rm" | 
#           spp == "ro" | spp == "sm" | spp == "eh") %>% 
#  filter(numb_yrs_grow >3) %>% 
    filter(!type %in% c("SUM", "other"))  %>%   
    group_by(season, TRMT, type) %>% 
  summarize(
    avg_tpa = mean(TOTAL_tpa),
    age = mean(age)
    )


#testing code for two lines per facet copied from z4 (this works!)
ggplot(data=x4, mapping = aes(x=age, y = avg_tpa, group=TRMT)) +
  geom_point(mapping = aes(color = TRMT), shape = 1) +
  geom_smooth(method=lm, se=F, aes(linetype = as.factor(TRMT))) +
  facet_wrap(~type) +
    stat_cor(label.x.npc="left",label.y.npc = "top", 
           method="pearson", r.digits=2, p.digits=4, p.accuracy = 0.001)+
    stat_regline_equation(label.x.npc="center",label.y.npc = "top")+  
          #stat_cor and stat_regline within library(ggpubr)
    theme_classic()+
  #theme_bw() +
  labs(
    x = "Number of Growing Seasons Post Harvest",
    y = "Stems per acre",
    title = "Annual Change in Abundance for Stems by 'Type' > 4.5 ft",
    subtitle = "Arnot Forest Slash Walls, summary_regen_veg_data.Rmd, fig=x4"
  ) 
  
  ggsave("figures/TPA_type_gt4.5.fig x4.jpg")

  
#
# code to provide annual totals by type and TRMT
#
regen_sb2 <- regen_all %>% 
  filter(spp == "sb" ) %>% 
  mutate(type = "sb")

regen_sb2_gone <- regen_all %>% 
  filter(!spp == "sb")

  
 regen_summary_type_exp_gt4.5ft_initial <- regen_type_exp_gt4.5ft_initial %>% 
  filter(tpa>0) %>% 
  mutate(
    describ = "type_summ"  #labeling data to be clear about the variables selected
  ) %>% 
  group_by(describ, age, TRMT, type) %>% arrange(TRMT, age, type) %>% 
    summarize(
      tpa = mean(tpa),  # the average tpa for each spp is aggregated across points
      tpha = mean(tpha),
      wall_denom = sum(wall_denom),  ####  !!  ## this calculation is correct (posted 3march2023)
      points_occupied = sum(points_occupied),
      n = n()
    ) %>% 
  # filter (age > 1) %>%   # only Recknagel has data from year 1
  mutate(
    PCT_pts_present = (points_occupied / wall_denom )*100 ####  !!  ## this calculation is correct (posted 3march2023)
     ) %>% 
   ungroup %>% 
   arrange(age, TRMT, type)

quicktest <- bind_rows(regen_sb2, regen_sb2_gone) %>% 
  filter(numb_yrs_grow == "1", TRMT == "control") %>% 
  filter(tpa>0) %>% 
  filter(size == "exp_med" | size == "exp_tall" | size == "exp_sapl")

 
  #this code provides the sum of stem density by harvest, giving totals plus type sum in next dataframe 
  regen_summ_type_exp_gt4.5ft <- regen_summary_type_exp_gt4.5ft_initial %>% 
    select(describ, age, TRMT, type, tpa, tpha, PCT_pts_present, wall_denom, age) %>% 
    group_by(describ, age, TRMT) %>% 
    summarize(
      tpa = sum(tpa),
      tpha=sum(tpha),
    ) %>% 
    ungroup %>% 
    mutate(type = "total")

    
regen_summary_type_exp_GT_4.5ft <- bind_rows( regen_summary_type_exp_gt4.5ft_initial, regen_summ_type_exp_gt4.5ft  ) %>% 
  arrange(type) %>% 
    ungroup %>% 
    select(describ, age, TRMT, type, tpa, tpha, PCT_pts_present, wall_denom, age) %>% 
    rename(
      averageL_tpa = tpa,
      average_tpha = tpha
    ) %>% 
    arrange(desc(age), TRMT, type) %>% 
    dplyr::mutate_if(is.numeric, round, 0)  #changes all numeric variable to round to 0 significant digits


#calculations confirmed 2/21/2023

write_csv(regen_summary_type_exp_gt4.5ft_initial, "regen_summary_type_exp_gt4.5ft_composition_tables.csv")
write_csv(regen_summary_type_exp_GT_4.5ft, "regen_summary_type_exp_gt4.5ft_5_dom.csv")
  

  col02 <- regen_summary_type_exp_gt4.5ft_initial %>% 
    ungroup %>% 
    select(type, age, TRMT, tpa, tpha, PCT_pts_present) %>% 
    mutate(
      age = as.character(age),
      ) %>%
    mutate(TRMT = fct_recode(TRMT,
      "protected" = "protected",
      "control" = "control"
    )) %>% 
    group_by(type, age, TRMT) %>% 
    filter(!type=="other") %>% 
    ggplot()+
    geom_col(aes(x=age, y = tpa, fill = type)) + #, position = "dodge2"
    facet_wrap(~TRMT)+
     theme_pubr() +
    labs(
      x = "Number of Growing Seasons Post Harvest (yrs)",
      y = "Stems Per Acre",
      title = "Abundance of Exposed Stems > 4.5 ft by 'Type'",
      subtitle = "Slash Walls, summary_regen_veg_data.Rmd, fig=col02"
    )

  set_palette(col02, "uchicago")
  ggsave("figures/type_summ_tpa_by_age_exp_gt4.5.jpg")


  col05 <- regen_summary_type_exp_gt4.5ft_initial %>% 
    ungroup %>% 
    select(type, age, TRMT, tpa, tpha, PCT_pts_present) %>% 
    mutate(
      age = as.character(age),
      ) %>%
    mutate(TRMT = fct_recode(TRMT,
      "protected" = "protected",
      "control" = "control"
    )) %>% 
    group_by(type, age, TRMT) %>% 
    filter(!type=="other") %>% 
    ggplot()+
    geom_col(aes(x=age, y = PCT_pts_present, fill = TRMT), position = "dodge2") + #
    facet_wrap(~type)+
     theme_pubr() +
    labs(
      x = "Number of Growing Seasons Post Harvest (yrs)",
      y = "% Points Occupied (average by harvest)",
      title = "Occupancy of Exposed Stems > 4.5 ft by 'Type'",
      subtitle = "Slash Walls, summary_regen_veg_data.Rmd, fig=col05"
    )

  set_palette(col05, "uchicago")
  ggsave("figures/type_summ_PCT_pts_occupy_by_age_exp_gt4.5.jpg")
  
```


#
# composition for medium and sapl size exposed stems by growing season and type (ignore harvest and spp)
# include all harvests


```{r}





```





  